### 2024-05-27 ms.qmd
Updates

---
docx: 
bibliography: "./bib/refs.bib"
csl: "./bib/biology-letters.csl"
reference-doc: "./bib/tmpl.docx"
execute:
  echo: false
  error: false
  cache: false
  warning: false
link-citations: true
crossref:  
  fig-title: Fig    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Table  # (default is "Table")
editor_options: 
  chunk_output_type: console
---

## Introduction

## Methods

#### Husbandry  
*Breeding colony* -- Lizards tested in out experiment came from a breeding colony established in the lab since 2019. The colony consists of approx. 270 adults of *L. delicata* and 180 adults of and *L. guichenoti*. All lizards are housed in commonal containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure. We provide enclosures with non-stick matting, shelter, and several small water dishes. Lizards are fed approx. 40 mid-size crickets (*Acheta domestica*) per enclosure three days a week, and water is sprayed daily. Crickets are dusted with calcium weekly and multivitamin and calcium biweekly. Room temperatures are set to 22-24ºC, but to ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle, keeping the warm side of enclosures at 32ºC.

*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we placed small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside in one extreme of the communal enclosures (see above) to provided females with a place to lay the eggs. We checked for the presence of eggs in the boxes three days a week. Eggs' length and width were measured with a digital caliper to the nearest 0.1 mm, while mass was recorded with a (OHAUS, Model spx123) digital scale ± 0.001g error. Eggs were assigned a clutch number and a individual identity. Then we treated eggs with CORT or vehicle (see CORT and Temperature manipulation below) and placed them in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite). We covered the cups with cling wrap to retain moisture and then left the eggs in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.

*Hatchlings* -- Incubators were checked three times a week for the presence of hatchlings. Once hatchlings were collected, we measured SVL and Tail Length (TL) with a ruler to the nearest mm, and weighed them with a (OHAUS, Model spx123) digital scale ± 0.001g error. Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) provided with non-stick matting and a small water dish. During this period, they were sprayed water every day and received 3-6 small *A. domestica* crickets three times a week. All care otherwise follows similar protocols to adults (see above).

#### Manipulating Early Thermal and CORT Environments

To test empirically the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A). We first allocated eggs onto one of two different treatments: in the CORT treatment, eggs were topically supplied with 5µL of CORT dissolved in 100% Ethanol (vehicle) at a final (10 pg CORT/mL) concentration; in the Control treatment eggs received an equal volume of the vehicle. CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data). Then, eggs were incubated in one of the two previously mentioned temperature regimes (Cold or Hot) until hatching. This temperatures represent the upper and lower limit of the natural incubation temperatures [@qualls2000post; @cheetham2011embryonic]. The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.

#### Numerical discrimination task

Two weeks before starting the experimen phase (see below), lizards were moved to the experimental arena for acclimatation. The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers provided with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the center (@fig-Methods B). The new enclosures were placed in 7 different racks located in two different rooms. Each rack was associated to its own CCTV systems (device model DVR-HP210475) allowing us to record lizards' behaviour during the experiment. The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack. During acclimatation and throughout the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum*. The temperature of both rooms was set to between 22-24ºC, but we also provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle.

The numerical discrimination test consisted in a series of two-choice trials where lizards where presented with different number of food items. Each choice was placed in one of two sides of a white 3D printed platform (see @fig-Methods B, C) and then covered with a transparent Petri dish to prevent the lizard from obtaining the food. We performed five tests that differed in the ratio between both choices: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4. We never used more than four items because higher numbers are considered to be encoded by different regions of the brain [**REFs**]. The food consisted in frozen small crickets (*A. domestica*) that were thawed and dusted with calcium and multivitamin. We decided to use frozen crickets to avoid movement to affect lizards' behaviour. Lizards were habituated to the frozen crickets by providing them with one per day during one month before the beginning of the experiments. During the acclimatation period, to habituate lizards to the platform and the Petri dish, the frozen cricket that was placed on top of the Petri dish in one of the sides of the platform, simulating the conditions of the experiment (see @fig-Methods B). The side where the food was placed was changed every day to avoid any side bias during the trials. The side where we placed the higher number of crickets was randomised between trials and individuals. We also counterbalanced the amount of individuals per treatment tested to the same type of trial everyday. The crickets used during the experiment were all the same size, but to control for other quantitative traits [**REFs**], we changed the orientation of the crickets as shown in @fig-Methods D. In addition, after the experiments, we made another set of trials to test whether the orientation of the crickets could affect lizards' choice, but we found no effect of orientation (see Supplementary Material).

All tests were done between 10 and 12 in the morning when the lizards werew most active. Each trial was performed daily for five days, between the 17^th^ to the 22^nd^ of May 2024. After the tests, videos were analysed to record: a) the first choice of the lizard (1 if it was the higher number, 0 if not), b) the latency, or time spent by the lizard to make a choice, c) the time spent close to a choice after five minutes of perceiving it, and d) the number of attacks towards a choice for 5 min after perceiving it. We considered a choice when the lizard touched the Petri dish with its snout, by then, we considered that the choice had been perceived. Finally, we considered an attack when the lizard tried to bite the Petri dish.

```{r, fig-Methods}
#| label: fig-Methods
#| fig.cap: "Experimental design of early environment manipulation (**A**), numerical discrimination arena an measures(**B**), measures of the platform used for the experiments (**C**), and types of numerical trial (**D**), where each line indicates the position and orientation of the food items in each test."

knitr::include_graphics("./Others/NUMDIS_FIG_1.png")

```

#### Statistical analyses  
**We performed the analyses for each species separately. We fitted two different Bayesian multilevel models using the brm function from the _brms_ package [@burkner2017brms] using an R extension (version 2.8.2) [@R]. We ran four parallel MCMC chains of 3000 iterations for each model, with a warmup period of 1000 iterations. We modelled correct choices [correct (1) or not (0)] as the response variable, and trial, hormone (CORT versus Control), incubation temperature (Cold versus Hot), and their interaction as fixed factors. We included a random intercept and slope (trial) for each lizard in our models. We also incorporated the clutch identity as a random factor. We used the posterior distributions of parameters from these models to test for differences in learning rate between treatments and species. Learning slopes were obtained using the ‘trial’ estimates and its interaction with hormone and temperature treatments. Slope estimates greater than zero were considered as evidence of learning, while those less or equal to zero were not. We calculated the probability that slopes or contrasts between slopes differed from zero (p~mcmc~) using the posterior distributions. We considered statistical significance if p~mcmc~  < 0.05.**

#### Ethical note
Subjects were collected from a breeding colony kept in the lab since 2019. All animals used in this experiment were provided humane laboratory housing, with thermorregulation opportunity, light, and humidity conditions (see above for details). Lizards in the colony received five-seven crickets per individual three times a week while the subjects in the experiment were fed once per day. All lizards were handled minimally to avoid stress. Experimental procedures and laboratory housing complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33). Upon completion of experimental trials, lizards were kept in their enclosure until they were euthanized following standard humane procedures.

## Results
```{r, sampleSize}
#| label: sampleSize
# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
source(here("R", "func.R"))
#
specie <- c("delicata", "guichenoti")
groups <- c("Red", "Blue")
hormone <- c("CORT", "Control")
temperature <- c("Cold", "Hot")
#
n_list <- list()
#
for(i in 1:length(specie)){
  for(j in 1:length(groups)){
    for(k in 1:length(hormone)){
      for(l in 1:length(temperature)){
        n_list[[paste(specie[i], groups[j], hormone[k], temperature[l], sep = "_")]] <- sample(specie[i], groups[j], hormone[k], temperature[l])
      }
    }
  }
}
```

Originally, we started with 96 lizards, 48 per species and 12 per treatment per species. However, due to natural mortality (n = 11), no completion of the training stage (n = 1), or no motivation during the learning tasks (n = 3; see above), we ended up with a total of 81 lizards. Final sample sizes per treatment and species are disclosed on @tbl-data.

```{r, models}
#| label: models
# Fitting the model and extraction of posteriors for both types of task and species using fit_m function (see func.r in R folder). The result everytime the function is used is a df with the posteriors of the model. The functions saves the model automatically in output/models; and when the parameter refit = FALSE then the posteriors are extracted from the model previously written instead of fitting the model again each time.
source(here("R", "func.R"))
# A) Associative task. Model formula: FC_associative ~ Associative_Trial*cort*temp + (1 + Associative_Trial|lizard_id)
## A.1) L. delicata
deli_asso_red <- fit_m("asso", "deli", "red", refit = FALSE)
write.csv(deli_asso_red, file= "./output/Checking/deli_asso_red.csv")
deli_asso_blue <- fit_m("asso", "deli", "blue", refit = FALSE)
write.csv(deli_asso_blue, file= "./output/Checking/deli_asso_blue.csv")
## A.2) L. guichenoti
guich_asso_red <- fit_m("asso", "guich", "red", refit = FALSE)
write.csv(guich_asso_red, file= "./output/Checking/guich_asso_red.csv")
guich_asso_blue <- fit_m("asso", "guich", "blue", refit = FALSE)
write.csv(guich_asso_blue, file= "./output/Checking/guich_asso_blue.csv")
# B) Reversal task. Model formula: FC_reversal ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)
## B.1) L. delicata
deli_rev_red <- fit_m("rev", "deli", "red", refit = FALSE)
write.csv(deli_rev_red, file= "./output/Checking/deli_rev_red.csv")
deli_rev_blue <- fit_m("rev", "deli", "blue", refit = FALSE)
write.csv(deli_rev_blue, file= "./output/Checking/deli_rev_blue.csv")
## B.2) L. guichenoti
guich_rev_red <- fit_m("rev", "guich", "red", refit = FALSE)
write.csv(guich_rev_red, file= "./output/Checking/guich_rev_red.csv")
guich_rev_blue <- fit_m("rev", "guich", "blue", refit = FALSE)
write.csv(guich_rev_blue, file= "./output/Checking/guich_rev_blue.csv")
```

```{r, resultsRev}
# Rename some of the posteriors and make new estimates for the learning rate for the Reversal task doing the same thing we did in the chunk above.
## 1) L. delicata
### Group = red
drr_CORTCold <- deli_rev_red$b_trial_reversal
drr_ControlCold <- (deli_rev_red$'b_trial_reversal:cortControl' + deli_rev_red$b_trial_reversal)
drr_CORTHot <- (deli_rev_red$'b_trial_reversal:tempHot' + deli_rev_red$b_trial_reversal)
drr_ControlHot <- (deli_rev_red$'b_trial_reversal:cortControl:tempHot' + deli_rev_red$b_trial_reversal + deli_rev_red$'b_trial_reversal:cortControl' + deli_rev_red$'b_trial_reversal:tempHot')
### Group = blue
drb_CORTCold <- deli_rev_blue$b_trial_reversal
drb_ControlCold <- (deli_rev_blue$'b_trial_reversal:cortControl' + deli_rev_blue$b_trial_reversal)
drb_CORTHot <- (deli_rev_blue$'b_trial_reversal:tempHot' + deli_rev_blue$b_trial_reversal)
drb_ControlHot <- (deli_rev_blue$'b_trial_reversal:cortControl:tempHot' + deli_rev_blue$b_trial_reversal + deli_rev_blue$'b_trial_reversal:cortControl' + deli_rev_blue$'b_trial_reversal:tempHot')
## 2) L. guichenoti
### Group = red
grr_CORTCold <- guich_rev_red$b_trial_reversal
grr_ControlCold <- (guich_rev_red$'b_trial_reversal:cortControl' + guich_rev_red$b_trial_reversal)
grr_CORTHot <- (guich_rev_red$'b_trial_reversal:tempHot' + guich_rev_red$b_trial_reversal)
grr_ControlHot <- (guich_rev_red$'b_trial_reversal:cortControl:tempHot' + guich_rev_red$b_trial_reversal + guich_rev_red$'b_trial_reversal:cortControl' + guich_rev_red$'b_trial_reversal:tempHot')
### Group = blue
grb_CORTCold <- guich_rev_blue$b_trial_reversal
grb_ControlCold <- (guich_rev_blue$'b_trial_reversal:cortControl' + guich_rev_blue$b_trial_reversal)
grb_CORTHot <- (guich_rev_blue$'b_trial_reversal:tempHot' + guich_rev_blue$b_trial_reversal)
grb_ControlHot <- (guich_rev_blue$'b_trial_reversal:cortControl:tempHot' + guich_rev_blue$b_trial_reversal + guich_rev_blue$'b_trial_reversal:cortControl' + guich_rev_blue$'b_trial_reversal:tempHot')
```

```{r, tbl-data}
#| label: tbl-data
#| tbl-cap: "Estimates of Associative learning slope for all the different treatments per each task, species and group. Mean shows the aritmetic mean of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean. All p-values were obtained using pmcmc and test the hypothesis that the mean is equal to zero. In bold, those values that are significant (p-value <0.05)"
source(here("R", "func.R"))
#
############################## CREATING BIG DF FOR TABLE ##############################
# Building the vectors for titles of rows and columns
specie <- c("L. delicata", "L. guichenoti")
groups <- c("Red", "Blue")
test <- c("Associative", "Reversal")
treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
values <- c("Mean", "95% CI", "p-value")
# Building the vectors for estimated means, co.intervals(95%), and p-values for the slopes obtained from posteriors. p-values are obtained using pmcmc function (see func.R), assuming a two-tailed test that testes the hypothesis that the value (slopes in this case) is 0.
#First get estimates for both tasks
estimates_asso <- list(
  dar_CORTCold, dar_ControlCold, dar_CORTHot, dar_ControlHot, 
  dab_CORTCold, dab_ControlCold, dab_CORTHot, dab_ControlHot, 
  gar_CORTCold, gar_ControlCold, gar_CORTHot, gar_ControlHot, 
  gab_CORTCold, gab_ControlCold, gab_CORTHot, gab_ControlHot
)
#
estimates_rev <- list(
  drr_CORTCold, drr_ControlCold, drr_CORTHot, drr_ControlHot, 
  drb_CORTCold, drb_ControlCold, drb_CORTHot, drb_ControlHot, 
  grr_CORTCold, grr_ControlCold, grr_CORTHot, grr_ControlHot, 
  grb_CORTCold, grb_ControlCold, grb_CORTHot, grb_ControlHot
)
# Then get the mean, co.intervals(95%), and p-values
asso_mean <- format_dec(sapply(estimates_asso, mean), 3)
asso_interval_025 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.025)), 3)
asso_interval_975 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.975)), 3)
asso_intervals <- paste(asso_interval_025, asso_interval_975, sep = " , ")
asso_pvalue <- format_dec(sapply(estimates_asso, pmcmc), 3)
#
rev_mean <- format_dec(sapply(estimates_rev, mean), 3)
rev_interval_025 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.025)), 3)
rev_interval_975 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.975)), 3)
rev_intervals <- paste(rev_interval_025, rev_interval_975, sep = " , ")
rev_pvalue <- format_dec(sapply(estimates_rev, pmcmc), 3)
#
# Building the df for the associative
asso_df <- data.frame(
  Specie = rep(specie, each = length(groups) * length(treatments)),
  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
  Mean = rep(asso_mean, each = 1),
  CI = rep(asso_intervals, each = 1),
  PValue = rep(asso_pvalue, each = 1),
  Task = rep("Associative", length(asso_mean))
)
# Building the df for the reversal
rev_df <- data.frame(
  Specie = rep(specie, each = length(groups) * length(treatments)),
  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
  Mean = rep(rev_mean, each = 1),
  CI = rep(rev_intervals, each = 1),
  PValue = rep(rev_pvalue, each = 1),
  Task = rep("Reversal", length(rev_mean))
)
# Joining both dfs
table_data <- rbind(asso_df, rev_df)
table_data[, sapply(table_data, is.numeric)] <- lapply(table_data[, sapply(table_data, is.numeric)], function(x) format(x, scientific = FALSE))
#
############################## ADDING SAMPLE SIZE TO DF FOR TABLE ##############################
# Make n_list into a df
n_df <- as.data.frame(do.call(rbind, n_list)) %>%
  rename("n" = V1) %>%
  rownames_to_column("model") %>%
  separate(model, into = c("Specie", "Group", "cort", "temp"), sep = "_") %>%
  unite("Treatment", c("cort", "temp"), sep = "-") %>%
  mutate(Specie = factor(Specie,
                  labels = c(delicata = "L. delicata", guichenoti = "L. guichenoti")),
        Treatment = factor(Treatment,
                   levels = c("CORT-Cold", "Control-Cold", "CORT-Hot","Control-Hot")))
# Merge both dfs, put sample size together with the treatment, and organize the new df to make it look like the table
new_table_data <- merge(table_data, n_df) %>%
  rename('p-value' = 'PValue', '95% CI' = 'CI') %>% #Change the names of the columns for the table
  pivot_wider(names_from = Task, values_from = c(Mean, `95% CI`, `p-value`)) %>% # to split between Asociative and Reversal
  select(Specie, Group, Treatment, Mean_Associative, `95% CI_Associative`, `p-value_Associative`, Mean_Reversal, `95% CI_Reversal`, `p-value_Reversal`, n) %>% #To order the columns in the way I want for the table
  mutate(Specie = factor(Specie,
                  levels = c("L. delicata", "L. guichenoti")),
        Group = factor(Group,
                  levels = c("Red", "Blue")),
        Treatment = factor(Treatment, 
                  levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")))%>%
  arrange(Specie, Group, Treatment) %>% # To arrange the rows the way I want
  unite("Treatment", c("Treatment", "n"), sep = " (n = ") %>%
  mutate(Treatment = paste0(Treatment, ")"))
write.csv(new_table_data, file= "./output/Checking/new_table_data.csv")
#
############################## MAKING THE TABLE ##############################
## Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 fint.size = 10)
# Split the table_data df by task
real_table <- flextable(new_table_data) %>%
    bold(~ `p-value_Associative` < 0.05, ~ `p-value_Associative` + Mean_Associative + `95% CI_Associative`) %>%
    bold(~ `p-value_Reversal` < 0.05, ~ `p-value_Reversal` + Mean_Reversal + `95% CI_Reversal`) %>%
    set_table_properties(width = 1) %>%
    align(align="center", part="all") %>% 
    add_header_row(values = c("", "Associative task", "Reversal task"), colwidths = c(3, 3, 3)) %>%
    set_header_labels(Mean_Associative = "Mean",
                      `95% CI_Associative` = "95% CI",
                      `p-value_Associative` = "p-value",
                      Mean_Reversal = "Mean",
                      `95% CI_Reversal` = "95% CI",
                      `p-value_Reversal` = "p-value") %>%
    italic(j = 1, italic = TRUE, part = "body") %>% # To have names od species in italics
    flextable::compose(i = c(2:8,10:16), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
    flextable::compose(i = c(2:4,6:8,10:12,14:16), j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
    hline(i = c(4,12), j = c(2:9), part = "body") %>% # To make some horizontal lines
    hline(i = c(8), j = c(1:9), part = "body") %>% # To make some horizontal lines
    vline(i = (1:16), j = c(3,6), part = "body") %>% # To make some vertical lines on body
    vline(j=c(3,6), part = "header") %>% # To make some vertical lines on header
    autofit() 
real_table
```

In the reversal tesk, we decided to pool both groups 'Blue' and 'Red' to make the between-treatment comparisons for two reasons. First, we did not find any significant differences caused group on the learning slopes (_L. delicata_: 'Blue' learning slope - 'Red'learning slope = `r format_dec(mean((drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drr_CORTCold + drr_ControlHot)), 3)`, p = `r format_dec(pmcmc((drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drr_CORTCold + drr_ControlHot)), 3)`; _L. guichenoti_: 'Blue' learning slope - 'Red'learning slope = `r format_dec(mean((grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTCold + grr_ControlHot)), 3)`, p = `r format_dec(pmcmc((grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTCold + grr_ControlHot)), 3)`). Second, we were expecting an acquired bias towards the colour assigned in the associative task as part of the design. Nonetheless, we did not find any significant differences between treatments in _L. delicata_ ('Control-Cold' - 'CORT-Cold' = `r format_dec(mean((drr_ControlCold + drb_ControlCold) - (drr_CORTCold + drb_CORTCold)), 3)`, p- value = `r format_dec(pmcmc((drr_ControlCold + drb_ControlCold) - (drr_CORTCold + drb_CORTCold)), 3)`; 'Control-Hot' - 'CORT-Hot' = `r format_dec(mean((drr_ControlHot + drb_ControlHot) - (drr_CORTHot + drb_CORTHot)), 3)`, p- value = `r format_dec(pmcmc((drr_ControlHot + drb_ControlHot) - (drr_CORTHot + drb_CORTHot)), 3)`; 'Control-Hot' - 'Control-Cold' = `r format_dec(mean((drr_ControlHot + drb_ControlHot) - (drr_ControlCold + drb_ControlCold)), 3)`, p-value = `r format_dec(pmcmc((drr_ControlHot + drb_ControlHot) - (drr_ControlCold + drb_ControlCold)), 3)`; 'CORT-Hot' - 'CORT-Cold' = `r format_dec(mean((drr_CORTHot + drb_CORTHot) - (drr_CORTCold + drb_CORTCold)), 3)`, p-value = `r format_dec(pmcmc((drr_CORTHot + drb_CORTHot) - (drr_CORTCold + drb_CORTCold)), 3)`) (see @Figdeli) or _L. guichenoti_ ('Control-Cold' - 'CORT-Cold' = `r format_dec(mean((grr_ControlCold + grb_ControlCold) - (grr_CORTCold + grb_CORTCold)), 3)`, p- value = `r format_dec(pmcmc((grr_ControlCold + grb_ControlCold) - (grr_CORTCold + grb_CORTCold)), 3)`; 'Control-Hot' - 'CORT-Hot' = `r format_dec(mean((grr_ControlHot + grb_ControlHot) - (grr_CORTHot + grb_CORTHot)), 3)`, p- value = `r format_dec(pmcmc((grr_ControlHot + grb_ControlHot) - (grr_CORTHot + grb_CORTHot)), 3)`; 'Control-Hot' - 'Control-Cold' = `r format_dec(mean((grr_ControlHot + grb_ControlHot) - (grr_ControlCold + grb_ControlCold)), 3)`, p-value = `r format_dec(pmcmc((grr_ControlHot + grb_ControlHot) - (grr_ControlCold + grb_ControlCold)), 3)`; 'CORT-Hot' - 'CORT-Cold' = `r format_dec(mean((grr_CORTHot + grb_CORTHot) - (grr_CORTCold + grb_CORTCold)), 3)`, p-value = `r format_dec(pmcmc((grr_CORTHot + grb_CORTHot) - (grr_CORTCold + grb_CORTCold)), 3)`) (see @Figguich). We did not find any effect of incubation temperature (_L. delicata_: 'Hot' learning slope - 'Cold' learning slope = `r format_dec(mean((drr_CORTHot + drr_ControlHot + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drb_CORTCold + drb_ControlCold)),3)`, p-value = `r format_dec(pmcmc((drr_CORTHot + drr_ControlHot + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drb_CORTCold + drb_ControlCold)),3)`; _L. guichenoti_: 'Hot' learning slope - 'Cold' learning slope = `r format_dec(mean((grr_CORTHot + grr_ControlHot + grb_CORTHot + grb_ControlHot) -(grr_CORTCold + grr_ControlCold + grb_CORTCold + grb_ControlCold)),3)`, p-value = `r format_dec(pmcmc((grr_CORTHot + grr_ControlHot + grb_CORTHot + grb_ControlHot) -(grr_CORTCold + grr_ControlCold + grb_CORTCold + grb_ControlCold)),3)`) or the hormone (_L. delicata_: 'Control' learning slope - 'CORT' learning slope = `r format_dec(mean((drr_ControlCold + drr_ControlHot + drb_ControlCold + drb_ControlHot) - (drr_CORTCold + drr_CORTHot + drb_CORTCold + drb_CORTHot)), 3)`, p-value = `r format_dec(pmcmc((drr_ControlCold + drr_ControlHot + drb_ControlCold + drb_ControlHot) - (drr_CORTCold + drr_CORTHot + drb_CORTCold + drb_CORTHot)), 3)`;_L. guichenoti_: 'Control' learning slope - 'CORT' learning slope = `r format_dec(mean((grr_ControlCold + grr_ControlHot + grb_ControlCold + grb_ControlHot)-(grr_CORTCold + grr_CORTHot + grb_CORTCold + grb_CORTHot)), 3)`, p-value = `r format_dec(pmcmc((grr_ControlCold + grr_ControlHot + grb_ControlCold + grb_ControlHot)-(grr_CORTCold + grr_CORTHot + grb_CORTCold + grb_CORTHot)), 3)`) when groups were pooled. Finally, we did not find any significant differences when we compared the estimated slopes between species for the reversal task (_L. delicata_ - _L. guichenoti_ = `r format_dec(mean((drr_CORTCold + drr_ControlCold + drr_CORTHot + drr_ControlHot + drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTHot + grr_ControlHot + grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot)), 3)`, p-value = `r format_dec(pmcmc((drr_CORTCold + drr_ControlCold + drr_CORTHot + drr_ControlHot + drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTHot + grr_ControlHot + grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot)), 3)`).

## Discussion

## References
<div id="refs"></div>

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

# Suplementary Material

#### Colour preference in the Associative task
To test if lizards were biased towards the assigned colour as our preliminary analyses suggested, we employed the intercepts  from our posterior distributions. We first estimated the predicted probability of choosing the correct feeder first in the first trial, by using the formula: 
$$\text{Probability} = \frac{1}{1 + e^{-Intercept}}$$

Second, we tested the hypothesis that the estimated probability was higher than 0.33 (the probability expected by chance of choosing the correct feeder) using pmcmc. If the estimated probability is above 0.33. we consider it as an indication that there was a preference towards that colour that could be affecting learning slopes. The results per treatment are summarized in @tbl-bias. 
```{r, tbl-bias}
#| label: tbl-bias
#| tbl-cap: "Probability of choosing the correct feeder in the first trial when the correct feeder was blue (Prob Blue) or red (Prob Red) for each species and each treatment. p-values were obtained using pmcmc and test the hypothesis that the probability is >0.33. In bold, those values that are significant (p-value <0.05)"
source(here("R", "func.R"))
# First we estimate the probability of choosing right in the first trial using the intercepts from the posteriors
## 1) L. delicata
### Group = red
probright_drCORTCold <- 1/(1+exp(-deli_asso_red$b_Intercept))
probright_drControlCold <- 1/(1+exp(-(deli_asso_red$b_cortControl + deli_asso_red$b_Intercept)))
probright_drCORTHot <- 1/(1+exp(-(deli_asso_red$b_tempHot + deli_asso_red$b_Intercept)))
probright_drControlHot <- 1/(1+exp(-(deli_asso_red$'b_cortControl:tempHot' + deli_asso_red$b_cortControl + deli_asso_red$b_tempHot + deli_asso_red$b_Intercept)))
### Group = blue
probright_dbCORTCold <- 1/(1+exp(-(deli_asso_blue$b_Intercept)))
probright_dbControlCold <- 1/(1+exp(-(deli_asso_blue$b_cortControl + deli_asso_blue$b_Intercept)))
probright_dbCORTHot <- 1/(1+exp(-(deli_asso_blue$b_tempHot + deli_asso_blue$b_Intercept)))
probright_dbControlHot <- 1/(1+exp(-(deli_asso_blue$'b_cortControl:tempHot' + deli_asso_blue$b_cortControl + deli_asso_blue$b_tempHot + deli_asso_blue$b_Intercept)))
## 2) L. guichenoti
### Group = red
probright_grCORTCold <- 1/(1+exp(-guich_asso_red$b_Intercept))
probright_grControlCold <- 1/(1+exp(-(guich_asso_red$b_cortControl + guich_asso_red$b_Intercept)))
probright_grCORTHot <- 1/(1+exp(-(guich_asso_red$b_tempHot + guich_asso_red$b_Intercept)))
probright_grControlHot <- 1/(1+exp(-(guich_asso_red$'b_cortControl:tempHot' + guich_asso_red$b_cortControl + guich_asso_red$b_tempHot + guich_asso_red$b_Intercept)))
### Group = blue
probright_gbCORTCold <- 1/(1+exp(-guich_asso_blue$b_Intercept))
probright_gbControlCold <- 1/(1+exp(-(guich_asso_blue$b_cortControl + guich_asso_blue$b_Intercept)))
probright_gbCORTHot <- 1/(1+exp(-(guich_asso_blue$b_tempHot + guich_asso_blue$b_Intercept)))
probright_gbControlHot <- 1/(1+exp(-(guich_asso_blue$'b_cortControl:tempHot' + guich_asso_blue$b_cortControl + guich_asso_blue$b_tempHot + guich_asso_blue$b_Intercept)))
# Second we build a df with all the mean probabilities of choosing right in the first trial for each treatment and species
prob_df <- data.frame(
  probright_drCORTCold_t = format_dec(mean(probright_drCORTCold), 3),
  probright_drControlCold_t = format_dec(mean(probright_drControlCold), 3),
  probright_drCORTHot_t = format_dec(mean(probright_drCORTHot), 3),
  probright_drControlHot_t = format_dec(mean(probright_drControlHot), 3),
  probright_dbCORTCold_t = format_dec(mean(probright_dbCORTCold), 3),
  probright_dbControlCold_t = format_dec(mean(probright_dbControlCold), 3),
  probright_dbCORTHot_t = format_dec(mean(probright_dbCORTHot), 3),
  probright_dbControlHot_t = format_dec(mean(probright_dbControlHot), 3),
  probright_grCORTCold_t = format_dec(mean(probright_grCORTCold), 3),
  probright_grControlCold_t = format_dec(mean(probright_grControlCold), 3),
  probright_grCORTHot_t = format_dec(mean(probright_grCORTHot), 3),
  probright_grControlHot_t = format_dec(mean(probright_grControlHot), 3),
  probright_gbCORTCold_t = format_dec(mean(probright_gbCORTCold), 3),
  probright_gbControlCold_t = format_dec(mean(probright_gbControlCold), 3),
  probright_gbCORTHot_t = format_dec(mean(probright_gbCORTHot), 3),
  probright_gbControlHot_t = format_dec(mean(probright_gbControlHot), 3)) %>%
gather(key = "column", value = "value", .) %>%
rename("prob_choice" = "value") %>%
data.frame()
write.csv(prob_df, file= "./output/Checking/prob_df.csv")
# Then we build a df with the statistical test (one-tailed pmcmc) that the probability is >0.33 (the probability of choosing right by chance)
comp_pval <- data.frame(
  probright_drCORTCold_t = format_dec(pmcmc(probright_drCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_drControlCold_t = format_dec(pmcmc(probright_drControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_drCORTHot_t = format_dec(pmcmc(probright_drCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_drControlHot_t = format_dec(pmcmc(probright_drControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_dbCORTCold_t = format_dec(pmcmc(probright_dbCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_dbControlCold_t = format_dec(pmcmc(probright_dbControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_dbCORTHot_t = format_dec(pmcmc(probright_dbCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_dbControlHot_t = format_dec(pmcmc(probright_dbControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_grCORTCold_t = format_dec(pmcmc(probright_grCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_grControlCold_t = format_dec(pmcmc(probright_grControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_grCORTHot_t = format_dec(pmcmc(probright_grCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_grControlHot_t = format_dec(pmcmc(probright_grControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_gbCORTCold_t = format_dec(pmcmc(probright_gbCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_gbControlCold_t = format_dec(pmcmc(probright_gbControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_gbCORTHot_t = format_dec(pmcmc(probright_gbCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_gbControlHot_t = format_dec(pmcmc(probright_gbControlHot, null = 0.33, twotail = FALSE, dir=">"), 3)) %>%
gather(key = "column", value = "value", .) %>%
rename("p_value" = "value") %>%
data.frame()
write.csv(comp_pval, file= "./output/Checking/comp_pval.csv")
# Now we merge them and organise them to have the final dt for the table
table_df <- merge(prob_df, comp_pval, by ="column") %>%
  mutate(Specie = gsub(".*_(d|g).*", "\\1",column)) %>%
  mutate(Group = gsub(".*_(dr|db|gr|gb).*", "\\1",column)) %>%
  mutate(Treatment = gsub(".*(CORTCold|ControlCold|CORTHot|ControlHot).*", "\\1",column)) %>% #Last three commands used to get the species, group and treatment from the column names
  mutate(Treatment = factor(Treatment, 
                    levels = c("CORTCold", "ControlCold", "CORTHot", "ControlHot"),
                    labels = c("CORTCold" = "CORT-Cold", "ControlCold" = "Control-Cold", "CORTHot" = "CORT-Hot", "ControlHot" = "Control-Hot"))) %>%
  mutate(Specie = factor(Specie,
                levels = c("d", "g"),
                labels = c("d" = "L. delicata", "g" = "L. guichenoti"))) %>%
  mutate(Group = recode(Group, "dr" = "Red", "db" = "Blue", "gr" = "Red", "gb" = "Blue")) %>%
  select(-column) %>%
  pivot_wider(names_from = Group, values_from = c(prob_choice, p_value)) %>% #To split the df in Red and Blue
  select(Specie, Treatment, prob_choice_Blue, p_value_Blue, prob_choice_Red, p_value_Red) %>%
  arrange(Specie, Treatment) %>% # To arrange the rows the way I wantdata.frame()
data.frame()
write.csv(table_df, file= "./output/Checking/table_df.csv")
#################
colour_table <- flextable(table_df) %>%
  bold(~ `p_value_Blue` < 0.05, ~ `p_value_Blue` + prob_choice_Blue) %>%
  bold(~ `p_value_Red` < 0.05, ~ `p_value_Red` + prob_choice_Red) %>%
  set_table_properties(width = 1) %>%
  align(align="center", part="all") %>% 
  set_header_labels(prob_choice_Blue = "Prob Blue", 
                    p_value_Blue = "p-value Blue", 
                    prob_choice_Red = "Prob Red", 
                    p_value_Red = "p-value Red") %>%
  italic(j = 1, italic = TRUE, part = "body") %>% # To have names of species in italics
  flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  hline(i = c(4), j = c(1:6), part = "body") %>% # To make some horizontal lines
  vline(j = c(2,4), part = "all") %>% # To make some vertical lines on body
autofit()
colour_table
```

On average, we found that, for both species, the proportion of correct choices in the first trial was significantly above chance when the correct feeder was blue (_L. delicata_: mean Prob choice = `r format_dec(mean((probright_dbCORTCold + probright_dbControlCold + probright_dbCORTHot + probright_dbControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_dbCORTCold + probright_dbControlCold + probright_dbCORTHot + probright_dbControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`; _L. guichenoti_: mean Prob choice = `r format_dec(mean((probright_gbCORTCold + probright_gbControlCold + probright_gbCORTHot + probright_gbControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_gbCORTCold + probright_gbControlCold + probright_gbCORTHot + probright_gbControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`), but not when it was red (_L. delicata_: mean Prob choice = `r format_dec(mean((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`; _L. guichenoti_: mean Prob choice = `r format_dec(mean((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`).

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

#### Husbandry  
*Breeding colony* -- We tested juveniles coming from a breeding colony established in the lab since 2019. There is a total of 270 and 180 adults of *L. delicata* and *L. guichenoti* respectively, housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure. Enclosures are provided with non-stick matting, shelter, and several small water dishes. Water is given daily, and they are fed approx. 40 mid-size crickets (*Acheta domestica*) per enclosure three days a week. Crickets are dusted with calcium weekly and multivitamin and calcium biweekly. To ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle. Room temperatures are set to 22-24 Celsius, and warm side of enclosures is usually at 32 Celsius.  
*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we provided females with a place to lay the eggs by means of small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside, that were placed in one extreme of the communal enclosures (see above). We checked for the presence of eggs in the boxes three days a week. After collection, we measured length and width of eggs with a digital caliper to the nearest 0.1 mm and weight them with a (OHAUS, Model spx123) digital scale ± 0.001g error. Then eggs were treated with CORT or vehicle (see CORT and Temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite). The cups were covered with cling wrap to retain moisture and left in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.  
*Hatchlings* -- Eggs in the incubator were checked three times a week for hatchlings. After hatchling, we measured juveniles' SVL and Tail Length (TL) with a rule to the nearest mm and weighted them with a (OHAUS, Model spx123) digital scale ± 0.001g error. We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with non-stick matting and a small water dish. During this period, they were sprayed water every day and received 3-6 small *A. domestica* crickets three times a week. All care otherwise follows similar protocols to adults (see above).  
Two weeks before we started the training phase (see below), lizards were moved to the experimental arena for acclimatation. The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the other. These new enclosures were placed in two rooms in 7 different racks associated to 7 different CCTV systems (device model DVR-HP210475) that allowed us to record their behaviour during the experiment (see details below). The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack. During acclimatation and all the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum*. We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle. The rooms temperature was set to between 22-24 Celsius.

#### CORT and Temperature manipulation  
To test empirically the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A). We first allocated eggs to one of two different treatments: CORT treatment, where eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO (vehicle) at a final (10 pg CORT/mL) concentration (CORT treatment); and a Control treatment, where eggs received an equal volume of the vehicle. CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data). Then, eggs were incubated in one of the two previously mentioned temperature regimes ('Cold' or 'Hot') until hatching. The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.

### Learning task
The habituation phase was divided into three stages: in the first stage, the reward, a small, frozen cricket (*A. domestica*) was placed in an opaque petri dish (3D x 1.6H cm) (@fig-Methods B, Stage 1); in the second stage, the petri dish with the cricket was placed on top of the white 3D printed ramps (@fig-Methods B, Stage 2); and in the third stage, the cricket was left inside a well (3D x 1.75H cm) on top of the ramp (Fig. @fig-Methods B, Stage 3). Trials in every stage lasted one hour, begining when the feeding block (petri dish, ramp, or both) was placed inside the enclosure and finished when we took the feeding block away. We recorded whether the cricket had been consumed or not and we considered an stage as completed if the lizards ate the crickets in at least 5 out of 6 trials. This phase lasted 38 days, and only in one case we decided not to use the lizard in the next phases because its behaviour was not consistent.  
In the associative phase, we used three ramps that were identical in shape and size but differ in the colour (Associative task in Fig. @fig-Methods B). As in the last stage of the habituation phase, we placed the cricket inside the wells of the ramps, but then two of them were covered with 3D-printed lids (3D x 0.5H cm) so prey was only accessible in "the correct" ramp. The food reward was placed in all three wells to avoid lizard using prey chemical cues, that were released in the closed wells thanks to a series of small holes on the top of the lids. The colours of the feeders were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision]. However, we decided to split the lizards in two groups to control for potential colour preference that could bias our results. One group of lizards was assigned the blue ramp as the correct choice (i.e., the ramp with the non-covered frozen cricket) in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal. In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association. Lizards were tested in this task once a day for 35 days. After the colour association phase, we performed a choice reversal task (Reversal task in @fig-Methods B). Here, the attainability of prey was indicated by a different colour from the reversal, requiring the lizards to form a novel association between the new colour and the food reward. This test was done once a day for 40 days.  
We performed the full experiment was between the 6^th^ of March until the 26^th^ of June 2023, between 11 to 12 am, when the lizards were active. Trials in the learning phases (colour associative task and reversal tasks) were recorded with different CCTV systems always using the same camera per individual. Videos were analysed manually using a standard video player (IINA) by PR, who recorded whether the first choice made by each subject was the correct feeder or not. We considered a choice was made if the head of the lizard was inside the well of one of the ramps. We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses. We excluded from our analyses those individuals with more than 15 trials failed (i.e. they did not make a choice), and we considered the first trial to be the first one where the individual made a choice. We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

#### Checking the models plots  
```{r, residuals}
# Chunk for calling all the models and making the residuals
# Associative task
## L. delicata
### Red
mod_dar <- readRDS(here("output/models/asso_deli_red.rds"))
resid_dar <- residuals(mod_dar)
### Blue
mod_dab <- readRDS(here("output/models/asso_deli_blue.rds"))
resid_dab <- residuals(mod_dab)
## L. guichenoti
### Red
mod_gar <- readRDS(here("output/models/asso_guich_red.rds"))
resid_gar <- residuals(mod_gar)
### Blue
mod_gab <- readRDS(here("output/models/asso_guich_blue.rds"))
resid_gab <- residuals(mod_gab)
# Reversal task
## L. delicata
### Red
mod_drr <- readRDS(here("output/models/rev_deli_red.rds"))
resid_drr <- residuals(mod_drr)
### Blue
mod_drb <- readRDS(here("output/models/rev_deli_blue.rds"))
resid_drb <- residuals(mod_drb)
## L. guichenoti
### Red
mod_grr <- readRDS(here("output/models/rev_guich_red.rds"))
resid_grr <- residuals(mod_grr)
### Blue
mod_grb <- readRDS(here("output/models/rev_guich_blue.rds"))
resid_grb <- residuals(mod_grb)
```

Model formula for the reversal task is:  
Choice  ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)  
Plots for the different models of the associative task:  
1.- _L. delicata_  
  1.a.- Red  
```{r , plotmod_drr, out.width="70%", fig.align="center"}
plot(mod_drr)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

  1.b.- Blue  
```{r , plotmod_drb, out.width="70%", fig.align="center"}  
plot(mod_drb)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

2.- _L. guichenoti_  
  2.a.- Red  
```{r , plotmod_grr, out.width="70%", fig.align="center"}
plot(mod_grr)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

  2.b.- Blue  
```{r , plotmod_grb, out.width="70%", fig.align="center"}
plot(mod_grb)
```

### 2024-05-27 .vdoc.r

@@ -1,0 +1,484 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: fig-Methods
+#| fig.cap: "Experimental design of early environment manipulation (**A**), numerical discrimination arena (**B**), and types of test (**C**)."
+
+knitr::include_graphics("./Others/NUMDIS_FIG_1.png")
+
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: sampleSize
+# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
+source(here("R", "func.R"))
+#
+specie <- c("delicata", "guichenoti")
+groups <- c("Red", "Blue")
+hormone <- c("CORT", "Control")
+temperature <- c("Cold", "Hot")
+#
+n_list <- list()
+#
+for(i in 1:length(specie)){
+  for(j in 1:length(groups)){
+    for(k in 1:length(hormone)){
+      for(l in 1:length(temperature)){
+        n_list[[paste(specie[i], groups[j], hormone[k], temperature[l], sep = "_")]] <- sample(specie[i], groups[j], hormone[k], temperature[l])
+      }
+    }
+  }
+}
+#
+#
+#
+#
+#
+#| label: models
+# Fitting the model and extraction of posteriors for both types of task and species using fit_m function (see func.r in R folder). The result everytime the function is used is a df with the posteriors of the model. The functions saves the model automatically in output/models; and when the parameter refit = FALSE then the posteriors are extracted from the model previously written instead of fitting the model again each time.
+source(here("R", "func.R"))
+# A) Associative task. Model formula: FC_associative ~ Associative_Trial*cort*temp + (1 + Associative_Trial|lizard_id)
+## A.1) L. delicata
+deli_asso_red <- fit_m("asso", "deli", "red", refit = FALSE)
+write.csv(deli_asso_red, file= "./output/Checking/deli_asso_red.csv")
[…]

### 2024-06-10 ms.qmd
Updates

@@
 ## Results
+
+**Preliminary table**
+```{r, tbl-data}
+source(here("R", "data_process.R"))
+summary_table <- data_num %>%
+  filter(choice == 1) %>% # Filter to keep only correct choices
+  group_by(trt, test_type) %>% # Group by treatment and test
+  summarise(count = n()) %>% # Count the number of correct choices
+  pivot_wider(names_from = test_type, values_from = count, values_fill = list(count = 0)) %>% # Spread tests to columns
+data.frame()
+write.csv(summary_table, file= "./output/Checking/summary_table.csv")
+data_num_bias <- data_num %>%
+  mutate(
+    actual_choice_lat = case_when(
+      pos_high_lat == "Right" & choice == 1 ~ "Right",
+      pos_high_lat == "Right" & choice == 0 ~ "Left",
+      pos_high_lat == "Left" & choice == 1 ~ "Left",
+      pos_high_lat == "Left" & choice == 0 ~ "Right",
+      TRUE ~ NA_character_
+    ),
+    actual_choice_dist = case_when(
+      pos_high_risk == "Close" & choice == 1 ~ "Close",
+      pos_high_risk == "Close" & choice == 0 ~ "Far",
+      pos_high_risk == "Far" & choice == 1 ~ "Far",
+      pos_high_risk == "Far" & choice == 0 ~ "Close",
+      TRUE ~ NA_character_
+    )
+  )
+summary_biases <- data_num_bias %>%
+  group_by(trt) %>% # Group by treatment and test
+  summarise(
+    right_choices = sum(actual_choice_lat == "Right", na.rm = TRUE),
+    left_choices = sum(actual_choice_lat == "Left", na.rm = TRUE),
+    close_choices = sum(actual_choice_dist == "Close", na.rm = TRUE),
+    far_choices = sum(actual_choice_dist == "Far", na.rm = TRUE)
+  ) %>% # Count the number of correct choices
+data.frame()
+write.csv(summary_biases, file= "./output/Checking/summary_biases.csv")
+```
+
 ```{r, sampleSize}

### 2024-06-10 .vdoc.r

@@ -1,484 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig.cap: "Experimental design of early environment manipulation (**A**), numerical discrimination arena (**B**), and types of test (**C**)."
-
-knitr::include_graphics("./Others/NUMDIS_FIG_1.png")
-
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
-source(here("R", "func.R"))
-#
-specie <- c("delicata", "guichenoti")
-groups <- c("Red", "Blue")
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(i in 1:length(specie)){
-  for(j in 1:length(groups)){
-    for(k in 1:length(hormone)){
-      for(l in 1:length(temperature)){
-        n_list[[paste(specie[i], groups[j], hormone[k], temperature[l], sep = "_")]] <- sample(specie[i], groups[j], hormone[k], temperature[l])
-      }
-    }
-  }
-}
-#
-#
-#
-#
-#
-#| label: models
-# Fitting the model and extraction of posteriors for both types of task and species using fit_m function (see func.r in R folder). The result everytime the function is used is a df with the posteriors of the model. The functions saves the model automatically in output/models; and when the parameter refit = FALSE then the posteriors are extracted from the model previously written instead of fitting the model again each time.
-source(here("R", "func.R"))
-# A) Associative task. Model formula: FC_associative ~ Associative_Trial*cort*temp + (1 + Associative_Trial|lizard_id)
-## A.1) L. delicata
-deli_asso_red <- fit_m("asso", "deli", "red", refit = FALSE)
-write.csv(deli_asso_red, file= "./output/Checking/deli_asso_red.csv")
[…]

### 2024-06-10 R/data_process.R

@@ -2,63 +2,20 @@
 # 1_data_process
 ####################################
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats)
-# A) We extract the lizards we are using from the main data_base
-meta_data <- read.csv("./data/lizard_database.csv") %>%
-  filter(lizard_id %in% c("LG2859_23", "LG2893_23", "LG2882_23", "LG2915_23", "LG2946_23", "LG2917_23", "LG2811_23",
-  "LG3182_23", "LG2975_23", "LG2869_23", "LG2898_23", "LG2897_23", "LG3205_23", "LG2812_23", "LG2934_23", "LG3208_23",
-  "LG3035_23", "LG3151_23", "LG3076_23", "LG2866_23", "LG2831_23", "LG3204_23", "LG2976_23", "LG2894_23", "LG3186_23",
-  "LG2810_23", "LG2947_23", "LG3022_23", "LG2941_23", "LG3150_23", "LG3161_23", "LG2991_23", "LG2880_23", "LG2914_23",
-  "LG2977_23", "LG2923_23", "LG2943_23", "LG2935_23", "LG2844_23", "LG2932_23", "LG3126_23", "LG2846_23", "LG3121_23",
-  "LG3119_23", "LG3168_23", "LG2813_23", "LG3075_23", "LG3175_23", "LG3209_23", "LG3117_23", "LG3134_23", "LG3146_23",
-  "LG3167_23", "LG3077_23", "LG3172_23", "LG3128_23", "LG3002_23", "LG3130_23", "LG2998_23", "LG3059_23", "LG3073_23",
-  "LG3133_23", "LG2817_23", "LG2988_23", "LG3000_23", "LG3019_23", "LG3061_23", "LG3004_23", "LG3058_23", "LG2972_23",
-  "LG2978_23", "LG3020_23", "LG3037_23", "LG2933_23", "LG3012_23", "LG3001_23", "LG2986_23", "LG2999_23")) %>%
-data.frame()
-write.csv(meta_data, "./data/numerical.csv") 
-#
-# B) Clean the meta_data
-# Remove individuals who did not participate (more than 15 NAs), remove trials [36-40] (Only in associative) and split treatment into Temp and Cort
-data_asso <- data %>%
+# Clean the data
+# Remove individuals who did not participate (more than 2 NAs) and split treatment into Temp and Cort
[…]

### 2024-06-18 ms.qmd
Update ms.qmd

@@
 Each trial was performed daily for five days, between the 17^th^ to the 22^nd^ of May 2024.
-After the tests, videos were analysed to record: a) the first choice of the lizard (1 if it was the higher number, 0 if not), b) the latency, or time spent by the lizard to make a choice, c) the time spent close to a choice after five minutes of perceiving it, and d) the number of attacks towards a choice for 5 min after perceiving it.
-We considered a choice when the lizard touched the Petri dish with its snout, by then, we considered that the choice had been perceived.
-Finally, we considered an attack when the lizard tried to bite the Petri dish.
+After the tests, videos were analysed manually using (**PROGRAM**).
+We recorded two main variables: the first choice of the lizard, and the interest shown for the higher amount of food.
+The first choice was obtained after the first interaction with one of the two options: 1 if they interacted with the higher number of items first, 0 if not.
+For us to determine the first choice, we considered that the lizard was interacting with the food when it touched the Petri dish with its snout or when the front half of the body was on top of the platform.
+For five minutes after the first interaction with each option, we recorded the time spent interacting with it; we did this for both amounts of food independently of what was the first choice.
+We measured the interest shown for the higher amount of food by substracting the time spent (for a max of 5 min) in the platform with less crickets to the time spent in the platform with more crickets.
+This gave as a gradient from -300 to +300 where the minimum represents interest in the lower amount of food and the maximum shows a preference towards higher number of crickets, while 0 shows no discrimination between them.
+If there was no interaction with one of the options during the test, we considered the time spent as 0 for that option.
+When there was no interaction with any of the choices, both choice and interest were noted as NA.
+We excluded from the analyses all the individuals that did not participate in less than three of the tests.
 

### 2024-08-18 ms.qmd
Updates

--- 2024-05-27 ms.qmd
@@
 bibliography: "./bib/refs.bib"
-csl: "./bib/biology-letters.csl"
+csl: "./bib/behavioral-ecology-and-sociobiology.csl"
 reference-doc: "./bib/tmpl.docx"
@@
 
+```{r, packages}
+#| label: packages
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS,
+fitdistrplus)
+```
+
 ## Introduction
+
+- Quantitative abilities and its importance
+The knowledge of quantities and the relation among them is one of the highest properties of cognition (Reznikova 2018).
+Among those numerical competences, the simplest ability is to make judgments and comparisons between “more” versus “less” (Reznikova 2007; Stancher et al.
+2015), which is widespread in the animal kingdom (Vonk and Beran 2012; Beran et al.
+2014).
+Sensitivity to relative differences in quantity might be adaptive in several situations (Nieder 2017, 2018).
+For instance, female lions (Panthera leo) seem to use numerical assessment of possible rivals to cope with or avoid social contests (McComb et al.
+1994).
+Moreover, the ability to correctly estimate quantities is also associated with, for example, navigation (Chitka and Geiger 1995), choosing larger groups of potential mates (Carazo et al.
+2009) and/or social companions (Bisazza et al.
+2010), or increasing foraging efficiency (Uller et al.
+2003).
+In this sense, the literature is plenty of studies employing the spontaneous choice approach in many animals (e.g., see Carazo et al.
+2009; Pahl et al.
+2013 for invertebrates; Bisazza et al.
+2010; Gómez-Laplaza and Gerlai 2013; Mehlis et al.
+2015 for fishes; Uller et al.
+2003; Krusche et al.
+2010; Stancher et al.
+2015 for amphibians; Garland et al.
+2012 for birds; Ferkin et al.
+2005; Abramson et al.
+2011; Mahamane et al.
+2014; Cox and Montrose 2016; Plotnik et al.
+2019 for mammals) and also of research that explores numerical abilities with training procedures (e.g., see Gross et al.
+2009; Pahl et al.
+2013; Howard et al.
+2018, 2019, for invertebrates; Bisazza et al.
+2014 for fishes; Nieder 2017 for birds; Kilian et al.
+2003; Addessi et al.
+2008; Vonk and Beran 2012; Vonk 2014; Beran and Parrish 2016; Nieder 2017 for mammals).
+
+- Different ways to test numerical abilities
+There are two main approaches to study numerical abilities: spontaneous choice tests and training procedures.
+In spontaneous choice tests, experimenters present to subjects two groups of biologically relevant stimuli that differ in numerosity, under the assumption that animals should select the most appropriate quantity (e.g.
+the larger amount of food) (Agrillo and Bisazza 2014; Cox and Montrose 2016).
+In training
+procedures, animals must learn a numerical rule to receive a reward (Agrillo and Bisazza 2014; Miletto-Petrazzini et al.
+2018).
+Both experimental approaches have some advantages and disadvantages (reviewed in Agrillo and Bisazza 2014).
+In spontaneous choice tests, animals exhibit their natural behavior, enabling interspecific comparisons and the acquirement of ecological issues related to numerousness (Krusche et al.
+2010; Abramson et al.
+2011).
+However, motivation plays a key-role in spontaneous choice tests (Agrillo and Bisazza 2014), which implies that the absence of results might not be due to the lack of skills to solve the numerical task.
+Also, individual preferences may arise in this kind of procedures and, hence, a large sample of subjects is needed to consider inter-individual variability.
+Finally, it could be difficult, but necessary, to control for continuous variables such as area, density or movement (Agrillo and Bisazza 2014; Mehlis et al.
+2015).
+Regardless, both methods are considered complementary and necessary to draw a firm picture of the numerical abilities of a species (Miletto-
+Petrazzini et al. 2018).
+
+- The role of early environment on the development of cogition
+The conditions in which organisms develop can be altered by the circumstances faced by their parents.
+For instance, exposing threespine sticklebacks females to a model predator during ovulation, affected their offspring’s tedency to use social cues (Feng et al., 2015).
+Under stressful situations, animal’s sensory systems detect and transmit environmental signals to instigate adaptive physiological and behavioural adjustments – responses that are mediated by glucocorticoids (GCs) (Sapolsky et al., 2000).
+Glucocorticoids can be transmitted to offspring
+and influence their phenotype and development t 46 hrough transgenerational effects (Crino et al., 2023).
+In addition to the environments experienced by parents, offspring also experience potentially stressful environmental conditions that can interact with or amplify parental effects.
+For example, human disturbance alters snapping turtles’ nest-site choice, ultimately impacting the temperature at which eggs are incubated (Kolbe & Janzen, 2002).
+Because animals’ responses to abrupt changes in temperature are mediated by GCs (Crino et al., 2023), and this response can affect animals’ decision-making while nesting, GCs transmission and early thermal environment are predicted to interact to shape offspring phenotype.
+Both GCs and temperature are known to exert potent effects on brain development (Amiel et al., 2017; Coomber et al., 1997; Jonson et al., 1976; Zhu et al., 2004) and cognition (Clark et al., 2014; Lui et al., 2017).
+    - The role of stress hormones on the development of numerical abilities
+      NOTHING DONE?
+    - The role of temperature on the development of numerical abilities
+      Port Jackson shark eggs were incubated and reared at current and projected end-of-century temperatures (+ 3 °C).
+We tested juvenile’s learning ability with a quantity discrimination task.
+The mortality rate of sharks reared in warm water was 41.7% compared with no mortality in the present-day sharks.
+Contrary to expectations, our results suggest that surviving hatchlings from the elevated-temperature group took fewer days to reach learning criterion and had a higher proportion of correct choice compared with hatchlings reared under present conditions.
+Additionally, this is the first data suggesting that sharks can discriminate different quantities (Vila-Pouca et al., 2019).
+
+- Numerical abilities in reptiles
+    - Learned num abilities
+    - Spontaneous discrimination
+In contrast, in reptiles, research of quantity discrimination is scarce and contradictory.
+Soldati et al.
+(2017) taught red-footed tortoises (Chelonoidis carbonaria) to associate cards of two colors to different volumes of reward and tested their choices afterward.
+In that succeeding experiment, tortoises showed a clear preference for the color card associated with a greater amount of food, thus concluding that discrimination between quantities is present in this species.
+Similarly, Hermann’s tortoises (Testudo hermanni) also proved to discriminate food by numerosity or size in spontaneous choice tests rating from 0.25 to 0.75 (Gazzola et al.
+2018).
+In contrast, other studies suggest different results; in the Italian wall-lizard (Podarcis siculus), spontaneous choice tests showed no numerical discrimination abilities by this lizard (Miletto-Petrazzini et al.
+2017), although learning procedures demonstrated an ability to discern between small close numbers (2 versus 4) (Miletto-Petrazzini et al.
+2018).
+**Check new papers: Lin in turtles, Szabo in lizards, etc.**
+
+- The role of early environment on reptiles cognition
+Both GCs and temperature are known to exert potent effects on brain development (Amiel et al., 2017; Coomber et al., 1997) and cognition (Clark et al., 2014), but little is known about how they affect numerical discrimination.
+Most studies have reported better performance in different associative learning and motor tasks when incubated at higher temperatures (Amiel et al., 2014; Amiel & Shine, 2012; Clark et al., 2014).
+Studies that have reported cold-incubated lizards to learning faster (Abayarathna & Webb, 2020; Dayananda & Webb, 2017) used incubation temperatures far beyond the natural range in their hot treatments and it is unclear how such conditions relate to wild environments.
+Furthermore, sensorial capacities: **Studies from Burger** | **CORT and cognition in reptiles** | **CORT and brain in reptiles**
+
+- Hypotheses
+    - We expect that lizards incubated at high temperatures will have a better performance in the numerical discrimination task than those incubated at low temperatures.
+    - We expect that lizards incubated with CORT will have a worse performance in the numerical discrimination task than those incubated with the vehicle.
+    - We expect that the effect of CORT on the performance of the numerical discrimination task will be more pronounced in lizards incubated at low temperatures than in those incubated at high temperatures.
+- Predictions:
+    - Lizards incubated at high temperatures will discriminate better quantities even with low ratios (lower latencies, choose large amount more often, more interested in larger amounts - i.e.
+"fail" less tests).
+    - Lizards incubated with CORT will discriminate worse quantities even with high ratios (higher latencies, choices not different from chance, equally interested in both amounts of food - i.e.
+"fail" more tests).
+    - The effect of CORT on the performance of the numerical discrimination task will be more pronounced in lizards incubated at low temperatures than in those incubated at high temperatures (i.e.
+CORT-Cold will "fail" more tests than anyone else).
 
@@
 The colony consists of approx.
-270 adults of *L.
-delicata* and 180 adults of and *L.
-guichenoti*.
-All lizards are housed in commonal containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
-We provide enclosures with non-stick matting, shelter, and several small water dishes.
-Lizards are fed approx.
-40 mid-size crickets (*Acheta domestica*) per enclosure three days a week, and water is sprayed daily.
+180 adults of and _L.
+guichenoti_ housed in commonal containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
+Enclosures are provided with non-stick matting, shelter, and several small water dishes filled daily; and lizards are fed mid-size crickets (*Acheta domestica*) three days a week.
 Crickets are dusted with calcium weekly and multivitamin and calcium biweekly.
@@
 
-*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we placed small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside in one extreme of the communal enclosures (see above) to provided females with a place to lay the eggs.
+*Eggs collection and incubation* -- Between mid-October 2023 to the end of February 2024, we placed small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside in one extreme of the communal enclosures (see above) to provided females with a place to lay the eggs.
 We checked for the presence of eggs in the boxes three days a week.
@@
 
-To test empirically the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A).
+We manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A).
 We first allocated eggs onto one of two different treatments: in the CORT treatment, eggs were topically supplied with 5µL of CORT dissolved in 100% Ethanol (vehicle) at a final (10 pg CORT/mL) concentration; in the Control treatment eggs received an equal volume of the vehicle.
@@
 Then, eggs were incubated in one of the two previously mentioned temperature regimes (Cold or Hot) until hatching.
-This temperatures represent the upper and lower limit of the natural incubation temperatures [@qualls2000post; @cheetham2011embryonic].
+This temperatures represent the upper and lower limit of the natural incubation temperatures [@qualls2000post].
 The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.
@@
 During acclimatation and throughout the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum*.
-The temperature of both rooms was set to between 22-24ºC, but we also provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle.
+The temperature of both rooms was set to between 22-24ºC, but we also provided a temperature gradient by means of a heat cord along the center of the arena, and heat lamps in a 12 h light: 12 h dark cycle.
 
@@
 We decided to use frozen crickets to avoid movement to affect lizards' behaviour.
-Lizards were habituated to the frozen crickets by providing them with one per day during one month before the beginning of the experiments.
-During the acclimatation period, to habituate lizards to the platform and the Petri dish, the frozen cricket that was placed on top of the Petri dish in one of the sides of the platform, simulating the conditions of the experiment (see @fig-Methods B).
+Lizards were fed frozen crickets during one month before the beginning of the experiments for habituation to frozen prey.
+Two weeks before the experiments, to habituate lizards to the platform and the Petri dish, the frozen cricket that was placed on top of the Petri dish in one of the sides of the platform, simulating the conditions of the experiment (see @fig-Methods B).
 The side where the food was placed was changed every day to avoid any side bias during the trials.
-The side where we placed the higher number of crickets was randomised between trials and individuals.
-We also counterbalanced the amount of individuals per treatment tested to the same type of trial everyday.
-The crickets used during the experiment were all the same size, but to control for other quantitative traits [**REFs**], we changed the orientation of the crickets as shown in @fig-Methods D.
-In addition, after the experiments, we made another set of trials to test whether the orientation of the crickets could affect lizards' choice, but we found no effect of orientation (see Supplementary Material).
+However, to further control for that, the side where we placed the higher number of crickets was randomised between trials and individuals.
+We also counterbalanced the amount of individuals per treatment tested in the same type of trial each day.
+The crickets used during the experiment were all the same size, but to control for other quantitative traits
+[**REFs In this regard, we must remark that, in our design, we did not control for other quantity-related continuous cues such as area, volume, density, weight, movement, or contour length that could be used for quantitative judgments (Uller et al.
+2003; Krusche et al.
+2010; Abramson et al.
+2011; Garland et al.
+2012; Vonk and Beran 2012; Gómez-Laplaza and Gerlai 2013; Mehlis et al.
+2015), neither for olfactory cues (Plotnik et al.
+2019).
+These continuous variables usually covariate with quantity, but discrimination among them does not necessarily imply processment of numerical information (Agrillo and Bissaza 2014).**]
+, we changed the orientation of the crickets as shown in @fig-Methods D.
+In addition, after the experiments, we made another set of trials to test whether the orientation of the crickets could affect lizards' choice, but we found no effect of orientation (see Supplementary Material - Prey orientation test).
 
-All tests were done between 10 and 12 in the morning when the lizards werew most active.
+All tests were done between 10 and 12 in the morning when the lizards were most active.
 Each trial was performed daily for five days, between the 17^th^ to the 22^nd^ of May 2024.
-After the tests, videos were analysed to record: a) the first choice of the lizard (1 if it was the higher number, 0 if not), b) the latency, or time spent by the lizard to make a choice, c) the time spent close to a choice after five minutes of perceiving it, and d) the number of attacks towards a choice for 5 min after perceiving it.
-We considered a choice when the lizard touched the Petri dish with its snout, by then, we considered that the choice had been perceived.
-Finally, we considered an attack when the lizard tried to bite the Petri dish.
+We recorded three main variables: i) Latency, defined as the time between the start of the test - when the platform was left in the enclosure - until the first interaction with one of the options; ii) First choice of the lizard, recorded as 1 if the lizard interacted first with the larger amount or 0 if not; and iii) Interest shown for the higher amount of food (detailed below).
+We considered that the lizard was interacting with the food when it touched the Petri dish with its snout or when the front half of the body was on top of the platform.
+For five minutes after the first interaction with one of the options, we recorded the time spent interacting with that option for a maximum of five minutes; we did this for both amounts of food independently of what was the first choice.
+We measured the interest shown for the higher amount of food by substracting the time interacting with the option with less crickets to the time spent interacting with the larger amount of food.
+If there was no interaction with one of the options during the test, we considered the time spent as 0 for that option.
+When there was no interaction with any of the choices, both Latency, First choice, and Interest were noted as NA.
+We also considered Interest as NA in some few cases where the animals managed to lift the dish and eat the prey; Latency and First choice were recorded and employed in the analyses.
+We excluded from the analyses all the individuals that did not make any choice in three or more tests.
 
@@
 #### Statistical analyses  
+INT This gave as a gradient from -300 to +300 where the minimum represents interest in the lower amount of food and the maximum shows a preference towards higher number of crickets, while 0 shows no discrimination between them.
 **We performed the analyses for each species separately.
@@
 ## Results
+
+```{r, data_process}
+#| label: data_process
+# Load the data and process it to get the main dfs
+source(here("R", "data_process.R"))
+```
+
 ```{r, sampleSize}
@@
 #
-specie <- c("delicata", "guichenoti")
-groups <- c("Red", "Blue")
 hormone <- c("CORT", "Control")
 temperature <- c("Cold", "Hot")
+side <- c("Right", "Left")
 #
@@
 #
-for(i in 1:length(specie)){
-  for(j in 1:length(groups)){
 for(k in 1:length(hormone)){
   for(l in 1:length(temperature)){
-        n_list[[paste(specie[i], groups[j], hormone[k], temperature[l], sep = "_")]] <- sample(specie[i], groups[j], hormone[k], temperature[l])
-      }
+    for(i in 1:length(side)){
+      n_list[[paste(hormone[k], temperature[l], side[i], sep = "_")]] <- sample(hormone[k], temperature[l], side [i])
     }
@@
 
-Originally, we started with 96 lizards, 48 per species and 12 per treatment per species.
-However, due to natural mortality (n = 11), no completion of the training stage (n = 1), or no motivation during the learning tasks (n = 3; see above), we ended up with a total of 81 lizards.
-Final sample sizes per treatment and species are disclosed on @tbl-data.
-
 ```{r, models}
 #| label: models
-# Fitting the model and extraction of posteriors for both types of task and species using fit_m function (see func.r in R folder).
-The result everytime the function is used is a df with the posteriors of the model.
-The functions saves the model automatically in output/models; and when the parameter refit = FALSE then the posteriors are extracted from the model previously written instead of fitting the model again each time.
-source(here("R", "func.R"))
-# A) Associative task. Model formula: FC_associative ~ Associative_Trial*cort*temp + (1 + Associative_Trial|lizard_id)
-## A.1) L. delicata
-deli_asso_red <- fit_m("asso", "deli", "red", refit = FALSE)
-write.csv(deli_asso_red, file= "./output/Checking/deli_asso_red.csv")
-deli_asso_blue <- fit_m("asso", "deli", "blue", refit = FALSE)
-write.csv(deli_asso_blue, file= "./output/Checking/deli_asso_blue.csv")
-## A.2) L. guichenoti
-guich_asso_red <- fit_m("asso", "guich", "red", refit = FALSE)
-write.csv(guich_asso_red, file= "./output/Checking/guich_asso_red.csv")
-guich_asso_blue <- fit_m("asso", "guich", "blue", refit = FALSE)
-write.csv(guich_asso_blue, file= "./output/Checking/guich_asso_blue.csv")
-# B) Reversal task. Model formula: FC_reversal ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)
-## B.1) L. delicata
-deli_rev_red <- fit_m("rev", "deli", "red", refit = FALSE)
-write.csv(deli_rev_red, file= "./output/Checking/deli_rev_red.csv")
-deli_rev_blue <- fit_m("rev", "deli", "blue", refit = FALSE)
-write.csv(deli_rev_blue, file= "./output/Checking/deli_rev_blue.csv")
-## B.2) L. guichenoti
-guich_rev_red <- fit_m("rev", "guich", "red", refit = FALSE)
-write.csv(guich_rev_red, file= "./output/Checking/guich_rev_red.csv")
-guich_rev_blue <- fit_m("rev", "guich", "blue", refit = FALSE)
-write.csv(guich_rev_blue, file= "./output/Checking/guich_rev_blue.csv")
+# Fitting the model and extraction of posteriors 
+refit = FALSE
+if(refit){
+  # Get the formula and family of variables
+  lat <- bf(latency ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = negbinomial())
+  choice <- bf(choice ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = bernoulli(link = "logit"))
+  int <- bf(compared_interest ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch),
+    family = student())
+  # Fit the model
+  model <- brm(lat + choice + int,
+                chains = 4, cores = 4, iter = 3000, warmup = 1000, control = list(adapt_delta = 0.99),
+                data = data_num)
+    # Write the model to a file
+    saveRDS(model, file = here("output/models/model_num.rds"))
+  } else {
+      # Read the model from a file
+      model <- readRDS(file = here("output/models/model_num.rds"))
+  } 
+  # Extract posteriors
+posteriors <- as_draws_df(model)
 ```
 
-```{r, resultsRev}
-# Rename some of the posteriors and make new estimates for the learning rate for the Reversal task doing the same thing we did in the chunk above.
-## 1) L. delicata
-### Group = red
-drr_CORTCold <- deli_rev_red$b_trial_reversal
-drr_ControlCold <- (deli_rev_red$'b_trial_reversal:cortControl' + deli_rev_red$b_trial_reversal)
-drr_CORTHot <- (deli_rev_red$'b_trial_reversal:tempHot' + deli_rev_red$b_trial_reversal)
-drr_ControlHot <- (deli_rev_red$'b_trial_reversal:cortControl:tempHot' + deli_rev_red$b_trial_reversal + deli_rev_red$'b_trial_reversal:cortControl' + deli_rev_red$'b_trial_reversal:tempHot')
-### Group = blue
-drb_CORTCold <- deli_rev_blue$b_trial_reversal
-drb_ControlCold <- (deli_rev_blue$'b_trial_reversal:cortControl' + deli_rev_blue$b_trial_reversal)
-drb_CORTHot <- (deli_rev_blue$'b_trial_reversal:tempHot' + deli_rev_blue$b_trial_reversal)
-drb_ControlHot <- (deli_rev_blue$'b_trial_reversal:cortControl:tempHot' + deli_rev_blue$b_trial_reversal + deli_rev_blue$'b_trial_reversal:cortControl' + deli_rev_blue$'b_trial_reversal:tempHot')
-## 2) L. guichenoti
-### Group = red
-grr_CORTCold <- guich_rev_red$b_trial_reversal
-grr_ControlCold <- (guich_rev_red$'b_trial_reversal:cortControl' + guich_rev_red$b_trial_reversal)
-grr_CORTHot <- (guich_rev_red$'b_trial_reversal:tempHot' + guich_rev_red$b_trial_reversal)
-grr_ControlHot <- (guich_rev_red$'b_trial_reversal:cortControl:tempHot' + guich_rev_red$b_trial_reversal + guich_rev_red$'b_trial_reversal:cortControl' + guich_rev_red$'b_trial_reversal:tempHot')
-### Group = blue
-grb_CORTCold <- guich_rev_blue$b_trial_reversal
-grb_ControlCold <- (guich_rev_blue$'b_trial_reversal:cortControl' + guich_rev_blue$b_trial_reversal)
-grb_CORTHot <- (guich_rev_blue$'b_trial_reversal:tempHot' + guich_rev_blue$b_trial_reversal)
-grb_ControlHot <- (guich_rev_blue$'b_trial_reversal:cortControl:tempHot' + guich_rev_blue$b_trial_reversal + guich_rev_blue$'b_trial_reversal:cortControl' + guich_rev_blue$'b_trial_reversal:tempHot')
+```{r, resultsorg}
+#| label: resultsorg
+# Organise the estimated results based on the posteriors
+##
+#
+# If I want to get the probability of choosing right per treatment and trial from the posteriors, I will need to take Intercept of each ch treatment and add it, then get the trial-treatment slope, and then model each probability as:
+# P = exp(intercept + slope*trial) / (1 + exp(intercept + slope*trial))
+# And to estimate the Latency it will have to be the same but
+# Lat = intercept * exp(slope*trial)
+# What is trial here?
+# What should I do to get the estimated compared_interest if it follows a t student distribution?
+# Would be:
+# Estimated interest = intercept + slope * trial 
+# correct?
+# How do I get each Intercept for each treatment/trial? And the slope?
+# 
+# E.g.: Getting estimated Latency for CORT-Hot in the 1 VS 3 trial (base level is CORT-Cold test 1 VS 4)
+# Intercept = posteriors$b_latency_Intercept + posteriors$b_latency_test_type1VS3 + posteriors$b_latency_tempHot 
+# Slope = b_latency_test_type1VS3:tempHot
+# Estimated Latency = Intercept * exp(Slope * Trial?)
+# 
 ```
@@
 source(here("R", "func.R"))
-#
-############################## CREATING BIG DF FOR TABLE ##############################
-# Building the vectors for titles of rows and columns
-specie <- c("L. delicata", "L. guichenoti")
-groups <- c("Red", "Blue")
-test <- c("Associative", "Reversal")
-treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
-values <- c("Mean", "95% CI", "p-value")
-# Building the vectors for estimated means, co.intervals(95%), and p-values for the slopes obtained from posteriors.
-p-values are obtained using pmcmc function (see func.R), assuming a two-tailed test that testes the hypothesis that the value (slopes in this case) is 0.
-#First get estimates for both tasks
-estimates_asso <- list(
-  dar_CORTCold, dar_ControlCold, dar_CORTHot, dar_ControlHot, 
-  dab_CORTCold, dab_ControlCold, dab_CORTHot, dab_ControlHot, 
-  gar_CORTCold, gar_ControlCold, gar_CORTHot, gar_ControlHot, 
-  gab_CORTCold, gab_ControlCold, gab_CORTHot, gab_ControlHot
-)
-#
-estimates_rev <- list(
-  drr_CORTCold, drr_ControlCold, drr_CORTHot, drr_ControlHot, 
-  drb_CORTCold, drb_ControlCold, drb_CORTHot, drb_ControlHot, 
-  grr_CORTCold, grr_ControlCold, grr_CORTHot, grr_ControlHot, 
-  grb_CORTCold, grb_ControlCold, grb_CORTHot, grb_ControlHot
-)
-# Then get the mean, co.intervals(95%), and p-values
-asso_mean <- format_dec(sapply(estimates_asso, mean), 3)
-asso_interval_025 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.025)), 3)
-asso_interval_975 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.975)), 3)
-asso_intervals <- paste(asso_interval_025, asso_interval_975, sep = " , ")
-asso_pvalue <- format_dec(sapply(estimates_asso, pmcmc), 3)
-#
-rev_mean <- format_dec(sapply(estimates_rev, mean), 3)
-rev_interval_025 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.025)), 3)
-rev_interval_975 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.975)), 3)
-rev_intervals <- paste(rev_interval_025, rev_interval_975, sep = " , ")
-rev_pvalue <- format_dec(sapply(estimates_rev, pmcmc), 3)
-#
-# Building the df for the associative
-asso_df <- data.frame(
-  Specie = rep(specie, each = length(groups) * length(treatments)),
-  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
-  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
-  Mean = rep(asso_mean, each = 1),
-  CI = rep(asso_intervals, each = 1),
-  PValue = rep(asso_pvalue, each = 1),
-  Task = rep("Associative", length(asso_mean))
-)
-# Building the df for the reversal
-rev_df <- data.frame(
-  Specie = rep(specie, each = length(groups) * length(treatments)),
-  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
-  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
-  Mean = rep(rev_mean, each = 1),
-  CI = rep(rev_intervals, each = 1),
-  PValue = rep(rev_pvalue, each = 1),
-  Task = rep("Reversal", length(rev_mean))
-)
-# Joining both dfs
-table_data <- rbind(asso_df, rev_df)
-table_data[, sapply(table_data, is.numeric)] <- lapply(table_data[, sapply(table_data, is.numeric)], function(x) format(x, scientific = FALSE))
-#
-############################## ADDING SAMPLE SIZE TO DF FOR TABLE ##############################
-# Make n_list into a df
-n_df <- as.data.frame(do.call(rbind, n_list)) %>%
-  rename("n" = V1) %>%
-  rownames_to_column("model") %>%
-  separate(model, into = c("Specie", "Group", "cort", "temp"), sep = "_") %>%
-  unite("Treatment", c("cort", "temp"), sep = "-") %>%
-  mutate(Specie = factor(Specie,
-                  labels = c(delicata = "L. delicata", guichenoti = "L. guichenoti")),
-        Treatment = factor(Treatment,
-                   levels = c("CORT-Cold", "Control-Cold", "CORT-Hot","Control-Hot")))
-# Merge both dfs, put sample size together with the treatment, and organize the new df to make it look like the table
-new_table_data <- merge(table_data, n_df) %>%
-  rename('p-value' = 'PValue', '95% CI' = 'CI') %>% #Change the names of the columns for the table
-  pivot_wider(names_from = Task, values_from = c(Mean, `95% CI`, `p-value`)) %>% # to split between Asociative and Reversal
-  select(Specie, Group, Treatment, Mean_Associative, `95% CI_Associative`, `p-value_Associative`, Mean_Reversal, `95% CI_Reversal`, `p-value_Reversal`, n) %>% #To order the columns in the way I want for the table
-  mutate(Specie = factor(Specie,
-                  levels = c("L. delicata", "L. guichenoti")),
-        Group = factor(Group,
-                  levels = c("Red", "Blue")),
-        Treatment = factor(Treatment, 
-                  levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")))%>%
-  arrange(Specie, Group, Treatment) %>% # To arrange the rows the way I want
-  unite("Treatment", c("Treatment", "n"), sep = " (n = ") %>%
-  mutate(Treatment = paste0(Treatment, ")"))
-write.csv(new_table_data, file= "./output/Checking/new_table_data.csv")
 #
@@
 
-In the reversal tesk, we decided to pool both groups 'Blue' and 'Red' to make the between-treatment comparisons for two reasons.
-First, we did not find any significant differences caused group on the learning slopes (_L.
-delicata_: 'Blue' learning slope - 'Red'learning slope = `r format_dec(mean((drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drr_CORTCold + drr_ControlHot)), 3)`, p = `r format_dec(pmcmc((drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drr_CORTCold + drr_ControlHot)), 3)`; _L.
-guichenoti_: 'Blue' learning slope - 'Red'learning slope = `r format_dec(mean((grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTCold + grr_ControlHot)), 3)`, p = `r format_dec(pmcmc((grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTCold + grr_ControlHot)), 3)`).
-Second, we were expecting an acquired bias towards the colour assigned in the associative task as part of the design.
-Nonetheless, we did not find any significant differences between treatments in _L.
-delicata_ ('Control-Cold' - 'CORT-Cold' = `r format_dec(mean((drr_ControlCold + drb_ControlCold) - (drr_CORTCold + drb_CORTCold)), 3)`, p- value = `r format_dec(pmcmc((drr_ControlCold + drb_ControlCold) - (drr_CORTCold + drb_CORTCold)), 3)`; 'Control-Hot' - 'CORT-Hot' = `r format_dec(mean((drr_ControlHot + drb_ControlHot) - (drr_CORTHot + drb_CORTHot)), 3)`, p- value = `r format_dec(pmcmc((drr_ControlHot + drb_ControlHot) - (drr_CORTHot + drb_CORTHot)), 3)`; 'Control-Hot' - 'Control-Cold' = `r format_dec(mean((drr_ControlHot + drb_ControlHot) - (drr_ControlCold + drb_ControlCold)), 3)`, p-value = `r format_dec(pmcmc((drr_ControlHot + drb_ControlHot) - (drr_ControlCold + drb_ControlCold)), 3)`; 'CORT-Hot' - 'CORT-Cold' = `r format_dec(mean((drr_CORTHot + drb_CORTHot) - (drr_CORTCold + drb_CORTCold)), 3)`, p-value = `r format_dec(pmcmc((drr_CORTHot + drb_CORTHot) - (drr_CORTCold + drb_CORTCold)), 3)`) (see @Figdeli) or _L.
-guichenoti_ ('Control-Cold' - 'CORT-Cold' = `r format_dec(mean((grr_ControlCold + grb_ControlCold) - (grr_CORTCold + grb_CORTCold)), 3)`, p- value = `r format_dec(pmcmc((grr_ControlCold + grb_ControlCold) - (grr_CORTCold + grb_CORTCold)), 3)`; 'Control-Hot' - 'CORT-Hot' = `r format_dec(mean((grr_ControlHot + grb_ControlHot) - (grr_CORTHot + grb_CORTHot)), 3)`, p- value = `r format_dec(pmcmc((grr_ControlHot + grb_ControlHot) - (grr_CORTHot + grb_CORTHot)), 3)`; 'Control-Hot' - 'Control-Cold' = `r format_dec(mean((grr_ControlHot + grb_ControlHot) - (grr_ControlCold + grb_ControlCold)), 3)`, p-value = `r format_dec(pmcmc((grr_ControlHot + grb_ControlHot) - (grr_ControlCold + grb_ControlCold)), 3)`; 'CORT-Hot' - 'CORT-Cold' = `r format_dec(mean((grr_CORTHot + grb_CORTHot) - (grr_CORTCold + grb_CORTCold)), 3)`, p-value = `r format_dec(pmcmc((grr_CORTHot + grb_CORTHot) - (grr_CORTCold + grb_CORTCold)), 3)`) (see @Figguich).
-We did not find any effect of incubation temperature (_L.
-delicata_: 'Hot' learning slope - 'Cold' learning slope = `r format_dec(mean((drr_CORTHot + drr_ControlHot + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drb_CORTCold + drb_ControlCold)),3)`, p-value = `r format_dec(pmcmc((drr_CORTHot + drr_ControlHot + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drb_CORTCold + drb_ControlCold)),3)`; _L.
-guichenoti_: 'Hot' learning slope - 'Cold' learning slope = `r format_dec(mean((grr_CORTHot + grr_ControlHot + grb_CORTHot + grb_ControlHot) -(grr_CORTCold + grr_ControlCold + grb_CORTCold + grb_ControlCold)),3)`, p-value = `r format_dec(pmcmc((grr_CORTHot + grr_ControlHot + grb_CORTHot + grb_ControlHot) -(grr_CORTCold + grr_ControlCold + grb_CORTCold + grb_ControlCold)),3)`) or the hormone (_L.
-delicata_: 'Control' learning slope - 'CORT' learning slope = `r format_dec(mean((drr_ControlCold + drr_ControlHot + drb_ControlCold + drb_ControlHot) - (drr_CORTCold + drr_CORTHot + drb_CORTCold + drb_CORTHot)), 3)`, p-value = `r format_dec(pmcmc((drr_ControlCold + drr_ControlHot + drb_ControlCold + drb_ControlHot) - (drr_CORTCold + drr_CORTHot + drb_CORTCold + drb_CORTHot)), 3)`;_L.
-guichenoti_: 'Control' learning slope - 'CORT' learning slope = `r format_dec(mean((grr_ControlCold + grr_ControlHot + grb_ControlCold + grb_ControlHot)-(grr_CORTCold + grr_CORTHot + grb_CORTCold + grb_CORTHot)), 3)`, p-value = `r format_dec(pmcmc((grr_ControlCold + grr_ControlHot + grb_ControlCold + grb_ControlHot)-(grr_CORTCold + grr_CORTHot + grb_CORTCold + grb_CORTHot)), 3)`) when groups were pooled.
-Finally, we did not find any significant differences when we compared the estimated slopes between species for the reversal task (_L.
-delicata_ - _L.
-guichenoti_ = `r format_dec(mean((drr_CORTCold + drr_ControlCold + drr_CORTHot + drr_ControlHot + drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTHot + grr_ControlHot + grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot)), 3)`, p-value = `r format_dec(pmcmc((drr_CORTCold + drr_ControlCold + drr_CORTHot + drr_ControlHot + drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTHot + grr_ControlHot + grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot)), 3)`).
+## Discussion
+- Animals do not discriminate independently of the treatment or the ratio of the numerical task.
+  - Other studies about discrimination
+  - The importance of context
+    - Learned VS spontaneous discrimination
+There are two main approaches to study numerical abilities: spontaneous choice tests and training procedures.
+In spontaneous choice tests, experimenters present to subjects two groups of biologically relevant stimuli that differ in numerosity, under the assumption that animals should select the most appropriate quantity (e.g.
+the larger amount of food) (Agrillo and Bisazza 2014; Cox and Montrose 2016).
+In training
+procedures, animals must learn a numerical rule to receive a reward (Agrillo and Bisazza 2014; Miletto-Petrazzini et al.
+2018).
+Both experimental approaches have some advantages and disadvantages (reviewed in Agrillo and Bisazza 2014).
+In spontaneous choice tests, animals exhibit their natural behavior, enabling interspecific comparisons and the acquirement of ecological issues related to numerousness (Krusche et al.
+2010; Abramson et al.
+2011).
+However, motivation plays a key-role in spontaneous choice tests (Agrillo and Bisazza 2014), which implies that the absence of results might not be due to the lack of skills to solve the numerical task.
+Also, individual preferences may arise in this kind of procedures and, hence, a large sample of subjects is needed to consider inter-individual variability.
+Finally, it could be difficult, but necessary, to control for continuous variables such as area, density or movement (Agrillo and Bisazza 2014; Mehlis et al.
+2015).
+Regardless, both methods are considered complementary and necessary to draw a firm picture of the numerical abilities of a species (Miletto-
+Petrazzini et al.
+2018).
+the Italian wall-lizard (Podarcis siculus), spontaneous choice tests showed no numerical discrimination abilities by this lizard (Miletto-Petrazzini et al.
+2017), although learning procedures demonstrated an ability to discern between small close numbers (2 versus 4) (Miletto-Petrazzini et al.
+2018)
+    - Ecological context: food VS predators/conspecifics/etc 
+Something with fishes, food, and conspecifics (**check our paper**)
 
-## Discussion
+- Treatment did not affect the discrimination capacity or the decision making
+  - Other studies showing the effect of temperature on discrimination
+  - Other studies about the effects of hormones on discrimination
+  - Studies showing the effect of temperature and CORT on lizards' cognition
+  - The 'two systems'
+Researchers have suggested that numerousness can be processed by two types of systems: one system for small sets (≤ 4) that works by tracking and encoding objects as individual entities; and an analogous system that operates with large numerosities by estimation processes (for a review see Cantlon et al.
+2009; Hyde 2011).
+**Are they encoded in different regions?
+Brin-nuclei-dependent effects of CORT and temp**
+
+- Conclusions
 
@@
 
-#### Colour preference in the Associative task
-To test if lizards were biased towards the assigned colour as our preliminary analyses suggested, we employed the intercepts  from our posterior distributions.
-We first estimated the predicted probability of choosing the correct feeder first in the first trial, by using the formula:
-$$\text{Probability} = \frac{1}{1 + e^{-Intercept}}$$
+#### Treatments effects
+Table S1.- statistical comparisson between treatments
 
-Second, we tested the hypothesis that the estimated probability was higher than 0.33 (the probability expected by chance of choosing the correct feeder) using pmcmc.
-If the estimated probability is above 0.33.
-we consider it as an indication that there was a preference towards that colour that could be affecting learning slopes.
-The results per treatment are summarized in @tbl-bias.
-```{r, tbl-bias}
-#| label: tbl-bias
-#| tbl-cap: "Probability of choosing the correct feeder in the first trial when the correct feeder was blue (Prob Blue) or red (Prob Red) for each species and each treatment.
-p-values were obtained using pmcmc and test the hypothesis that the probability is >0.33.
-In bold, those values that are significant (p-value <0.05)"
-source(here("R", "func.R"))
-# First we estimate the probability of choosing right in the first trial using the intercepts from the posteriors
-## 1) L. delicata
-### Group = red
-probright_drCORTCold <- 1/(1+exp(-deli_asso_red$b_Intercept))
-probright_drControlCold <- 1/(1+exp(-(deli_asso_red$b_cortControl + deli_asso_red$b_Intercept)))
-probright_drCORTHot <- 1/(1+exp(-(deli_asso_red$b_tempHot + deli_asso_red$b_Intercept)))
-probright_drControlHot <- 1/(1+exp(-(deli_asso_red$'b_cortControl:tempHot' + deli_asso_red$b_cortControl + deli_asso_red$b_tempHot + deli_asso_red$b_Intercept)))
-### Group = blue
-probright_dbCORTCold <- 1/(1+exp(-(deli_asso_blue$b_Intercept)))
-probright_dbControlCold <- 1/(1+exp(-(deli_asso_blue$b_cortControl + deli_asso_blue$b_Intercept)))
-probright_dbCORTHot <- 1/(1+exp(-(deli_asso_blue$b_tempHot + deli_asso_blue$b_Intercept)))
-probright_dbControlHot <- 1/(1+exp(-(deli_asso_blue$'b_cortControl:tempHot' + deli_asso_blue$b_cortControl + deli_asso_blue$b_tempHot + deli_asso_blue$b_Intercept)))
-## 2) L. guichenoti
-### Group = red
-probright_grCORTCold <- 1/(1+exp(-guich_asso_red$b_Intercept))
-probright_grControlCold <- 1/(1+exp(-(guich_asso_red$b_cortControl + guich_asso_red$b_Intercept)))
-probright_grCORTHot <- 1/(1+exp(-(guich_asso_red$b_tempHot + guich_asso_red$b_Intercept)))
-probright_grControlHot <- 1/(1+exp(-(guich_asso_red$'b_cortControl:tempHot' + guich_asso_red$b_cortControl + guich_asso_red$b_tempHot + guich_asso_red$b_Intercept)))
-### Group = blue
-probright_gbCORTCold <- 1/(1+exp(-guich_asso_blue$b_Intercept))
-probright_gbControlCold <- 1/(1+exp(-(guich_asso_blue$b_cortControl + guich_asso_blue$b_Intercept)))
-probright_gbCORTHot <- 1/(1+exp(-(guich_asso_blue$b_tempHot + guich_asso_blue$b_Intercept)))
-probright_gbControlHot <- 1/(1+exp(-(guich_asso_blue$'b_cortControl:tempHot' + guich_asso_blue$b_cortControl + guich_asso_blue$b_tempHot + guich_asso_blue$b_Intercept)))
-# Second we build a df with all the mean probabilities of choosing right in the first trial for each treatment and species
-prob_df <- data.frame(
-  probright_drCORTCold_t = format_dec(mean(probright_drCORTCold), 3),
-  probright_drControlCold_t = format_dec(mean(probright_drControlCold), 3),
-  probright_drCORTHot_t = format_dec(mean(probright_drCORTHot), 3),
-  probright_drControlHot_t = format_dec(mean(probright_drControlHot), 3),
-  probright_dbCORTCold_t = format_dec(mean(probright_dbCORTCold), 3),
-  probright_dbControlCold_t = format_dec(mean(probright_dbControlCold), 3),
-  probright_dbCORTHot_t = format_dec(mean(probright_dbCORTHot), 3),
-  probright_dbControlHot_t = format_dec(mean(probright_dbControlHot), 3),
-  probright_grCORTCold_t = format_dec(mean(probright_grCORTCold), 3),
-  probright_grControlCold_t = format_dec(mean(probright_grControlCold), 3),
-  probright_grCORTHot_t = format_dec(mean(probright_grCORTHot), 3),
-  probright_grControlHot_t = format_dec(mean(probright_grControlHot), 3),
-  probright_gbCORTCold_t = format_dec(mean(probright_gbCORTCold), 3),
-  probright_gbControlCold_t = format_dec(mean(probright_gbControlCold), 3),
-  probright_gbCORTHot_t = format_dec(mean(probright_gbCORTHot), 3),
-  probright_gbControlHot_t = format_dec(mean(probright_gbControlHot), 3)) %>%
-gather(key = "column", value = "value", .) %>%
-rename("prob_choice" = "value") %>%
-data.frame()
-write.csv(prob_df, file= "./output/Checking/prob_df.csv")
-# Then we build a df with the statistical test (one-tailed pmcmc) that the probability is >0.33 (the probability of choosing right by chance)
-comp_pval <- data.frame(
-  probright_drCORTCold_t = format_dec(pmcmc(probright_drCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_drControlCold_t = format_dec(pmcmc(probright_drControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_drCORTHot_t = format_dec(pmcmc(probright_drCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_drControlHot_t = format_dec(pmcmc(probright_drControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_dbCORTCold_t = format_dec(pmcmc(probright_dbCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_dbControlCold_t = format_dec(pmcmc(probright_dbControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_dbCORTHot_t = format_dec(pmcmc(probright_dbCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_dbControlHot_t = format_dec(pmcmc(probright_dbControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_grCORTCold_t = format_dec(pmcmc(probright_grCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_grControlCold_t = format_dec(pmcmc(probright_grControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_grCORTHot_t = format_dec(pmcmc(probright_grCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_grControlHot_t = format_dec(pmcmc(probright_grControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_gbCORTCold_t = format_dec(pmcmc(probright_gbCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_gbControlCold_t = format_dec(pmcmc(probright_gbControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_gbCORTHot_t = format_dec(pmcmc(probright_gbCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_gbControlHot_t = format_dec(pmcmc(probright_gbControlHot, null = 0.33, twotail = FALSE, dir=">"), 3)) %>%
-gather(key = "column", value = "value", .) %>%
-rename("p_value" = "value") %>%
-data.frame()
-write.csv(comp_pval, file= "./output/Checking/comp_pval.csv")
-# Now we merge them and organise them to have the final dt for the table
-table_df <- merge(prob_df, comp_pval, by ="column") %>%
-  mutate(Specie = gsub(".*_(d|g).*", "\\1",column)) %>%
-  mutate(Group = gsub(".*_(dr|db|gr|gb).*", "\\1",column)) %>%
-  mutate(Treatment = gsub(".*(CORTCold|ControlCold|CORTHot|ControlHot).*", "\\1",column)) %>% #Last three commands used to get the species, group and treatment from the column names
-  mutate(Treatment = factor(Treatment, 
-                    levels = c("CORTCold", "ControlCold", "CORTHot", "ControlHot"),
-                    labels = c("CORTCold" = "CORT-Cold", "ControlCold" = "Control-Cold", "CORTHot" = "CORT-Hot", "ControlHot" = "Control-Hot"))) %>%
-  mutate(Specie = factor(Specie,
-                levels = c("d", "g"),
-                labels = c("d" = "L. delicata", "g" = "L. guichenoti"))) %>%
-  mutate(Group = recode(Group, "dr" = "Red", "db" = "Blue", "gr" = "Red", "gb" = "Blue")) %>%
-  select(-column) %>%
-  pivot_wider(names_from = Group, values_from = c(prob_choice, p_value)) %>% #To split the df in Red and Blue
-  select(Specie, Treatment, prob_choice_Blue, p_value_Blue, prob_choice_Red, p_value_Red) %>%
-  arrange(Specie, Treatment) %>% # To arrange the rows the way I wantdata.frame()
-data.frame()
-write.csv(table_df, file= "./output/Checking/table_df.csv")
-#################
-colour_table <- flextable(table_df) %>%
-  bold(~ `p_value_Blue` < 0.05, ~ `p_value_Blue` + prob_choice_Blue) %>%
-  bold(~ `p_value_Red` < 0.05, ~ `p_value_Red` + prob_choice_Red) %>%
-  set_table_properties(width = 1) %>%
-  align(align="center", part="all") %>% 
-  set_header_labels(prob_choice_Blue = "Prob Blue", 
-                    p_value_Blue = "p-value Blue", 
-                    prob_choice_Red = "Prob Red", 
-                    p_value_Red = "p-value Red") %>%
-  italic(j = 1, italic = TRUE, part = "body") %>% # To have names of species in italics
-  flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  hline(i = c(4), j = c(1:6), part = "body") %>% # To make some horizontal lines
-  vline(j = c(2,4), part = "all") %>% # To make some vertical lines on body
-autofit()
-colour_table
-```
-
-On average, we found that, for both species, the proportion of correct choices in the first trial was significantly above chance when the correct feeder was blue (_L.
-delicata_: mean Prob choice = `r format_dec(mean((probright_dbCORTCold + probright_dbControlCold + probright_dbCORTHot + probright_dbControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_dbCORTCold + probright_dbControlCold + probright_dbCORTHot + probright_dbControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`; _L.
-guichenoti_: mean Prob choice = `r format_dec(mean((probright_gbCORTCold + probright_gbControlCold + probright_gbCORTHot + probright_gbControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_gbCORTCold + probright_gbControlCold + probright_gbCORTHot + probright_gbControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`), but not when it was red (_L.
-delicata_: mean Prob choice = `r format_dec(mean((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`; _L.
-guichenoti_: mean Prob choice = `r format_dec(mean((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`).
+Table S2.- Evaluating whether choice diff from 0.5 and interest diff from 0
 
@@
 
-#### Husbandry  
-*Breeding colony* -- We tested juveniles coming from a breeding colony established in the lab since 2019.
-There is a total of 270 and 180 adults of *L.
-delicata* and *L.
-guichenoti* respectively, housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
-Enclosures are provided with non-stick matting, shelter, and several small water dishes.
-Water is given daily, and they are fed approx.
-40 mid-size crickets (*Acheta domestica*) per enclosure three days a week.
-Crickets are dusted with calcium weekly and multivitamin and calcium biweekly.
-To ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle.
-Room temperatures are set to 22-24 Celsius, and warm side of enclosures is usually at 32 Celsius.
-*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we provided females with a place to lay the eggs by means of small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside, that were placed in one extreme of the communal enclosures (see above).
-We checked for the presence of eggs in the boxes three days a week.
-After collection, we measured length and width of eggs with a digital caliper to the nearest 0.1 mm and weight them with a (OHAUS, Model spx123) digital scale ± 0.001g error.
-Then eggs were treated with CORT or vehicle (see CORT and Temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
-The cups were covered with cling wrap to retain moisture and left in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.
-*Hatchlings* -- Eggs in the incubator were checked three times a week for hatchlings.
-After hatchling, we measured juveniles' SVL and Tail Length (TL) with a rule to the nearest mm and weighted them with a (OHAUS, Model spx123) digital scale ± 0.001g error.
-We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with non-stick matting and a small water dish.
-During this period, they were sprayed water every day and received 3-6 small *A.
-domestica* crickets three times a week.
-All care otherwise follows similar protocols to adults (see above).
-Two weeks before we started the training phase (see below), lizards were moved to the experimental arena for acclimatation.
-The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the other.
-These new enclosures were placed in two rooms in 7 different racks associated to 7 different CCTV systems (device model DVR-HP210475) that allowed us to record their behaviour during the experiment (see details below).
-The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack.
-During acclimatation and all the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum*.
-We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle.
-The rooms temperature was set to between 22-24 Celsius.
+#### Testing potential side biases
 
-#### CORT and Temperature manipulation  
-To test empirically the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A).
-We first allocated eggs to one of two different treatments: CORT treatment, where eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO (vehicle) at a final (10 pg CORT/mL) concentration (CORT treatment); and a Control treatment, where eggs received an equal volume of the vehicle.
-CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data).
-Then, eggs were incubated in one of the two previously mentioned temperature regimes ('Cold' or 'Hot') until hatching.
-The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.
+_Table S3.- Number of individuals per treatment that chose the right (R) or left (L) side in each of the numerical tests.
+p-value indicates the result of the binomial test comparing the number of choices between sides._
+```{r, tbl-sides}
+#| label: tbl-sides
+source(here("R", "func.R"))
+#
+# 1) MAKING THE DF FOR THE TABLE
+data_sides <- data_num %>%
+  group_by(trt, test_type) %>% # Group by treatment and test
+  summarise(
+    right_choices = sum(side_choice == "Right", na.rm = TRUE),
+    left_choices = sum(side_choice == "Left", na.rm = TRUE),
+  ) %>% # Count the number of correct choices
+  rowwise() %>%  # Ensure that mutate works row by row
+  mutate(
+    total_choices = right_choices + left_choices,  # Calculate total choices
+    p_value = format_dec(binom.test(right_choices, total_choices)$p.value, 2)  # Perform binom test
+  ) %>% # Test for differences between sides
+  ungroup() %>%  # Remove rowwise grouping
+  mutate(choices_Right_Left_p = paste0("R = ", right_choices, " | L = ", left_choices, " , p = ", p_value), # Create a column with the choices
+        trt = recode(trt, 
+                 "B_23" = "CORT-Cold (n = 20)", 
+                 "A_23" = "Control-Cold (n = 20)", 
+                 "B_28" = "CORT-Hot (n = 20)", 
+                 "A_28" = "Control-Hot (n = 20)")) %>% # To order the treatments
+  dplyr::select(trt, test_type, choices_Right_Left_p) %>% # Remove unused columns
+  pivot_wider(names_from = test_type, values_from = c(choices_Right_Left_p)) %>% # To split between tests
+data.frame()
+#
+# 2) MAKING THE TABLE
+#
+# Split the table_data df by task
+sides_table <- flextable(data_sides) %>%
+  compose(j = 1, value = as_paragraph(""), part = "header") %>%
+  compose(j = 2, value = as_paragraph("1 VS 4"), part = "header") %>%
+  compose(j = 3, value = as_paragraph("1 VS 3"), part = "header") %>%
+  compose(j = 4, value = as_paragraph("2 VS 4"), part = "header") %>%
+  compose(j = 5, value = as_paragraph("2 VS 3"), part = "header") %>%
+  compose(j = 6, value = as_paragraph("3 VS 4"), part = "header") %>%
+  set_table_properties( width = 0.9, layout = "autofit") %>%
+  align(align = "center", part = "all")
+sides_table
+```
 
-### Learning task
-The habituation phase was divided into three stages: in the first stage, the reward, a small, frozen cricket (*A.
-domestica*) was placed in an opaque petri dish (3D x 1.6H cm) (@fig-Methods B, Stage 1); in the second stage, the petri dish with the cricket was placed on top of the white 3D printed ramps (@fig-Methods B, Stage 2); and in the third stage, the cricket was left inside a well (3D x 1.75H cm) on top of the ramp (Fig.
-@fig-Methods B, Stage 3).
-Trials in every stage lasted one hour, begining when the feeding block (petri dish, ramp, or both) was placed inside the enclosure and finished when we took the feeding block away.
-We recorded whether the cricket had been consumed or not and we considered an stage as completed if the lizards ate the crickets in at least 5 out of 6 trials.
-This phase lasted 38 days, and only in one case we decided not to use the lizard in the next phases because its behaviour was not consistent.
-In the associative phase, we used three ramps that were identical in shape and size but differ in the colour (Associative task in Fig.
-@fig-Methods B).
-As in the last stage of the habituation phase, we placed the cricket inside the wells of the ramps, but then two of them were covered with 3D-printed lids (3D x 0.5H cm) so prey was only accessible in "the correct" ramp.
-The food reward was placed in all three wells to avoid lizard using prey chemical cues, that were released in the closed wells thanks to a series of small holes on the top of the lids.
-The colours of the feeders were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision].
-However, we decided to split the lizards in two groups to control for potential colour preference that could bias our results.
-One group of lizards was assigned the blue ramp as the correct choice (i.e., the ramp with the non-covered frozen cricket) in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal.
-In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association.
-Lizards were tested in this task once a day for 35 days.
-After the colour association phase, we performed a choice reversal task (Reversal task in @fig-Methods B).
-Here, the attainability of prey was indicated by a different colour from the reversal, requiring the lizards to form a novel association between the new colour and the food reward.
-This test was done once a day for 40 days.
-We performed the full experiment was between the 6^th^ of March until the 26^th^ of June 2023, between 11 to 12 am, when the lizards were active.
-Trials in the learning phases (colour associative task and reversal tasks) were recorded with different CCTV systems always using the same camera per individual.
-Videos were analysed manually using a standard video player (IINA) by PR, who recorded whether the first choice made by each subject was the correct feeder or not.
-We considered a choice was made if the head of the lizard was inside the well of one of the ramps.
-We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses.
-We excluded from our analyses those individuals with more than 15 trials failed (i.e.
-they did not make a choice), and we considered the first trial to be the first one where the individual made a choice.
-We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+#### Prey orientation test
+Because we were concerned that the orientation of the crickets could affect the lizards' choice, we performed one simple choice test to investigate this.
+We used the same platform and Petri dish as in the numerical discrimination task, but we placed one cricket on each side of the platform, one oriented vertically in respect to the shelter position (see @fig-Methods B) and the other one horizontal to the shelter.
+We perform a unique trial randomising the position of the orientation of the cricket.
+We recorded the amount of individuals that chose the cricket oriented vertically and horizontally.
+Results are shown in the table below.
+
+_Table S4.
+Number of individuals per treatment that chose the cricket oriente horizontally (Horizontal) or vertically (Vertical) in our control tests.
+p-value indicates the result of the binomial test comparing the number of choices between both choices._
+```{r, tbl-orientation}
+#| label: tbl-sides
+source(here("R", "func.R"))
+#
+# 1) MAKING THE DF FOR THE TABLE
+data_or_res <- data_orient %>%
+  mutate(choice_or = ifelse(choice_horizontal == 0, "Vertical", "Horizontal")) %>%
+  dplyr::select(trt, choice_horizontal, choice_or) %>%
+  group_by(trt) %>% # Group by treatment and test
+  summarize(
+    Horizontal_Count = sum(choice_or == "Horizontal"),
+    Vertical_Count = sum(choice_or == "Vertical")
+  ) %>% # Count the number of horizontal and vertical choices
+  rowwise() %>%  # Ensure that mutate works row by row
+  mutate(
+    total_choices = Horizontal_Count + Vertical_Count,  # Calculate total choices
+    p_value = format_dec(binom.test(Horizontal_Count, total_choices)$p.value, 2)  # Perform binom test
+  ) %>% # Test for differences between sides
+  ungroup() %>%  # Remove rowwise grouping
+  mutate(trt = recode(trt, 
+            "B_23" = "CORT-Cold (n = 20)", 
+            "A_23" = "Control-Cold (n = 20)", 
+            "B_28" = "CORT-Hot (n = 20)", 
+            "A_28" = "Control-Hot (n = 20)")) %>% # To order the treatments
+  dplyr::select(trt, Horizontal_Count, Vertical_Count, p_value) %>% # Remove unused columns
+data.frame()
+#
+# 2) MAKING THE TABLE
+#
+# Split the table_data df by task
+orient_table <- flextable(data_or_res) %>%
+  compose(j = 1, value = as_paragraph(""), part = "header") %>%
+  compose(j = 2, value = as_paragraph("Horizontal"), part = "header") %>%
+  compose(j = 3, value = as_paragraph("Vertical"), part = "header") %>%
+  compose(j = 4, value = as_paragraph("p-value"), part = "header") %>%
+  set_table_properties( width = 0.9, layout = "autofit") %>%
+  align(align = "center", part = "all")
+orient_table
+```
 
@@
 #### Checking the models plots  
-```{r, residuals}
-# Chunk for calling all the models and making the residuals
-# Associative task
-## L. delicata
-### Red
-mod_dar <- readRDS(here("output/models/asso_deli_red.rds"))
-resid_dar <- residuals(mod_dar)
-### Blue
-mod_dab <- readRDS(here("output/models/asso_deli_blue.rds"))
-resid_dab <- residuals(mod_dab)
-## L. guichenoti
-### Red
-mod_gar <- readRDS(here("output/models/asso_guich_red.rds"))
-resid_gar <- residuals(mod_gar)
-### Blue
-mod_gab <- readRDS(here("output/models/asso_guich_blue.rds"))
-resid_gab <- residuals(mod_gab)
-# Reversal task
-## L. delicata
-### Red
-mod_drr <- readRDS(here("output/models/rev_deli_red.rds"))
-resid_drr <- residuals(mod_drr)
-### Blue
-mod_drb <- readRDS(here("output/models/rev_deli_blue.rds"))
-resid_drb <- residuals(mod_drb)
-## L. guichenoti
-### Red
-mod_grr <- readRDS(here("output/models/rev_guich_red.rds"))
-resid_grr <- residuals(mod_grr)
-### Blue
-mod_grb <- readRDS(here("output/models/rev_guich_blue.rds"))
-resid_grb <- residuals(mod_grb)
+```{r, checking_model, out.width="70%", fig.align="center"}
+# Chunk for plotting the model
+#
+plot(model)
 ```
-
-Model formula for the reversal task is:  
-Choice  ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)  
-Plots for the different models of the associative task:  
-1.- _L. delicata_  
-  1.a.- Red  
-```{r , plotmod_drr, out.width="70%", fig.align="center"}
-plot(mod_drr)
-```
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-  1.b.- Blue  
-```{r , plotmod_drb, out.width="70%", fig.align="center"}  
-plot(mod_drb)
-```
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-2.- _L. guichenoti_  
-  2.a.- Red  
-```{r , plotmod_grr, out.width="70%", fig.align="center"}
-plot(mod_grr)
-```
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-  2.b.- Blue  
-```{r , plotmod_grb, out.width="70%", fig.align="center"}
-plot(mod_grb)
-```

### 2024-08-18 .vdoc.r

@@ -1,0 +1,357 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: packages
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS,
+fitdistrplus)
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: fig-Methods
+#| fig.cap: "Experimental design of early environment manipulation (**A**), numerical discrimination arena an measures(**B**), measures of the platform used for the experiments (**C**), and types of numerical trial (**D**), where each line indicates the position and orientation of the food items in each test."
+
+knitr::include_graphics("./Others/NUMDIS_FIG_1.png")
+
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: data_process
+# Load the data and process it to get the main dfs
+source(here("R", "data_process.R"))
+#
+#
+#
+#| label: sampleSize
+# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
+source(here("R", "func.R"))
+#
+hormone <- c("CORT", "Control")
+temperature <- c("Cold", "Hot")
+side <- c("Right", "Left")
+#
+n_list <- list()
+#
+for(k in 1:length(hormone)){
+  for(l in 1:length(temperature)){
+    for(i in 1:length(side)){
+      n_list[[paste(hormone[k], temperature[l], side[i], sep = "_")]] <- sample(hormone[k], temperature[l], side [i])
+    }
+  }
+}
+#
+#
+#
+#| label: models
+# Fitting the model and extraction of posteriors 
+refit = FALSE
+if(refit){
+  # Get the formula and family of variables
[…]

### 2024-08-18 R/data_process.R

@@ -1,12 +1,14 @@
 ####################################
 # 1_data_process
 ####################################
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats)
-# Clean the data
-# Remove individuals who did not participate (more than 2 NAs) and split treatment into Temp and Cort
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS,
+fitdistrplus)
+# 
+#### CLEAN THE MAIN DATABASE
+# Remove individuals who did not participate (more than 2 NAs), split treatment into Temp and Cort, and relevel test_type
 data_num <- read.csv(here("data/lizard_database_number.csv")) %>%
   group_by(lizard_id) %>%
-    filter(sum(is.na(choice)) <= 3) %>%
+    filter(sum(is.na(choice)) <= 2) %>%
   ungroup()  %>%
   mutate(temp = gsub("[AB]_", "", trt),
          cort = gsub("_[2][38]", "", trt))  %>%
@@ -16,6 +18,21 @@ data_num <- read.csv(here("data/lizard_database_number.csv")) %>%
   mutate(cort = factor(cort,
     levels = c("B", "A"),
     labels = c("B" = "CORT", "A" = "Control"))) %>%
-  data.frame()
-
+  mutate(test_type = factor(test_type,
+    levels = c("1 VS 4", "1 VS 3", "2 VS 4", "2 VS 3", "3 VS 4"))) %>%
+data.frame()
+#
+#
+#
+#### CLEAN THE DATABASE FOR THE ORIENTATION TEST
+data_orient <- read.csv(here("data/orientation_test.csv")) %>%
+  mutate(temp = gsub("[AB]_", "", trt),
+         cort = gsub("_[2][38]", "", trt))  %>%
+  mutate(temp = factor(temp,
+    levels = c("23", "28"),
+    labels = c("23" = "Cold", "28" = "Hot"))) %>%
+  mutate(cort = factor(cort,
+    levels = c("B", "A"),
+    labels = c("B" = "CORT", "A" = "Control"))) %>%
+data.frame()
 

### 2024-08-18 R/func.r

@@ -7,119 +7,20 @@ pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4,
 # Extract sample size per group
 #' @title sample
 #' @description Estract sample size per group
-#' @param sp To select the species of interest ("deli"/"guich")
-#' @param bias To select group depending on the colour assigned as correct for each task ("blue"/"red")
 #' @param corti To select cort treatment ("CORT"/"Control")
 #' @param therm To select temp treatment ("Cold"/"Hot")
-sample <- function(sp, bias, corti, therm){
+#' @param bias To select the side selected first ("Left"/"Right")
+sample <- function(corti, therm, bias){
   #Specify database
-  data <- data_asso
+  data <- data_num
   # Count sample
   sample_size <- data %>%
-                filter(species == sp, group == bias, cort == corti, temp == therm, Associative_Trial == 1) %>%
+                filter(cort == corti, temp == therm, side_choice == bias, test_type == "1 VS 4") %>%
                 group_by(lizard_id) %>%
                 summarise(n = n()) %>%
                 summarise(total_count = sum(n)) %>%
                 pull(total_count)
   return(sample_size)
-}
-####################
-####################
-# Fit models and extract posteriors both per each treatment:
-#' @title fit_m Function
-#' @description Fit brm models for different the associative task
-#' @param type To define if we are analysing the associative ("asso) or the reversal ("rev")
-#' @param sp To select the species of interest ("deli"/"guich")
-#' @param bias To select group depending on the colour assigned as correct for each task ("blue"/"red")
-#' @param refir To choose whether to refit the models (TRUE, default) or use the ones already made (FALSE)
-#' @return Raw posteriors of fitted brm model for each treatment, species, and group (df)
-fit_m <- function(type, sp, bias, refit = TRUE) {
-  #Specify the type
-  if (type == "asso"){
-    data <- data_asso
-    formula <- FC_associative ~ Associative_Trial*cort*temp + (1 + Associative_Trial|lizard_id)
-  }else{
-    if(type == "rev"){
-      data <- data_rev
-      formula <- FC_reversal ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)
-    } else {
-      stop("Type not valid")
-      return(NULL)  # Return NULL in case of an invalid option
-    }
-  }
-  #Specify species
[…]

### 2024-08-19 ms.qmd
updates

@@
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS,
-fitdistrplus)
+fitdistrplus, DHARMa)
 ```
@@
   # Get the formula and family of variables
-  lat <- bf(latency ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = negbinomial())
+  lat <- bf(log(latency) ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = gaussian())
   choice <- bf(choice ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = bernoulli(link = "logit"))
   int <- bf(compared_interest ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch),
-    family = student())
+    family = gaussian())
   # Fit the model
   model <- brm(lat + choice + int,
-                chains = 4, cores = 4, iter = 3000, warmup = 1000, control = list(adapt_delta = 0.99),
+                chains = 4, cores = 4, iter = 3000, warmup = 1000, control = list(adapt_delta = 0.99, max_treedepth =12),
                 data = data_num)
@@
 #| label: resultsorg
-# Organise the estimated results based on the posteriors
-##
+# Get predictions for every test and treatment based on the posteriors of the model. We used the function org_post (see func.R) to get the predictions.
+source(here("R", "func.R"))
+## A) Get predictions
+post_lat <- org_post("lat")
+post_choice <- org_post("choice")
+post_int <- org_post("int")
 #
-# If I want to get the probability of choosing right per treatment and trial from the posteriors, I will need to take Intercept of each ch treatment and add it, then get the trial-treatment slope, and then model each probability as:
-# P = exp(intercept + slope*trial) / (1 + exp(intercept + slope*trial))
-# And to estimate the Latency it will have to be the same but
-# Lat = intercept * exp(slope*trial)
-# What is trial here?
-# What should I do to get the estimated compared_interest if it follows a t student distribution?
-# Would be:
-# Estimated interest = intercept + slope * trial 
-# correct?
-# How do I get each Intercept for each treatment/trial? And the slope?
-# 
-# E.g.: Getting estimated Latency for CORT-Hot in the 1 VS 3 trial (base level is CORT-Cold test 1 VS 4)
-# Intercept = posteriors$b_latency_Intercept + posteriors$b_latency_test_type1VS3 + posteriors$b_latency_tempHot 
-# Slope = b_latency_test_type1VS3:tempHot
-# Estimated Latency = Intercept * exp(Slope * Trial?)
+write.csv(post_lat, here("output/Checking/post_lat.csv"))
 # 
@@
 # Chunk for plotting the model
+## Initial plot
+plot(model)
 #
-plot(model)
+## pp_checks
+pp_check(model, resp = "loglatency")
+pp_check(model, resp = "choice")
+pp_check(model, resp = "comparedinterest")
 ```

### 2024-08-19 .vdoc.r

@@ -21,7 +21,7 @@
 #
 #| label: packages
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS,
-fitdistrplus)
+fitdistrplus, DHARMa)
 #
 #
 #
@@ -134,13 +134,13 @@ for(k in 1:length(hormone)){
 refit = FALSE
 if(refit){
   # Get the formula and family of variables
-  lat <- bf(latency ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = negbinomial())
+  lat <- bf(log(latency) ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = gaussian())
   choice <- bf(choice ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = bernoulli(link = "logit"))
   int <- bf(compared_interest ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch),
-    family = student())
+    family = gaussian())
   # Fit the model
   model <- brm(lat + choice + int,
-                chains = 4, cores = 4, iter = 3000, warmup = 1000, control = list(adapt_delta = 0.99),
+                chains = 4, cores = 4, iter = 3000, warmup = 1000, control = list(adapt_delta = 0.99, max_treedepth =12),
                 data = data_num)
     # Write the model to a file
     saveRDS(model, file = here("output/models/model_num.rds"))
@@ -148,30 +148,20 @@ if(refit){
       # Read the model from a file
       model <- readRDS(file = here("output/models/model_num.rds"))
   } 
-  # Extract posteriors
+# Extract posteriors
 posteriors <- as_draws_df(model)
 #
 #
 #
 #| label: resultsorg
-# Organise the estimated results based on the posteriors
-##
+# Get predictions for every test and treatment based on the posteriors of the model. We used the function org_post (see func.R) to get the predictions.
+source(here("R", "func.R"))
+## A) Get predictions
+post_lat <- org_post("lat")
+post_choice <- org_post("choice")
+post_int <- org_post("int")
 #
-# If I want to get the probability of choosing right per treatment and trial from the posteriors, I will need to take Intercept of each ch treatment and add it, then get the trial-treatment slope, and then model each probability as:
[…]

### 2024-08-19 R/data_process.R

@@ -20,6 +20,7 @@ data_num <- read.csv(here("data/lizard_database_number.csv")) %>%
     labels = c("B" = "CORT", "A" = "Control"))) %>%
   mutate(test_type = factor(test_type,
     levels = c("1 VS 4", "1 VS 3", "2 VS 4", "2 VS 3", "3 VS 4"))) %>%
+  mutate(age = scale(age, center = TRUE, scale =FALSE)) %>%
 data.frame()
 #
 #

### 2024-08-19 R/func.r

@@ -24,6 +24,45 @@ sample <- function(corti, therm, bias){
 }
 ####################
 ####################
+# Get predictions based on posteriors
+#' @title org_post
+#' @param var The response variable we are interested in ("lat", "choice", "int")
+org_post <- function(var){
+  df <- data.frame() #create empty df
+  sex <- unique(data_num$sex)
+  test <- unique(data_num$test_type)
+  temp <- unique(data_num$temp)
+  horm <- unique(data_num$cort) # create the vectors employed for the loops
+  # create names for each variable and each df:
+  if(var == "lat"){
+    name <- "loglatency"
+  } else if(var == "choice"){
+    name <- "choice"
+  } else if(var == "int"){
+    name <- comparedinterest
+  } else{
+    stop("Variable not valid")
+  } 
+  data <- posteriors %>% dplyr::select(contains(paste0("b_", name)))
+  for(i in sex){
+    if(i == "m"){
+      df_m <- data %>%
+        mutate(!!paste0("b_", name, "_Intercept") := 
+                 !!sym(paste0("b_", name, "_Intercept")) + 
+                 !!sym(paste0("b_", name, "_sexm"))) %>%
+        dplyr::select(-contains(paste0("b_", name, "_sexm")))
+    } else if(i == "f"){
+      df_f <- data %>% 
+        dplyr::select(-contains(paste0("b_", name, "_sexm")))
+    }
+  }
+df <- rbind(df_m, df_f) 
+return(df)
+}
+print(paste0("b_", name, "_Intercept"))
+print(paste0("b_", name, "sexm"))
+####################
+####################
 # Estimate p-values using pmcm
 #' @title pMCMC Function
 #' @param x The vector for the posterior distribution. Note that this will test the null hypothesis that the parameter of interest is significantly different from 0. 

### 2024-08-20 ms.qmd
Updates

@@
   } 
-# Extract posteriors
+# 
+## Extract posteriors
 posteriors <- as_draws_df(model)
-```
-
-```{r, resultsorg}
-#| label: resultsorg
-# Get predictions for every test and treatment based on the posteriors of the model. We used the function org_post (see func.R) to get the predictions.
+# 
+## Get predictions for every test and treatment based on the posteriors of the model. We used the function org_post (see func.R) to get the predictions.
 source(here("R", "func.R"))
@@
 # 
-write.csv(post_lat, here("output/Checking/post_lat.csv"))
 #

### 2024-08-20 .vdoc.r

@@ -148,21 +148,23 @@ if(refit){
       # Read the model from a file
       model <- readRDS(file = here("output/models/model_num.rds"))
   } 
-# Extract posteriors
+# 
+## Extract posteriors
 posteriors <- as_draws_df(model)
-#
-#
-#
-#| label: resultsorg
-# Get predictions for every test and treatment based on the posteriors of the model. We used the function org_post (see func.R) to get the predictions.
+# 
+## Get predictions for every test and treatment based on the posteriors of the model. We used the function org_post (see func.R) to get the predictions.
 source(here("R", "func.R"))
 ## A) Get predictions
 post_lat <- org_post("lat")
 post_choice <- org_post("choice")
 post_int <- org_post("int")
-#
-write.csv(post_lat, here("output/Checking/post_lat.csv"))
 # 
+# probabilities
+a <- exp(post_choice$X1VS4_Cold_CORT)/(1 + exp(post_choice$X1VS4_Cold_CORT))
+b <- exp(post_choice$X1VS4_Cold_Control)/(1 + exp(post_choice$X1VS4_Cold_Control))
+c <- exp(post_choice$X3VS4_Hot_CORT)/(1 + exp(post_choice$X3VS4_Hot_CORT))
+d <- exp(post_choice$X2VS4_Hot_CORT)/(1 + exp(post_choice$X2VS4_Hot_CORT))
+#
 #
 #
 #

### 2024-08-20 R/func.r

@@ -30,8 +30,12 @@ sample <- function(corti, therm, bias){
 org_post <- function(var){
   df <- data.frame() #create empty df
   sex <- unique(data_num$sex)
-  test <- unique(data_num$test_type)
-  temp <- unique(data_num$temp)
+  test <- data_num$test_type %>%
+    unique() %>%
+    levels() %>%
+    gsub(" ", "", .) %>%
+    factor()
+  therm <- unique(data_num$temp)
   horm <- unique(data_num$cort) # create the vectors employed for the loops
   # create names for each variable and each df:
   if(var == "lat"){
@@ -39,11 +43,51 @@ org_post <- function(var){
   } else if(var == "choice"){
     name <- "choice"
   } else if(var == "int"){
-    name <- comparedinterest
+    name <- "comparedinterest"
   } else{
     stop("Variable not valid")
-  } 
-  data <- posteriors %>% dplyr::select(contains(paste0("b_", name)))
+  }
+  # Modify the main df depending on the selected variable
+  data <- posteriors %>% 
+    dplyr::select(contains(paste0("b_", name))) %>%
+    dplyr::select(-matches("age"))
+  # Create function to extract the columns based on the test and treatment
+  choose_col <- function(initial_df, test_type, temperature, hormone){
+    if(test_type == "1VS4"){
+      col_df_1 <- initial_df %>%
+        dplyr::select(-dplyr::matches("1VS3|2VS4|2VS3|3VS4"))
+    } else if (test_type == "1VS3"){
+      col_df_1 <- initial_df %>%
+        dplyr::select(-dplyr::matches("2VS4|2VS3|3VS4"))
+    } else if(test_type == "2VS4"){
+      col_df_1 <- initial_df %>%
+        dplyr::select(-dplyr::matches("1VS3|2VS3|3VS4"))
+    } else if(test_type == "2VS3"){
+      col_df_1 <- initial_df %>%
+        dplyr::select(-dplyr::matches("1VS3|2VS4|3VS4"))
+    } else if(test_type == "3VS4"){
+      col_df_1 <- initial_df %>%
+        dplyr::select(-dplyr::matches("1VS3|2VS4|2VS3"))
+    }
+    if(temperature == "Hot"){
+      col_df_2 <- col_df_1
+    } else if(temperature == "Cold"){
+      col_df_2 <- col_df_1 %>%
+        dplyr::select(-dplyr::matches("Hot")) %>%
+      data.frame()
+    }
+    if(hormone == "Control"){
+      col_df_final <- col_df_2
[…]

### 2024-08-20 ms.qmd
Updates

@@
 # 
+```
+
+```{r, posteriors}
+#| label: posteriors
+# Code to extract posteriors from the model and get predicted values
+source(here("R", "func.R"))
+#
 ## Extract posteriors
@@
 ## Get predictions for every test and treatment based on the posteriors of the model. We used the function org_post (see func.R) to get the predictions.
-source(here("R", "func.R"))
-## A) Get predictions
+#
+# Latency
 post_lat <- org_post("lat")
+# Choice
 post_choice <- org_post("choice")
+prob_choice <- exp(post_choice)/(1 + exp(post_choice))
+# Estimated interest for the higher amount of food
 post_int <- org_post("int")
-# 
 #

### 2024-08-20 .vdoc.r

@@ -149,21 +149,21 @@ if(refit){
       model <- readRDS(file = here("output/models/model_num.rds"))
   } 
 # 
+#
+#
+#
+#| label: posteriors
+# Code to extract posteriors from the model and get predicted values
+#
 ## Extract posteriors
 posteriors <- as_draws_df(model)
 # 
 ## Get predictions for every test and treatment based on the posteriors of the model. We used the function org_post (see func.R) to get the predictions.
-source(here("R", "func.R"))
+
 ## A) Get predictions
 post_lat <- org_post("lat")
 post_choice <- org_post("choice")
 post_int <- org_post("int")
-# 
-# probabilities
-a <- exp(post_choice$X1VS4_Cold_CORT)/(1 + exp(post_choice$X1VS4_Cold_CORT))
-b <- exp(post_choice$X1VS4_Cold_Control)/(1 + exp(post_choice$X1VS4_Cold_Control))
-c <- exp(post_choice$X3VS4_Hot_CORT)/(1 + exp(post_choice$X3VS4_Hot_CORT))
-d <- exp(post_choice$X2VS4_Hot_CORT)/(1 + exp(post_choice$X2VS4_Hot_CORT))
 #
 #
 #

### 2024-08-20 R/func.r

@@ -23,7 +23,7 @@ sample <- function(corti, therm, bias){
   return(sample_size)
 }
 ####################
-####################
+####################  
 # Get predictions based on posteriors
 #' @title org_post
 #' @param var The response variable we are interested in ("lat", "choice", "int")
@@ -51,7 +51,7 @@ org_post <- function(var){
   data <- posteriors %>% 
     dplyr::select(contains(paste0("b_", name))) %>%
     dplyr::select(-matches("age"))
-  # Create function to extract the columns based on the test and treatment
+  # Create function to select the columns
   choose_col <- function(initial_df, test_type, temperature, hormone){
     if(test_type == "1VS4"){
       col_df_1 <- initial_df %>%
@@ -78,7 +78,7 @@ org_post <- function(var){
     }
     if(hormone == "Control"){
       col_df_final <- col_df_2
-    } else if(temperature == "CORT"){
+    } else if(hormone == "CORT"){
       col_df_final <- col_df_2 %>%
         dplyr::select(-dplyr::matches("Control")) %>%
       data.frame()

### 2024-08-21 ms.qmd
Updates

@@
 
+Include here general text: Results are in Table 1 and figure X.
+
+Preliminary results showed: talk about sex and age, refer to summary (prepare that), and explain how did we control by that.
+
+_Table 1.
+Estimates of the effects of each treatment and their interaction on the latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
+The table shows the comparissons for each predictor and the p~mcmcm~ value of the comparissons.
+In bold, the estimates with p~mcmc~ < 0.05._
 ```{r, tbl-data}
 #| label: tbl-data
-#| tbl-cap: "Estimates of Associative learning slope for all the different treatments per each task, species and group.
-Mean shows the aritmetic mean of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean.
-All p-values were obtained using pmcmc and test the hypothesis that the mean is equal to zero.
-In bold, those values that are significant (p-value <0.05)"
+#| tbl-cap: "Estimates of the effects of each treatment and their interaction on the latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
+The table shows the comparissons for each predictor and the p~mcmcm~ value of the comparissons.
+In bold, the estimates with p~mcmc~ < 0.05."
 source(here("R", "func.R"))
 #
-############################## MAKING THE TABLE ##############################
+# Making the df for each variable
+#
+## Initialize df_org as a list to store data frames for each variable
+df_org <- list()
+## Define variables and tests
+variable <- c("Latency", "Choice", "Interest")
+test <- c("1VS4", "1VS3", "2VS4", "2VS3", "3VS4")
+## Loop over each variable
+for(a in variable){
+  # Select the original df to use depending on the variable
+  if(a == "Latency"){
+    post <- post_lat
+  } else if(a == "Choice"){
+    post <- prob_choice
+  } else if(a == "Interest"){
+    post <- post_int
+  }
+  df_tables <- data.frame()
+  for(b in test){
+    # Modify the df to get the values per each test
+    post_mod <- post %>%
+      dplyr::select(dplyr::matches(b))
+    # Extract the required columns
+    Cold_CORT <- post_mod[[paste0("X", b, "_Cold_CORT")]]
+    Cold_Control <- post_mod[[paste0("X", b, "_Cold_Control")]]
+    Hot_CORT <- post_mod[[paste0("X", b, "_Hot_CORT")]]
+    Hot_Control <- post_mod[[paste0("X", b, "_Hot_Control")]]
+    # Calculate effects and pmcmc values
+    effect_temp <- format_dec(mean(c(Hot_CORT, Hot_Control)) - mean(c(Cold_CORT, Cold_Control)), 3)
+    effect_hormone <- format_dec(mean(c(Hot_Control, Cold_Control)) - mean(c(Hot_CORT, Cold_CORT)), 3)
+    interaction <- format_dec((mean(Hot_Control) - mean(Cold_Control)) - (mean(Hot_CORT) - mean(Cold_CORT)), 3)
+    #
+    pmcmc_temp <- format_p(pmcmc(c(Hot_CORT, Hot_Control) - c(Cold_CORT, Cold_Control)), 3)
+    pmcmc_hormone <- format_p(pmcmc(c(Hot_Control, Cold_Control) - c(Hot_CORT, Cold_CORT)), 3)
+    pmcmc_interaction <- format_p(pmcmc((Hot_Control - Cold_Control) - (Hot_CORT - Cold_CORT)), 3)
+    # Define labels and values
+    effects_predictors_names <- c("Temperature", 
+                          "Hormone", 
+                          "Interaction")
+    values_predictors <- c(effect_temp, effect_hormone, interaction)
+    values_pmcmc <- c(pmcmc_temp, pmcmc_hormone, pmcmc_interaction)
+    # Create a temporary data frame for this loop iteration
+    temporal_df <- data.frame(Predictor = effects_predictors_names,
+                        Effect = values_predictors,
+                        pmcmc = values_pmcmc,
+                        Test = b,
+                        stringsAsFactors = FALSE)
+    # Merge the temporary data frame to the main df_tables
+    df_tables <- rbind(df_tables, temporal_df)
+  }
+  df_tables$Variable <- a
+  df_org[[a]] <- df_tables
+}
+# Combine all data frames in df_org into one data frame
+table_df <- do.call(rbind, df_org)
+# Convert the combined list into a data frame and modify for the table
+table_df <- as.data.frame(table_df, stringsAsFactors = FALSE) %>%
+  mutate(results = paste(Effect, "(p", pmcmc, ")")) %>%
+  dplyr::select(Variable, Test, Predictor, results) %>%
+  pivot_wider(names_from = Test, values_from = c(results), names_prefix = "")
+#
+#
+# Making the table
+#
 ## Table format
 set_flextable_defaults(
- font.family = "Times New Roman",
+ font.family = "Arial",
  fint.size = 10)
 # Split the table_data df by task
-real_table <- flextable(new_table_data) %>%
-    bold(~ `p-value_Associative` < 0.05, ~ `p-value_Associative` + Mean_Associative + `95% CI_Associative`) %>%
-    bold(~ `p-value_Reversal` < 0.05, ~ `p-value_Reversal` + Mean_Reversal + `95% CI_Reversal`) %>%
+real_table <- flextable(table_df) %>%
     set_table_properties(width = 1) %>%
     align(align="center", part="all") %>% 
-    add_header_row(values = c("", "Associative task", "Reversal task"), colwidths = c(3, 3, 3)) %>%
-    set_header_labels(Mean_Associative = "Mean",
-                      `95% CI_Associative` = "95% CI",
-                      `p-value_Associative` = "p-value",
-                      Mean_Reversal = "Mean",
-                      `95% CI_Reversal` = "95% CI",
-                      `p-value_Reversal` = "p-value") %>%
-    italic(j = 1, italic = TRUE, part = "body") %>% # To have names od species in italics
-    flextable::compose(i = c(2:8,10:16), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-    flextable::compose(i = c(2:4,6:8,10:12,14:16), j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
-    hline(i = c(4,12), j = c(2:9), part = "body") %>% # To make some horizontal lines
-    hline(i = c(8), j = c(1:9), part = "body") %>% # To make some horizontal lines
-    vline(i = (1:16), j = c(3,6), part = "body") %>% # To make some vertical lines on body
-    vline(j=c(3,6), part = "header") %>% # To make some vertical lines on header
+    add_header_row(values = c("", "Tests"), colwidths = c(2, 5)) %>%
+    flextable::compose(i = c(2,3,5,6,8,9), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+    hline(i = c(3,6), j = c(1:7), part = "body") %>% # To make some horizontal lines
+    vline(i = (1:9), j = c(2:6), part = "body") %>% # To make some vertical lines on body
+    vline(i = 2, j=c(2:6), part = "header") %>% # To make some vertical lines on header
     autofit() 
 real_table
+```
+
+Results: Latency, Choice, and Interest for each test and treatment, references to fig x and Table 1.
+Basically, no effects of treatment on the performance of the numerical discrimination task.
+no discrimination either (refer to Table in Suppl.
+Material).
+
+```{r, fig-results}
+#| label: fig-results
+#| fig.cap: "" 
 ```
@@
 
+#### Preliminary results
+
+_Table S1.- Summary of the model_
+
+```{r, tbl-summary}
+#| label: tbl-summary
+#| tbl-cap: "Summary of the model fitted to the data."
+```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
 #### Treatments effects
-Table S1.- statistical comparisson between treatments
 
-Table S2.- Evaluating whether choice diff from 0.5 and interest diff from 0
+_Table S2.- Evaluating whether choice diff from 0.5 and interest diff from 0_
+```{r, tbl-summary}
+#| label: tbl-summary
+#| tbl-cap: "Summary of the model fitted to the data."
+```
 
@@
     total_choices = Horizontal_Count + Vertical_Count,  # Calculate total choices
-    p_value = format_dec(binom.test(Horizontal_Count, total_choices)$p.value, 2)  # Perform binom test
+    p_value = format_dec(binom.test(Horizontal_Count, total_choices)$p.value, 3)  # Perform binom test
   ) %>% # Test for differences between sides

### 2024-08-21 .vdoc.r

@@ -154,52 +154,119 @@ if(refit){
 #
 #| label: posteriors
 # Code to extract posteriors from the model and get predicted values
+source(here("R", "func.R"))
 #
 ## Extract posteriors
 posteriors <- as_draws_df(model)
 # 
 ## Get predictions for every test and treatment based on the posteriors of the model. We used the function org_post (see func.R) to get the predictions.
-
-## A) Get predictions
+#
+# Latency
 post_lat <- org_post("lat")
+# Choice
 post_choice <- org_post("choice")
+prob_choice <- exp(post_choice)/(1 + exp(post_choice))
+# Estimated interest for the higher amount of food
 post_int <- org_post("int")
 #
 #
 #
 #
+#
+#
+#
+#
+#
 #| label: tbl-data
-#| tbl-cap: "Estimates of Associative learning slope for all the different treatments per each task, species and group. Mean shows the aritmetic mean of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean. All p-values were obtained using pmcmc and test the hypothesis that the mean is equal to zero. In bold, those values that are significant (p-value <0.05)"
+#| tbl-cap: "Estimates of the effects of each treatment and their interaction on the latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests. The table shows the comparissons for each predictor and the p~mcmcm~ value of the comparissons. In bold, the estimates with p~mcmc~ < 0.05."
 source(here("R", "func.R"))
 #
-############################## MAKING THE TABLE ##############################
+# Making the df for each variable
+#
+## Initialize df_org as a list to store data frames for each variable
+df_org <- list()
+## Define variables and tests
+variable <- c("Latency", "Choice", "Interest")
+test <- c("1VS4", "1VS3", "2VS4", "2VS3", "3VS4")
+## Loop over each variable
+for(a in variable){
+  # Select the original df to use depending on the variable
+  if(a == "Latency"){
+    post <- post_lat
+  } else if(a == "Choice"){
+    post <- prob_choice
+  } else if(a == "Interest"){
+    post <- post_int
+  }
+  df_tables <- data.frame()
+  for(b in test){
[…]

### 2024-08-21 R/func.r

@@ -166,3 +166,16 @@ format_dec <- function(x, n) {
   z <- sprintf(paste0("%.",n,"f"), x)
   return(as.numeric(z))
 }
+####################
+####################
+# Function to format p_values with n decimal places
+#' @title format_p
+#' @param x The object
+#' @param n The number of decimals
+format_p <- function(x, n) {
+  z <- sprintf(paste0("%.",n,"f"), x)
+  tmp <- ifelse(as.numeric(z) <= 0.001, "< 0.001",
+         ifelse(as.numeric(z) <= 0.05 & as.numeric(z) > 0.001, "< 0.05",
+                paste0("= ", as.character(z))))
+  return(tmp)
+}

### 2024-08-21 ms.qmd
Updates

@@
 _Table 1.
-Estimates of the effects of each treatment and their interaction on the latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
+Estimates of the effects of temperature, CORT, and their interaction on latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
 The table shows the comparissons for each predictor and the p~mcmcm~ value of the comparissons.
@@
     # Calculate effects and pmcmc values
-    effect_temp <- format_dec(mean(c(Hot_CORT, Hot_Control)) - mean(c(Cold_CORT, Cold_Control)), 3)
-    effect_hormone <- format_dec(mean(c(Hot_Control, Cold_Control)) - mean(c(Hot_CORT, Cold_CORT)), 3)
-    interaction <- format_dec((mean(Hot_Control) - mean(Cold_Control)) - (mean(Hot_CORT) - mean(Cold_CORT)), 3)
+    effect_temp <- format_dec(mean(c(Hot_CORT, Hot_Control)) - mean(c(Cold_CORT, Cold_Control)), 2)
+    effect_hormone <- format_dec(mean(c(Hot_Control, Cold_Control)) - mean(c(Hot_CORT, Cold_CORT)), 2)
+    interaction <- format_dec((mean(Hot_Control) - mean(Cold_Control)) - (mean(Hot_CORT) - mean(Cold_CORT)), 2)
     #
-    pmcmc_temp <- format_p(pmcmc(c(Hot_CORT, Hot_Control) - c(Cold_CORT, Cold_Control)), 3)
-    pmcmc_hormone <- format_p(pmcmc(c(Hot_Control, Cold_Control) - c(Hot_CORT, Cold_CORT)), 3)
-    pmcmc_interaction <- format_p(pmcmc((Hot_Control - Cold_Control) - (Hot_CORT - Cold_CORT)), 3)
+    pmcmc_temp <- format_p(pmcmc(c(Hot_CORT, Hot_Control) - c(Cold_CORT, Cold_Control)), 2)
+    pmcmc_hormone <- format_p(pmcmc(c(Hot_Control, Cold_Control) - c(Hot_CORT, Cold_CORT)), 2)
+    pmcmc_interaction <- format_p(pmcmc((Hot_Control - Cold_Control) - (Hot_CORT - Cold_CORT)), 2)
     # Define labels and values
@@
  font.family = "Arial",
- fint.size = 10)
-# Split the table_data df by task
+ font.size = 10,
+ layout = "autofit")
+# Create the table
 real_table <- flextable(table_df) %>%
@@
     flextable::compose(i = c(2,3,5,6,8,9), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-    hline(i = c(3,6), j = c(1:7), part = "body") %>% # To make some horizontal lines
-    vline(i = (1:9), j = c(2:6), part = "body") %>% # To make some vertical lines on body
-    vline(i = 2, j=c(2:6), part = "header") %>% # To make some vertical lines on header
-    autofit() 
+    hline(i = c(3,6), j = c(1:7), part = "body") # To make some horizontal lines 
 real_table
@@
 
-_Table S1.- Summary of the model_
+_Table S1.- Summary of the model fitted_
 
-```{r, tbl-summary}
-#| label: tbl-summary
-#| tbl-cap: "Summary of the model fitted to the data."
+```{r, summary_org}
+#| label: summary_org
+#
+## Get the summary of the posteriors to get the estimates of each parameter of the model
+sum <- summary(posteriors)
+## Extract the values of the summary
+sum_df <- as.data.frame(sum) %>%
+  dplyr::select(-mad) %>%
+  filter(!str_detect(variable, "^r_"))
+```
+
+For loglatency
+```{r, tbl-summary_lat}
+#| label: tbl-summary_lat
+#| tbl-cap: "Summary of the model fitted for the latency variable."
+#
+source(here("R", "func.R"))
+## Get df only for latency
+df_sum_lat <- sum_table_org(sum_df, "lat")
+## Make table
+table_sum_lat <- flextable(df_sum_lat)
+table_sum_lat
+```
+
+For choice
+```{r, tbl-summary_choice}
+#| label: tbl-summary_choice
+#| tbl-cap: "Summary of the model fitted for the choice variable."
+#
+source(here("R", "func.R"))
+## Get df only for choice
+df_sum_choice <- sum_table_org(sum_df, "choice")
+## Make table
+table_sum_choice <- flextable(df_sum_choice)
+table_sum_choice
+```
+
+For compared interest
+```{r, tbl-summary_int}
+#| label: tbl-summary_int
+#| tbl-cap: "Summary of the model fitted for the interest variable."
+#
+source(here("R", "func.R"))
+## Get df only for interest
+df_sum_int <- sum_table_org(sum_df, "int")
+## Make table
+table_sum_int <- flextable(df_sum_int)
+table_sum_int
+```
+
+Other effects
+```{r, tbl-summary_other}
+#| label: tbl-summary_other
+#| tbl-cap: "Summary of the model fitted for other effects."
+#
+source(here("R", "func.R"))
+## Get df only for interest
+df_sum_other <- sum_df %>%
+  filter(!str_detect(variable, "^b_"))
+## Make table
+table_sum_other <- flextable(df_sum_other)
+table_sum_other
 ```
@@
 
-_Table S2.- Evaluating whether choice diff from 0.5 and interest diff from 0_
-```{r, tbl-summary}
-#| label: tbl-summary
-#| tbl-cap: "Summary of the model fitted to the data."
+_Table S2.- Estimated mean probability of choosing first the higher amount (Choice) and the mean interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
+p indicates the p~mcmcm~ value (p) of testing the hypothesis Choice ≠ 0.5, and Interest ≠ 0, which would indicate a preference towards one of choices (> towards higher amounts, < towards lower amounts) of the comparissons._
+```{r, tbl-disc}
+#| label: tbl-disc
+#| tbl-cap: "Estimated probability of choosing first the higher amount (Choice) and the interest for the higher amount of food (Interest) per treatment group for each os the tests performed.
+p indicates the p~mcmcm~ value of testing the hypothesis Choice ≠ 0.5, and Interest ≠ 0, which would indicate a preference towards one of choices (> towards higher amounts, < towards lower amounts) of the comparissons."
+#
+source(here("R", "func.R"))
+#
+# Making the df for each variable
+#
+## Initialize df_org as a list to store data frames for each variable
+disc_list <- list()
+## Define variables and tests
+variable <- c("Choice", "Interest")
+test <- c("1VS4", "1VS3", "2VS4", "2VS3", "3VS4")
+## Loop over each variable
+for(c in variable){
+  # Select the original df to use depending on the variable
+  if(c == "Choice"){
+    post <- prob_choice
+    val <- 0.5
+  } else if(c == "Interest"){
+    post <- post_int
+    val <- 0
+  }
+  df_disc <- data.frame()
+  for(d in test){
+    # Modify the df to get the values per each test
+    post_mod <- post %>%
+      dplyr::select(dplyr::matches(d))
+    # Extract the required columns
+    Cold_CORT <- post_mod[[paste0("X", d, "_Cold_CORT")]]
+    Cold_Control <- post_mod[[paste0("X", d, "_Cold_Control")]]
+    Hot_CORT <- post_mod[[paste0("X", d, "_Hot_CORT")]]
+    Hot_Control <- post_mod[[paste0("X", d, "_Hot_Control")]]
+    # Calculate means and pmcmc values
+    Cold_CORT_mean <- format_dec(mean(Cold_CORT), 2)
+    Cold_Control_mean <- format_dec(mean(Cold_Control), 2)
+    Hot_CORT_mean <- format_dec(mean(Hot_CORT), 2)
+    Hot_Control_mean <- format_dec(mean(Hot_Control), 2)
+    #
+    pmcmc_Cold_CORT <- format_p(pmcmc(Cold_CORT, null = val), 2)
+    pmcmc_Cold_Control <- format_p(pmcmc(Cold_Control, null = val), 2)
+    pmcmc_Hot_CORT <- format_p(pmcmc(Hot_CORT, null = val), 2)
+    pmcmc_Hot_Control <- format_p(pmcmc(Hot_Control, null = val), 2)
+    # Define labels and values
+    treatment_names <- c("Cold-CORT", "Cold-Control", "Hot-CORT", "Hot-Control")
+    values_means <- c(Cold_CORT_mean, Cold_Control_mean, Hot_CORT_mean, Hot_Control_mean)
+    values_pmcmc <- c(pmcmc_Cold_CORT, pmcmc_Cold_Control, pmcmc_Hot_CORT, pmcmc_Hot_Control)
+    # Create a temporary data frame for this loop iteration
+    temporal_df <- data.frame(Treatment = treatment_names,
+                        Effect = values_means,
+                        pmcmc = values_pmcmc,
+                        Test = d,
+                        stringsAsFactors = FALSE)
+    # Merge the temporary data frame to the main df_tables
+    df_disc <- rbind(df_disc, temporal_df)
+  }
+  df_disc$Variable <- c
+  disc_list[[c]] <- df_disc
+}
+# Combine all data frames in df_org into one data frame
+disc_df <- do.call(rbind, disc_list)
+# Convert the combined list into a data frame and modify for the table
+disc_df_final <- as.data.frame(disc_df, stringsAsFactors = FALSE) %>%
+  mutate(pmcmc = gsub("^= ", "", pmcmc)) %>%
+  pivot_wider(names_from = Test, values_from = c(Effect, pmcmc), names_prefix = "") %>%
+  dplyr::select(Variable, Treatment, 
+                Effect_1VS4, pmcmc_1VS4,
+                Effect_1VS3, pmcmc_1VS3,
+                Effect_2VS4, pmcmc_2VS4,
+                Effect_2VS3, pmcmc_2VS3,
+                Effect_3VS4, pmcmc_3VS4)
+#
+#
+# Making the table
+#
+## Table format
+set_flextable_defaults(
+ font.family = "Arial",
+ fint.size = 10)
+# Create the table
+disc_table <- flextable(disc_df_final) %>%
+    set_table_properties(width = 1) %>%
+    align(align="center", part="all") %>% 
+    add_header_row(values = c("", "1VS4", "1VS3", "2VS4", "2VS3", "3VS4"), colwidths = c(2, 2, 2, 2, 2, 2)) %>%
+    add_header_row(values = c("", "Tests"), colwidths = c(2, 10)) %>%
+    flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+    flextable::compose(i = 3, j = c(3,5,7,9,11), value = as_paragraph("Est. mean"), part = "head") %>% 
+    flextable::compose(i = 3, j = c(4,6,8,10,12), value = as_paragraph("p"), part = "head") %>% 
+    hline(i = 4, j = c(1:12), part = "body") %>% # To make some horizontal lines
+    vline(i = c(1:8), j = c(4,6,8,10), part = "body") %>% # To make some vertical lines
+    vline(i = c(2,3), j = c(4,6,8,10), part = "head") 
+disc_table
 ```
@@
 #| label: tbl-sides
+#| tbl-cap: "Number of individuals per treatment that chose the right (R) or left (L) side in each of the numerical tests.
+p-value indicates the result of the binomial test comparing the number of choices between sides."
 source(here("R", "func.R"))
@@
 ```{r, tbl-orientation}
-#| label: tbl-sides
+#| label: tbl-orientation
+#| tbl-cap: "Number of individuals per treatment that chose the cricket oriente horizontally (Horizontal) or vertically (Vertical) in our control tests.
+p-value indicates the result of the binomial test comparing the number of choices between both choices."
 source(here("R", "func.R"))

### 2024-08-21 .vdoc.r

@@ -209,13 +209,13 @@ for(a in variable){
     Hot_CORT <- post_mod[[paste0("X", b, "_Hot_CORT")]]
     Hot_Control <- post_mod[[paste0("X", b, "_Hot_Control")]]
     # Calculate effects and pmcmc values
-    effect_temp <- format_dec(mean(c(Hot_CORT, Hot_Control)) - mean(c(Cold_CORT, Cold_Control)), 3)
-    effect_hormone <- format_dec(mean(c(Hot_Control, Cold_Control)) - mean(c(Hot_CORT, Cold_CORT)), 3)
-    interaction <- format_dec((mean(Hot_Control) - mean(Cold_Control)) - (mean(Hot_CORT) - mean(Cold_CORT)), 3)
+    effect_temp <- format_dec(mean(c(Hot_CORT, Hot_Control)) - mean(c(Cold_CORT, Cold_Control)), 2)
+    effect_hormone <- format_dec(mean(c(Hot_Control, Cold_Control)) - mean(c(Hot_CORT, Cold_CORT)), 2)
+    interaction <- format_dec((mean(Hot_Control) - mean(Cold_Control)) - (mean(Hot_CORT) - mean(Cold_CORT)), 2)
     #
-    pmcmc_temp <- format_p(pmcmc(c(Hot_CORT, Hot_Control) - c(Cold_CORT, Cold_Control)), 3)
-    pmcmc_hormone <- format_p(pmcmc(c(Hot_Control, Cold_Control) - c(Hot_CORT, Cold_CORT)), 3)
-    pmcmc_interaction <- format_p(pmcmc((Hot_Control - Cold_Control) - (Hot_CORT - Cold_CORT)), 3)
+    pmcmc_temp <- format_p(pmcmc(c(Hot_CORT, Hot_Control) - c(Cold_CORT, Cold_Control)), 2)
+    pmcmc_hormone <- format_p(pmcmc(c(Hot_Control, Cold_Control) - c(Hot_CORT, Cold_CORT)), 2)
+    pmcmc_interaction <- format_p(pmcmc((Hot_Control - Cold_Control) - (Hot_CORT - Cold_CORT)), 2)
     # Define labels and values
     effects_predictors_names <- c("Temperature", 
                           "Hormone", 
@@ -248,17 +248,15 @@ table_df <- as.data.frame(table_df, stringsAsFactors = FALSE) %>%
 ## Table format
 set_flextable_defaults(
  font.family = "Arial",
- fint.size = 10)
[…]

### 2024-08-21 R/func.r

@@ -179,3 +179,23 @@ format_p <- function(x, n) {
                 paste0("= ", as.character(z))))
   return(tmp)
 }
+####################
+####################
+# Function to organise the posteriors for the summary (Suppl) tables
+#' @title sum_table_org
+#' @param data_post the df employed
+#' @param beh the variable selected for each table ("lat", "choice", "int")
+sum_table_org <- function(data_post, beh){
+  if(beh == "lat"){
+    name_var <- "b_loglatency"
+  } else if(beh == "choice"){
+    name_var <- "b_choice"
+  } else if(beh == "int"){
+    name_var <- "b_comparedinterest"
+  } else {
+    stop("Incorrect variable")
+  }
+  data_sum <- data_post %>%
+    filter(str_detect(variable, name_var))
+return(data_sum)
+}

### 2024-08-22 ms.qmd
Updates

@@
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS,
-fitdistrplus, DHARMa)
+fitdistrplus, ggh4x, PupillometryR, cowplot, png, grid, remotes, ggthemes)
 ```
@@
 
+```{r, fig-results}
+#| label: fig-results
+#| fig.cap: "Mean estimates of loglatency (Latency), the probability of choosing first the larger number of crickets (Choice), and the mean estimated interest for the larger number of crickets (Interest) per each of the treatments for each of the numerical tests performed.
+The x-axis represents the estimate, and in the y-axis are the density of the estimates.
+The different colours indicate the different treatments.
+Points and bars represent the mean and standard deviation of the mean of the estimates, respectively.
+Vertical dashed lines in Choice and Interest graphs indicate p = 0.5 and int = 0, respectively."
+#
+source(here("R", "func.R"))
+#
+#### A) Organise the df used for both parts of the figure for each variable
+#
+# First the df for violin plots for each variable using df_violin_plot function from func.r
+df_viol_lat <- df_violin_plot("lat")
+df_viol_choice <- df_violin_plot("choice")
+df_viol_int <- df_violin_plot("int")
+#
+# Second the df for the points and bars plot for each variable using df_bars_plot function from func.r
+df_bars_lat <- df_bars_plot("lat")
+df_bars_choice <- df_bars_plot("choice")
+df_bars_int <- df_bars_plot("int")
+#
+#### B) Make the plots for each variable using plotting function from func.r
+#
+fig_lat <- plotting("lat")
+fig_choice <- plotting("choice")
+fig_int <- plotting("int")
+#
+#### C) Combine the plots and make final figure
+fig_results <- plot_grid(fig_lat, fig_choice, fig_int, ncol = 1)
+ggsave("./output/figures/fig_results.png", plot=fig_results, width = 25, height = 25, units = "cm", dpi = 600)
+knitr::include_graphics("./output/figures/fig_results.png")
+```
+
+Results: Latency, Choice, and Interest for each test and treatment, references to fig x and Table 1.
+Basically, no effects of treatment on the performance of the numerical discrimination task.
+no discrimination either (refer to Table in Suppl.
+Material).
+
 _Table 1.
-Estimates of the effects of temperature, CORT, and their interaction on latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
-The table shows the comparissons for each predictor and the p~mcmcm~ value of the comparissons.
-In bold, the estimates with p~mcmc~ < 0.05._
+Effects of temperature, CORT, and their interaction on latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
+The table shows the comparissons for each predictor (Temperature = [meanHot - meanCold]; Hormone = [meanControl - meanCORT]; Interaction = [(meanHot-Control - meanHot-CORT) - (meanCold-Control - meanCold-CORT)]) and the p~mcmcm~ value of the comparisons.
+In bold, the effects where p~mcmc~ < 0.05._
 ```{r, tbl-data}
 #| label: tbl-data
-#| tbl-cap: "Estimates of the effects of each treatment and their interaction on the latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
-The table shows the comparissons for each predictor and the p~mcmcm~ value of the comparissons.
-In bold, the estimates with p~mcmc~ < 0.05."
+#| tbl-cap: "Effects of temperature, CORT, and their interaction on latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
+The table shows the comparissons for each predictor (Temperature = [meanHot - meanCold]; Hormone = [meanControl - meanCORT]; Interaction = [(meanHot-Control - meanHot-CORT) - (meanCold-Control - meanCold-CORT)]) and the p~mcmcm~ value of the comparisons.
+In bold, the effects where p~mcmc~ < 0.05."
 source(here("R", "func.R"))
@@
     # Calculate effects and pmcmc values
-    effect_temp <- format_dec(mean(c(Hot_CORT, Hot_Control)) - mean(c(Cold_CORT, Cold_Control)), 2)
-    effect_hormone <- format_dec(mean(c(Hot_Control, Cold_Control)) - mean(c(Hot_CORT, Cold_CORT)), 2)
-    interaction <- format_dec((mean(Hot_Control) - mean(Cold_Control)) - (mean(Hot_CORT) - mean(Cold_CORT)), 2)
+    effect_temp <- format_dec(mean(c(Hot_CORT, Hot_Control)) - mean(c(Cold_CORT, Cold_Control)), 1)
+    effect_hormone <- format_dec(mean(c(Hot_Control, Cold_Control)) - mean(c(Hot_CORT, Cold_CORT)), 1)
+    interaction <- format_dec((mean(Hot_Control) - mean(Cold_Control)) - (mean(Hot_CORT) - mean(Cold_CORT)), 1)
     #
@@
 table_df <- as.data.frame(table_df, stringsAsFactors = FALSE) %>%
-  mutate(results = paste(Effect, "(p", pmcmc, ")")) %>%
+  mutate(results = trimws(paste0(Effect, " (p", pmcmc,")"))) %>%
   dplyr::select(Variable, Test, Predictor, results) %>%
@@
  font.family = "Arial",
- font.size = 10,
- layout = "autofit")
+ font.size = 9)
 # Create the table
 real_table <- flextable(table_df) %>%
-    set_table_properties(width = 1) %>%
     align(align="center", part="all") %>% 
@@
     flextable::compose(i = c(2,3,5,6,8,9), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-    hline(i = c(3,6), j = c(1:7), part = "body") # To make some horizontal lines 
+    hline(i = c(3,6), j = c(1:7), part = "body") %>% # To make some horizontal lines 
+    set_table_properties(layout = "autofit", width = 1)
 real_table
-```
-
-Results: Latency, Choice, and Interest for each test and treatment, references to fig x and Table 1.
-Basically, no effects of treatment on the performance of the numerical discrimination task.
-no discrimination either (refer to Table in Suppl.
-Material).
-
-```{r, fig-results}
-#| label: fig-results
-#| fig.cap: "" 
 ```
@@
 
-#### Preliminary results
+#### Model results
 
@@
 ## Extract the values of the summary
-sum_df <- as.data.frame(sum) %>%
+sum_2 <- as.data.frame(sum) %>%
   dplyr::select(-mad) %>%
   filter(!str_detect(variable, "^r_"))
+sum_df <- sum_2 %>%
+  mutate(across(where(is.numeric), ~ sprintf(paste0("%.", 2, "f"), .)))  
 ```
@@
 _Table S2.- Estimated mean probability of choosing first the higher amount (Choice) and the mean interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
-p indicates the p~mcmcm~ value (p) of testing the hypothesis Choice ≠ 0.5, and Interest ≠ 0, which would indicate a preference towards one of choices (> towards higher amounts, < towards lower amounts) of the comparissons._
+p indicates the p~mcmcm~ value (p) of testing the hypothesis Choice ≠ 0.5, and Interest ≠ 0, which would indicate a preference towards one of choices (> towards higher amounts, < towards lower amounts) of the comparissons.
+In bold, the groups where p~mcmcm~ < 0.05_
 ```{r, tbl-disc}
@@
   mutate(pmcmc = gsub("^= ", "", pmcmc)) %>%
-  pivot_wider(names_from = Test, values_from = c(Effect, pmcmc), names_prefix = "") %>%
-  dplyr::select(Variable, Treatment, 
-                Effect_1VS4, pmcmc_1VS4,
-                Effect_1VS3, pmcmc_1VS3,
-                Effect_2VS4, pmcmc_2VS4,
-                Effect_2VS3, pmcmc_2VS3,
-                Effect_3VS4, pmcmc_3VS4)
+  pivot_wider(names_from = Variable, values_from = c(Effect, pmcmc), names_prefix = "") %>%
+  dplyr::select(Test, Treatment, 
+                Effect_Choice, pmcmc_Choice,
+                Effect_Interest, pmcmc_Interest)
 #
@@
     align(align="center", part="all") %>% 
-    add_header_row(values = c("", "1VS4", "1VS3", "2VS4", "2VS3", "3VS4"), colwidths = c(2, 2, 2, 2, 2, 2)) %>%
-    add_header_row(values = c("", "Tests"), colwidths = c(2, 10)) %>%
-    flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-    flextable::compose(i = 3, j = c(3,5,7,9,11), value = as_paragraph("Est. mean"), part = "head") %>% 
-    flextable::compose(i = 3, j = c(4,6,8,10,12), value = as_paragraph("p"), part = "head") %>% 
-    hline(i = 4, j = c(1:12), part = "body") %>% # To make some horizontal lines
-    vline(i = c(1:8), j = c(4,6,8,10), part = "body") %>% # To make some vertical lines
-    vline(i = c(2,3), j = c(4,6,8,10), part = "head") 
+    # Bold the cell containing "< 0.05" in the `pmcmc_Choice` column and the next cell
+    bold(i = which(disc_df_final$pmcmc_Choice == "< 0.05"), j = "pmcmc_Choice", part = "body") %>%
+    bold(i = which(disc_df_final$pmcmc_Choice == "< 0.05"), j = "Effect_Choice", part = "body") %>%
+    bold(i = which(disc_df_final$pmcmc_Choice == "< 0.05"), j = "Treatment", part = "body") %>%
+    # Bold the cell containing "< 0.05" in the `pmcmc_Interest` column and the next cell
+    bold(i = which(disc_df_final$pmcmc_Interest == "< 0.05"), j = "pmcmc_Interest", part = "body") %>%
+    bold(i = which(disc_df_final$pmcmc_Interest == "< 0.05"), j = "Effect_Interest", part = "body") %>%
+    bold(i = which(disc_df_final$pmcmc_Interest == "< 0.05"), j = "Treatment", part = "body") %>%
+    flextable::compose(i = c(2:4,6:8,10:12,14:16,18:20), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+    flextable::compose(i = 1, j = c(3,5), value = as_paragraph("Est. mean"), part = "head") %>% 
+    flextable::compose(i = 1, j = c(4,6), value = as_paragraph("p"), part = "head") %>% 
+    hline(i = c(4,8,12,16), j = c(1:6), part = "body") %>% # To make some horizontal lines
+    vline(i = c(1:20), j = 4, part = "body") %>% # To make some vertical lines
+    vline(i = 1, j = 4, part = "head") 
 disc_table
@@
 _Table S3.- Number of individuals per treatment that chose the right (R) or left (L) side in each of the numerical tests.
-p-value indicates the result of the binomial test comparing the number of choices between sides._
+p-value indicates the result of the binomial test comparing the number of choices between sides.
+In bold, the groups where p-value < 0.05_
 ```{r, tbl-sides}

### 2024-08-22 R/func.r

@@ -199,3 +199,142 @@ sum_table_org <- function(data_post, beh){
     filter(str_detect(variable, name_var))
 return(data_sum)
 }
+####################
+####################
+# Function to create the df for the violin plots
+#' @title df_violin_plot
+#' @param data_fig for choosing df based on the variable selected ("lat", "choice", "int")
+df_violin_plot <- function(data_fig){
+  test <- c("1VS4", "1VS3", "2VS4", "2VS3", "3VS4")
+  fig_df_1 <- data.frame()
+  if(data_fig == "lat"){
+    fig_df_0 <- post_lat
+  } else if(data_fig == "choice"){
+    fig_df_0 <- prob_choice
+  } else if(data_fig == "int"){
+    fig_df_0 <- post_int
+  } else {
+    stop("Incorrect data")
+  }
+  for(m in test){
+    # Modify the df to get the values per each test
+    fig_df_sub <- fig_df_0 %>%
+      dplyr::select(dplyr::matches(m))
+    # Extract the required columns
+    Cold_CORT <- fig_df_sub[[paste0("X", m, "_Cold_CORT")]]
+    Cold_Control <- fig_df_sub[[paste0("X", m, "_Cold_Control")]]
+    Hot_CORT <- fig_df_sub[[paste0("X", m, "_Hot_CORT")]]
+    Hot_Control <- fig_df_sub[[paste0("X", m, "_Hot_Control")]]
+    # Create a temporary data frame for this loop iteration
+    temporal_df <- data.frame(cbind(Test = m,
+                                    Cold_CORT,
+                                    Cold_Control,
+                                    Hot_CORT,
+                                    Hot_Control))
+    # Merge the temporary data frame to the main df_tables
+    fig_df_1 <- rbind(fig_df_1, temporal_df)
+  }
+  # Modify the obtain df to make it adequate for the plot
+  df_viol_plot <- fig_df_1 %>%
+    pivot_longer(cols = c(Cold_CORT, Cold_Control, Hot_CORT, Hot_Control),
+               names_to = "Treatment",
+               values_to = "Estimate") %>%
+    mutate(Treatment = factor(Treatment, levels = c("Cold_CORT", "Cold_Control", "Hot_CORT", "Hot_Control"), 
+                                      labels = c("Cold_CORT" = "Cold-CORT (n = 20)",
+                                                  "Cold_Control" = "Cold-Control (n = 20)",
[…]

### 2024-08-22 .vdoc.r
Delete .vdoc.r

@@ -1,585 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: packages
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS,
-fitdistrplus, DHARMa)
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig.cap: "Experimental design of early environment manipulation (**A**), numerical discrimination arena an measures(**B**), measures of the platform used for the experiments (**C**), and types of numerical trial (**D**), where each line indicates the position and orientation of the food items in each test."
-
-knitr::include_graphics("./Others/NUMDIS_FIG_1.png")
-
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: data_process
-# Load the data and process it to get the main dfs
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-side <- c("Right", "Left")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    for(i in 1:length(side)){
-      n_list[[paste(hormone[k], temperature[l], side[i], sep = "_")]] <- sample(hormone[k], temperature[l], side [i])
-    }
-  }
-}
-#
-#
-#
-#| label: models
-# Fitting the model and extraction of posteriors 
-refit = FALSE
-if(refit){
-  # Get the formula and family of variables
[…]

### 2024-08-26 ms.qmd
Updates

@@
 
-- Quantitative abilities and its importance
-The knowledge of quantities and the relation among them is one of the highest properties of cognition (Reznikova 2018).
-Among those numerical competences, the simplest ability is to make judgments and comparisons between “more” versus “less” (Reznikova 2007; Stancher et al.
-2015), which is widespread in the animal kingdom (Vonk and Beran 2012; Beran et al.
-2014).
-Sensitivity to relative differences in quantity might be adaptive in several situations (Nieder 2017, 2018).
-For instance, female lions (Panthera leo) seem to use numerical assessment of possible rivals to cope with or avoid social contests (McComb et al.
-1994).
-Moreover, the ability to correctly estimate quantities is also associated with, for example, navigation (Chitka and Geiger 1995), choosing larger groups of potential mates (Carazo et al.
-2009) and/or social companions (Bisazza et al.
-2010), or increasing foraging efficiency (Uller et al.
-2003).
-In this sense, the literature is plenty of studies employing the spontaneous choice approach in many animals (e.g., see Carazo et al.
-2009; Pahl et al.
-2013 for invertebrates; Bisazza et al.
-2010; Gómez-Laplaza and Gerlai 2013; Mehlis et al.
-2015 for fishes; Uller et al.
-2003; Krusche et al.
-2010; Stancher et al.
-2015 for amphibians; Garland et al.
-2012 for birds; Ferkin et al.
-2005; Abramson et al.
-2011; Mahamane et al.
-2014; Cox and Montrose 2016; Plotnik et al.
-2019 for mammals) and also of research that explores numerical abilities with training procedures (e.g., see Gross et al.
-2009; Pahl et al.
-2013; Howard et al.
-2018, 2019, for invertebrates; Bisazza et al.
-2014 for fishes; Nieder 2017 for birds; Kilian et al.
-2003; Addessi et al.
-2008; Vonk and Beran 2012; Vonk 2014; Beran and Parrish 2016; Nieder 2017 for mammals).
+The ability to discern between more and less quantities is a basic cognitive skill widely observed across the animal kingdom (**Stancher et al.
+2015; Vonk and Beran 2012; Beran et al.
+2014**), and is considered to be adaptive in different ecological contexts (**Nieder 2017, 2018**).
+Female lions (_Panthera leo_), for example, can use numerical information to assess the risk of confronting rival groups (**McComb et al.
+1994**); but other animals can employ quantitative abilities for navigation (**Chitka and Geiger 1995**), mate selection (**Carazo et al.
+2009**) or foraging (**Uller et al.
+2003**).
+Given the pervasive advantages of numerical and quantity discrimination, research has extensively explored the ability of different species to differentiate between quantities under different contexts and paradigmas.
+Some studies focus on the type of cues involved in quantitative competences (**REFs**), others on the context where animals may use these abilities (**REFs above**), and others in the neural mechanisms underlying quantitative and numerical discrimination (**REFs**).
+The later have demonstrated that non-verbal numerical representations activate an extensive cortical network, with the prefrontal cortex and the intraparietal sulcus being particularly relevant (**Nieder, 2005**).
+This link between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.
 
-- Different ways to test numerical abilities
-There are two main approaches to study numerical abilities: spontaneous choice tests and training procedures.
-In spontaneous choice tests, experimenters present to subjects two groups of biologically relevant stimuli that differ in numerosity, under the assumption that animals should select the most appropriate quantity (e.g.
-the larger amount of food) (Agrillo and Bisazza 2014; Cox and Montrose 2016).
-In training
-procedures, animals must learn a numerical rule to receive a reward (Agrillo and Bisazza 2014; Miletto-Petrazzini et al.
-2018).
-Both experimental approaches have some advantages and disadvantages (reviewed in Agrillo and Bisazza 2014).
-In spontaneous choice tests, animals exhibit their natural behavior, enabling interspecific comparisons and the acquirement of ecological issues related to numerousness (Krusche et al.
-2010; Abramson et al.
-2011).
-However, motivation plays a key-role in spontaneous choice tests (Agrillo and Bisazza 2014), which implies that the absence of results might not be due to the lack of skills to solve the numerical task.
-Also, individual preferences may arise in this kind of procedures and, hence, a large sample of subjects is needed to consider inter-individual variability.
-Finally, it could be difficult, but necessary, to control for continuous variables such as area, density or movement (Agrillo and Bisazza 2014; Mehlis et al.
-2015).
-Regardless, both methods are considered complementary and necessary to draw a firm picture of the numerical abilities of a species (Miletto-
-Petrazzini et al. 2018).
+Animal's brain is particularly sensitive to environmental inputs during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004], with long-lasting effects on cognitive abilities [@zhu_prenatal_2004; @amiel_egg_2014].
+For instance, increases in glucocorticoids (GCs) - the main groups of hormones mediating the response to stressors [@sapolsky_how_2000] - during early life can alter animals' performance in spatial, associative, or reversal learning tasks [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
+These effects have been shown to vary depending on factors such as dose, subjects' sex, or type of cognitive test employed [@szuran_water_1994 @farrell_developmental_2015-learn; @bebus_associative_2016], but little is known about how exposure to GCs during development can affect quantitative abilities.
+Similarly, thermal early environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
+For example, Port Jackson shark incubated at elevated-temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019].
+However, the effects of early thermal environment on numerical abilities in other taxa remain unknown.
+In this sense, not only is it well-established that early-life GCs and thermal environment independently affect cognitive abilities, but these factors are also predicted to interact.
+In vertebrates, abrupt changes in temperature can lead to increased GC levels which, in turn, may elevate GCs in their offspring [@racic_effects_2020; @Crino_2023].
+Simultaneously, these offspring are directly influenced by the same thermal environment, resulting in significant interactions that can impact cognitive development.
+As such, and given the ecological importance of quantitative abilities in nature [**Nieder 2017, 2018**], it is crucial to study the combined effects of early-life GCs and thermal environment on quantity and numerical discrimination.
+Here, we investigate the effects of prenatal exposure to corticosterone (CORT) - the main GC in reptiles [@Crino_2023] - and thermal environment on the ability of lizards to discriminate between different number of food items using the Common garden skink (_Lampropholis guichenoti)_ as a model species.
 
-- The role of early environment on the development of cogition
-The conditions in which organisms develop can be altered by the circumstances faced by their parents.
-For instance, exposing threespine sticklebacks females to a model predator during ovulation, affected their offspring’s tedency to use social cues (Feng et al., 2015).
-Under stressful situations, animal’s sensory systems detect and transmit environmental signals to instigate adaptive physiological and behavioural adjustments – responses that are mediated by glucocorticoids (GCs) (Sapolsky et al., 2000).
-Glucocorticoids can be transmitted to offspring
-and influence their phenotype and development t 46 hrough transgenerational effects (Crino et al., 2023).
-In addition to the environments experienced by parents, offspring also experience potentially stressful environmental conditions that can interact with or amplify parental effects.
-For example, human disturbance alters snapping turtles’ nest-site choice, ultimately impacting the temperature at which eggs are incubated (Kolbe & Janzen, 2002).
-Because animals’ responses to abrupt changes in temperature are mediated by GCs (Crino et al., 2023), and this response can affect animals’ decision-making while nesting, GCs transmission and early thermal environment are predicted to interact to shape offspring phenotype.
-Both GCs and temperature are known to exert potent effects on brain development (Amiel et al., 2017; Coomber et al., 1997; Jonson et al., 1976; Zhu et al., 2004) and cognition (Clark et al., 2014; Lui et al., 2017).
-    - The role of stress hormones on the development of numerical abilities
-      NOTHING DONE?
-    - The role of temperature on the development of numerical abilities
-      Port Jackson shark eggs were incubated and reared at current and projected end-of-century temperatures (+ 3 °C).
-We tested juvenile’s learning ability with a quantity discrimination task.
-The mortality rate of sharks reared in warm water was 41.7% compared with no mortality in the present-day sharks.
-Contrary to expectations, our results suggest that surviving hatchlings from the elevated-temperature group took fewer days to reach learning criterion and had a higher proportion of correct choice compared with hatchlings reared under present conditions.
-Additionally, this is the first data suggesting that sharks can discriminate different quantities (Vila-Pouca et al., 2019).
+Traditionally, the study of numerical abilities has been mainly focused on mammals, birds, and fishes (**Agrillo and Bisazza 2014; Nieder 2017**), but recent research has shown that reptiles are also competent in numbers.
+Red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour to higher volumes of food (**Soldati et al.
+2017**), Hermann’s tortoises (_Testudo hermanni_) proved to discriminate between big and small amounts off food in a spontaneous choice test (**Gazzola et al.
+2018**), and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) showed high skills when tested in a trained quantity discrimination task (**Lin et al.
+2021, 2024**).
+In lizards, __Podarcis siculus_ did not discriminate between different amounts of food in a spontaneous choice test (**Miletto-Petrazzini et al.
+2017**), but were successful in a trained quantity discrimination task (**Miletto-Petrazzini et al.
+2018**), rock lizards (_Iberolacerta cyreni_) chose bigger quantities of crickets when presented 5 versus 10 items [@recio_prey_2021], and Gidgee skinks (_Egernia stokesii_) used numbers to select the larger amount of food in a spontaneous choice test [@szabo_spontaneous_2021].
+Inded, the literature evidences basic quantitative abilities in reptiles, that may vary with the type of cue and test employed.
+However, the question of how early environment can affect numerical abilities remains open as, to our knowledge, there is no study investigating how early life conditions impact numerical discrimination in this group.
+Incubating lizards at elevated temperatures has been shown to improve cognitive abilities in consistency with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017]; and thermal environment impaired cognition only under incubation temperatures beyond lizards' natural range [@dayananda_incubation_2017 or @abayarathna_effects_2020].
+Our understanding of the effects of prenatal CORT on cognitive abilities in reptiles is less understood, but in other taxa, high levels of GCs are tipically related to cognitive impairments [@szuran_water_1994; @zhu_prenatal_2004].
 
-- Numerical abilities in reptiles
-    - Learned num abilities
-    - Spontaneous discrimination
-In contrast, in reptiles, research of quantity discrimination is scarce and contradictory.
-Soldati et al.
-(2017) taught red-footed tortoises (Chelonoidis carbonaria) to associate cards of two colors to different volumes of reward and tested their choices afterward.
-In that succeeding experiment, tortoises showed a clear preference for the color card associated with a greater amount of food, thus concluding that discrimination between quantities is present in this species.
-Similarly, Hermann’s tortoises (Testudo hermanni) also proved to discriminate food by numerosity or size in spontaneous choice tests rating from 0.25 to 0.75 (Gazzola et al.
-2018).
-In contrast, other studies suggest different results; in the Italian wall-lizard (Podarcis siculus), spontaneous choice tests showed no numerical discrimination abilities by this lizard (Miletto-Petrazzini et al.
-2017), although learning procedures demonstrated an ability to discern between small close numbers (2 versus 4) (Miletto-Petrazzini et al.
-2018).
-**Check new papers: Lin in turtles, Szabo in lizards, etc.**
-
-- The role of early environment on reptiles cognition
-Both GCs and temperature are known to exert potent effects on brain development (Amiel et al., 2017; Coomber et al., 1997) and cognition (Clark et al., 2014), but little is known about how they affect numerical discrimination.
-Most studies have reported better performance in different associative learning and motor tasks when incubated at higher temperatures (Amiel et al., 2014; Amiel & Shine, 2012; Clark et al., 2014).
-Studies that have reported cold-incubated lizards to learning faster (Abayarathna & Webb, 2020; Dayananda & Webb, 2017) used incubation temperatures far beyond the natural range in their hot treatments and it is unclear how such conditions relate to wild environments.
-Furthermore, sensorial capacities: **Studies from Burger** | **CORT and cognition in reptiles** | **CORT and brain in reptiles**
-
-- Hypotheses
-    - We expect that lizards incubated at high temperatures will have a better performance in the numerical discrimination task than those incubated at low temperatures.
-    - We expect that lizards incubated with CORT will have a worse performance in the numerical discrimination task than those incubated with the vehicle.
-    - We expect that the effect of CORT on the performance of the numerical discrimination task will be more pronounced in lizards incubated at low temperatures than in those incubated at high temperatures.
-- Predictions:
-    - Lizards incubated at high temperatures will discriminate better quantities even with low ratios (lower latencies, choose large amount more often, more interested in larger amounts - i.e.
-"fail" less tests).
-    - Lizards incubated with CORT will discriminate worse quantities even with high ratios (higher latencies, choices not different from chance, equally interested in both amounts of food - i.e.
-"fail" more tests).
-    - The effect of CORT on the performance of the numerical discrimination task will be more pronounced in lizards incubated at low temperatures than in those incubated at high temperatures (i.e.
-CORT-Cold will "fail" more tests than anyone else).
+In our study, we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 28ºC ± 3ºC) in a 2x2 factorial design.
+After hatching, we assessed the lizards' numerical discrimination abilities in a spontaneous choice test, where they selected between different numbers of crickets.
+We hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task.
+Additionally, we predicted that increased CORT levels would impair numerical competence.
+Moreover, we expected the negative impact of CORT on numerical performance to be more pronounced in lizards incubated at lower temperatures compared to those in higher temperatures.
+In line with Weber's law [**REFs**], we further anticipated that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g.
+3 VS 4), whereas event the lowest ratios (1 VS 4) could be challenging for those incubated with CORT.
 
@@
 The crickets used during the experiment were all the same size, but to control for other quantitative traits
-[**REFs In this regard, we must remark that, in our design, we did not control for other quantity-related continuous cues such as area, volume, density, weight, movement, or contour length that could be used for quantitative judgments (Uller et al.
+(**Uller et al.
 2003; Krusche et al.
@@
 2012; Vonk and Beran 2012; Gómez-Laplaza and Gerlai 2013; Mehlis et al.
-2015), neither for olfactory cues (Plotnik et al.
-2019).
-These continuous variables usually covariate with quantity, but discrimination among them does not necessarily imply processment of numerical information (Agrillo and Bissaza 2014).**]
-, we changed the orientation of the crickets as shown in @fig-Methods D.
+2015**), we changed the orientation of the crickets as shown in @fig-Methods D.
 In addition, after the experiments, we made another set of trials to test whether the orientation of the crickets could affect lizards' choice, but we found no effect of orientation (see Supplementary Material - Prey orientation test).
@@
 #### Statistical analyses  
-INT This gave as a gradient from -300 to +300 where the minimum represents interest in the lower amount of food and the maximum shows a preference towards higher number of crickets, while 0 shows no discrimination between them.
-**We performed the analyses for each species separately.
-We fitted two different Bayesian multilevel models using the brm function from the _brms_ package [@burkner2017brms] using an R extension (version 2.8.2) [@R].
-We ran four parallel MCMC chains of 3000 iterations for each model, with a warmup period of 1000 iterations.
-We modelled correct choices [correct (1) or not (0)] as the response variable, and trial, hormone (CORT versus Control), incubation temperature (Cold versus Hot), and their interaction as fixed factors.
-We included a random intercept and slope (trial) for each lizard in our models.
-We also incorporated the clutch identity as a random factor.
-We used the posterior distributions of parameters from these models to test for differences in learning rate between treatments and species.
-Learning slopes were obtained using the ‘trial’ estimates and its interaction with hormone and temperature treatments.
-Slope estimates greater than zero were considered as evidence of learning, while those less or equal to zero were not.
-We calculated the probability that slopes or contrasts between slopes differed from zero (p~mcmc~) using the posterior distributions.
-We considered statistical significance if p~mcmc~  < 0.05.**
+In our analyses, we fitted a multivariate Bayesian multilevel model using the brm function from the _brms_ package [@burkner2017brms] in R extension (version 2.8.2) [@R].
+We ran four parallel MCMC chains of 4000 iterations with a warmup period of 1000 iterations.
+We modelled the three main recorded behaviour: Latency (log-transformed), Choice, and Interest (see above) as the response variables, and test (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4), hormone (CORT versus Control), incubation temperature (Cold versus Hot) and their interaction as fixed factors.
+We also included in our predictors sex, and age (centered to 0).
+We included lizard indentity and clutch as random factors.
+We used the posterior distributions of parameters from these models to test for differences in the ability to discriminate between different number of crickets between treatments and species, for each of the numerical tests.
+We calculated the contrast between treatments for each test using p~mcmc~ values, and we also estimated whether predicted Choice and Interest differed from 0.5 or zero respectively using p~mcmc~.
+We considered statistical significance if p~mcmc~ < 0.05.
 
@@
 Experimental procedures and laboratory housing complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33).
-Upon completion of experimental trials, lizards were kept in their enclosure until they were euthanized following standard humane procedures.
+Upon completion of experimental trials, lizards were kept in their enclosure and will be euthanized following standard humane procedures once they reach the human-ending point.
 
@@
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
-#| label: sampleSize
-# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task.
-We used function sample (see func.R) to estimate the sample size per treatment and species.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-side <- c("Right", "Left")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    for(i in 1:length(side)){
-      n_list[[paste(hormone[k], temperature[l], side[i], sep = "_")]] <- sample(hormone[k], temperature[l], side [i])
-    }
-  }
-}
 ```
@@
   model <- brm(lat + choice + int,
-                chains = 4, cores = 4, iter = 3000, warmup = 1000, control = list(adapt_delta = 0.99, max_treedepth =12),
+                chains = 4, cores = 4, iter = 4000, warmup = 1000, control = list(adapt_delta = 0.99, max_treedepth =12),
                 data = data_num)
@@
 
-Include here general text: Results are in Table 1 and figure X.
-
-Preliminary results showed: talk about sex and age, refer to summary (prepare that), and explain how did we control by that.
+Our results showed an effect of sex and age on Latency (mean estimated effect of sex = `r {format_dec(mean(posteriors$b_loglatency_sexm), 2)}`, 95% CI = [`r {format_dec(quantile(posteriors$b_loglatency_sexm, 0.025), 2)}`, `r {format_dec(quantile(posteriors$b_loglatency_sexm, 0.975), 2)}`]; mean estimated effect of age = `r {format_dec(mean(posteriors$b_loglatency_age),2)}`, 95% CI = [`r {format_dec(quantile(posteriors$b_loglatency_age, 0.025), 2)}`, `r {format_dec(quantile(posteriors$b_loglatency_age, 0.975),2)}`]; see Table S1 in Supplementary Material), where males and older individials made, overall, faster decisions.
+However, we did not see such effects on Choice or Interest (see Table S1 in Supplementary Material).
+Nevertheless, when we analyse the effect of treatment on performance, for every variable recorded, we control for sex by averaging the predicted values for males and females, and we control for age by including the centered age in the model.
+These results are summarized in @fig-results and Table 1.
 
@@
 
-Results: Latency, Choice, and Interest for each test and treatment, references to fig x and Table 1.
-Basically, no effects of treatment on the performance of the numerical discrimination task.
-no discrimination either (refer to Table in Suppl.
-Material).
+Temperature, CORT, or their interaction contrasts did not yield any significant result in any of the numerical tests (Table 1) and we did not observe any clear pattern indicating relevant effects of early environment on any of the tests (@fig-results).
+Except for Hot-CORT lizards in 1 VS 3 tests or Cold-Control lizards in 2 VS 4 tests, the predicted probability of choosing the greater number of crickets first was not different from 0.5 (@fig-results; but also see Table S2 in Supplementary Material).
+Similarly, the predicted interest for the larger number of crickets was not different from 0 in any of the tests (@fig-results; but also see Table S2 in Supplementary Material).
 
@@
 ## Discussion
-- Animals do not discriminate independently of the treatment or the ratio of the numerical task.
-  - Other studies about discrimination
-  - The importance of context
-    - Learned VS spontaneous discrimination
-There are two main approaches to study numerical abilities: spontaneous choice tests and training procedures.
-In spontaneous choice tests, experimenters present to subjects two groups of biologically relevant stimuli that differ in numerosity, under the assumption that animals should select the most appropriate quantity (e.g.
-the larger amount of food) (Agrillo and Bisazza 2014; Cox and Montrose 2016).
-In training
-procedures, animals must learn a numerical rule to receive a reward (Agrillo and Bisazza 2014; Miletto-Petrazzini et al.
-2018).
-Both experimental approaches have some advantages and disadvantages (reviewed in Agrillo and Bisazza 2014).
-In spontaneous choice tests, animals exhibit their natural behavior, enabling interspecific comparisons and the acquirement of ecological issues related to numerousness (Krusche et al.
+Our results indicate that lizards were not able to discriminate between different number of crickets in any of the tests, and that their performance did not vary across treatments.
+We did not see changes in Latency in any of the tests, and treatment did not affect lizards' time to make a decision either.
+Furthermore, in most of the cases, the estimated probability of choosing the option with more crickets first did not differ from random (0.5), with significant differences emerging only in two tests for two of the four treatments; yet no clear pattern was observed.
+Similar findings were obtained on the interest for the larger number of food items, that did not differ from zero in any of the tests or groups.
+These results imply that _L.
+guichenoti_ did not use numerical or other potential quantitative cues we were not controlling for to make foraging decisions.
+
+In out experiment, we employed a spontaneous choice test to investigate _L.
+guichenoti_'s numerical abilities.
+This test is considered a reliable method to assess numeosity in animals, as it allows for the observation of natural behaviour and the acquirement of ecological issues related to numerousness and quantity discrimination (**REFs**).
+However, one of the main caveats of this approach is that motivation plays a key-role in animals' performance (**Agrillo and Bisazza, 2014**).
+Here, we discarded the trials where the lizards did not interact with any of the options, so the results observed are based exclusively in those where the lizards made a choice, and, as such, we do not consider motivation to be a factor affecting our results.
+Spontaneous choice tests are also based on the assumption that animals should select the most appropriate quantity, which may not be necessarily the largest one (**Agrillo and Bisazza, 2014**).
+Although the Optimal Foraging Theory predicts animals to prefer larger amounts of food, this may not hold true if increased handling costs are involved (**REFs**).
+Nonetheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of age on their choice or interest, as older animals are typically larger and may be more adept at handling a greater number of prey.
+Finally, spontaneous choice tests are constraint by the researchers' ability to control for other quantitative cues that may correlate with numerosity (**Agrillo and Bisazza, 2014; Uller et al.
+2003; Krusche et al.
 2010; Abramson et al.
-2011).
-However, motivation plays a key-role in spontaneous choice tests (Agrillo and Bisazza 2014), which implies that the absence of results might not be due to the lack of skills to solve the numerical task.
-Also, individual preferences may arise in this kind of procedures and, hence, a large sample of subjects is needed to consider inter-individual variability.
-Finally, it could be difficult, but necessary, to control for continuous variables such as area, density or movement (Agrillo and Bisazza 2014; Mehlis et al.
-2015).
-Regardless, both methods are considered complementary and necessary to draw a firm picture of the numerical abilities of a species (Miletto-
-Petrazzini et al.
-2018).
-the Italian wall-lizard (Podarcis siculus), spontaneous choice tests showed no numerical discrimination abilities by this lizard (Miletto-Petrazzini et al.
-2017), although learning procedures demonstrated an ability to discern between small close numbers (2 versus 4) (Miletto-Petrazzini et al.
-2018)
-    - Ecological context: food VS predators/conspecifics/etc 
-Something with fishes, food, and conspecifics (**check our paper**)
+2011; Garland et al.
+2012; Vonk and Beran 2012; Gómez-Laplaza and Gerlai 2013; Mehlis et al.
+2015**).
+Although we controlled for some quantitative information by rotating crickets in the option with less items to increase total area, we can not rule out the possibility that lizards could be using other quantitative cues.
+If lizards were using other cues exclusively for making their decision, we would still expect to see some type of discrimination between choices unless the cues were conflicting [**Cooper et al_Journal of Herpetology**].
+Indeed, it could be that _L.
+guichenoti_ uses different quantitative cues at the same time in their foraging decisions, and, when conflicted, the preference towards one or the other varies between individuals, which may produce the same results we observed here.
+Alternatively, it could be that _L.
+guichenoti_ does not use numerical cues in their foraging decisions, and that they rely on other type of information to make their choices.
+Half of the studies showing numerosity in reptiles employed trained quantity discrimination tasks [**Soldati et al.
+2017; Lin et al.
+2021, 2024; Miletto-Petrazzini et al.
+2018; Tomonaga et al.
+2023 Animal Cognition**].
+In the remaining studies, the relevant stimuli involved either vegetables [**Gazzola et al.
+2018**; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
+10)[@recio_prey_2021].
+In contrast, _Podarcis siculus_ failed to discriminate between different numbers of _Musca domestica_ larvae in a spontaneous choice test, but they showed a clear preference towards bigger larvae in a similar task [**Miletto-Petrazzini et al.
+2017**]; and they were able to learn to discriminate between small close numbers in a training procedure [**Miletto-Petrazzini et al.
+2018**].
+This suggest that numerosity may not be relevant in foraging contexts for reptiles that eat live prey.
+In this regard, numerical abilities may have evolved under different situations.
+[**EXAMPLE FISHES THAT NOT FORAGE BUT YES SOCIAL**].
+Further studies should explore numerosity in _L.
+guichenoti_ with a different methodology and accross different contexts.
+**For instance: things about higher numbers (our paper), use different type of relevant stimuli (females, conspecifics?
+Carazo et al and other), use training procedures.**
 
@@
   - Other studies about the effects of hormones on discrimination
+For instance, mammals’ hippocampi contain more GCs receptors (GRs) than other regions of the brain (Siegel et al., 2006), which may be the main cause of overall impairments inspatial learning in rats but not in other cognitive domains (Haq et al., 2021).
   - Studies showing the effect of temperature and CORT on lizards' cognition
   - The 'two systems'
+The second non-verbal system is thought to be an object tracking system (also known as ‘subitizing’) that yields relatively discrete representations.
+This system keeps track of a small number of items (originally set to 6 (REF.
+32), but later restricted to 3 or 4 items) by assign-ing ‘markers’ (also known as ‘fingers’,‘pointers’,‘symbols’, 'tokens’ or ‘files’) to individual elements (**Nieder, 2005 Nature Reviews**)
 Researchers have suggested that numerousness can be processed by two types of systems: one system for small sets (≤ 4) that works by tracking and encoding objects as individual entities; and an analogous system that operates with large numerosities by estimation processes (for a review see Cantlon et al.
@@
 - Conclusions
+Our results indicate that _L.
+guichenoti_ does not use numerical to make foraging decisions.
+This outcome, however, does not imply lack of numerical abilities, since other types of experimental approaches can yield different results.
+We also observed that lizards' performance was not different between treatments, which suggests that the early environment did not affect their numerical abilities.
+However, other tests and paradigmas should be used to test this further since both CORT and temperature can have different effects on different brain nuclei.
 

### 2024-08-26 R/func.r

@@ -3,26 +3,6 @@
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes)
 #
 ####################
-####################
-# Extract sample size per group
-#' @title sample
-#' @description Estract sample size per group
-#' @param corti To select cort treatment ("CORT"/"Control")
-#' @param therm To select temp treatment ("Cold"/"Hot")
-#' @param bias To select the side selected first ("Left"/"Right")
-sample <- function(corti, therm, bias){
-  #Specify database
-  data <- data_num
-  # Count sample
-  sample_size <- data %>%
-                filter(cort == corti, temp == therm, side_choice == bias, test_type == "1 VS 4") %>%
-                group_by(lizard_id) %>%
-                summarise(n = n()) %>%
-                summarise(total_count = sum(n)) %>%
-                pull(total_count)
-  return(sample_size)
-}
-####################
 ####################  
 # Get predictions based on posteriors
 #' @title org_post
@@ -101,7 +81,7 @@ org_post <- function(var){
           for(l in horm){
             vector_m_df <- choose_col(df_m, j, k, l)
             vector_m <- rowSums(vector_m_df)
-            sum_vector_m <- base::sample(vector_m, size = 2500, replace = FALSE)
+            sum_vector_m <- base::sample(vector_m, size = 6000, replace = FALSE)
             column_name_m <- paste0(j, "_", k, "_", l)
             assign(column_name_m, sum_vector_m)
             males[[column_name_m]] <- get(column_name_m)
@@ -117,7 +97,7 @@ org_post <- function(var){
           for(l in horm){
             vector_f_df <- choose_col(df_f, j, k, l)
             vector_f <- rowSums(vector_f_df)
-            sum_vector_f <- base::sample(vector_f, size = 2500, replace = FALSE)
+            sum_vector_f <- base::sample(vector_f, size = 6000, replace = FALSE)
             column_name_f <- paste0(j, "_", k, "_", l)
             assign(column_name_f, sum_vector_f)
             females[[column_name_f]] <- get(column_name_f)

### 2024-08-27 ms.qmd
Updates

@@
 ---
+title: "The effects of prenatal corticosterone and thermal environment on numerical discrimination abilities in a lizard"
 docx: 
@@
 
+Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Chris Friesen^3^, Amelia Y. Peardon ^1^, Daniel W.A. Noble^1^
+
+^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
+^2^ Flinder's University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
+^3^ University of Wollongong, Wollongong, NSW 2500, Australia  
+$\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au
+
+## Abstract
+
+## Significant statement
+
 ## Introduction  
+The ability to discern between more and fewer quantities is a basic cognitive skill widely observed across the animal kingdom [@stancher2015numerical; @vonk2012bears; @beran2016capuchin].
+It is considered adaptive in various ecological contexts, contributing to behaviors such as social decisions, mate selection, or foraging [@lucon2017individual; @nieder2018evolution].
+For instance, female lions (_Panthera leo_) use numerical information to assess the risk of confronting rival groups [@mccomb1994roaring], while other animals use these abilities for navigating their environment [@chittka1995can], locate potential mates [@carazo2009quantity], or to make optimal foraging decisions [@uller2003salamanders].
+Given the adaptive advantages of numerical and quantity discrimination, research has extensively explored the ability of different species to differentiate between quantities under different contexts and paradigmas.
 
-The ability to discern between more and less quantities is a basic cognitive skill widely observed across the animal kingdom (**Stancher et al.
-2015; Vonk and Beran 2012; Beran et al.
-2014**), and is considered to be adaptive in different ecological contexts (**Nieder 2017, 2018**).
-Female lions (_Panthera leo_), for example, can use numerical information to assess the risk of confronting rival groups (**McComb et al.
-1994**); but other animals can employ quantitative abilities for navigation (**Chitka and Geiger 1995**), mate selection (**Carazo et al.
-2009**) or foraging (**Uller et al.
-2003**).
-Given the pervasive advantages of numerical and quantity discrimination, research has extensively explored the ability of different species to differentiate between quantities under different contexts and paradigmas.
-Some studies focus on the type of cues involved in quantitative competences (**REFs**), others on the context where animals may use these abilities (**REFs above**), and others in the neural mechanisms underlying quantitative and numerical discrimination (**REFs**).
-The later have demonstrated that non-verbal numerical representations activate an extensive cortical network, with the prefrontal cortex and the intraparietal sulcus being particularly relevant (**Nieder, 2005**).
-This link between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.
+Numerical competences are often evaluated by considering the types of cues animals use [@plotnik2019elephants; @cooper2024tell], the ecological contexts in which these abilities are deployed [@mccomb1994roaring; @chittka1995can; @uller2003salamanders; @carazo2009quantity], and the neural mechanisms that underlie them [@nieder2005counting].
+The later have demonstrated that non-verbal numerical representations in mammals activate an extensive cortical network, with the prefrontal cortex and the intraparietal sulcus being particularly relevant [@nieder2005counting].
+This connection between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.
 
@@
 For instance, increases in glucocorticoids (GCs) - the main groups of hormones mediating the response to stressors [@sapolsky_how_2000] - during early life can alter animals' performance in spatial, associative, or reversal learning tasks [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
-These effects have been shown to vary depending on factors such as dose, subjects' sex, or type of cognitive test employed [@szuran_water_1994 @farrell_developmental_2015-learn; @bebus_associative_2016], but little is known about how exposure to GCs during development can affect quantitative abilities.
-Similarly, thermal early environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
+These effects have been shown to vary depending on factors such as dose, subjects' sex, or type of cognitive test employed [@szuran_water_1994 @farrell_developmental_2015-learn; @bebus_associative_2016; @haq_prenatal_2021], but little is known about how exposure to GCs during development can affect quantitative abilities.
+At the same time, thermal early environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
 For example, Port Jackson shark incubated at elevated-temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019].
-However, the effects of early thermal environment on numerical abilities in other taxa remain unknown.
-In this sense, not only is it well-established that early-life GCs and thermal environment independently affect cognitive abilities, but these factors are also predicted to interact.
-In vertebrates, abrupt changes in temperature can lead to increased GC levels which, in turn, may elevate GCs in their offspring [@racic_effects_2020; @Crino_2023].
-Simultaneously, these offspring are directly influenced by the same thermal environment, resulting in significant interactions that can impact cognitive development.
-As such, and given the ecological importance of quantitative abilities in nature [**Nieder 2017, 2018**], it is crucial to study the combined effects of early-life GCs and thermal environment on quantity and numerical discrimination.
-Here, we investigate the effects of prenatal exposure to corticosterone (CORT) - the main GC in reptiles [@Crino_2023] - and thermal environment on the ability of lizards to discriminate between different number of food items using the Common garden skink (_Lampropholis guichenoti)_ as a model species.
+However, the effects of early thermal environment on numerical abilities in other taxa remains largely unexplored.
 
-Traditionally, the study of numerical abilities has been mainly focused on mammals, birds, and fishes (**Agrillo and Bisazza 2014; Nieder 2017**), but recent research has shown that reptiles are also competent in numbers.
-Red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour to higher volumes of food (**Soldati et al.
-2017**), Hermann’s tortoises (_Testudo hermanni_) proved to discriminate between big and small amounts off food in a spontaneous choice test (**Gazzola et al.
-2018**), and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) showed high skills when tested in a trained quantity discrimination task (**Lin et al.
-2021, 2024**).
-In lizards, __Podarcis siculus_ did not discriminate between different amounts of food in a spontaneous choice test (**Miletto-Petrazzini et al.
-2017**), but were successful in a trained quantity discrimination task (**Miletto-Petrazzini et al.
-2018**), rock lizards (_Iberolacerta cyreni_) chose bigger quantities of crickets when presented 5 versus 10 items [@recio_prey_2021], and Gidgee skinks (_Egernia stokesii_) used numbers to select the larger amount of food in a spontaneous choice test [@szabo_spontaneous_2021].
+Given that abrupt temperature changes can increase GC levels in vertebrates, leading to elevated GCs in offspring [@racic_effects_2020; @Crino_2023], it is essential to study the combined effects of early-life GCs and thermal conditions on cognitive development.
+Understanding how these factors influence quantitative abilities could reveal important insights into the ecological relevance of numerosity.
+To fill this gap, we used the common garden skink (_Lampropholis guichenoti_) as a model to test the effects of prenatal corticosterone (CORT) - the main GC in reptiles [@Crino_2023] - exposure and thermal environment on numerical discrimination abilities.
+
+Traditionally, research on numerical abilities has been mainly focused on mammals, birds, and fishes [@nieder2005counting; @agrillo2014spontaneous], but recent studies demonstrate that reptiles are also competent in numbers.
+For example, red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour to higher volumes of food [@soldati2017long], Hermann’s tortoises (_Testudo hermanni_) proved to discriminate between big and small food quantities in a spontaneous choice test [@gazzola2018continuous], and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) showed high skills when tested in a trained quantity discrimination task [@lin2021superior; @lin2024trained].
+In lizards, __Podarcis siculus_ failed to discriminate between different number of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity], whereas rock lizards (_Iberolacerta cyreni_) exhibited a clear preference for larger quantities of crickets when presented 5 versus 10 items [@recio_prey_2021].
 Inded, the literature evidences basic quantitative abilities in reptiles, that may vary with the type of cue and test employed.
-However, the question of how early environment can affect numerical abilities remains open as, to our knowledge, there is no study investigating how early life conditions impact numerical discrimination in this group.
+
+At the same time, reptiles' cognition is known to be significantly influenced by the environment in which they develop.
 Incubating lizards at elevated temperatures has been shown to improve cognitive abilities in consistency with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017]; and thermal environment impaired cognition only under incubation temperatures beyond lizards' natural range [@dayananda_incubation_2017 or @abayarathna_effects_2020].
-Our understanding of the effects of prenatal CORT on cognitive abilities in reptiles is less understood, but in other taxa, high levels of GCs are tipically related to cognitive impairments [@szuran_water_1994; @zhu_prenatal_2004].
+The effects of prenatal CORT on cognitive abilities in reptiles is less understood, but in other taxa, high levels of GCs are tipically related to cognitive impairments [@szuran_water_1994; @zhu_prenatal_2004].
+However, the combined influence of prenatal CORT and thermal environment remains unknown, particularly on numerical abilities.
 
-In our study, we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 28ºC ± 3ºC) in a 2x2 factorial design.
-After hatching, we assessed the lizards' numerical discrimination abilities in a spontaneous choice test, where they selected between different numbers of crickets.
-We hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task.
-Additionally, we predicted that increased CORT levels would impair numerical competence.
-Moreover, we expected the negative impact of CORT on numerical performance to be more pronounced in lizards incubated at lower temperatures compared to those in higher temperatures.
-In line with Weber's law [**REFs**], we further anticipated that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g.
-3 VS 4), whereas event the lowest ratios (1 VS 4) could be challenging for those incubated with CORT.
+Our study aimed to address this gap by manipulating CORT concentration in _L.
+guichenoti_ eggs and incubating them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 28ºC ± 3ºC) in a 2x2 factorial design.
+After hatching, we assessed the lizards' numerical discrimination abilities in a spontaneous choice test, where they were presented with different numbers of crickets.
+Based in other studies that explored the cognitive effect of prenatal GCs or thermal early environment in reptiles and other taxa, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
+Additionally, we predicted that increased CORT levels would impair numerical competence [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures temperatures.
+Consistent with Weber's law [@agrillo2014spontaneous], we expected that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g.
+3 VS 4), whereas event the lowest ratios (e.g.
+1 VS 4) could be challenging for those exposed to CORT and/or incubated at lower temperatures.
 
@@
 We performed five tests that differed in the ratio between both choices: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4.
-We never used more than four items because higher numbers are considered to be encoded by different regions of the brain [**REFs**].
+We never used more than four items because higher numbers are considered to be represented by a different system [@nieder2005counting; @cantlon2009beyond; @hyde2011two].
 The food consisted in frozen small crickets (*A.
@@
 We also counterbalanced the amount of individuals per treatment tested in the same type of trial each day.
-The crickets used during the experiment were all the same size, but to control for other quantitative traits
-(**Uller et al.
-2003; Krusche et al.
-2010; Abramson et al.
-2011; Garland et al.
-2012; Vonk and Beran 2012; Gómez-Laplaza and Gerlai 2013; Mehlis et al.
-2015**), we changed the orientation of the crickets as shown in @fig-Methods D.
+The crickets used during the experiment were all the same size, but to control for other quantitative traits [@agrillo2014spontaneous], we changed the orientation of the crickets as shown in @fig-Methods D.
 In addition, after the experiments, we made another set of trials to test whether the orientation of the crickets could affect lizards' choice, but we found no effect of orientation (see Supplementary Material - Prey orientation test).
@@
 Furthermore, in most of the cases, the estimated probability of choosing the option with more crickets first did not differ from random (0.5), with significant differences emerging only in two tests for two of the four treatments; yet no clear pattern was observed.
-Similar findings were obtained on the interest for the larger number of food items, that did not differ from zero in any of the tests or groups.
-These results imply that _L.
+These findings were further reinforced by similar results regarding the interest for the larger number of food items, that did not differ from zero in any of the tests or groups.
+Altogether, these findings suggest that _L.
 guichenoti_ did not use numerical or other potential quantitative cues we were not controlling for to make foraging decisions.
 
-In out experiment, we employed a spontaneous choice test to investigate _L.
+In our experiment, we employed a spontaneous choice test to investigate _L.
 guichenoti_'s numerical abilities.
-This test is considered a reliable method to assess numeosity in animals, as it allows for the observation of natural behaviour and the acquirement of ecological issues related to numerousness and quantity discrimination (**REFs**).
-However, one of the main caveats of this approach is that motivation plays a key-role in animals' performance (**Agrillo and Bisazza, 2014**).
-Here, we discarded the trials where the lizards did not interact with any of the options, so the results observed are based exclusively in those where the lizards made a choice, and, as such, we do not consider motivation to be a factor affecting our results.
-Spontaneous choice tests are also based on the assumption that animals should select the most appropriate quantity, which may not be necessarily the largest one (**Agrillo and Bisazza, 2014**).
-Although the Optimal Foraging Theory predicts animals to prefer larger amounts of food, this may not hold true if increased handling costs are involved (**REFs**).
-Nonetheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of age on their choice or interest, as older animals are typically larger and may be more adept at handling a greater number of prey.
-Finally, spontaneous choice tests are constraint by the researchers' ability to control for other quantitative cues that may correlate with numerosity (**Agrillo and Bisazza, 2014; Uller et al.
-2003; Krusche et al.
-2010; Abramson et al.
-2011; Garland et al.
-2012; Vonk and Beran 2012; Gómez-Laplaza and Gerlai 2013; Mehlis et al.
-2015**).
+This test is considered a reliable method to assess numerosity in animals, as it allows for the observation of natural behaviour and the acquirement of ecological issues related to numerousness and quantity discrimination [@nieder2005counting; @agrillo2014spontaneous].
+However, one of the main caveats of this approach is that motivation plays a key-role in animals' performance [@agrillo2014spontaneous; @bisazza2014experim].
+Here, we discarded the trials where the lizards did not interact with any of the options, so the results observed are based exclusively in those where the lizards made a choice.
+Therefore we do not consider motivation to be a factor affecting our results.
+Moreover, spontaneous choice tests are based on the assumption that animals should select the most appropriate quantity, which may not be necessarily the largest one [@agrillo2014spontaneous].
+Although spontaneous choice tests employing food assume animals to prefer larger amounts of food, this may not hold true if increased handling costs are involved [@agrillo2014spontaneous].
+Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of age on their choice or interest, as older animals are typically larger and may be more adept at handling a greater number of prey.
+
+Another consideration is the limitation of spontaneous choice tests to control for other quantitative cues that may correlate with numerosity [@agrillo2014spontaneous].
 Although we controlled for some quantitative information by rotating crickets in the option with less items to increase total area, we can not rule out the possibility that lizards could be using other quantitative cues.
-If lizards were using other cues exclusively for making their decision, we would still expect to see some type of discrimination between choices unless the cues were conflicting [**Cooper et al_Journal of Herpetology**].
-Indeed, it could be that _L.
-guichenoti_ uses different quantitative cues at the same time in their foraging decisions, and, when conflicted, the preference towards one or the other varies between individuals, which may produce the same results we observed here.
-Alternatively, it could be that _L.
-guichenoti_ does not use numerical cues in their foraging decisions, and that they rely on other type of information to make their choices.
-Half of the studies showing numerosity in reptiles employed trained quantity discrimination tasks [**Soldati et al.
-2017; Lin et al.
-2021, 2024; Miletto-Petrazzini et al.
-2018; Tomonaga et al.
-2023 Animal Cognition**].
-In the remaining studies, the relevant stimuli involved either vegetables [**Gazzola et al.
-2018**; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
+If that were the case, however, we would still see some type of discrimination between choices unless the cues were conflicting [@cooper2024tell].
+Indeed, it is possible that _L.
+guichenoti_ uses multiple quantitative cues simoultaneously during foraging decisions, and, when those cues conflict, the preference for one cue over another might vary between individuals, potentially producing the results observed here.
+Alternatively, _L.
+guichenoti_ may not rely on numerical cues for foraging decisions, instead utilizing other type of information.
+
+Interestingly, half of the studies demonstrating numerosity in reptiles employed trained quantity discrimination tasks [@soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained].
+In the remaining successful studies, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
 10)[@recio_prey_2021].
-In contrast, _Podarcis siculus_ failed to discriminate between different numbers of _Musca domestica_ larvae in a spontaneous choice test, but they showed a clear preference towards bigger larvae in a similar task [**Miletto-Petrazzini et al.
-2017**]; and they were able to learn to discriminate between small close numbers in a training procedure [**Miletto-Petrazzini et al.
-2018**].
-This suggest that numerosity may not be relevant in foraging contexts for reptiles that eat live prey.
-In this regard, numerical abilities may have evolved under different situations.
-[**EXAMPLE FISHES THAT NOT FORAGE BUT YES SOCIAL**].
-Further studies should explore numerosity in _L.
-guichenoti_ with a different methodology and accross different contexts.
-**For instance: things about higher numbers (our paper), use different type of relevant stimuli (females, conspecifics?
-Carazo et al and other), use training procedures.**
+However, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities [@miletto2017quantitative] despite showing high quantitative skills when size of prey was the main cue [@miletto2017quantitative], or when tested using training procedures [@miletto2018quantity].
+Hence, it is possible that further studies exploring numerosity in _L.
+guichenoti_ using different methodologies or contexts, such as larger quantities [@nieder2005counting; @recio_prey_2021], inclusion of chemical stimuli [@carazo2009quantity; @plotnik2019elephants; @recio_prey_2021], using relevant stimuli like conspecifics [@carazo2009quantity; @lucon2017individual], or use training procedures [@agrillo2014spontaneous; @bisazza2014experim; @soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained], may yield more conclusive results.
 
-- Treatment did not affect the discrimination capacity or the decision making
-  - Other studies showing the effect of temperature on discrimination
-  - Other studies about the effects of hormones on discrimination
-For instance, mammals’ hippocampi contain more GCs receptors (GRs) than other regions of the brain (Siegel et al., 2006), which may be the main cause of overall impairments inspatial learning in rats but not in other cognitive domains (Haq et al., 2021).
-  - Studies showing the effect of temperature and CORT on lizards' cognition
-  - The 'two systems'
-The second non-verbal system is thought to be an object tracking system (also known as ‘subitizing’) that yields relatively discrete representations.
-This system keeps track of a small number of items (originally set to 6 (REF.
-32), but later restricted to 3 or 4 items) by assign-ing ‘markers’ (also known as ‘fingers’,‘pointers’,‘symbols’, 'tokens’ or ‘files’) to individual elements (**Nieder, 2005 Nature Reviews**)
-Researchers have suggested that numerousness can be processed by two types of systems: one system for small sets (≤ 4) that works by tracking and encoding objects as individual entities; and an analogous system that operates with large numerosities by estimation processes (for a review see Cantlon et al.
-2009; Hyde 2011).
-**Are they encoded in different regions?
-Brin-nuclei-dependent effects of CORT and temp**
+We did not find any effect of prenatal conditions on _L.
+guichenoti_'s ability to use numerical information for making foraging decisions.
+We anticipated that lower incubation temperatures and exposure to CORT will impair lizards' performance, especially in those tests with higher ratios following Weber's Law [@agrillo2014spontaneous; @nieder2018evolution].
+However, lizards' behaviour was not impacted by our treatments, suggesting that early environmental conditions do not affect numerical competence in this species.
+Our predictions were based in previous studies showing the effects of prenatal temperature and GCs on cognitive abilities in reptiles and other groups [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic].
+However, these effects can be species-specific, varying across cognitive domains and test types [@szuran_water_1994; @farrell_developmental_2015-learn; @bebus_associative_2016;@dayananda_incubation_2017; @abayarathna_effects_2020; @haq_prenatal_2021].
 
-- Conclusions
-Our results indicate that _L.
-guichenoti_ does not use numerical to make foraging decisions.
-This outcome, however, does not imply lack of numerical abilities, since other types of experimental approaches can yield different results.
-We also observed that lizards' performance was not different between treatments, which suggests that the early environment did not affect their numerical abilities.
-However, other tests and paradigmas should be used to test this further since both CORT and temperature can have different effects on different brain nuclei.
+In this context, it is important to consider that both GCs and thermal environment can exert differential effects on distinct regions of the brain.
+For example, incubation temperature impacts mitochondrial activity and cell density in specific brain nuclei [@coomber_independent_1997; @amiel_effects_2017].
+Similarly, impairments in spatial learning - but not in other learning tasks - in rats [@haq_prenatal_2021] are likely associated with the greater density of GCs receptors mammals' hippocampi compared to other regions of the brain [@siegel_2006].
+While it remains unknown which regions of the reptiles' brain encode numerical information, our results suggest that relevant regions may be unaffected by the prenatal conditions tested here.
+However, researchers have proposed that non-verbal number representation is processed by two types of systems: one for small sets (≤ 4) that tracks and encodes objects as individual entities, and another and another for larger numerosities that operates through estimation processes [@nieder2005counting; @cantlon2009beyond; @hyde2011two].
+Our data indicates that at least those brain regions involved in discrete numerical representation might be robust to the effect of prenatal CORT and temperature in _L.
+guichenoti_.
+Studies using higher quantities of food items may help elucidate the effects of prenatal conditions on _L.
+guichenoti_'s numerical abilities.
+Alternatively, it is also plausible that treatment effects are masked by the overall poor performance across most of the tests; which could be a consequence of the species' relying in other type of information for foraging rather than numerical information.
+Thus, future research should investigate the effects of prenatal CORT and thermal environment on numerical competence by employing trained quantity discrimination tasks [@soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained].
+
+In conclusion, our study is, to the best of our knowledge, the first to explore the impact of prenatal conditions on numerical abilities in reptiles.
+Our findings suggest that _L.
+guichenoti_ does not rely on numerical information for foraging decisions and that their performance does not vary across different prenatal treatments.
+This implies that early environmental conditions, such as incubation temperature and CORT exposure, may not affect numerosity in this species.
+However, it is critical to note that the methodologies employed in this study may not have fully captured the nuances of numerical cognition in _L.
+guichenoti_.
+Therefore, we recommend that future studies employ a broader array of tests and paradigms, including training-based tasks, larger sets of stimuli, and other sensory modalities, to more comprehensively assess numerical competence in this and related species.
+Such work will be pivotal in unraveling the effects of early environmental factors on cognitive abilities in reptiles and advancing our understanding of numerosity across taxa.
 
@@
 
-#### Checking the models plots  
+#### Checking the model plots  
 ```{r, checking_model, out.width="70%", fig.align="center"}

### 2024-08-27 ms.qmd
Updates

@@
 docx: 
+  reference-doc: "./bib/tmpl.docx"
 bibliography: "./bib/refs.bib"
 csl: "./bib/behavioral-ecology-and-sociobiology.csl"
-reference-doc: "./bib/tmpl.docx"
 execute:
@@
 
+ORCID:
+
+Pablo Recio ORCID: 0000-0002-5890-0218  
+Dalton C. Leibold ORCID: 0000-0001-9645-2033  
+Ondi L. Crino ORCID: 0000-0001-5700-1387  
+Chris Friesen ORCID: 0000-0001-5338-7454  
+Amelia Y. Peardon ORCID: 0009-0005-6227-8379  
+Daniel W.A. Noble ORCID: 0000-0001-9460-8743
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+## Keywords
+Numerical discrimination, Corticosterone, Temperature, Reptiles, Early life stress, Cognitive abilities
+
 ## Abstract
+Animals use quantitative judgements in several ecological contexts.
+From social behaviour to navigating their own environment, quantitative abilities can be crucial in animals' behaviour.
+Since prenatal conditions are predicted to impact brain development, quantitative competence may be as well affected by the early environment.
+Factors like temperature or stress-related hormones transmitted by the parents can play a prominent role in shaping cognitive traits.
+In adition, abrupt temperature changes can lead to the transmittion of stress-related hormones to the offspring, making essential to study the combined effects of stress hormones and thermal conditions during development.
+Here, we investigated the effects of prenatal corticosterone (CORT) and temperature on numerical discrimination in the Common garden skink (_Lampropholis guichenoti_).
+We manipulated CORT levels and incubation temperature of eggs employing a 2x2 factorial design, and then assessed numerousness through a spontaneous choice test with food as the relevant stimulus.
+Animals were subjected to five different tests with differen number of items and ratios (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 2 VS 4) in a repeated measures design.
+We hypothesized that CORT exposure and cold temperatures would constraint performance in the numerical task.
+Contrary to our expectations, we did not find any evidence of numerical discrimination for any of the treatments.
+These findings suggest that _L.
+guichenoti_ may not use numerical cues to make foraging decisions, and that early conditions does not affect their behaviour towards different quantities of food.
 
 ## Significant statement
+Discriminating between quantities is crucial for animal behaviour, yet the impact of early environmental factors on the development of quantitative competence remains largely unexplored.
+In this study, we investigated for the first time the effects of prenatal corticosterone exposure and incubation temperature on numerical discrimination in a reptile.
+Contrary to our predictions, we did not observe any evidence of numerical discrimination regardless of incubation conditions.
+Our results align with previous studies showing that reptiles consuming live prey struggle to differentiate prey numbers in spontaneous choice tests.
+These results open new avenues for exploring the evolution of numerical abilities in reptiles and identifying factors influencing the development of quantitative skills in this group.
+Further research should explore alternative methods, such as trained tasks, to better understand the cognitive mechanisms underlying numerical competence in reptiles.
 
@@
 #| label: fig-Methods
-#| fig.cap: "Experimental design of early environment manipulation (**A**), numerical discrimination arena an measures(**B**), measures of the platform used for the experiments (**C**), and types of numerical trial (**D**), where each line indicates the position and orientation of the food items in each test."
+#| fig.cap: "Experimental design of the study.
+Panel (**A**) shows the early environment manipulation procedures.
+(**B**) illustrates the arena where the tests were performed, while panel (**C**) indicates the measurements of the platform used for the experiments.
+Finally, panel (**D**) displays the types of numerical tests used and the orientation of the crickets in each test."
 
@@
 #| label: fig-results
-#| fig.cap: "Mean estimates of loglatency (Latency), the probability of choosing first the larger number of crickets (Choice), and the mean estimated interest for the larger number of crickets (Interest) per each of the treatments for each of the numerical tests performed.
+#| fig.cap: "Mean estimates of loglatency (Latency), the probability of choosing first the larger number of crickets (Choice), and the estimated interest for the larger number of crickets (Interest) per each of the treatments for each of the numerical tests performed.
 The x-axis represents the estimate, and in the y-axis are the density of the estimates.
@@
 Points and bars represent the mean and standard deviation of the mean of the estimates, respectively.
-Vertical dashed lines in Choice and Interest graphs indicate p = 0.5 and int = 0, respectively."
+Vertical dashed lines in Choice and Interest graphs values 0.5 and 0, respectively."
 #

### 2024-08-27 ms.qmd
Updates

@@
 Effects of temperature, CORT, and their interaction on latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
-The table shows the comparissons for each predictor (Temperature = [meanHot - meanCold]; Hormone = [meanControl - meanCORT]; Interaction = [(meanHot-Control - meanHot-CORT) - (meanCold-Control - meanCold-CORT)]) and the p~mcmcm~ value of the comparisons.
+The table shows the comparissons for each predictor (Temperature = [meanHot - meanCold]; Hormone = [meanControl - meanCORT]; Interaction = [(meanHot-Control - meanHot-CORT) - (meanCold-Control - meanCold-CORT)]) and the p~mcmc~ value of the comparisons.
 In bold, the effects where p~mcmc~ < 0.05._
@@
 Such work will be pivotal in unraveling the effects of early environmental factors on cognitive abilities in reptiles and advancing our understanding of numerosity across taxa.
+
+## Data accessibility  
+All data, data description, and R code are available in public repository [https://github.com/Pablo-Recio/CORT-Temp_Numerical](https://github.com/Pablo-Recio/CORT-Temp_Numerical).
+
+## Declaration of AI use  
+We declare Chat GPT was used for questions related to coding and data analyses.
+All other parts of the manuscript were written by the authors and those parts where AI was used were checked and modified accordingly by the authors.
+
+## Authors' contributions  
+P.R.: conceptualization, methodology, data collection, data curation, formal analysis, writing—original draft, writing—review and editing; D.C.L.: data collection, writing—review and editing; O.C.: conceptualization, writing—review and editing; C.F.: conceptualization, funding acquisition, writing—review and editing; A.Y.P.: data collection, writing—review and editing; D.N.: conceptualization, methodology, funding acquisition, project administration, resources, supervision, writing—review and editing.
+All authors gave final approval for publication and agreed to be held accountable for the work performed therein.
+
+## Conflict of interest declaration  
+We declare we have no competing interests.
+
+## Funding  
+This work was supported by a National Australian University fellowship (P.R.), and the Australian Research Council (grant no. DP210101152) to D.N. and C.F.
+
+## Acknowledgements  
+We thank the help and assistance of our lab technicians Benjamin Durant and Michelle Stephens for taking care of the lizards.
+We are also grateful ANU MakerSpace, where we designed and built the prototypes of the 3D-printed platforms.
+Finally, we wish to acknowledge the anonymous reviewers for their valuable feedback on the manuscript.
 

### 2024-09-22 ms.qmd
Updates

@@
 ---
-title: "The effects of prenatal corticosterone and thermal environment on numerical discrimination abilities in a lizard"
+title: "Early environment and numerical discrimination in a lizard"
 docx: 
@@
 
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Chris Friesen^3^, Amelia Y. Peardon ^1^, Daniel W.A. Noble^1^
+Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen^3^, Amelia Y. Peardon ^1^, Daniel W.A. Noble^1^
 
@@
 Ondi L. Crino ORCID: 0000-0001-5700-1387  
-Chris Friesen ORCID: 0000-0001-5338-7454  
+Christopher R. Friesen ORCID: 0000-0001-5338-7454  
 Amelia Y. Peardon ORCID: 0009-0005-6227-8379  
@@
 ## Abstract
-Animals use quantitative judgements in several ecological contexts.
-From social behaviour to navigating their own environment, quantitative abilities can be crucial in animals' behaviour.
-Since prenatal conditions are predicted to impact brain development, quantitative competence may be as well affected by the early environment.
-Factors like temperature or stress-related hormones transmitted by the parents can play a prominent role in shaping cognitive traits.
-In adition, abrupt temperature changes can lead to the transmittion of stress-related hormones to the offspring, making essential to study the combined effects of stress hormones and thermal conditions during development.
-Here, we investigated the effects of prenatal corticosterone (CORT) and temperature on numerical discrimination in the Common garden skink (_Lampropholis guichenoti_).
-We manipulated CORT levels and incubation temperature of eggs employing a 2x2 factorial design, and then assessed numerousness through a spontaneous choice test with food as the relevant stimulus.
-Animals were subjected to five different tests with differen number of items and ratios (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 2 VS 4) in a repeated measures design.
-We hypothesized that CORT exposure and cold temperatures would constraint performance in the numerical task.
-Contrary to our expectations, we did not find any evidence of numerical discrimination for any of the treatments.
+From social behaviour to navigating their environment, quantitative abilities can be crucial in animals' life.
+Early life conditions are predicted to impact brain development, and quantitative competence may not be exent of these effects.
+Early thermal environment or glucocorticoids (GCs) - stress-related hormones - transmitted by the parents can play a prominent role in shaping cognition through effects on brain physiology.
+Furthermore, temperature can lead to increased levels of GCs that are passed on to the offspring, emphasizing the need to study the combined effects of these two factors.
+Here, we investigated the effects of prenatal corticosterone (CORT) - the main GC in reptiles - and incubation temperature on numerical discrimination in the common garden skink (_Lampropholis guichenoti_).
+We manipulated CORT levels and temperature during incubation and then assessed numerical abilities through a spontaneous choice test with food as the relevant stimulus.
+Employing a repeated measures design, we subjected lizards to five numerical tests each differing in the total number of items and the ratios between the two choices (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 2 VS 4).
+Contrary to our expectations, we found no evidence of numerical discrimination or any effect of treatments on lizards' behaviour.
 These findings suggest that _L.
-guichenoti_ may not use numerical cues to make foraging decisions, and that early conditions does not affect their behaviour towards different quantities of food.
+guichenoti_ may not use numerical cues to make foraging decisions and that early conditions do not affect their behaviour towards different quantities of food.
 
-## Significant statement
+## Significance statement
 Discriminating between quantities is crucial for animal behaviour, yet the impact of early environmental factors on the development of quantitative competence remains largely unexplored.
@@
 These results open new avenues for exploring the evolution of numerical abilities in reptiles and identifying factors influencing the development of quantitative skills in this group.
-Further research should explore alternative methods, such as trained tasks, to better understand the cognitive mechanisms underlying numerical competence in reptiles.
+Further research should explore alternative methods, such as trained tasks, to better understand the cognitive mechanisms underlying quantitative competence in reptiles.
 
 ## Introduction  
-The ability to discern between more and fewer quantities is a basic cognitive skill widely observed across the animal kingdom [@stancher2015numerical; @vonk2012bears; @beran2016capuchin].
-It is considered adaptive in various ecological contexts, contributing to behaviors such as social decisions, mate selection, or foraging [@lucon2017individual; @nieder2018evolution].
-For instance, female lions (_Panthera leo_) use numerical information to assess the risk of confronting rival groups [@mccomb1994roaring], while other animals use these abilities for navigating their environment [@chittka1995can], locate potential mates [@carazo2009quantity], or to make optimal foraging decisions [@uller2003salamanders].
-Given the adaptive advantages of numerical and quantity discrimination, research has extensively explored the ability of different species to differentiate between quantities under different contexts and paradigmas.
+The ability to discern between quantities is a fundamental cognitive skill widely observed across the animal kingdom [@stancher2015numerical; @vonk2012bears; @beran2016capuchin].
+Numerical cognition is adaptive in various ecological contexts, contributing to social decisions, mate selection, or foraging [@lucon2017individual; @nieder2018evolution].
+For instance, female lions (_Panthera leo_) use numerical information to assess the risk of confronting rival groups [@mccomb1994roaring], male adult mealworms (_Tenebrio molitor_) select sites with scents of more females [@carazo2009quantity], and red-backed salamanders (_Plethodon cinereus_) prefer tubes containing larger numbers of prey [@uller2003salamanders].
+Given the adaptive advantages of numerical and quantity discrimination, research has extensively explored the ability of different species to differentiate between quantities under different contexts and paradigms.
 
-Numerical competences are often evaluated by considering the types of cues animals use [@plotnik2019elephants; @cooper2024tell], the ecological contexts in which these abilities are deployed [@mccomb1994roaring; @chittka1995can; @uller2003salamanders; @carazo2009quantity], and the neural mechanisms that underlie them [@nieder2005counting].
-The later have demonstrated that non-verbal numerical representations in mammals activate an extensive cortical network, with the prefrontal cortex and the intraparietal sulcus being particularly relevant [@nieder2005counting].
+Numerical competence is often evaluated by considering the types of cues animals use [@plotnik2019elephants; @cooper2024tell], the ecological contexts in which these abilities are deployed [@mccomb1994roaring; @chittka1995can; @uller2003salamanders; @carazo2009quantity], and the neural mechanisms that underlie them [@nieder2005counting].
+The latter have demonstrated that non-verbal numerical representations in mammals activate an extensive cortical network, with the prefrontal cortex and the intraparietal sulcus being particularly relevant [@nieder2005counting].
 This connection between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.
 
-Animal's brain is particularly sensitive to environmental inputs during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004], with long-lasting effects on cognitive abilities [@zhu_prenatal_2004; @amiel_egg_2014].
-For instance, increases in glucocorticoids (GCs) - the main groups of hormones mediating the response to stressors [@sapolsky_how_2000] - during early life can alter animals' performance in spatial, associative, or reversal learning tasks [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
-These effects have been shown to vary depending on factors such as dose, subjects' sex, or type of cognitive test employed [@szuran_water_1994 @farrell_developmental_2015-learn; @bebus_associative_2016; @haq_prenatal_2021], but little is known about how exposure to GCs during development can affect quantitative abilities.
-At the same time, thermal early environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
-For example, Port Jackson shark incubated at elevated-temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019].
+Animals' brains are particularly sensitive to environmental inputs during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004], with long-lasting effects on cognitive abilities [@zhu_prenatal_2004; @amiel_egg_2014].
+For instance, glucocorticoids (GCs), the main groups of hormones mediating the physiological stress response [@sapolsky_how_2000], can alter animals' performance in spatial, associative, or reversal learning tasks if elevated during early stages of development [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
+Prenatal exposure to GCs is typically associated with impaired learning skills, but these effects have been shown to vary depending on factors such as dose, subjects' sex, or type of cognitive test employed [@szuran_water_1994; @farrell_developmental_2015-learn; @bebus_associative_2016; @haq_prenatal_2021].
+Nevertheless, little is known about how exposure to GCs during development can affect quantitative abilities.
+
+At the same time, the early thermal environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
+For example, Port-Jackson sharks (_Heterodontus portusjacksoni_) incubated at elevated temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019].
 However, the effects of early thermal environment on numerical abilities in other taxa remains largely unexplored.
 
-Given that abrupt temperature changes can increase GC levels in vertebrates, leading to elevated GCs in offspring [@racic_effects_2020; @Crino_2023], it is essential to study the combined effects of early-life GCs and thermal conditions on cognitive development.
+Given that abrupt temperature changes can increase GC levels in vertebrates, leading to elevated GCs transmitted to the offspring [@racic_effects_2020; @Crino_2023], it is essential to study the combined effects of early-life GCs and thermal conditions on cognitive development.
 Understanding how these factors influence quantitative abilities could reveal important insights into the ecological relevance of numerosity.
-To fill this gap, we used the common garden skink (_Lampropholis guichenoti_) as a model to test the effects of prenatal corticosterone (CORT) - the main GC in reptiles [@Crino_2023] - exposure and thermal environment on numerical discrimination abilities.
+To fill this gap, we used the common garden skink (_Lampropholis guichenoti_) as a model to test the effects of elevated prenatal corticosterone (CORT) - the main GC in reptiles [@Crino_2023] - and incubation temperature on numerical discrimination abilities.
 
-Traditionally, research on numerical abilities has been mainly focused on mammals, birds, and fishes [@nieder2005counting; @agrillo2014spontaneous], but recent studies demonstrate that reptiles are also competent in numbers.
-For example, red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour to higher volumes of food [@soldati2017long], Hermann’s tortoises (_Testudo hermanni_) proved to discriminate between big and small food quantities in a spontaneous choice test [@gazzola2018continuous], and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) showed high skills when tested in a trained quantity discrimination task [@lin2021superior; @lin2024trained].
-In lizards, __Podarcis siculus_ failed to discriminate between different number of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity], whereas rock lizards (_Iberolacerta cyreni_) exhibited a clear preference for larger quantities of crickets when presented 5 versus 10 items [@recio_prey_2021].
-Inded, the literature evidences basic quantitative abilities in reptiles, that may vary with the type of cue and test employed.
+Traditionally, research on numerical abilities has been mainly focused on mammals, birds, and fishes [@nieder2005counting; @agrillo2014spontaneous], but recent studies demonstrate that reptiles are also competent in numberical abilities.
+For example, red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour with higher volumes of food [@soldati2017long], Hermann’s tortoises (_Testudo hermanni_) proved to discriminate between large and small food quantities in a spontaneous choice test [@gazzola2018continuous], and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) rapidly learned a quantity discrimination task [@lin2021superior; @lin2024trained].
+The italian wall lizard (_Podarcis siculus_) failed to discriminate between different numbers of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity].
+In contrast, Iberian rock lizards (_Iberolacerta cyreni_) preferred larger quantities of crickets when presented five versus ten items [@recio_prey_2021].
+Indeed, the literature shows evidence for basic quantitative abilities in reptiles that may vary with the type of cue and the test employed.
 
-At the same time, reptiles' cognition is known to be significantly influenced by the environment in which they develop.
-Incubating lizards at elevated temperatures has been shown to improve cognitive abilities in consistency with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017]; and thermal environment impaired cognition only under incubation temperatures beyond lizards' natural range [@dayananda_incubation_2017 or @abayarathna_effects_2020].
-The effects of prenatal CORT on cognitive abilities in reptiles is less understood, but in other taxa, high levels of GCs are tipically related to cognitive impairments [@szuran_water_1994; @zhu_prenatal_2004].
-However, the combined influence of prenatal CORT and thermal environment remains unknown, particularly on numerical abilities.
+Reptiles' cognition is significantly influenced by the environment in which they develop.
+For example, incubating lizards at elevated temperatures improves cognitive abilities consistent with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017], but impairs cognition at temperatures above lizards’ natural incubation thermal range [@dayananda_incubation_2017; @abayarathna_effects_2020].
+The effects of prenatal CORT on cognitive abilities in reptiles is less understood, but in other taxa, high levels of GCs are typically related to cognitive impairments [@szuran_water_1994; @zhu_prenatal_2004].
+However, the combined influence of prenatal CORT and thermal environment on cognition remains unknown, particularly on numerical abilities.
 
-Our study aimed to address this gap by manipulating CORT concentration in _L.
-guichenoti_ eggs and incubating them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 28ºC ± 3ºC) in a 2x2 factorial design.
+Our study aimed to address this question by manipulating CORT concentration (CORT-treated or a sham control) in _L.
+guichenoti_ eggs and incubating them under one of two temperature regimes (Cold - 23 ± 3ºC or Hot - 28 ± 3ºC) in a 2x2 factorial design.
 After hatching, we assessed the lizards' numerical discrimination abilities in a spontaneous choice test, where they were presented with different numbers of crickets.
-Based in other studies that explored the cognitive effect of prenatal GCs or thermal early environment in reptiles and other taxa, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
-Additionally, we predicted that increased CORT levels would impair numerical competence [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures temperatures.
-Consistent with Weber's law [@agrillo2014spontaneous], we expected that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g.
-3 VS 4), whereas event the lowest ratios (e.g.
-1 VS 4) could be challenging for those exposed to CORT and/or incubated at lower temperatures.
+Based on other studies that explored the cognitive effects of prenatal GCs or thermal early environment in reptiles and other taxa, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
+Additionally, we predicted that increased CORT levels would impair numerical competence [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
+In this sense, a common observation in nature is that the ability to detect a change in stimulus intensity depends on the relative, not absolute, difference between the stimuli ['Weber's law'; @agrillo2014spontaneous].
+Consistent with Weber's law, we predicted that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g.
+3 VS 4), whereas even the lowest ratios (e.g.
+1 VS 4) could be challenging for those exposed to CORT or incubated at lower temperatures.
 
@@
 
-#### Husbandry  
-*Breeding colony* -- Lizards tested in out experiment came from a breeding colony established in the lab since 2019.
-The colony consists of approx.
-180 adults of and _L.
-guichenoti_ housed in commonal containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
-Enclosures are provided with non-stick matting, shelter, and several small water dishes filled daily; and lizards are fed mid-size crickets (*Acheta domestica*) three days a week.
-Crickets are dusted with calcium weekly and multivitamin and calcium biweekly.
-Room temperatures are set to 22-24ºC, but to ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle, keeping the warm side of enclosures at 32ºC.
+#### Animal husbandry  
+*Breeding colony* -- The lizards tested in our experiment came from a breeding colony established in the lab in 2019.
+The colony consisted of approximately 180 adults of _L.
+guichenoti_ housed in communal containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
+Enclosures were provided with non-stick matting, shelter, and several small water dishes filled daily, and lizards were fed mid-size crickets (*Acheta domestica*) three days a week.
+Crickets were dusted with calcium weekly and multivitamins and calcium biweekly.
+Room temperatures were set to 22-24ºC, but to ensure a temperature gradient, we employed a heat chord and a heat lamp following a 12 h light:12 h dark cycle, keeping the warm side of enclosures at 32-34ºC.
 
-*Eggs collection and incubation* -- Between mid-October 2023 to the end of February 2024, we placed small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside in one extreme of the communal enclosures (see above) to provided females with a place to lay the eggs.
-We checked for the presence of eggs in the boxes three days a week.
-Eggs' length and width were measured with a digital caliper to the nearest 0.1 mm, while mass was recorded with a (OHAUS, Model spx123) digital scale ± 0.001g error.
-Eggs were assigned a clutch number and a individual identity.
-Then we treated eggs with CORT or vehicle (see CORT and Temperature manipulation below) and placed them in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
-We covered the cups with cling wrap to retain moisture and then left the eggs in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.
+*Egg collection and incubation* -- Between mid-October 2023 to the end of February 2024, we placed small boxes (12.5 L x 8.3 W x 5 H cm) containing vermiculite on one side of the communal enclosures (see above) to provide females with a place to lay the eggs.
+We checked for eggs in the boxes three days a week.
+Eggs' length and width were measured with a digital calliper to the nearest mm (± 0.1 mm), while mass was recorded with a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
+Eggs were assigned a clutch number and an individual identity.
+Then we treated eggs with CORT or vehicle (see CORT and Temperature manipulation below) and placed them in individual cups (80 mL) with moist vermiculite (12 g water to 4 g vermiculite).
+We covered the cups with plastic wrap to retain moisture and left the eggs in incubators (LATWIT 2X5D-R1160) programmed to two different thermal regimes (see CORT and Temperature manipulation below) until hatching.
+Incubators were checked for the presence of hatchlings three times a week.
 
-*Hatchlings* -- Incubators were checked three times a week for the presence of hatchlings.
-Once hatchlings were collected, we measured SVL and Tail Length (TL) with a ruler to the nearest mm, and weighed them with a (OHAUS, Model spx123) digital scale ± 0.001g error.
+*Hatchlings* -- Immediately after hatching, we measured Snout-Vent Length (SVL) and Tail Length (TL) with a ruler to the nearest mm (± 0.1 mm), and mass using a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
 Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) provided with non-stick matting and a small water dish.
-During this period, they were sprayed water every day and received 3-6 small *A.
+During this period, they were given water daily and received 3-6 small *A.
 domestica* crickets three times a week.
@@
 
-We manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A).
-We first allocated eggs onto one of two different treatments: in the CORT treatment, eggs were topically supplied with 5µL of CORT dissolved in 100% Ethanol (vehicle) at a final (10 pg CORT/mL) concentration; in the Control treatment eggs received an equal volume of the vehicle.
-CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data).
+We manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23 ± 3ºC or Hot -- 30 ± 3ºC) in a 2x2 factorial design (@fig-Methods A).
+Eggs were allocated to one of two different developmental stress treatments: CORT-treated eggs were topically supplied with 5µL of CORT dissolved in 100% Ethanol at a final concentration of 10 pg CORT/mL.
+In the Control treatment eggs received an equal volume of the 100% Ethanol vehicle.
+We selected those doses based on previous studies where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs in _Lampropholis delicata_ [@crino2024eggs].
 Then, eggs were incubated in one of the two previously mentioned temperature regimes (Cold or Hot) until hatching.
-This temperatures represent the upper and lower limit of the natural incubation temperatures [@qualls2000post].
-The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.
+These temperatures represent the upper and lower limit of the natural incubation temperatures for this species [@qualls2000post].
+The number of eggs per clutch assigned to each hormone and temperature treatment was counterbalanced per clutch.
 
@@
 
-Two weeks before starting the experimen phase (see below), lizards were moved to the experimental arena for acclimatation.
-The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers provided with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the center (@fig-Methods B).
-The new enclosures were placed in 7 different racks located in two different rooms.
-Each rack was associated to its own CCTV systems (device model DVR-HP210475) allowing us to record lizards' behaviour during the experiment.
-The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack.
-During acclimatation and throughout the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum*.
-The temperature of both rooms was set to between 22-24ºC, but we also provided a temperature gradient by means of a heat cord along the center of the arena, and heat lamps in a 12 h light: 12 h dark cycle.
+Two weeks before starting the numerical discrimination task (see below), lizards were moved to the experimental arenas for acclimatisation.
+The arenas were individual medium-size (41 L x 29.7 W x 22 H cm) plastic containers provided with a shelter (9 L x 6 W x 1.5 H cm) on one of the sides and a water dish in the center (@fig-Methods B).
+These arenas were placed in 7 different racks located in two experimental rooms.
+Each rack was associated with its own CCTV system (device model DVR-HP210475), allowing us to record lizards' behaviour during the experiment.
+Despite the conditions in the experimental rooms were the same as in the colony room, we counterbalanced the number of lizards per treatment in each rack to control for any potential effects of the room or the position of the lizard in the rack.
+During acclimatisation and throughout the experiment, lizards were fed with only one cricket daily, dusted with calcium and multivitamins (see protocol below), and water was supplied *ad libitum*.
+The temperature of both rooms was set to 22-24ºC, but we also provided a temperature gradient (24-32°C) using a heat cord along the center of the arena and heat lamps in a 12 h light: 12 h dark cycle.
 
-The numerical discrimination test consisted in a series of two-choice trials where lizards where presented with different number of food items.
-Each choice was placed in one of two sides of a white 3D printed platform (see @fig-Methods B, C) and then covered with a transparent Petri dish to prevent the lizard from obtaining the food.
-We performed five tests that differed in the ratio between both choices: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4.
-We never used more than four items because higher numbers are considered to be represented by a different system [@nieder2005counting; @cantlon2009beyond; @hyde2011two].
-The food consisted in frozen small crickets (*A.
-domestica*) that were thawed and dusted with calcium and multivitamin.
-We decided to use frozen crickets to avoid movement to affect lizards' behaviour.
-Lizards were fed frozen crickets during one month before the beginning of the experiments for habituation to frozen prey.
-Two weeks before the experiments, to habituate lizards to the platform and the Petri dish, the frozen cricket that was placed on top of the Petri dish in one of the sides of the platform, simulating the conditions of the experiment (see @fig-Methods B).
+The numerical discrimination test consisted of a series of two-choice trials where lizards were presented with different numbers of food items.
+Each choice was placed on one of the two lateral platforms of a white 3D printed device (see @fig-Methods B, C) and then covered with a transparent Petri dish to prevent the lizard from obtaining the food.
+We performed five tests that differed in the ratio between both choices: 1 cricket VS 4 crickets, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4.
+The food consisted of small frozen crickets (*A.
+domestica*) that were thawed and dusted with calcium and multivitamins.
+We decided to use frozen crickets to avoid movement that would affect lizards' behaviour [@cooper2024tell].
+
+Lizards were fed frozen crickets for one month prior to the experiments to habituate them to eating frozen prey.
+To habituate lizards to the experimental conditions, frozen crickets were placed on top of one of the platforms and Petri dishes during each feeding for two weeks prior to experimental trials.
 The side where the food was placed was changed every day to avoid any side bias during the trials.
-However, to further control for that, the side where we placed the higher number of crickets was randomised between trials and individuals.
-We also counterbalanced the amount of individuals per treatment tested in the same type of trial each day.
+However, to further control for that, the side where we placed the higher number of crickets was randomised between trials and individuals during the tests.
+We also counterbalanced the number of individuals per treatment tested in the same type of trial each day.
 The crickets used during the experiment were all the same size, but to control for other quantitative traits [@agrillo2014spontaneous], we changed the orientation of the crickets as shown in @fig-Methods D.
-In addition, after the experiments, we made another set of trials to test whether the orientation of the crickets could affect lizards' choice, but we found no effect of orientation (see Supplementary Material - Prey orientation test).
-
-All tests were done between 10 and 12 in the morning when the lizards were most active.
-Each trial was performed daily for five days, between the 17^th^ to the 22^nd^ of May 2024.
-We recorded three main variables: i) Latency, defined as the time between the start of the test - when the platform was left in the enclosure - until the first interaction with one of the options; ii) First choice of the lizard, recorded as 1 if the lizard interacted first with the larger amount or 0 if not; and iii) Interest shown for the higher amount of food (detailed below).
-We considered that the lizard was interacting with the food when it touched the Petri dish with its snout or when the front half of the body was on top of the platform.
-For five minutes after the first interaction with one of the options, we recorded the time spent interacting with that option for a maximum of five minutes; we did this for both amounts of food independently of what was the first choice.
-We measured the interest shown for the higher amount of food by substracting the time interacting with the option with less crickets to the time spent interacting with the larger amount of food.
-If there was no interaction with one of the options during the test, we considered the time spent as 0 for that option.
-When there was no interaction with any of the choices, both Latency, First choice, and Interest were noted as NA.
-We also considered Interest as NA in some few cases where the animals managed to lift the dish and eat the prey; Latency and First choice were recorded and employed in the analyses.
-We excluded from the analyses all the individuals that did not make any choice in three or more tests.
+In addition, after the experiments, we conducted another set of trials to test whether the orientation of the crickets could affect lizards’ choice, but we found no effect of orientation (see Supplementary Material - Prey orientation test).
 
@@
 
+All tests were done between 1000-1200 hours when the lizards were most active.
+Each trial was performed daily for five days between the 17^th^ and 22^nd^ of May 2024.
+We recorded three main variables: i) Latency, defined as the time between the start of the test - when the platform was left in the enclosure - until the first interaction with one of the options; ii) First choice of the lizard (Choice), recorded as 1 if the lizard interacted first with the larger amount or 0 if not; and iii) Interest shown for the higher amount of food (detailed below).
+We considered that the lizard interacted with the food when it touched the Petri dish with its snout or when the front half of the body was on top of the platform.
+For five minutes after the first interaction with one of the options, we recorded the time spent interacting with that option for a maximum of five minutes; we did this for both amounts of food independently of what was the first choice.
+We measured the interest shown for the higher amount of food (Interest) by substracting the time interacting with the option with fewer crickets to the time spent interacting with the larger amount of food.
+If there was no interaction with one of the options during the test, we considered the time spent as 0 for that option.
+When there was no interaction with any of the choices, both Latency, Choice, and Interest were noted as NA.
+We also considered Interest as NA in a few cases where the animals managed to lift the dish and eat the prey; Latency and Choice were recorded and employed in the analyses.
+We excluded from the analyses all the individuals that did not make any choice in three or more tests.
+
 #### Statistical analyses  
 In our analyses, we fitted a multivariate Bayesian multilevel model using the brm function from the _brms_ package [@burkner2017brms] in R extension (version 2.8.2) [@R].
-We ran four parallel MCMC chains of 4000 iterations with a warmup period of 1000 iterations.
-We modelled the three main recorded behaviour: Latency (log-transformed), Choice, and Interest (see above) as the response variables, and test (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4), hormone (CORT versus Control), incubation temperature (Cold versus Hot) and their interaction as fixed factors.
+We ran four parallel MCMC chains of 3000 iterations for each model, with a warmup period of 1000 iterations.
+We checked that all MCMC chains converged (Rhat < 1.2) and were mixing effectively to ensure we had >1000 effective samples from the posterior distribution.
+
+We modelled the three main recorded behaviours: Latency, Choice, and Interest as the response variables.
+We assumed normal distributions for Latency (log-transformed) and Interest [family = gaussian()].
+The error structure for Choice was modeled using a Bernoulli distribution with a logit link function [family = Bernoulli(link = ‘logit’)].
+We included test (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4), hormone (CORT versus Control), incubation temperature (Cold versus Hot), and the three-way interaction as fixed factors.
 We also included in our predictors sex, and age (centered to 0).
-We included lizard indentity and clutch as random factors.
-We used the posterior distributions of parameters from these models to test for differences in the ability to discriminate between different number of crickets between treatments and species, for each of the numerical tests.
-We calculated the contrast between treatments for each test using p~mcmc~ values, and we also estimated whether predicted Choice and Interest differed from 0.5 or zero respectively using p~mcmc~.
+We included lizard identity and clutch as random factors.
+_L.
+guichenoti_ lays up to two clutches per year [@joss1985reproductive].
+Since egg collection was done during half of the breeding season, each clutch likely came from a unique mother, and so, clutch identity captures potential maternal effects.
+
+We used the posterior distributions of parameters from these models to test for between treatment differences in the ability to discriminate different numbers of crickets for each of the numerical tests.
+We used p~mcmc~ to tests whether the contrast between treatments for each test differed from zero and whether Choice and Interest differed from 0.5 or zero respectively.
 We considered statistical significance if p~mcmc~ < 0.05.
@@
 #### Ethical note
-Subjects were collected from a breeding colony kept in the lab since 2019.
-All animals used in this experiment were provided humane laboratory housing, with thermorregulation opportunity, light, and humidity conditions (see above for details).
-Lizards in the colony received five-seven crickets per individual three times a week while the subjects in the experiment were fed once per day.
+Subjects were collected from a breeding colony that had been kept in the lab since 2019.
+All animals in this experiment were provided humane laboratory housing, thermoregulation opportunities, and light and humidity conditions (see above for details).
+Lizards in the colony received five to seven crickets per individual three times a week while the subjects in the experiment were fed once per day.
 All lizards were handled minimally to avoid stress.
 Experimental procedures and laboratory housing complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33).
-Upon completion of experimental trials, lizards were kept in their enclosure and will be euthanized following standard humane procedures once they reach the human-ending point.
+Upon completion of experimental trials, lizards were kept in their enclosure and will be euthanized following standard humane procedures once they reach the humane endpoint.
 
@@
 
-Our results showed an effect of sex and age on Latency (mean estimated effect of sex = `r {format_dec(mean(posteriors$b_loglatency_sexm), 2)}`, 95% CI = [`r {format_dec(quantile(posteriors$b_loglatency_sexm, 0.025), 2)}`, `r {format_dec(quantile(posteriors$b_loglatency_sexm, 0.975), 2)}`]; mean estimated effect of age = `r {format_dec(mean(posteriors$b_loglatency_age),2)}`, 95% CI = [`r {format_dec(quantile(posteriors$b_loglatency_age, 0.025), 2)}`, `r {format_dec(quantile(posteriors$b_loglatency_age, 0.975),2)}`]; see Table S1 in Supplementary Material), where males and older individials made, overall, faster decisions.
+Latency was affected by sex (mean estimated effect of sex = `r {format_dec(mean(posteriors$b_loglatency_sexm), 2)}`, 95% CI = [`r {format_dec(quantile(posteriors$b_loglatency_sexm, 0.025), 2)}`, `r {format_dec(quantile(posteriors$b_loglatency_sexm, 0.975), 2)}`]) and age (mean estimated effect of age = `r {format_dec(mean(posteriors$b_loglatency_age),2)}`, 95% CI = [`r {format_dec(quantile(posteriors$b_loglatency_age, 0.025), 2)}`, `r {format_dec(quantile(posteriors$b_loglatency_age, 0.975),2)}`]; see Table S1 in Supplementary Material).
+Males and older individuals made, overall, faster decisions.
 However, we did not see such effects on Choice or Interest (see Table S1 in Supplementary Material).
@@
 #| label: fig-results
-#| fig.cap: "Mean estimates of loglatency (Latency), the probability of choosing first the larger number of crickets (Choice), and the estimated interest for the larger number of crickets (Interest) per each of the treatments for each of the numerical tests performed.
-The x-axis represents the estimate, and in the y-axis are the density of the estimates.
+#| fig.cap: "Mean estimates of log-latency (Latency), the probability of choosing first the larger number of crickets (Choice), and the estimated interest for the larger number of crickets (Interest) per each of the treatments for each of the numerical tests performed.
+The x-axis represents the estimate, and the y-axis is the density of the estimates.
 The different colours indicate the different treatments.
@@
 
-Temperature, CORT, or their interaction contrasts did not yield any significant result in any of the numerical tests (Table 1) and we did not observe any clear pattern indicating relevant effects of early environment on any of the tests (@fig-results).
-Except for Hot-CORT lizards in 1 VS 3 tests or Cold-Control lizards in 2 VS 4 tests, the predicted probability of choosing the greater number of crickets first was not different from 0.5 (@fig-results; but also see Table S2 in Supplementary Material).
-Similarly, the predicted interest for the larger number of crickets was not different from 0 in any of the tests (@fig-results; but also see Table S2 in Supplementary Material).
+Neither temperature, CORT, nor their interaction contrasts affected performance in any of the tests, nor were there any clear patterns suggestive of an effect of the developmental environment on numerical discrimination (see Table 1 and @fig-results).
+Except for Hot-CORT lizards in 1 VS 3 tests or Cold-Control lizards in 2 VS 3 tests, the predicted probability of choosing the larger number of crickets first was not different from 0.5 (@fig-results; but also see Table S2 in Supplementary Material).
+Similarly, Interest was not different from 0 in any of the tests (@fig-results; but also see Table S2 in Supplementary Material).
 
 _Table 1.
-Effects of temperature, CORT, and their interaction on latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
-The table shows the comparissons for each predictor (Temperature = [meanHot - meanCold]; Hormone = [meanControl - meanCORT]; Interaction = [(meanHot-Control - meanHot-CORT) - (meanCold-Control - meanCold-CORT)]) and the p~mcmc~ value of the comparisons.
+Effects of temperature, CORT, and their interaction on Latency, Choice, and Interest in each of the numerical discrimination tests.
+The table shows the contrasts for each predictor (Temperature = [meanHot - meanCold]; Hormone = [meanControl - meanCORT]; Interaction = [(meanHot-Control - meanHot-CORT) - (meanCold-Control - meanCold-CORT)]) and the p~mcmc~ value of the contrasts.
 In bold, the effects where p~mcmc~ < 0.05._
@@
 #| label: tbl-data
-#| tbl-cap: "Effects of temperature, CORT, and their interaction on latency, choice, and interest shown for the higher amount of food in each of the numerical discrimination tests.
-The table shows the comparissons for each predictor (Temperature = [meanHot - meanCold]; Hormone = [meanControl - meanCORT]; Interaction = [(meanHot-Control - meanHot-CORT) - (meanCold-Control - meanCold-CORT)]) and the p~mcmcm~ value of the comparisons.
+#| tbl-cap: "Effects of temperature, CORT, and their interaction on Latency, Choice, and Interest in each of the numerical discrimination tests.
+The table shows the contrasts for each predictor (Temperature = [meanHot - meanCold]; Hormone = [meanControl - meanCORT]; Interaction = [(meanHot-Control - meanHot-CORT) - (meanCold-Control - meanCold-CORT)]) and the p~mcmcm~ value of the contrasts.
 In bold, the effects where p~mcmc~ < 0.05."
@@
 ## Discussion
-Our results indicate that lizards were not able to discriminate between different number of crickets in any of the tests, and that their performance did not vary across treatments.
+Our results indicate that lizards did not discriminate between different numbers of crickets and that their performance did not vary across treatments.
+We predicted that lizards would employ nummerical cues to make foraging decisions, and that their ability to discriminate between those cues would be impaired with exposure to CORT during development or cold incubation temperatures.
+Our findings suggest otherwise.
 We did not see changes in Latency in any of the tests, and treatment did not affect lizards' time to make a decision either.
-Furthermore, in most of the cases, the estimated probability of choosing the option with more crickets first did not differ from random (0.5), with significant differences emerging only in two tests for two of the four treatments; yet no clear pattern was observed.
-These findings were further reinforced by similar results regarding the interest for the larger number of food items, that did not differ from zero in any of the tests or groups.
+Furthermore, in most cases, Choice did not differ from random (0.5), with significant differences emerging only in two tests for two of the four treatments; yet no clear pattern was observed.
+These findings were further reinforced by similar results regarding Interest, that did not differ from zero in any of the tests or treatments.
 Altogether, these findings suggest that _L.
@@
 
-In our experiment, we employed a spontaneous choice test to investigate _L.
+Our experiment employed a spontaneous choice test to investigate _L.
 guichenoti_'s numerical abilities.
-This test is considered a reliable method to assess numerosity in animals, as it allows for the observation of natural behaviour and the acquirement of ecological issues related to numerousness and quantity discrimination [@nieder2005counting; @agrillo2014spontaneous].
-However, one of the main caveats of this approach is that motivation plays a key-role in animals' performance [@agrillo2014spontaneous; @bisazza2014experim].
-Here, we discarded the trials where the lizards did not interact with any of the options, so the results observed are based exclusively in those where the lizards made a choice.
-Therefore we do not consider motivation to be a factor affecting our results.
-Moreover, spontaneous choice tests are based on the assumption that animals should select the most appropriate quantity, which may not be necessarily the largest one [@agrillo2014spontaneous].
-Although spontaneous choice tests employing food assume animals to prefer larger amounts of food, this may not hold true if increased handling costs are involved [@agrillo2014spontaneous].
-Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of age on their choice or interest, as older animals are typically larger and may be more adept at handling a greater number of prey.
+This test is considered a reliable method to animal numerosity, as it allows for observing instictive behaviours [@nieder2005counting; @agrillo2014spontaneous].
+However, one of the main caveats of this approach is that motivation plays a key role in animals' performance [@agrillo2014spontaneous; @bisazza2014experim].
+Here, we discarded the trials in which the lizards did not interact with any of the options, so the observed results are based exclusively on those trials in which a choice was made.
+Therefore, we do not consider motivation a factor affecting our results.
 
 Another consideration is the limitation of spontaneous choice tests to control for other quantitative cues that may correlate with numerosity [@agrillo2014spontaneous].
-Although we controlled for some quantitative information by rotating crickets in the option with less items to increase total area, we can not rule out the possibility that lizards could be using other quantitative cues.
+Although we controlled for some quantitative information by rotating crickets in the option with fewer items to increase total area, we can not rule out the possibility that lizards could use other quantitative cues.
 If that were the case, however, we would still see some type of discrimination between choices unless the cues were conflicting [@cooper2024tell].
 Indeed, it is possible that _L.
-guichenoti_ uses multiple quantitative cues simoultaneously during foraging decisions, and, when those cues conflict, the preference for one cue over another might vary between individuals, potentially producing the results observed here.
+guichenoti_ uses multiple quantitative cues simultaneously during foraging decisions.
+When those cues conflict, the preference for one cue over another might vary between individuals, potentially producing the results observed here.
 Alternatively, _L.
-guichenoti_ may not rely on numerical cues for foraging decisions, instead utilizing other type of information.
+guichenoti_ may not rely on numerical cues for foraging decisions, instead utilising other types of information.
 
-Interestingly, half of the studies demonstrating numerosity in reptiles employed trained quantity discrimination tasks [@soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained].
-In the remaining successful studies, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
+Spontaneous choice tests assume that animals should select the most appropriate quantity, which may not necessarily be the largest one [@agrillo2014spontaneous].
+This occurrs if the costs of selecting the larger option outweigh the benefits.
+For example, when presented two groups with different numbers of conspecifics or if prey handling costs are related to larger quantitative traits like size [@agrillo2014spontaneous].
+It could be that the perceived handling costs of larger number of crickets is affecting lizards' behaviour in our experiment.
+Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of age on Choice or Interest, as older animals are typically larger and may be more adept at handling a greater number of prey items.
+The absence of choice in most of the tests suggests that the benefits of selecting larger number of crickets are not different from selecting fewer.
+For instance, when considering hunting groups of 1 VS 4 or 2 VS 4 crickets, the probability of success might be similar.
+As a result, _L.
+guichenoti_ might not have evolved a preference for prey groups with such small differences.
+
+Interestingly, in those studies where reptiles discriminated between different amounts of food, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
 10)[@recio_prey_2021].
-However, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities [@miletto2017quantitative] despite showing high quantitative skills when size of prey was the main cue [@miletto2017quantitative], or when tested using training procedures [@miletto2018quantity].
-Hence, it is possible that further studies exploring numerosity in _L.
-guichenoti_ using different methodologies or contexts, such as larger quantities [@nieder2005counting; @recio_prey_2021], inclusion of chemical stimuli [@carazo2009quantity; @plotnik2019elephants; @recio_prey_2021], using relevant stimuli like conspecifics [@carazo2009quantity; @lucon2017individual], or use training procedures [@agrillo2014spontaneous; @bisazza2014experim; @soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained], may yield more conclusive results.
+In contrast, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities despite showing high quantitative skills when prey size was the main cue [@miletto2017quantitative] or when tested using training procedures [@miletto2018quantity].
+Further studies on numerosity in _L.
+guichenoti_ should explore different methods, stimuli or contexts.
+For example, using larger quantities [@nieder2005counting; @recio_prey_2021], testing numerical discrimination in the context of mate selection [@carazo2009quantity; @lucon2017individual], or employing training procedures [@agrillo2014spontaneous; @bisazza2014experim; @soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained] may yield more conclusive results.
 
 We did not find any effect of prenatal conditions on _L.
-guichenoti_'s ability to use numerical information for making foraging decisions.
+guichenoti_'s ability to use numerical information to make foraging decisions.
 We anticipated that lower incubation temperatures and exposure to CORT will impair lizards' performance, especially in those tests with higher ratios following Weber's Law [@agrillo2014spontaneous; @nieder2018evolution].
-However, lizards' behaviour was not impacted by our treatments, suggesting that early environmental conditions do not affect numerical competence in this species.
-Our predictions were based in previous studies showing the effects of prenatal temperature and GCs on cognitive abilities in reptiles and other groups [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic].
+However, our treatments did not impact lizards' behaviour.
+Our predictions were based on previous studies showing the effects of prenatal temperature and GCs on cognitive abilities in reptiles and other groups [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic].
 However, these effects can be species-specific, varying across cognitive domains and test types [@szuran_water_1994; @farrell_developmental_2015-learn; @bebus_associative_2016;@dayananda_incubation_2017; @abayarathna_effects_2020; @haq_prenatal_2021].
@@
 For example, incubation temperature impacts mitochondrial activity and cell density in specific brain nuclei [@coomber_independent_1997; @amiel_effects_2017].
-Similarly, impairments in spatial learning - but not in other learning tasks - in rats [@haq_prenatal_2021] are likely associated with the greater density of GCs receptors mammals' hippocampi compared to other regions of the brain [@siegel_2006].
-While it remains unknown which regions of the reptiles' brain encode numerical information, our results suggest that relevant regions may be unaffected by the prenatal conditions tested here.
-However, researchers have proposed that non-verbal number representation is processed by two types of systems: one for small sets (≤ 4) that tracks and encodes objects as individual entities, and another and another for larger numerosities that operates through estimation processes [@nieder2005counting; @cantlon2009beyond; @hyde2011two].
-Our data indicates that at least those brain regions involved in discrete numerical representation might be robust to the effect of prenatal CORT and temperature in _L.
+Similarly, impairments in spatial learning - but not in other learning tasks - in rats [@haq_prenatal_2021] are likely associated with the greater density of GCs receptors in mammals' hippocampus compared to other regions of the brain [@siegel_2006].
+Researchers have proposed that non-verbal number representation is processed by two types of systems: one for small sets (≤ 4) that tracks and encodes objects as individual entities, and another for larger numerosities that operate through estimation processes [@nieder2005counting; @cantlon2009beyond; @hyde2011two].
+While it remains unknown which areas of the reptiles' brain encode numerical information, our results suggest that those brain regions involved in discrete numerical representation might be robust to the effect of prenatal CORT and temperature in _L.
 guichenoti_.
-Studies using higher quantities of food items may help elucidate the effects of prenatal conditions on _L.
-guichenoti_'s numerical abilities.
-Alternatively, it is also plausible that treatment effects are masked by the overall poor performance across most of the tests; which could be a consequence of the species' relying in other type of information for foraging rather than numerical information.
-Thus, future research should investigate the effects of prenatal CORT and thermal environment on numerical competence by employing trained quantity discrimination tasks [@soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained].
+Alternatively, it is also plausible that treatment effects are masked by the overall poor performance across most of the tests, which could be a consequence of the species’ reliance on another type of cues for foraging other than numerical information.
+Thus, future research should investigate the effects of prenatal CORT and thermal environment on numerical competence by employing trained quantity discrimination tasks and higher quantities [@soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained].
 
-In conclusion, our study is, to the best of our knowledge, the first to explore the impact of prenatal conditions on numerical abilities in reptiles.
+Our study is, to the best of our knowledge, the first to explore the impact of prenatal conditions on numerical abilities in reptiles.
 Our findings suggest that _L.
 guichenoti_ does not rely on numerical information for foraging decisions and that their performance does not vary across different prenatal treatments.
-This implies that early environmental conditions, such as incubation temperature and CORT exposure, may not affect numerosity in this species.
+These results indicate that early environmental conditions, such as incubation temperature and CORT exposure, may not affect numerosity in this species.
 However, it is critical to note that the methodologies employed in this study may not have fully captured the nuances of numerical cognition in _L.
@@
 Therefore, we recommend that future studies employ a broader array of tests and paradigms, including training-based tasks, larger sets of stimuli, and other sensory modalities, to more comprehensively assess numerical competence in this and related species.
-Such work will be pivotal in unraveling the effects of early environmental factors on cognitive abilities in reptiles and advancing our understanding of numerosity across taxa.
+Such work will be fundamental in unravelling early environmental factors' effects on reptiles' cognitive abilities and advancing our understanding of numerical discrimination across taxa.
 
@@
 ## Authors' contributions  
-P.R.: conceptualization, methodology, data collection, data curation, formal analysis, writing—original draft, writing—review and editing; D.C.L.: data collection, writing—review and editing; O.C.: conceptualization, writing—review and editing; C.F.: conceptualization, funding acquisition, writing—review and editing; A.Y.P.: data collection, writing—review and editing; D.N.: conceptualization, methodology, funding acquisition, project administration, resources, supervision, writing—review and editing.
+PR: conceptualization, methodology, data collection, data curation, formal analysis, writing—original draft, writing—review and editing; DL: data collection, writing—review and editing; OC: conceptualization, writing—review and editing; CF: conceptualization, funding acquisition, writing—review and editing; AP: data collection, writing—review and editing; DN: conceptualization, methodology, funding acquisition, project administration, resources, supervision, writing—review and editing.
 All authors gave final approval for publication and agreed to be held accountable for the work performed therein.
@@
 ## Funding  
-This work was supported by a National Australian University fellowship (P.R.), and the Australian Research Council (grant no. DP210101152) to D.N. and C.F.
+This work was supported by a National Australian University fellowship (PR), and the Australian Research Council (grant no. DP210101152) to DN and CF.
 
@@
 
-For loglatency
+For loglatency (Latency)
 ```{r, tbl-summary_lat}
@@
 
-For choice
+For Choice
 ```{r, tbl-summary_choice}
@@
 
-For compared interest
+For Interest
 ```{r, tbl-summary_int}
@@
 _Table S2.- Estimated mean probability of choosing first the higher amount (Choice) and the mean interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
-p indicates the p~mcmcm~ value (p) of testing the hypothesis Choice ≠ 0.5, and Interest ≠ 0, which would indicate a preference towards one of choices (> towards higher amounts, < towards lower amounts) of the comparissons.
+p indicates the p~mcmcm~ value (p) of testing the hypothesis Choice ≠ 0.5, and Interest ≠ 0, which would indicate a preference towards one of choices (> towards higher amounts, < towards lower amounts) of the contrasts.
 In bold, the groups where p~mcmcm~ < 0.05_
@@
 #| tbl-cap: "Estimated probability of choosing first the higher amount (Choice) and the interest for the higher amount of food (Interest) per treatment group for each os the tests performed.
-p indicates the p~mcmcm~ value of testing the hypothesis Choice ≠ 0.5, and Interest ≠ 0, which would indicate a preference towards one of choices (> towards higher amounts, < towards lower amounts) of the comparissons."
+p indicates the p~mcmcm~ value of testing the hypothesis Choice ≠ 0.5, and Interest ≠ 0, which would indicate a preference towards one of choices (> towards higher amounts, < towards lower amounts) of the contrasts."
 #
@@
 We used the same platform and Petri dish as in the numerical discrimination task, but we placed one cricket on each side of the platform, one oriented vertically in respect to the shelter position (see @fig-Methods B) and the other one horizontal to the shelter.
-We perform a unique trial randomising the position of the orientation of the cricket.
-We recorded the amount of individuals that chose the cricket oriented vertically and horizontally.
+We performed a unique trial randomising the position of the orientation of the cricket.
+We recorded the number of individuals that chose the cricket oriented vertically and horizontally.
 Results are shown in the table below.
@@
 _Table S4.
-Number of individuals per treatment that chose the cricket oriente horizontally (Horizontal) or vertically (Vertical) in our control tests.
+Number of individuals per treatment that chose the cricket oriented horizontally (Horizontal) or vertically (Vertical) in our control tests.
 p-value indicates the result of the binomial test comparing the number of choices between both choices._
@@
 #| label: tbl-orientation
-#| tbl-cap: "Number of individuals per treatment that chose the cricket oriente horizontally (Horizontal) or vertically (Vertical) in our control tests.
+#| tbl-cap: "Number of individuals per treatment that chose the cricket oriented horizontally (Horizontal) or vertically (Vertical) in our control tests.
 p-value indicates the result of the binomial test comparing the number of choices between both choices."

### 2024-09-22 R/func.r

@@ -270,10 +270,10 @@ plotting <- function(type_var){
   # Plot the predicted estimates for each treatment and test
   plot <- ggplot(df_violin, aes(x = Treatment, y = Estimate, fill = Treatment)) +
   geom_flat_violin(alpha = 0.6) + 
-  scale_fill_manual(values = c("Cold-CORT (n = 20)" = "#68bde1",
-                              "Cold-Control (n = 20)" = "darkblue",
-                              "Hot-CORT (n = 20)" = "#fa927d",
-                              "Hot-Control (n = 20)" = "#b50101"),
+  scale_fill_manual(values = c("Cold-CORT (n = 20)" = "#00008B",
+                              "Cold-Control (n = 20)" = "#68bde1",
+                              "Hot-CORT (n = 20)" = "#b50101",
+                              "Hot-Control (n = 20)" = "#fa927d"),
                     breaks = c("Hot-Control (n = 20)",
                               "Hot-CORT (n = 20)",
                               "Cold-Control (n = 20)",

### 2024-10-09 ms.qmd
Updates

@@
 #| label: packages
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS,
-fitdistrplus, ggh4x, PupillometryR, cowplot, png, grid, remotes, ggthemes)
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS, nortest, fitdistrplus, ggh4x, PupillometryR, cowplot, png, grid, remotes, ggthemes, bayestestR, HDInterval)
 ```
@@
 ## Abstract
-From social behaviour to navigating their environment, quantitative abilities can be crucial in animals' life.
-Early life conditions are predicted to impact brain development, and quantitative competence may not be exent of these effects.
-Early thermal environment or glucocorticoids (GCs) - stress-related hormones - transmitted by the parents can play a prominent role in shaping cognition through effects on brain physiology.
-Furthermore, temperature can lead to increased levels of GCs that are passed on to the offspring, emphasizing the need to study the combined effects of these two factors.
-Here, we investigated the effects of prenatal corticosterone (CORT) - the main GC in reptiles - and incubation temperature on numerical discrimination in the common garden skink (_Lampropholis guichenoti_).
-We manipulated CORT levels and temperature during incubation and then assessed numerical abilities through a spontaneous choice test with food as the relevant stimulus.
-Employing a repeated measures design, we subjected lizards to five numerical tests each differing in the total number of items and the ratios between the two choices (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 2 VS 4).
-Contrary to our expectations, we found no evidence of numerical discrimination or any effect of treatments on lizards' behaviour.
+From social behaviour to navigating their environment, quantitative abilities can be crucial to fitness.
+Early life conditions can impact brain development affecting quantitative competence.
+In anthropogenically modified environments, early thermal conditions or glucocorticoid concentrations (GCs) – stress-related hormones transmitted by the parents – can play a prominent role in shaping cognition through their effects on brain physiology.
+Furthermore, temperature can lead to increased levels of GCs that are passed on to offspring, emphasizing the need to study the combined effects of these two factors.
+Here, we investigated the effects of elevated prenatal corticosterone (CORT) – the main GC in reptiles – and incubation temperature on numerical discrimination in the common garden skink (_Lampropholis guichenoti_).
+We then assessed numerical abilities through a spontaneous choice test with food as the relevant stimulus.
+Employing a repeated measures design, we subjected lizards to five numerical tests each differing in the ratios between the two choices (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 2 VS 4).
+Contrary to our expectations, we found no evidence of numerical discrimination or any effect of treatments on lizard behaviour.
 These findings suggest that _L.
-guichenoti_ may not use numerical cues to make foraging decisions and that early conditions do not affect their behaviour towards different quantities of food.
+guichenoti_ may not use numerical cues to make foraging decisions.
+Our findings align with previous studies indicating that reptiles feeding on live prey may have difficulty distinguishing between different prey quantities in spontaneous choice tests.
 
 ## Significance statement
-Discriminating between quantities is crucial for animal behaviour, yet the impact of early environmental factors on the development of quantitative competence remains largely unexplored.
-In this study, we investigated for the first time the effects of prenatal corticosterone exposure and incubation temperature on numerical discrimination in a reptile.
-Contrary to our predictions, we did not observe any evidence of numerical discrimination regardless of incubation conditions.
+Discriminating between quantities can be crucial for individual fitness, yet the impact of early environmental factors on the development of quantitative competence remains largely unexplored.
+In this study, we investigated, for the first time, the effects of prenatal corticosterone exposure and incubation temperature on numerical discrimination in a reptile.
+Contrary to our predictions, we did not observe any evidence of numerical discrimination and therefore no impact of early developmental environment.
 Our results align with previous studies showing that reptiles consuming live prey struggle to differentiate prey numbers in spontaneous choice tests.
@@
 The ability to discern between quantities is a fundamental cognitive skill widely observed across the animal kingdom [@stancher2015numerical; @vonk2012bears; @beran2016capuchin].
-Numerical cognition is adaptive in various ecological contexts, contributing to social decisions, mate selection, or foraging [@lucon2017individual; @nieder2018evolution].
+Numerical cognition can contribute to social decisions, mate selection, and foraging and thus is adaptive in various ecological contexts [@lucon2017individual; @nieder2018evolution].
 For instance, female lions (_Panthera leo_) use numerical information to assess the risk of confronting rival groups [@mccomb1994roaring], male adult mealworms (_Tenebrio molitor_) select sites with scents of more females [@carazo2009quantity], and red-backed salamanders (_Plethodon cinereus_) prefer tubes containing larger numbers of prey [@uller2003salamanders].
-Given the adaptive advantages of numerical and quantity discrimination, research has extensively explored the ability of different species to differentiate between quantities under different contexts and paradigms.
+Given the adaptive advantages of numerical and quantity discrimination, research has extensively has focused on exploring the ability of different species to differentiate between quantities under different contexts and paradigms.
 
 Numerical competence is often evaluated by considering the types of cues animals use [@plotnik2019elephants; @cooper2024tell], the ecological contexts in which these abilities are deployed [@mccomb1994roaring; @chittka1995can; @uller2003salamanders; @carazo2009quantity], and the neural mechanisms that underlie them [@nieder2005counting].
-The latter have demonstrated that non-verbal numerical representations in mammals activate an extensive cortical network, with the prefrontal cortex and the intraparietal sulcus being particularly relevant [@nieder2005counting].
+The latter have demonstrated that non-verbal numerical representations in mammals activates an extensive neural pathway including regions involved in eye movement, attention control, or complex learning [@nieder2005counting].
 This connection between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.
 
-Animals' brains are particularly sensitive to environmental inputs during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004], with long-lasting effects on cognitive abilities [@zhu_prenatal_2004; @amiel_egg_2014].
-For instance, glucocorticoids (GCs), the main groups of hormones mediating the physiological stress response [@sapolsky_how_2000], can alter animals' performance in spatial, associative, or reversal learning tasks if elevated during early stages of development [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
-Prenatal exposure to GCs is typically associated with impaired learning skills, but these effects have been shown to vary depending on factors such as dose, subjects' sex, or type of cognitive test employed [@szuran_water_1994; @farrell_developmental_2015-learn; @bebus_associative_2016; @haq_prenatal_2021].
+Brains are particularly sensitive to environmental inputs during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004], with long-lasting effects on cognitive abilities [@zhu_prenatal_2004; @amiel_egg_2014].
+For instance, glucocorticoids (GCs), steroid hormones that mediate the physiological stress response [@sapolsky_how_2000], can alter animals' performance in spatial, associative, or reversal learning tasks if elevated during early stages of development [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
+Prenatal exposure to GCs is typically associated with impaired learning skills, but these effects have been shown to vary depending on factors such as dose, sex, or type of cognitive test employed [@szuran_water_1994; @farrell_developmental_2015-learn; @bebus_associative_2016; @haq_prenatal_2021].
 Nevertheless, little is known about how exposure to GCs during development can affect quantitative abilities.
 
-At the same time, the early thermal environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
-For example, Port-Jackson sharks (_Heterodontus portusjacksoni_) incubated at elevated temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019].
+The effects of prenatal GCs on learning may be modulated by other developmental effects or environmental conditions.
+For example, the early thermal environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
+In lizards, incubating eggs at elevated temperatures improves cognitive abilities consistent with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017], but impairs cognition at temperatures above lizards’ natural incubation thermal range [@dayananda_incubation_2017; @abayarathna_effects_2020].
+In addition, when @vila_pouca_quantity_2019 tested the effects of elevated incubation temperature on numerical discrimination in  Port-Jackson sharks (_Heterodontus portusjacksoni_), they found that sharks incubated at elevated temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019].
 However, the effects of early thermal environment on numerical abilities in other taxa remains largely unexplored.
 
-Given that abrupt temperature changes can increase GC levels in vertebrates, leading to elevated GCs transmitted to the offspring [@racic_effects_2020; @Crino_2023], it is essential to study the combined effects of early-life GCs and thermal conditions on cognitive development.
+Temperature can affect GC levels in vertebrates and can indirectly affect developing offspring through maternal effects [@jessop2016multiscale; @racic_effects_2020; @mentesana2022glucocorticoids; @Crino_2023].
+Early-life GCs and thermal conditions could have interactive effects that influence cognitive development.
 Understanding how these factors influence quantitative abilities could reveal important insights into the ecological relevance of numerosity.
 To fill this gap, we used the common garden skink (_Lampropholis guichenoti_) as a model to test the effects of elevated prenatal corticosterone (CORT) - the main GC in reptiles [@Crino_2023] - and incubation temperature on numerical discrimination abilities.
+However, the combined influence of prenatal CORT and thermal environment on cognition remains unknown, particularly on numerical abilities.
 
-Traditionally, research on numerical abilities has been mainly focused on mammals, birds, and fishes [@nieder2005counting; @agrillo2014spontaneous], but recent studies demonstrate that reptiles are also competent in numberical abilities.
-For example, red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour with higher volumes of food [@soldati2017long], Hermann’s tortoises (_Testudo hermanni_) proved to discriminate between large and small food quantities in a spontaneous choice test [@gazzola2018continuous], and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) rapidly learned a quantity discrimination task [@lin2021superior; @lin2024trained].
+Given the influence of numerical cognition on fitness, we would expect evolutionary pressures to be similar across taxa, and yet, research on numerical discrimination has been traditionally focused on mammals, birds, and fishes [@nieder2005counting; @agrillo2014spontaneous].
+However, recent studies demonstrate that reptiles are also competent in numbers.
+For example, red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour with higher volumes of food [@soldati2017long], Hermann’s tortoises (_Testudo hermanni_) discriminate between large and small food quantities in a spontaneous choice test [@gazzola2018continuous], and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) rapidly learned a quantity discrimination task [@lin2021superior; @lin2024trained].
 The italian wall lizard (_Podarcis siculus_) failed to discriminate between different numbers of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity].
@@
 
-Reptiles' cognition is significantly influenced by the environment in which they develop.
-For example, incubating lizards at elevated temperatures improves cognitive abilities consistent with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017], but impairs cognition at temperatures above lizards’ natural incubation thermal range [@dayananda_incubation_2017; @abayarathna_effects_2020].
-The effects of prenatal CORT on cognitive abilities in reptiles is less understood, but in other taxa, high levels of GCs are typically related to cognitive impairments [@szuran_water_1994; @zhu_prenatal_2004].
-However, the combined influence of prenatal CORT and thermal environment on cognition remains unknown, particularly on numerical abilities.
-
-Our study aimed to address this question by manipulating CORT concentration (CORT-treated or a sham control) in _L.
-guichenoti_ eggs and incubating them under one of two temperature regimes (Cold - 23 ± 3ºC or Hot - 28 ± 3ºC) in a 2x2 factorial design.
-After hatching, we assessed the lizards' numerical discrimination abilities in a spontaneous choice test, where they were presented with different numbers of crickets.
-Based on other studies that explored the cognitive effects of prenatal GCs or thermal early environment in reptiles and other taxa, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
-Additionally, we predicted that increased CORT levels would impair numerical competence [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
-In this sense, a common observation in nature is that the ability to detect a change in stimulus intensity depends on the relative, not absolute, difference between the stimuli ['Weber's law'; @agrillo2014spontaneous].
+Here, we investigated the effect of prenatal CORT and incubation temeprature on numerical discrimination in a lizard.
+We manipulated CORT concentration (CORT-treated or a sham control) in _L.
+guichenoti_ eggs and incubated them under one of two temperature regimes (Cold - 23 ± 3ºC or Hot - 28 ± 3ºC) in a 2x2 factorial design.
+Then, we assessed numerical discrimination abilities of lizards unsing a spontaneous choice test approach.
+The numerical discrimination task consisted of five tests that differed in the relative number of food items: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
+We based this experiment on observations in nature that the ability to detect a change in stimulus intensity depends on the relative, not absolute, difference between the stimuli ['Weber's law'; @agrillo2014spontaneous].
+Built upon other studies that explored the cognitive effects of prenatal GCs or thermal early environment in reptiles and other taxa, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
+We predicted that increased CORT levels would impair numerical competence [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
 Consistent with Weber's law, we predicted that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g.
@@
 We manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23 ± 3ºC or Hot -- 30 ± 3ºC) in a 2x2 factorial design (@fig-Methods A).
-Eggs were allocated to one of two different developmental stress treatments: CORT-treated eggs were topically supplied with 5µL of CORT dissolved in 100% Ethanol at a final concentration of 10 pg CORT/mL.
-In the Control treatment eggs received an equal volume of the 100% Ethanol vehicle.
+Eggs were allocated to one of two different developmental stress treatments: CORT-treated eggs were topically supplied with 5µL of a CORT solution (10pg/mL) Control treatment eggs received an equal volume of 100% Ethanol.
+Corticosterone treatments were made by dissolving crystalline corticosterone (Sigma, Cat.
+No.
+C2505) in 100% ethanol.
 We selected those doses based on previous studies where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs in _Lampropholis delicata_ [@crino2024eggs].
@@
 The arenas were individual medium-size (41 L x 29.7 W x 22 H cm) plastic containers provided with a shelter (9 L x 6 W x 1.5 H cm) on one of the sides and a water dish in the center (@fig-Methods B).
-These arenas were placed in 7 different racks located in two experimental rooms.
+Arenas were placed on seven different racks located in two experimental rooms.
 Each rack was associated with its own CCTV system (device model DVR-HP210475), allowing us to record lizards' behaviour during the experiment.
@@
 Each choice was placed on one of the two lateral platforms of a white 3D printed device (see @fig-Methods B, C) and then covered with a transparent Petri dish to prevent the lizard from obtaining the food.
-We performed five tests that differed in the ratio between both choices: 1 cricket VS 4 crickets, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4.
+We performed five tests that differed in the ratio between both choices of crickets.
+Specifically, we compared: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
 The food consisted of small frozen crickets (*A.
 domestica*) that were thawed and dusted with calcium and multivitamins.
-We decided to use frozen crickets to avoid movement that would affect lizards' behaviour [@cooper2024tell].
+We decided to use frozen crickets to avoid movement that would affect lizard behaviour [@cooper2024tell].
 
@@
 The side where the food was placed was changed every day to avoid any side bias during the trials.
-However, to further control for that, the side where we placed the higher number of crickets was randomised between trials and individuals during the tests.
+However, to further control for side biases, the higher number of crickets was randomly placed on the right or the left platform.
 We also counterbalanced the number of individuals per treatment tested in the same type of trial each day.
-The crickets used during the experiment were all the same size, but to control for other quantitative traits [@agrillo2014spontaneous], we changed the orientation of the crickets as shown in @fig-Methods D.
-In addition, after the experiments, we conducted another set of trials to test whether the orientation of the crickets could affect lizards’ choice, but we found no effect of orientation (see Supplementary Material - Prey orientation test).
+
+The crickets used during the experiment were all the same size, but to control for other quantitative traits we changed the orientation of the crickets as illustrated in @fig-Methods D.
+This approach created a conflict of information between choices, as horizontally oriented crickets could occupy the same space as a greater number of vertically oriented crickets.
+This allowed us to determine whether the lizards relied on discrete numerical units rather than non-numerical cues, such as total length, area, or volume [@agrillo2014spontaneous].
+In addition, after the experiments, we conducted another set of trials to test whether the orientation of the crickets could affect lizards’ choice, but we found no effect of orientation (see _Supplementary Material - Prey orientation test_).
 
@@
 #| label: fig-Methods
-#| fig.cap: "Experimental design of the study.
+#| fig.cap: "Experimental design.
 Panel (**A**) shows the early environment manipulation procedures.
@@
 Each trial was performed daily for five days between the 17^th^ and 22^nd^ of May 2024.
-We recorded three main variables: i) Latency, defined as the time between the start of the test - when the platform was left in the enclosure - until the first interaction with one of the options; ii) First choice of the lizard (Choice), recorded as 1 if the lizard interacted first with the larger amount or 0 if not; and iii) Interest shown for the higher amount of food (detailed below).
+We recorded three main variables: i) _Latency_, defined as the time between the start of the test - when the platform was left in the enclosure - until the first interaction with one of the options; ii) First choice of the lizard (_Choice_), recorded as 1 if the lizard interacted first with the larger amount or 0 if not; and iii) _Interest_ shown for the higher amount of food (detailed below).
 We considered that the lizard interacted with the food when it touched the Petri dish with its snout or when the front half of the body was on top of the platform.
 For five minutes after the first interaction with one of the options, we recorded the time spent interacting with that option for a maximum of five minutes; we did this for both amounts of food independently of what was the first choice.
-We measured the interest shown for the higher amount of food (Interest) by substracting the time interacting with the option with fewer crickets to the time spent interacting with the larger amount of food.
-If there was no interaction with one of the options during the test, we considered the time spent as 0 for that option.
-When there was no interaction with any of the choices, both Latency, Choice, and Interest were noted as NA.
+We measured the interest shown for the higher amount of food (_Interest_) by subtracting the time interacting with the option with fewer crickets to the time spent interacting with the larger amount of food.
+If there was no interaction with one of the options during the test, we considered the time spent as zero for that option.
+When there was no interaction with any of the choices, both _Latency_, _Choice_, and _Interest_ were noted as NA.
 We also considered Interest as NA in a few cases where the animals managed to lift the dish and eat the prey; Latency and Choice were recorded and employed in the analyses.
 We excluded from the analyses all the individuals that did not make any choice in three or more tests.
+
+Motivation plays a key role in animals' performance in spontaneous choice tests [@agrillo2014spontaneous; @bisazza2014experim].
+To ensure that lizards remained motivated to forage, they were provided with only one cricket during both acclimatization and the numerical task (see above).
+During acclimatization, lizards were fed at the same time each day as the experiments were conducted.
+In the numerical task, lizards received a cricket immediately after the test and then fasted for more than 20 hours, which we expected would help standardize fasting levels across individuals.
+To further control for motivation, we excluded from our analyses any trials in which lizards did not interact with any options or did not eat the cricket provided after the test (n = 0 in both cases).
+Consequently, we do not consider motivation to be a major factor influencing our results (but see Discussion).
 
@@
 
-We modelled the three main recorded behaviours: Latency, Choice, and Interest as the response variables.
+We modelled the three main recorded behaviours: _Latency_, _Choice_, and _Interest_ as the response variables.
 We assumed normal distributions for Latency (log-transformed) and Interest [family = gaussian()].
@@
 We included test (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4), hormone (CORT versus Control), incubation temperature (Cold versus Hot), and the three-way interaction as fixed factors.
-We also included in our predictors sex, and age (centered to 0).
+We also included in our predictors sex, and age.
+We did not find any effect of sex or age on Choice or Interest, but we found an effect on Latency (see Results).
+However, we decided to include both factors in all models.
+Age was centered to zero in the model, and we averaged the predicted values for males and females.
+
 We included lizard identity and clutch as random factors.
@@
 We used the posterior distributions of parameters from these models to test for between treatment differences in the ability to discriminate different numbers of crickets for each of the numerical tests.
-We used p~mcmc~ to tests whether the contrast between treatments for each test differed from zero and whether Choice and Interest differed from 0.5 or zero respectively.
-We considered statistical significance if p~mcmc~ < 0.05.
+Because the posteriors did not follow a normal distribution, we report the median as the main measure of central tendency.
+We also employed Highest Density Intervals using the hdi function in bayestestR [@bayestestR] to test if the contrasts between treatments for all the variables, or the _Interest_ per each test was different form zero.
+Given Choice was Bernoulli, we modified our null hypotheses to test if differed from a probability of 0.5 as there were only two possible options.
 
@@
 post_lat <- org_post("lat")
+normality_results_lat <- check_normality(post_lat)
 # Choice
@@
 prob_choice <- exp(post_choice)/(1 + exp(post_choice))
+normality_results_choice <- check_normality(prob_choice)
+write.csv(normality_results_choice, file = here("output/Checking/normality_results.csv"))
 # Estimated interest for the higher amount of food
 post_int <- org_post("int")
+normality_results_int <- check_normality(post_int)
 #
@@
 
-Latency was affected by sex (mean estimated effect of sex = `r {format_dec(mean(posteriors$b_loglatency_sexm), 2)}`, 95% CI = [`r {format_dec(quantile(posteriors$b_loglatency_sexm, 0.025), 2)}`, `r {format_dec(quantile(posteriors$b_loglatency_sexm, 0.975), 2)}`]) and age (mean estimated effect of age = `r {format_dec(mean(posteriors$b_loglatency_age),2)}`, 95% CI = [`r {format_dec(quantile(posteriors$b_loglatency_age, 0.025), 2)}`, `r {format_dec(quantile(posteriors$b_loglatency_age, 0.975),2)}`]; see Table S1 in Supplementary Material).
-Males and older individuals made, overall, faster decisions.
-However, we did not see such effects on Choice or Interest (see Table S1 in Supplementary Material).
-Nevertheless, when we analyse the effect of treatment on performance, for every variable recorded, we control for sex by averaging the predicted values for males and females, and we control for age by including the centered age in the model.
-These results are summarized in @fig-results and Table 1.
+Overall, males showed lower latencies than females (median of the estimated effect of sex = `r {format_dec(median(posteriors$b_loglatency_sexm), 2)}`, Higest Density Interval = [`r {format_dec(hdi(posteriors$b_loglatency_sexm, credMass = 0.95)[1], 2)}`, `r {format_dec(hdi(posteriors$b_loglatency_sexm, credMass = 0.95)[2], 2)}`]), and older individuals also made decisions faster (mean estimated effect of age = `r {format_dec(median(posteriors$b_loglatency_age),2)}`, Higest Density Interval = [`r {format_dec(hdi(posteriors$b_loglatency_age, credMass = 0.95)[1], 2)}`, `r {format_dec(hdi(posteriors$b_loglatency_age, credMass = 0.95)[2],2)}`]; see Table S1 in _Supplementary Material_).
+However, we did not see any effect of age or sex on _Choice_ or _Interest_ (see _Supplementary Material_).
 
@@
 #| label: fig-results
-#| fig.cap: "Mean estimates of log-latency (Latency), the probability of choosing first the larger number of crickets (Choice), and the estimated interest for the larger number of crickets (Interest) per each of the treatments for each of the numerical tests performed.
-The x-axis represents the estimate, and the y-axis is the density of the estimates.
+#| fig.cap: "Estimates of log-latency (Latency), the probability of choosing the larger number of crickets first (Choice), and the estimated interest in the larger number of crickets (Interest) for each of the numerical tests performed across different developmental treatments.
+The x-axis represents the estimate, and the y-axis is the posterior density of the estimates.
 The different colours indicate the different treatments.
-Points and bars represent the mean and standard deviation of the mean of the estimates, respectively.
-Vertical dashed lines in Choice and Interest graphs values 0.5 and 0, respectively."
+Points and bars represent the median and Higest Density Intervals of the estimates, respectively.
+Vertical dashed lines in Choice and Interest graphs values 0.5 and 0, respectively.
+Asterisks indicate values significantly different from 0."
 #
@@
 #### C) Combine the plots and make final figure
-fig_results <- plot_grid(fig_lat, fig_choice, fig_int, ncol = 1)
+fig_results <- plot_grid(fig_lat, fig_choice, fig_int, ncol = 1) +
+  annotate("text", x = 0.305, y = 0.56, label = "*", size = 7, color = "black") + 
+  annotate("text", x = 0.61, y = 0.505, label = "*", size = 7, color = "black")
 ggsave("./output/figures/fig_results.png", plot=fig_results, width = 25, height = 25, units = "cm", dpi = 600)
@@
 
-Neither temperature, CORT, nor their interaction contrasts affected performance in any of the tests, nor were there any clear patterns suggestive of an effect of the developmental environment on numerical discrimination (see Table 1 and @fig-results).
-Except for Hot-CORT lizards in 1 VS 3 tests or Cold-Control lizards in 2 VS 3 tests, the predicted probability of choosing the larger number of crickets first was not different from 0.5 (@fig-results; but also see Table S2 in Supplementary Material).
-Similarly, Interest was not different from 0 in any of the tests (@fig-results; but also see Table S2 in Supplementary Material).
+Neither temperature, CORT, or their interaction affected performance in any of the tests, nor were there any clear patterns suggestive of an effect of the developmental environment on numerical discrimination (see Table 1 and @fig-results).
+Except for Hot-CORT lizards in 1 VS 3 tests or Cold-Control lizards in 2 VS 3 tests, the predicted probability of choosing the larger number of crickets first was not different from 0.5 (@fig-results; but also see Table S2 in _Supplementary Material_).
+Similarly, interest in the larger number of crickets (_Interest_) was not different from 0 in any of the tests (@fig-results; but also see Table S2 in _Supplementary Material_).
 
@@
 Effects of temperature, CORT, and their interaction on Latency, Choice, and Interest in each of the numerical discrimination tests.
-The table shows the contrasts for each predictor (Temperature = [meanHot - meanCold]; Hormone = [meanControl - meanCORT]; Interaction = [(meanHot-Control - meanHot-CORT) - (meanCold-Control - meanCold-CORT)]) and the p~mcmc~ value of the contrasts.
-In bold, the effects where p~mcmc~ < 0.05._
+The table shows the contrasts for each predictor (Temperature = [medianHot - medianCold]; Hormone = [medianControl - medianCORT]; and their Interaction = [(medianHot-Control - medianHot-CORT) - (medianCold-Control - medianCold-CORT)]).
+Higest Density Intervals (HDI) test the hypothesis that contrasts are different from zero._
 ```{r, tbl-data}
@@
     Hot_Control <- post_mod[[paste0("X", b, "_Hot_Control")]]
-    # Calculate effects and pmcmc values
-    effect_temp <- format_dec(mean(c(Hot_CORT, Hot_Control)) - mean(c(Cold_CORT, Cold_Control)), 1)
-    effect_hormone <- format_dec(mean(c(Hot_Control, Cold_Control)) - mean(c(Hot_CORT, Cold_CORT)), 1)
-    interaction <- format_dec((mean(Hot_Control) - mean(Cold_Control)) - (mean(Hot_CORT) - mean(Cold_CORT)), 1)
+    # Calculate effects and HDI values
+    effect_hormone <- format_dec(median(c(Hot_Control, Cold_Control)) - median(c(Hot_CORT, Cold_CORT)), 1)
+    effect_temp <- format_dec(median(c(Hot_CORT, Hot_Control)) - median(c(Cold_CORT, Cold_Control)), 1)
+    interaction <- format_dec((median(Hot_Control) - median(Cold_Control)) - (median(Hot_CORT) - median(Cold_CORT)), 1)
     #
-    pmcmc_temp <- format_p(pmcmc(c(Hot_CORT, Hot_Control) - c(Cold_CORT, Cold_Control)), 2)
-    pmcmc_hormone <- format_p(pmcmc(c(Hot_Control, Cold_Control) - c(Hot_CORT, Cold_CORT)), 2)
-    pmcmc_interaction <- format_p(pmcmc((Hot_Control - Cold_Control) - (Hot_CORT - Cold_CORT)), 2)
+    hdi_hormone <- paste0("[", format_dec(hdi(c(Hot_Control, Cold_Control) - c(Hot_CORT, Cold_CORT))[1], 2), " , ",
+                          format_dec(hdi(c(Hot_Control, Cold_Control) - c(Hot_CORT, Cold_CORT))[2], 2), "]")
+    hdi_temp <- paste0("[", format_dec(hdi(c(Hot_CORT, Hot_Control) - c(Cold_CORT, Cold_Control))[1], 2), " , ",
+                        format_dec(hdi(c(Hot_CORT, Hot_Control) - c(Cold_CORT, Cold_Control))[2], 2), "]")
+    hdi_interaction <- paste0("[", format_dec(hdi((Hot_Control - Cold_Control) - (Hot_CORT - Cold_CORT))[1], 2), " , ",
+                              format_dec(hdi((Hot_Control - Cold_Control) - (Hot_CORT - Cold_CORT))[2], 2), "]")
     # Define labels and values
-    effects_predictors_names <- c("Temperature", 
-                          "Hormone", 
+    effects_predictors_names <- c("Hormone",
+                          "Temperature",
                           "Interaction")
-    values_predictors <- c(effect_temp, effect_hormone, interaction)
-    values_pmcmc <- c(pmcmc_temp, pmcmc_hormone, pmcmc_interaction)
+    values_predictors <- c(effect_hormone, effect_temp, interaction)
+    values_hdi <- c(hdi_hormone, hdi_temp, hdi_interaction)
     # Create a temporary data frame for this loop iteration
@@
                         Effect = values_predictors,
-                        pmcmc = values_pmcmc,
+                        hdi = values_hdi,
                         Test = b,
@@
 table_df <- as.data.frame(table_df, stringsAsFactors = FALSE) %>%
-  mutate(results = trimws(paste0(Effect, " (p", pmcmc,")"))) %>%
+  mutate(results = trimws(paste0(Effect, "\n", hdi))) %>%
   dplyr::select(Variable, Test, Predictor, results) %>%
@@
 ## Discussion
-Our results indicate that lizards did not discriminate between different numbers of crickets and that their performance did not vary across treatments.
-We predicted that lizards would employ nummerical cues to make foraging decisions, and that their ability to discriminate between those cues would be impaired with exposure to CORT during development or cold incubation temperatures.
-Our findings suggest otherwise.
-We did not see changes in Latency in any of the tests, and treatment did not affect lizards' time to make a decision either.
-Furthermore, in most cases, Choice did not differ from random (0.5), with significant differences emerging only in two tests for two of the four treatments; yet no clear pattern was observed.
-These findings were further reinforced by similar results regarding Interest, that did not differ from zero in any of the tests or treatments.
-Altogether, these findings suggest that _L.
-guichenoti_ did not use numerical or other potential quantitative cues we were not controlling for to make foraging decisions.
+Contrary to our predictions that numerical cues would influence foraging decisions, and that elevated prenatal CORT or cold incubation temperatures would impair numerical discrimination, we found no significant differences across tests or treatments.
+Our results indicate that lizards did not discriminate between different numbers of crickets, nor did their performance vary under our experimental conditions.
+Overall, these findings suggest that L.
+guichenoti did not utilize numerical or other potential quantitative cues for foraging decisions.
+This highlights the need for further research into the ecological implications of such behaviors in lizards and their foraging strategies.
 
-Our experiment employed a spontaneous choice test to investigate _L.
-guichenoti_'s numerical abilities.
-This test is considered a reliable method to animal numerosity, as it allows for observing instictive behaviours [@nieder2005counting; @agrillo2014spontaneous].
-However, one of the main caveats of this approach is that motivation plays a key role in animals' performance [@agrillo2014spontaneous; @bisazza2014experim].
-Here, we discarded the trials in which the lizards did not interact with any of the options, so the observed results are based exclusively on those trials in which a choice was made.
-Therefore, we do not consider motivation a factor affecting our results.
+#### Foraging context may influence prey number discrimination in reptiles  
+Lack of numerical discrimination in _L.
+guichenoti_ may be due to insectivorous lizards relying on other cues, such as movement or size, for foraging decisions.
+In those studies where reptiles discriminated between different amounts of food, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
+10)[@recio_prey_2021].
+In contrast, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities despite showing high quantitative skills when prey size was the main cue [@miletto2017quantitative] or when tested using training procedures [@miletto2018quantity].
+Remarkably, the numbers of food items employed in @miletto2017quantitative were always fewer than four.
+Spontaneous choice tests assume that animals should select the most appropriate quantity, which may not necessarily be the largest one [@agrillo2014spontaneous].
+This occurrs if the costs of selecting the larger option outweigh the benefits.
+For example, when presented two groups with different numbers of conspecifics or if prey handling costs are related to larger quantitative traits like size [@agrillo2014spontaneous].
+It could be that the perceived handling costs of larger numbers of prey is affecting lizards' response.
+Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of lizards' size on their choices, as larger animals may be more adept at handling a greater number of prey items.
+@miletto2017quantitative did not report any effect of lizards' size, and age - a proxy for size - did not affect _L.
+guichenoti_'s _Choice_ or _Interest_ in our study.
+The absence of preference in _P.
+siculus_ and _L.
+guichenoti_ suggests that the benefits of selecting larger number of prey are not different from selecting fewer.
+For instance, when considering hunting groups of 1 VS 4 or 2 VS 4 crickets, the probability of success might be similar.
+Furthermore, while one individual handles a prey item in the patch, other conspecifics could capture the rest.
+As a result, insectivorous lizards might not have evolved a preference for larger patches of prey, at least with small numbers [see @recio_prey_2021].
 
-Another consideration is the limitation of spontaneous choice tests to control for other quantitative cues that may correlate with numerosity [@agrillo2014spontaneous].
-Although we controlled for some quantitative information by rotating crickets in the option with fewer items to increase total area, we can not rule out the possibility that lizards could use other quantitative cues.
-If that were the case, however, we would still see some type of discrimination between choices unless the cues were conflicting [@cooper2024tell].
+Conversely, our findings may be influenced by the experimental design.
+In our study, we rotated the position of the crickets in the choice with fewer items to create a situation where the numbers and other quantitative cues associatied with numerosity were in conflict.
+This approach was intended to control for other quantitative cues that could be used by lizards to make decisions [@cooper2024tell].
+However, this manipulation could have made the task more challenging for lizards, potentially affecting their performance.
 Indeed, it is possible that _L.
@@
 When those cues conflict, the preference for one cue over another might vary between individuals, potentially producing the results observed here.
-Alternatively, _L.
-guichenoti_ may not rely on numerical cues for foraging decisions, instead utilising other types of information.
+Future studies should explore the use of other quantitative cues in _L.
+guichenoti_ and how they interact with numerical information.
 
-Spontaneous choice tests assume that animals should select the most appropriate quantity, which may not necessarily be the largest one [@agrillo2014spontaneous].
-This occurrs if the costs of selecting the larger option outweigh the benefits.
-For example, when presented two groups with different numbers of conspecifics or if prey handling costs are related to larger quantitative traits like size [@agrillo2014spontaneous].
-It could be that the perceived handling costs of larger number of crickets is affecting lizards' behaviour in our experiment.
-Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of age on Choice or Interest, as older animals are typically larger and may be more adept at handling a greater number of prey items.
-The absence of choice in most of the tests suggests that the benefits of selecting larger number of crickets are not different from selecting fewer.
-For instance, when considering hunting groups of 1 VS 4 or 2 VS 4 crickets, the probability of success might be similar.
-As a result, _L.
-guichenoti_ might not have evolved a preference for prey groups with such small differences.
+#### Foraging behaviour is robust to prenatal CORT and temeprature  
+Although the effects of early environment on cognition is widely studied, little atention has been paid to its impacts on numerical discrimination.
+To the best of our knowledge, there is only one study examining how developmental conditions influence numerical abilities.
+@vila_pouca_quantity_2019 investigated the effects of incubating and rearing Port Jackson sharks (_Heterodontus portusjacksoni_) at different temperatures on their ability to learn a quantity discrimination task.
+They found that those animals reared in warm water took fewer days to reach the learning criterion.
+Given the potent effects of early environment on cognition and brain development [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic], it is crucial to explore how early environmental conditions shape numerical abilities across taxa.
 
-Interestingly, in those studies where reptiles discriminated between different amounts of food, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
-10)[@recio_prey_2021].
-In contrast, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities despite showing high quantitative skills when prey size was the main cue [@miletto2017quantitative] or when tested using training procedures [@miletto2018quantity].
-Further studies on numerosity in _L.
-guichenoti_ should explore different methods, stimuli or contexts.
-For example, using larger quantities [@nieder2005counting; @recio_prey_2021], testing numerical discrimination in the context of mate selection [@carazo2009quantity; @lucon2017individual], or employing training procedures [@agrillo2014spontaneous; @bisazza2014experim; @soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained] may yield more conclusive results.
+In our experimental design, prenatal treatments did not influence numerical discrimination in _L.
+guichenoti_, but we cannot discount the possibility that early environmental factors may affect numerical discrimination under other approaches or contexts.
+Based on previous studies showing the effects of prenatal temperature and GCs on cognitive abilities in reptiles and other groups [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic], we predicted that lower incubation temperatures and exposure to CORT will impair lizards' performance, especially in those tests with higher ratios [@agrillo2014spontaneous; @nieder2018evolution].
+However, we did not find any effect of the experimental treatments on _L.
+guichenoti_'s behaviour when tested in a numerical task.
+Both GCs and thermal environment can exert differential effects on distinct areas of the brain.
+For example, incubating eggs of the eastern three-lined skink (_Bassiana dupeyerri_) at elevated temperatures increased cell density in the medial cortex - linked to learning process and memory - but did not impact cell density in other nuclei [@amiel_effects_2017].
+Similarly, impairments in spatial learning in rats [@haq_prenatal_2021] are likely associated with the greater density of GCs receptors in mammals' hippocampus compared to other regions of the brain [@siegel_2006].
+Researchers have proposed that non-verbal number representation is processed by two types of systems: one for small sets (≤ 4) that tracks and encodes objects as individual entities, and another for larger numerosities that operate through estimation processes [@nieder2005counting; @cantlon2009beyond; @hyde2011two].
+While it remains unknown which areas of the reptiles' brain encode numerical information, our results suggest that those brain regions involved in discrete numerical representation (≤ 4 items) might be robust to the effect of prenatal CORT and temperature in _L.
+guichenoti_.
 
-We did not find any effect of prenatal conditions on _L.
-guichenoti_'s ability to use numerical information to make foraging decisions.
-We anticipated that lower incubation temperatures and exposure to CORT will impair lizards' performance, especially in those tests with higher ratios following Weber's Law [@agrillo2014spontaneous; @nieder2018evolution].
-However, our treatments did not impact lizards' behaviour.
-Our predictions were based on previous studies showing the effects of prenatal temperature and GCs on cognitive abilities in reptiles and other groups [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic].
-However, these effects can be species-specific, varying across cognitive domains and test types [@szuran_water_1994; @farrell_developmental_2015-learn; @bebus_associative_2016;@dayananda_incubation_2017; @abayarathna_effects_2020; @haq_prenatal_2021].
+Alternatively, it is plausible that treatment effects are masked by the overall poor performance across most of the tests, which could be a consequence of the species’ reliance on another type of cues for foraging other than numerical information.
+In this regard, motivation is a key factor in spontaneous choice tests [@agrillo2014spontaneous; @bisazza2014experim].
+Both developmental CORT and incubation temperature can affect metabolic rate (hereafter MR) which could affect motivation in food based learning paradigms [@crino_corticosterone_2014-learn; @noble_developmental_2018; @cossin2022effect].
+Here, we  controlled for motivation by providing only one cricket per day and fasting lizards for more than 20 hours before the test.
+In addition, we excluded those tests where lizards did not interact with any of the options or did not eat the cricket provided after the test.
+However, it is possible that the fasting period was not long enough to standardize hunger levels across individuals.
+Further studies should explore the effects of prenatal conditions on motivation and how it interacts with cognitive abilities in reptiles.
 
-In this context, it is important to consider that both GCs and thermal environment can exert differential effects on distinct regions of the brain.
-For example, incubation temperature impacts mitochondrial activity and cell density in specific brain nuclei [@coomber_independent_1997; @amiel_effects_2017].
-Similarly, impairments in spatial learning - but not in other learning tasks - in rats [@haq_prenatal_2021] are likely associated with the greater density of GCs receptors in mammals' hippocampus compared to other regions of the brain [@siegel_2006].
-Researchers have proposed that non-verbal number representation is processed by two types of systems: one for small sets (≤ 4) that tracks and encodes objects as individual entities, and another for larger numerosities that operate through estimation processes [@nieder2005counting; @cantlon2009beyond; @hyde2011two].
-While it remains unknown which areas of the reptiles' brain encode numerical information, our results suggest that those brain regions involved in discrete numerical representation might be robust to the effect of prenatal CORT and temperature in _L.
-guichenoti_.
-Alternatively, it is also plausible that treatment effects are masked by the overall poor performance across most of the tests, which could be a consequence of the species’ reliance on another type of cues for foraging other than numerical information.
-Thus, future research should investigate the effects of prenatal CORT and thermal environment on numerical competence by employing trained quantity discrimination tasks and higher quantities [@soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained].
-
+#### Conclusion
 Our study is, to the best of our knowledge, the first to explore the impact of prenatal conditions on numerical abilities in reptiles.
@@
 
-_Table S2.- Estimated mean probability of choosing first the higher amount (Choice) and the mean interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
-p indicates the p~mcmcm~ value (p) of testing the hypothesis Choice ≠ 0.5, and Interest ≠ 0, which would indicate a preference towards one of choices (> towards higher amounts, < towards lower amounts) of the contrasts.
-In bold, the groups where p~mcmcm~ < 0.05_
+_Table S2.- Median and Higest Density Intervals of the estimated probability of choosing first the higher amount (Choice) and the estimated interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
+Higest Density Intervals test the hypothesis that Choice = 0.5, and Interest = 0, which would indicate a preference towards one of choices._
 ```{r, tbl-disc}
 #| label: tbl-disc
-#| tbl-cap: "Estimated probability of choosing first the higher amount (Choice) and the interest for the higher amount of food (Interest) per treatment group for each os the tests performed.
-p indicates the p~mcmcm~ value of testing the hypothesis Choice ≠ 0.5, and Interest ≠ 0, which would indicate a preference towards one of choices (> towards higher amounts, < towards lower amounts) of the contrasts."
+#| tbl-cap: "Median and Higest Density Intervals of the estimated probability of choosing first the higher amount (Choice) and the estimated interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
+Higest Density Intervals test the hypothesis that Choice = 0.5, and Interest = 0, which would indicate a preference towards one of choices."
 #
@@
     Hot_Control <- post_mod[[paste0("X", d, "_Hot_Control")]]
-    # Calculate means and pmcmc values
-    Cold_CORT_mean <- format_dec(mean(Cold_CORT), 2)
-    Cold_Control_mean <- format_dec(mean(Cold_Control), 2)
-    Hot_CORT_mean <- format_dec(mean(Hot_CORT), 2)
-    Hot_Control_mean <- format_dec(mean(Hot_Control), 2)
+    # Calculate medians and Higest Density Intervals
+    Cold_CORT_median <- format_dec(median(Cold_CORT), 2)
+    Cold_Control_median <- format_dec(median(Cold_Control), 2)
+    Hot_CORT_median <- format_dec(median(Hot_CORT), 2)
+    Hot_Control_median <- format_dec(median(Hot_Control), 2)
     #
-    pmcmc_Cold_CORT <- format_p(pmcmc(Cold_CORT, null = val), 2)
-    pmcmc_Cold_Control <- format_p(pmcmc(Cold_Control, null = val), 2)
-    pmcmc_Hot_CORT <- format_p(pmcmc(Hot_CORT, null = val), 2)
-    pmcmc_Hot_Control <- format_p(pmcmc(Hot_Control, null = val), 2)
+    int_Cold_CORT <- paste0("[", format_dec(hdi(Cold_CORT, credMass = 0.95)[1], 2), " , ",
+                            format_dec(hdi(Cold_CORT, credMass = 0.95)[2], 2), "]")
+    int_Cold_Control <- paste0("[", format_dec(hdi(Cold_Control, credMass = 0.95)[1], 2), " , ",
+                            format_dec(hdi(Cold_Control, credMass = 0.95)[2], 2), "]")
+    int_Hot_CORT <- paste0("[", format_dec(hdi(Hot_CORT, credMass = 0.95)[1], 2), " , ",
+                            format_dec(hdi(Hot_CORT, credMass = 0.95)[2], 2), "]")
+    int_Hot_Control <- paste0("[", format_dec(hdi(Hot_Control, credMass = 0.95)[1], 2), " , ",
+                            format_dec(hdi(Hot_Control, credMass = 0.95)[2], 2), "]")
     # Define labels and values
     treatment_names <- c("Cold-CORT", "Cold-Control", "Hot-CORT", "Hot-Control")
-    values_means <- c(Cold_CORT_mean, Cold_Control_mean, Hot_CORT_mean, Hot_Control_mean)
-    values_pmcmc <- c(pmcmc_Cold_CORT, pmcmc_Cold_Control, pmcmc_Hot_CORT, pmcmc_Hot_Control)
+    values_medians <- c(Cold_CORT_median, Cold_Control_median, Hot_CORT_median, Hot_Control_median)
+    values_int <- c(int_Cold_CORT, int_Cold_Control, int_Hot_CORT, int_Hot_Control)
     # Create a temporary data frame for this loop iteration
     temporal_df <- data.frame(Treatment = treatment_names,
-                        Effect = values_means,
-                        pmcmc = values_pmcmc,
+                        Effect = values_medians,
+                        int = values_int,
                         Test = d,
@@
 disc_df_final <- as.data.frame(disc_df, stringsAsFactors = FALSE) %>%
-  mutate(pmcmc = gsub("^= ", "", pmcmc)) %>%
-  pivot_wider(names_from = Variable, values_from = c(Effect, pmcmc), names_prefix = "") %>%
+  mutate(int = gsub("^= ", "", int)) %>%
+  pivot_wider(names_from = Variable, values_from = c(Effect, int), names_prefix = "") %>%
   dplyr::select(Test, Treatment, 
-                Effect_Choice, pmcmc_Choice,
-                Effect_Interest, pmcmc_Interest)
+                Effect_Choice, int_Choice,
+                Effect_Interest, int_Interest)
 #
@@
     align(align="center", part="all") %>% 
-    # Bold the cell containing "< 0.05" in the `pmcmc_Choice` column and the next cell
-    bold(i = which(disc_df_final$pmcmc_Choice == "< 0.05"), j = "pmcmc_Choice", part = "body") %>%
-    bold(i = which(disc_df_final$pmcmc_Choice == "< 0.05"), j = "Effect_Choice", part = "body") %>%
-    bold(i = which(disc_df_final$pmcmc_Choice == "< 0.05"), j = "Treatment", part = "body") %>%
-    # Bold the cell containing "< 0.05" in the `pmcmc_Interest` column and the next cell
-    bold(i = which(disc_df_final$pmcmc_Interest == "< 0.05"), j = "pmcmc_Interest", part = "body") %>%
-    bold(i = which(disc_df_final$pmcmc_Interest == "< 0.05"), j = "Effect_Interest", part = "body") %>%
-    bold(i = which(disc_df_final$pmcmc_Interest == "< 0.05"), j = "Treatment", part = "body") %>%
     flextable::compose(i = c(2:4,6:8,10:12,14:16,18:20), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-    flextable::compose(i = 1, j = c(3,5), value = as_paragraph("Est. mean"), part = "head") %>% 
-    flextable::compose(i = 1, j = c(4,6), value = as_paragraph("p"), part = "head") %>% 
+    flextable::add_header_row(values = c("", "Choice", "Interest"), colwidths = c(2, 2, 2)) %>%
+    flextable::compose(i = 2, j = c(3,5), value = as_paragraph("Median"), part = "head") %>% 
+    flextable::compose(i = 2, j = c(4,6), value = as_paragraph("HPD Interval"), part = "head") %>% 
     hline(i = c(4,8,12,16), j = c(1:6), part = "body") %>% # To make some horizontal lines
+    vline(i = c(1:20), j = 2, part = "body") %>% # To make some vertical lines
     vline(i = c(1:20), j = 4, part = "body") %>% # To make some vertical lines
-    vline(i = 1, j = 4, part = "head") 
+    vline(i = c(1,2), j = 2, part = "head") %>%
+    vline(i = c(1,2), j = 4, part = "head") 
 disc_table
@@
 #### Prey orientation test
-Because we were concerned that the orientation of the crickets could affect the lizards' choice, we performed one simple choice test to investigate this.
+Because the orientation of the crickets could affect the lizards' choice, we performed one simple choice test to investigate this.
 We used the same platform and Petri dish as in the numerical discrimination task, but we placed one cricket on each side of the platform, one oriented vertically in respect to the shelter position (see @fig-Methods B) and the other one horizontal to the shelter.
@@
 
+#### Searching for relevant literature
+To loof for relevant literature about the effects of early environment on numerical discrimination, we searched in the Web of Science and Scopus using the terms on respectively.
+
+We found 24 articles on Web of Science and 22 on Scopus, plus only one preprint on Scopus.
+Out of all the documents, only one study was relevant to our search: @vila_pouca_quantity_2019.
+That study appeared in both databases and was the only one that explored the effects of early environment on numerical abilities.
+
+```{r, fig-searchWOS}
+#| label: fig-searchWOS
+#| fig.cap: "Search query on Web of Science."
+
+knitr::include_graphics("./Others/Web of Science Search results.png")
+
+```
+
+```{r, fig-searchScopus}
+#| label: fig-searchScopus
+#| fig.cap: "Search query on Scopus."
+
+knitr::include_graphics("./Others/Scopus - Document search results.png")
+
+```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
 #### Checking the model plots  

### 2024-10-09 R/func.r

@@ -115,6 +115,17 @@ return(df)
 }
 ####################
 ####################
+# Function to check normality using Anderson-Darling normality test
+check_normality <- function(data_frame) {
+  results <- list()  # Create a list to store results
+  for (col in names(data_frame)) {
+    normality_test <- ad.test(data_frame[[col]])  # Perform Shapiro-Wilk test
+    results[[col]] <- normality_test$p.value  # Store the p-value
+  }
+  return(results)  # Return the results
+}
+####################
+####################
 # Estimate p-values using pmcm
 #' @title pMCMC Function
 #' @param x The vector for the posterior distribution. Note that this will test the null hypothesis that the parameter of interest is significantly different from 0. 
@@ -248,8 +259,9 @@ df_bars_plot <- function(data_fig){
   # Modify the df for the bar plots
   df_bars_plot <- fig_df_2 %>%
     group_by(Test, Treatment) %>%
-    summarize(mean_estimate = mean(Estimate, na.rm = TRUE),
-              sd_estimate = sd(Estimate, na.rm = TRUE)) %>%
+    summarize(median_estimate = median(Estimate, na.rm = TRUE),
+              low_lim = hdi(Estimate, credMass = 0.95)[1],
+              up_lim = hdi(Estimate, credMass = 0.95)[2]) %>%
     ungroup()
   data.frame()
 #
@@ -279,9 +291,9 @@ plotting <- function(type_var){
                               "Cold-Control (n = 20)",
                               "Cold-CORT (n = 20)")) +
   facet_wrap(~ Test, scales = "free_x", nrow = 1) +  # Facet by Treatment
-  geom_point(data = df_points, aes(y = mean_estimate, x = Treatment), position = position_dodge(width = 0.75), 
+  geom_point(data = df_points, aes(y = median_estimate, x = Treatment), position = position_dodge(width = 0.75), 
               color = "black", fill = "black", size = 3) +
-  geom_segment(data = df_points, aes(y = mean_estimate - sd_estimate, yend = mean_estimate + sd_estimate, 
+  geom_segment(data = df_points, aes(y = low_lim, yend = up_lim, 
                 x = Treatment, xend = Treatment), size = 1.5, color = "black") +
   ylim(min(df_violin$Estimate), max(df_violin$Estimate)) +
   coord_flip() +

### 2024-11-07 ms.qmd
Updates

@@
 ---
-title: "Early environment and numerical discrimination in a lizard"
+title: "Do early environmental experiences impact numerical discrimination abilities in a lizard?"
 docx: 
@@
 ## Abstract
-From social behaviour to navigating their environment, quantitative abilities can be crucial to fitness.
-Early life conditions can impact brain development affecting quantitative competence.
+From social behaviour to navigating complex environments, quantitative abilities can be crucial to fitness.
+Early life conditions can impact brain development to affect quantitative competence.
 In anthropogenically modified environments, early thermal conditions or glucocorticoid concentrations (GCs) – stress-related hormones transmitted by the parents – can play a prominent role in shaping cognition through their effects on brain physiology.
@@
 Employing a repeated measures design, we subjected lizards to five numerical tests each differing in the ratios between the two choices (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 2 VS 4).
-Contrary to our expectations, we found no evidence of numerical discrimination or any effect of treatments on lizard behaviour.
-These findings suggest that _L.
-guichenoti_ may not use numerical cues to make foraging decisions.
+Contrary to our expectations, we found no evidence for treatment effects on lizard behaviour and no use of numerical discrimination during foraging in _L.
+guichenoti_.
 Our findings align with previous studies indicating that reptiles feeding on live prey may have difficulty distinguishing between different prey quantities in spontaneous choice tests.
@@
 The ability to discern between quantities is a fundamental cognitive skill widely observed across the animal kingdom [@stancher2015numerical; @vonk2012bears; @beran2016capuchin].
-Numerical cognition can contribute to social decisions, mate selection, and foraging and thus is adaptive in various ecological contexts [@lucon2017individual; @nieder2018evolution].
+Numerical cognition can contribute to social decisions, mate selection, and foraging behaviour, which can be adaptive in various ecological contexts [@lucon2017individual; @nieder2018evolution].
 For instance, female lions (_Panthera leo_) use numerical information to assess the risk of confronting rival groups [@mccomb1994roaring], male adult mealworms (_Tenebrio molitor_) select sites with scents of more females [@carazo2009quantity], and red-backed salamanders (_Plethodon cinereus_) prefer tubes containing larger numbers of prey [@uller2003salamanders].
-Given the adaptive advantages of numerical and quantity discrimination, research has extensively has focused on exploring the ability of different species to differentiate between quantities under different contexts and paradigms.
+Given the adaptive advantages of numerical and quantity discrimination, research has focused extensively on exploring the ability of different species to differentiate between quantities under different contexts and paradigms.
 
@@
 The latter have demonstrated that non-verbal numerical representations in mammals activates an extensive neural pathway including regions involved in eye movement, attention control, or complex learning [@nieder2005counting].
-This connection between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.
+The connection between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.
 
 Brains are particularly sensitive to environmental inputs during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004], with long-lasting effects on cognitive abilities [@zhu_prenatal_2004; @amiel_egg_2014].
-For instance, glucocorticoids (GCs), steroid hormones that mediate the physiological stress response [@sapolsky_how_2000], can alter animals' performance in spatial, associative, or reversal learning tasks if elevated during early stages of development [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
+For instance, glucocorticoids (GCs), steroid hormones that mediate the physiological stress response [@sapolsky_how_2000], can alter performance in spatial, associative, or reversal learning tasks if elevated during early stages of development [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
+In ectotherms, temperature can affect GC levels in parents to affect developing offspring (i.e., maternal effects [@jessop2016multiscale; @racic_effects_2020; @mentesana2022glucocorticoids; @Crino_2023].
 Prenatal exposure to GCs is typically associated with impaired learning skills, but these effects have been shown to vary depending on factors such as dose, sex, or type of cognitive test employed [@szuran_water_1994; @farrell_developmental_2015-learn; @bebus_associative_2016; @haq_prenatal_2021].
@@
 For example, the early thermal environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
-In lizards, incubating eggs at elevated temperatures improves cognitive abilities consistent with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017], but impairs cognition at temperatures above lizards’ natural incubation thermal range [@dayananda_incubation_2017; @abayarathna_effects_2020].
-In addition, when @vila_pouca_quantity_2019 tested the effects of elevated incubation temperature on numerical discrimination in  Port-Jackson sharks (_Heterodontus portusjacksoni_), they found that sharks incubated at elevated temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019].
+In lizards, incubating eggs at elevated temperatures improves cognitive abilities consistent with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017], but impairs cognition at extreme temperatures [@dayananda_incubation_2017; @abayarathna_effects_2020].
+In addition, when @vila_pouca_quantity_2019 tested the effects of elevated incubation temperature on numerical discrimination in  Port-Jackson sharks (_Heterodontus portusjacksoni_) and found that sharks incubated at elevated temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019].
 However, the effects of early thermal environment on numerical abilities in other taxa remains largely unexplored.
 
-Temperature can affect GC levels in vertebrates and can indirectly affect developing offspring through maternal effects [@jessop2016multiscale; @racic_effects_2020; @mentesana2022glucocorticoids; @Crino_2023].
 Early-life GCs and thermal conditions could have interactive effects that influence cognitive development.
 Understanding how these factors influence quantitative abilities could reveal important insights into the ecological relevance of numerosity.
-To fill this gap, we used the common garden skink (_Lampropholis guichenoti_) as a model to test the effects of elevated prenatal corticosterone (CORT) - the main GC in reptiles [@Crino_2023] - and incubation temperature on numerical discrimination abilities.
-However, the combined influence of prenatal CORT and thermal environment on cognition remains unknown, particularly on numerical abilities.
-
 Given the influence of numerical cognition on fitness, we would expect evolutionary pressures to be similar across taxa, and yet, research on numerical discrimination has been traditionally focused on mammals, birds, and fishes [@nieder2005counting; @agrillo2014spontaneous].
@@
 
-Here, we investigated the effect of prenatal CORT and incubation temeprature on numerical discrimination in a lizard.
+Here, we investigated the effect of prenatal corticosterone (CORT) - the main GC in reptiles - and incubation temperature on numerical discrimination in a lizard.
 We manipulated CORT concentration (CORT-treated or a sham control) in _L.
@@
 We checked for eggs in the boxes three days a week.
-Eggs' length and width were measured with a digital calliper to the nearest mm (± 0.1 mm), while mass was recorded with a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
+Egg length and width were measured with a digital calliper to the nearest mm (± 0.1 mm), while mass was recorded with a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
 Eggs were assigned a clutch number and an individual identity.
-Then we treated eggs with CORT or vehicle (see CORT and Temperature manipulation below) and placed them in individual cups (80 mL) with moist vermiculite (12 g water to 4 g vermiculite).
-We covered the cups with plastic wrap to retain moisture and left the eggs in incubators (LATWIT 2X5D-R1160) programmed to two different thermal regimes (see CORT and Temperature manipulation below) until hatching.
+Then we treated eggs with CORT or a vehicle control (see _Manipulating Early Thermal and CORT Environments_ details below) and placed them in individual cups (80 mL) with moist vermiculite (12 g water to 4 g vermiculite).
+We covered the cups with plastic wrap to retain moisture and left the eggs in incubators (LATWIT 2X5D-R1160) programmed to two different thermal regimes (see _Manipulating Early Thermal and CORT Environments_ details below) until hatching.
 Incubators were checked for the presence of hatchlings three times a week.
@@
 
-We manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23 ± 3ºC or Hot -- 30 ± 3ºC) in a 2x2 factorial design (@fig-Methods A).
-Eggs were allocated to one of two different developmental stress treatments: CORT-treated eggs were topically supplied with 5µL of a CORT solution (10pg/mL) Control treatment eggs received an equal volume of 100% Ethanol.
+We manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23 ± 3 ºC or Hot -- 30 ± 3 ºC) in a 2x2 factorial design (@fig-Methods A).
+Eggs were allocated to one of two different developmental stress treatments: CORT-treated eggs were topically supplied with 5 µL of a CORT solution (10 pg/mL).
+Control treatment eggs received an equal volume of 100% Ethanol.
 Corticosterone treatments were made by dissolving crystalline corticosterone (Sigma, Cat.
@@
 C2505) in 100% ethanol.
-We selected those doses based on previous studies where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs in _Lampropholis delicata_ [@crino2024eggs].
+We selected doses based on previous studies where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs in _Lampropholis delicata_ [@crino2024eggs].
 Then, eggs were incubated in one of the two previously mentioned temperature regimes (Cold or Hot) until hatching.
 These temperatures represent the upper and lower limit of the natural incubation temperatures for this species [@qualls2000post].
-The number of eggs per clutch assigned to each hormone and temperature treatment was counterbalanced per clutch.
+
+The number of eggs per clutch assigned to each hormone and temperature treatment was counterbalanced in a partial split-clutch design.
+At least one egg per clutch to each treatment, and the remaining eggs were randomly assigned to one of the treatments.
+When we found less than four eggs in a cluthc, we assigned each egg randomly to one of the treatments.
 
@@
 #### Statistical analyses  
-In our analyses, we fitted a multivariate Bayesian multilevel model using the brm function from the _brms_ package [@burkner2017brms] in R extension (version 2.8.2) [@R].
+In our analyses, we fitted a multivariate Bayesian multilevel model using the brm function from the _brms_ package [@burkner2017brms] in R (version 2.8.2) [@R].
 We ran four parallel MCMC chains of 3000 iterations for each model, with a warmup period of 1000 iterations.
@@
 We did not find any effect of sex or age on Choice or Interest, but we found an effect on Latency (see Results).
-However, we decided to include both factors in all models.
+Regardless, we included both factors in all models.
 Age was centered to zero in the model, and we averaged the predicted values for males and females.
@@
 We used the posterior distributions of parameters from these models to test for between treatment differences in the ability to discriminate different numbers of crickets for each of the numerical tests.
-Because the posteriors did not follow a normal distribution, we report the median as the main measure of central tendency.
-We also employed Highest Density Intervals using the hdi function in bayestestR [@bayestestR] to test if the contrasts between treatments for all the variables, or the _Interest_ per each test was different form zero.
-Given Choice was Bernoulli, we modified our null hypotheses to test if differed from a probability of 0.5 as there were only two possible options.
+Because the posterior distributions were not always normally distributed, we report the median as the main measure of central tendency.
+We used the 95% Highest Posterior Density Intervals (95% HPDI) using the hdi function in bayestestR [@bayestestR] to test if the contrasts between treatments for all the variables, or the _Interest_ per each test was different form zero.
+Given Choice was Bernoulli, we modified our null hypotheses to test if differed from a probability of 0.5 as there were only two possible choices and if animals were choosing randomly we would predict each choice would have a 50% probability.
 
@@
 
-Overall, males showed lower latencies than females (median of the estimated effect of sex = `r {format_dec(median(posteriors$b_loglatency_sexm), 2)}`, Higest Density Interval = [`r {format_dec(hdi(posteriors$b_loglatency_sexm, credMass = 0.95)[1], 2)}`, `r {format_dec(hdi(posteriors$b_loglatency_sexm, credMass = 0.95)[2], 2)}`]), and older individuals also made decisions faster (mean estimated effect of age = `r {format_dec(median(posteriors$b_loglatency_age),2)}`, Higest Density Interval = [`r {format_dec(hdi(posteriors$b_loglatency_age, credMass = 0.95)[1], 2)}`, `r {format_dec(hdi(posteriors$b_loglatency_age, credMass = 0.95)[2],2)}`]; see Table S1 in _Supplementary Material_).
+Overall, males showed lower latencies than females (median of the estimated effect of sex = `r {format_dec(median(posteriors$b_loglatency_sexm), 2)}`, 95% HPDI = [`r {format_dec(hdi(posteriors$b_loglatency_sexm, credMass = 0.95)[1], 2)}`, `r {format_dec(hdi(posteriors$b_loglatency_sexm, credMass = 0.95)[2], 2)}`]), and older individuals also made decisions faster (mean estimated effect of age = `r {format_dec(median(posteriors$b_loglatency_age),2)}`, 95% HPDI = [`r {format_dec(hdi(posteriors$b_loglatency_age, credMass = 0.95)[1], 2)}`, `r {format_dec(hdi(posteriors$b_loglatency_age, credMass = 0.95)[2],2)}`]; see Table S3 in _Supplementary Material_).
 However, we did not see any effect of age or sex on _Choice_ or _Interest_ (see _Supplementary Material_).
@@
 The different colours indicate the different treatments.
-Points and bars represent the median and Higest Density Intervals of the estimates, respectively.
+Points and bars represent the median and 95% Higest Density Intervals (95% HPDI) of the estimates, respectively.
 Vertical dashed lines in Choice and Interest graphs values 0.5 and 0, respectively.
@@
 Neither temperature, CORT, or their interaction affected performance in any of the tests, nor were there any clear patterns suggestive of an effect of the developmental environment on numerical discrimination (see Table 1 and @fig-results).
-Except for Hot-CORT lizards in 1 VS 3 tests or Cold-Control lizards in 2 VS 3 tests, the predicted probability of choosing the larger number of crickets first was not different from 0.5 (@fig-results; but also see Table S2 in _Supplementary Material_).
-Similarly, interest in the larger number of crickets (_Interest_) was not different from 0 in any of the tests (@fig-results; but also see Table S2 in _Supplementary Material_).
+Except for Hot-CORT lizards in 1 VS 3 tests or Cold-Control lizards in 2 VS 3 tests, the predicted probability of choosing the larger number of crickets first was not different from 0.5 (@fig-results; but also see Table S1 and Table S2 in _Supplementary Material_).
+Similarly, interest in the larger number of crickets (_Interest_) was not different from 0 in any of the tests (@fig-results; but also see Table S1 and Table S2 in _Supplementary Material_).
 
@@
 The table shows the contrasts for each predictor (Temperature = [medianHot - medianCold]; Hormone = [medianControl - medianCORT]; and their Interaction = [(medianHot-Control - medianHot-CORT) - (medianCold-Control - medianCold-CORT)]).
-Higest Density Intervals (HDI) test the hypothesis that contrasts are different from zero._
+95% Higest Density Intervals (95% HPDI) test the hypothesis that contrasts are different from zero._
 ```{r, tbl-data}
@@
 #| tbl-cap: "Effects of temperature, CORT, and their interaction on Latency, Choice, and Interest in each of the numerical discrimination tests.
-The table shows the contrasts for each predictor (Temperature = [meanHot - meanCold]; Hormone = [meanControl - meanCORT]; Interaction = [(meanHot-Control - meanHot-CORT) - (meanCold-Control - meanCold-CORT)]) and the p~mcmcm~ value of the contrasts.
-In bold, the effects where p~mcmc~ < 0.05."
+The table shows the contrasts for each predictor (Temperature = [medianHot - medianCold]; Hormone = [medianControl - medianCORT]; and their Interaction = [(medianHot-Control - medianHot-CORT) - (medianCold-Control - medianCold-CORT)]).
+95% Higest Density Intervals (95% HPDI) test the hypothesis that contrasts are different from zero."
 source(here("R", "func.R"))
@@
     add_header_row(values = c("", "Tests"), colwidths = c(2, 5)) %>%
-    flextable::compose(i = c(2,3,5,6,8,9), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+    flextable::compose(i = c(3,6,8,9), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+    flextable::compose(i = 2, j = 1, value = as_paragraph("log(latency)"), part = "body") %>%
+    flextable::compose(i = 5, j = 1, value = as_paragraph("log(odds)"), part = "body") %>%
     hline(i = c(3,6), j = c(1:7), part = "body") %>% # To make some horizontal lines 
@@
 Our results indicate that lizards did not discriminate between different numbers of crickets, nor did their performance vary under our experimental conditions.
-Overall, these findings suggest that L.
-guichenoti did not utilize numerical or other potential quantitative cues for foraging decisions.
-This highlights the need for further research into the ecological implications of such behaviors in lizards and their foraging strategies.
+Overall, these findings suggest that _L.
+guichenoti_ did not use numerical or other potential quantitative cues for foraging decisions.
 
@@
 guichenoti_ may be due to insectivorous lizards relying on other cues, such as movement or size, for foraging decisions.
-In those studies where reptiles discriminated between different amounts of food, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
+In studies where reptiles discriminated between different amounts of food, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
 10)[@recio_prey_2021].
@@
 For example, when presented two groups with different numbers of conspecifics or if prey handling costs are related to larger quantitative traits like size [@agrillo2014spontaneous].
-It could be that the perceived handling costs of larger numbers of prey is affecting lizards' response.
-Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of lizards' size on their choices, as larger animals may be more adept at handling a greater number of prey items.
-@miletto2017quantitative did not report any effect of lizards' size, and age - a proxy for size - did not affect _L.
+It could be that the perceived handling costs of larger numbers of prey is affecting lizard response in our study.
+Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of lizard size on their choices, as larger animals may be more adept at handling a greater number of prey items.
+@miletto2017quantitative did not report any effect of body size, and age - a proxy for size - did not affect _L.
 guichenoti_'s _Choice_ or _Interest_ in our study.
@@
 siculus_ and _L.
-guichenoti_ suggests that the benefits of selecting larger number of prey are not different from selecting fewer.
+guichenoti_ suggests that the benefits of selecting a larger number of prey are not different from selecting fewer.
 For instance, when considering hunting groups of 1 VS 4 or 2 VS 4 crickets, the probability of success might be similar.
@@
 guichenoti_ and how they interact with numerical information.
+Such studies would enhance our understanding of how _L.
+guichenoti_ integrates different quantitative cues during foraging, providing insights into their decision-making processes and overall quantitative abilities.
 
-#### Foraging behaviour is robust to prenatal CORT and temeprature  
-Although the effects of early environment on cognition is widely studied, little atention has been paid to its impacts on numerical discrimination.
-To the best of our knowledge, there is only one study examining how developmental conditions influence numerical abilities.
-@vila_pouca_quantity_2019 investigated the effects of incubating and rearing Port Jackson sharks (_Heterodontus portusjacksoni_) at different temperatures on their ability to learn a quantity discrimination task.
-They found that those animals reared in warm water took fewer days to reach the learning criterion.
+#### Foraging behaviour is robust to early environmental experiences  
+Although the effects of early environment on cognition is widely studied, little attention has been paid to its impacts on numerical discrimination.
+To the best of our knowledge, there is only one study examining how developmental conditions influence numerical abilities [@vila_pouca_quantity_2019].
 Given the potent effects of early environment on cognition and brain development [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic], it is crucial to explore how early environmental conditions shape numerical abilities across taxa.
@@
 For example, incubating eggs of the eastern three-lined skink (_Bassiana dupeyerri_) at elevated temperatures increased cell density in the medial cortex - linked to learning process and memory - but did not impact cell density in other nuclei [@amiel_effects_2017].
-Similarly, impairments in spatial learning in rats [@haq_prenatal_2021] are likely associated with the greater density of GCs receptors in mammals' hippocampus compared to other regions of the brain [@siegel_2006].
+Similarly, impairments in spatial learning in rats [@haq_prenatal_2021] are likely associated with the greater density of GCs receptors in the hippocampus compared to other regions of the brain [@siegel_2006].
 Researchers have proposed that non-verbal number representation is processed by two types of systems: one for small sets (≤ 4) that tracks and encodes objects as individual entities, and another for larger numerosities that operate through estimation processes [@nieder2005counting; @cantlon2009beyond; @hyde2011two].
@@
 
-Alternatively, it is plausible that treatment effects are masked by the overall poor performance across most of the tests, which could be a consequence of the species’ reliance on another type of cues for foraging other than numerical information.
+Alternatively, it is plausible that treatment effects are masked by the overall poor performance across most of the tests, which could be a consequence of the species’ reliance on other types of cues for foraging other than numerical information.
 In this regard, motivation is a key factor in spontaneous choice tests [@agrillo2014spontaneous; @bisazza2014experim].
 Both developmental CORT and incubation temperature can affect metabolic rate (hereafter MR) which could affect motivation in food based learning paradigms [@crino_corticosterone_2014-learn; @noble_developmental_2018; @cossin2022effect].
-Here, we  controlled for motivation by providing only one cricket per day and fasting lizards for more than 20 hours before the test.
-In addition, we excluded those tests where lizards did not interact with any of the options or did not eat the cricket provided after the test.
-However, it is possible that the fasting period was not long enough to standardize hunger levels across individuals.
-Further studies should explore the effects of prenatal conditions on motivation and how it interacts with cognitive abilities in reptiles.
+Although still possible, this explanation is less likely in our study because we controlled for motivation by providing only one cricket per day and fasting lizards for more than 20 hours before the test.
+In addition, we excluded tests where lizards did not interact with any of the options or did not eat the cricket provided after the test.
+Further studies should explore the effects of prenatal conditions on motivation and how it may affect the outcome of cognitive choice tests in reptiles.
 
@@
 ## Funding  
-This work was supported by a National Australian University fellowship (PR), and the Australian Research Council (grant no. DP210101152) to DN and CF.
+This work was supported by a National Australian University PhD scholarship (PR), and the Australian Research Council (grant no. DP210101152) to DN and CRF.
 
@@
 
-#### Model results
+#### Effects of treatments on numerical discrimination using the estimates and the raw data
 
-_Table S1.- Summary of the model fitted_
-
-```{r, summary_org}
-#| label: summary_org
-#
-## Get the summary of the posteriors to get the estimates of each parameter of the model
-sum <- summary(posteriors)
-## Extract the values of the summary
-sum_2 <- as.data.frame(sum) %>%
-  dplyr::select(-mad) %>%
-  filter(!str_detect(variable, "^r_"))
-sum_df <- sum_2 %>%
-  mutate(across(where(is.numeric), ~ sprintf(paste0("%.", 2, "f"), .)))  
-```
-
-For loglatency (Latency)
-```{r, tbl-summary_lat}
-#| label: tbl-summary_lat
-#| tbl-cap: "Summary of the model fitted for the latency variable."
-#
-source(here("R", "func.R"))
-## Get df only for latency
-df_sum_lat <- sum_table_org(sum_df, "lat")
-## Make table
-table_sum_lat <- flextable(df_sum_lat)
-table_sum_lat
-```
-
-For Choice
-```{r, tbl-summary_choice}
-#| label: tbl-summary_choice
-#| tbl-cap: "Summary of the model fitted for the choice variable."
-#
-source(here("R", "func.R"))
-## Get df only for choice
-df_sum_choice <- sum_table_org(sum_df, "choice")
-## Make table
-table_sum_choice <- flextable(df_sum_choice)
-table_sum_choice
-```
-
-For Interest
-```{r, tbl-summary_int}
-#| label: tbl-summary_int
-#| tbl-cap: "Summary of the model fitted for the interest variable."
-#
-source(here("R", "func.R"))
-## Get df only for interest
-df_sum_int <- sum_table_org(sum_df, "int")
-## Make table
-table_sum_int <- flextable(df_sum_int)
-table_sum_int
-```
-
-Other effects
-```{r, tbl-summary_other}
-#| label: tbl-summary_other
-#| tbl-cap: "Summary of the model fitted for other effects."
-#
-source(here("R", "func.R"))
-## Get df only for interest
-df_sum_other <- sum_df %>%
-  filter(!str_detect(variable, "^b_"))
-## Make table
-table_sum_other <- flextable(df_sum_other)
-table_sum_other
-```
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-#### Treatments effects
-
-_Table S2.- Median and Higest Density Intervals of the estimated probability of choosing first the higher amount (Choice) and the estimated interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
-Higest Density Intervals test the hypothesis that Choice = 0.5, and Interest = 0, which would indicate a preference towards one of choices._
+_Table S1.- Median and 95% Higest Density Intervals (95% HPDI) of the estimated probability of choosing first the higher amount (Choice) and the estimated interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
+95% HPDI test the hypothesis that Choice = 0.5, and Interest = 0, which would indicate a preference towards one of choices._
 ```{r, tbl-disc}
 #| label: tbl-disc
-#| tbl-cap: "Median and Higest Density Intervals of the estimated probability of choosing first the higher amount (Choice) and the estimated interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
-Higest Density Intervals test the hypothesis that Choice = 0.5, and Interest = 0, which would indicate a preference towards one of choices."
+#| tbl-cap: "Median and 95% HPDI of the estimated probability of choosing first the higher amount (Choice) and the estimated interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
+95% HPDI test the hypothesis that Choice = 0.5, and Interest = 0, which would indicate a preference towards one of choices."
 #
@@
     Hot_Control <- post_mod[[paste0("X", d, "_Hot_Control")]]
-    # Calculate medians and Higest Density Intervals
+    # Calculate medians and 95% HPDI
     Cold_CORT_median <- format_dec(median(Cold_CORT), 2)
@@
     # Define labels and values
-    treatment_names <- c("Cold-CORT", "Cold-Control", "Hot-CORT", "Hot-Control")
+    treatment_names <- c("Cold-CORT (n = 20)", "Cold-Control", "Hot-CORT", "Hot-Control")
     values_medians <- c(Cold_CORT_median, Cold_Control_median, Hot_CORT_median, Hot_Control_median)
@@
                 Effect_Choice, int_Choice,
-                Effect_Interest, int_Interest)
+                Effect_Interest, int_Interest) %>%
+  mutate(Treatment = factor(Treatment, levels = c("Hot-Control (n = 20)",
+                                      "Hot-CORT (n = 20)",
+                                      "Cold-Control (n = 20)",
+                                      "Cold-CORT (n = 20)"))) %>%
+data.frame()
 #
@@
  font.family = "Arial",
- fint.size = 10)
+ font.size = 10)
 # Create the table
@@
     flextable::compose(i = 2, j = c(3,5), value = as_paragraph("Median"), part = "head") %>% 
-    flextable::compose(i = 2, j = c(4,6), value = as_paragraph("HPD Interval"), part = "head") %>% 
+    flextable::compose(i = 2, j = c(4,6), value = as_paragraph("95% HPDI"), part = "head") %>% 
     hline(i = c(4,8,12,16), j = c(1:6), part = "body") %>% # To make some horizontal lines
@@
 
+_Table S2.- Performance of each treatment in each of the numerical tests using the raw data.
+For the variable Latency and Interest, we show the median and the 95% CI.
+For the varaible Choice, we show the proportion of individuals that chose the higher number of crickets first._
+
+```{r, tbl-discraw}
+#| label: tbl-discraw
+#| tbl-cap: "Performance of each treatment in each of the numerical tests using the raw data.
+For the variable Latency and Interest, we show the median and the 95% CI.
+For the varaible Choice, we show the proportion of individuals that chose the higher number of crickets first."
+#
+data_table_raw <- data_num %>%
+  dplyr::select(trt, test_type, latency, choice, compared_interest) %>%
+  group_by(trt, test_type) %>%
+  summarize(
+    mean_latency = format_dec(mean(latency, na.rm = TRUE), 2),
+    ci_latency_lower = format_dec(quantile(latency, 0.025, na.rm = TRUE), 2),
+    ci_latency_upper = format_dec(quantile(latency, 0.975, na.rm = TRUE), 2),
+    mean_interest = format_dec(mean(compared_interest, na.rm = TRUE), 2),
+    ci_interest_lower = format_dec(quantile(compared_interest, 0.025, na.rm = TRUE), 2),
+    ci_interest_upper = format_dec(quantile(compared_interest, 0.975, na.rm = TRUE), 2),
+    Choice = as.character(sum(choice == 1, na.rm = TRUE))
+  ) %>%
+  ungroup() %>%
+  mutate(
+    ci_latency = paste0("[", ci_latency_lower, " , ", ci_latency_upper, "]"),
+    ci_interest = paste0("[", ci_interest_lower, " , ", ci_interest_upper, "]"),
+  ) %>%
+  mutate(
+    Latency = trimws(paste0(mean_latency, "\n", ci_latency)),
+    Interest = trimws(paste0(mean_interest, "\n", ci_interest))
+  ) %>%
+  mutate(trt = recode(trt, 
+            "B_23" = "Cold-CORT (n = 20)", 
+            "A_23" = "Cold-Control (n = 20)", 
+            "B_28" = "Hot-CORT (n = 20)", 
+            "A_28" = "Hot-Control (n = 20)")) %>% # To order the treatments
+  dplyr::select(test_type, trt, Latency, Choice, Interest) %>% 
+  pivot_longer(cols = c(Latency, Choice, Interest), 
+               names_to = "variable", 
+               values_to = "value") %>%
+  pivot_wider(names_from = test_type, 
+              values_from = value) %>%
+  dplyr::select(variable, trt, "1 VS 4", "1 VS 3", "2 VS 4", "2 VS 3", "3 VS 4") %>%
+  mutate(variable = factor(variable, levels = c("Latency", "Choice", "Interest")),
+         trt = factor(trt, levels = c("Hot-Control (n = 20)",
+                                      "Hot-CORT (n = 20)",
+                                      "Cold-Control (n = 20)",
+                                      "Cold-CORT (n = 20)"))) %>%
+  arrange(variable, trt) %>%
+data.frame()
+data_table_raw_b <- data_table_raw %>%
+  rename(
+    Variable = variable, 
+    Treatment = trt, 
+    "1 VS 4" = X1.VS.4,
+    "1 VS 3" = X1.VS.3,
+    "2 VS 4" = X2.VS.4,
+    "2 VS 3" = X2.VS.3,
+    "3 VS 4" = X3.VS.4) 
+#
+#
+# Table format
+set_flextable_defaults(
+ font.family = "Arial",
+ font.size = 10)
+# Create the table
+table_raw <- flextable(data_table_raw_b) %>%
+  # Add your custom header rows
+  flextable::add_header_row(values = c("", "Tests", ""), colwidths = c(4, 1, 2)) %>%
+  # Take away some text
+  flextable::compose(i = c(2,3,4,6,7,8,10,11,12), j = 1, value = as_paragraph(""), part = "body") %>%
+  hline(i = c(4,8), j = c(1:7), part = "body") # To make some horizontal lines
+table_raw
+```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+#### Model results
+
+_Table S3A.- Summary of the model fitted for loglatency (Latency)_
+
+```{r, summary_org}
+#| label: summary_org
+#
+## Get the summary of the posteriors to get the estimates of each parameter of the model
+sum <- summary(posteriors)
+## Extract the values of the summary
+sum_2 <- as.data.frame(sum) %>%
+  dplyr::select(-mad) %>%
+  filter(!str_detect(variable, "^r_"))
+sum_df <- sum_2 %>%
+  mutate(across(where(is.numeric), ~ sprintf(paste0("%.", 2, "f"), .)))  
+```
+
+```{r, tbl-summary_lat}
+#| label: tbl-summary_lat
+#| tbl-cap: "Summary of the model fitted for the latency variable."
+#
+source(here("R", "func.R"))
+## Get df only for latency
+df_sum_lat <- sum_table_org(sum_df, "lat")
+## Make table
+table_sum_lat <- flextable(df_sum_lat)
+table_sum_lat
+```
+
+_Table S3B.- Summary of the model fitted for Choice_   
+```{r, tbl-summary_choice}
+#| label: tbl-summary_choice
+#| tbl-cap: "Summary of the model fitted for the choice variable."
+#
+source(here("R", "func.R"))
+## Get df only for choice
+df_sum_choice <- sum_table_org(sum_df, "choice")
+## Make table
+table_sum_choice <- flextable(df_sum_choice)
+table_sum_choice
+```
+
+_Table S3C.- Summary of the model fitted for Interest_  
+```{r, tbl-summary_int}
+#| label: tbl-summary_int
+#| tbl-cap: "Summary of the model fitted for the interest variable."
+#
+source(here("R", "func.R"))
+## Get df only for interest
+df_sum_int <- sum_table_org(sum_df, "int")
+## Make table
+table_sum_int <- flextable(df_sum_int)
+table_sum_int
+```
+
+_Table S3D.- Summary of the model fitted for other effects_  
+```{r, tbl-summary_other}
+#| label: tbl-summary_other
+#| tbl-cap: "Summary of the model fitted for other effects."
+#
+source(here("R", "func.R"))
+## Get df only for interest
+df_sum_other <- sum_df %>%
+  filter(!str_detect(variable, "^b_"))
+## Make table
+table_sum_other <- flextable(df_sum_other)
+table_sum_other
+```
+
 ```{r, results='asis', echo=FALSE}
@@
 
-_Table S3.- Number of individuals per treatment that chose the right (R) or left (L) side in each of the numerical tests.
+_Table S4.- Number of individuals per treatment that chose the right (R) or left (L) side in each of the numerical tests.
 p-value indicates the result of the binomial test comparing the number of choices between sides.
-In bold, the groups where p-value < 0.05_
 ```{r, tbl-sides}
@@
 
-_Table S4.
-Number of individuals per treatment that chose the cricket oriented horizontally (Horizontal) or vertically (Vertical) in our control tests.
+_Table S4.- Number of individuals per treatment that chose the cricket oriented horizontally (Horizontal) or vertically (Vertical) in our control tests.
 p-value indicates the result of the binomial test comparing the number of choices between both choices._
@@
   mutate(trt = recode(trt, 
-            "B_23" = "CORT-Cold (n = 20)", 
-            "A_23" = "Control-Cold (n = 20)", 
-            "B_28" = "CORT-Hot (n = 20)", 
-            "A_28" = "Control-Hot (n = 20)")) %>% # To order the treatments
+            "B_23" = "Cold-CORT (n = 20)", 
+            "A_23" = "Cold-Control (n = 20)", 
+            "B_28" = "Hot-CORT (n = 20)", 
+            "A_28" = "Hot-Control (n = 20)"),
+          trt = factor(trt, levels = c("Hot-Control (n = 20)",
+                                  "Hot-CORT (n = 20)",
+                                  "Cold-Control (n = 20)",
+                                  "Cold-CORT (n = 20)"))) %>% # To order the treatments
   dplyr::select(trt, Horizontal_Count, Vertical_Count, p_value) %>% # Remove unused columns
@@
 orient_table <- flextable(data_or_res) %>%
-  compose(j = 1, value = as_paragraph(""), part = "header") %>%
-  compose(j = 2, value = as_paragraph("Horizontal"), part = "header") %>%
-  compose(j = 3, value = as_paragraph("Vertical"), part = "header") %>%
-  compose(j = 4, value = as_paragraph("p-value"), part = "header") %>%
+  flextable::compose(j = 1, value = as_paragraph(""), part = "header") %>%
+  flextable::compose(j = 2, value = as_paragraph("Horizontal"), part = "header") %>%
+  flextable::compose(j = 3, value = as_paragraph("Vertical"), part = "header") %>%
+  flextable::compose(j = 4, value = as_paragraph("p-value"), part = "header") %>%
   set_table_properties( width = 0.9, layout = "autofit") %>%
@@
 #### Searching for relevant literature
-To loof for relevant literature about the effects of early environment on numerical discrimination, we searched in the Web of Science and Scopus using the terms on respectively.
+To look for relevant literature about the effects of early environment on numerical discrimination, we searched in the Web of Science and Scopus using the terms on respectively.
+The queries employed are detailed below.
+We looked for manuscripts for all years available.
+The last search was conducted on the 11^th^ of October 2024; the results below refer to this last search.
 
-We found 24 articles on Web of Science and 22 on Scopus, plus only one preprint on Scopus.
+We found 24 articles on Web of Science and 22 on Scopus, plus one preprint on Scopus.
 Out of all the documents, only one study was relevant to our search: @vila_pouca_quantity_2019.

### 2024-11-07 .vdoc.r

@@ -1,0 +1,747 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: packages
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS, nortest, fitdistrplus, ggh4x, PupillometryR, cowplot, png, grid, remotes, ggthemes, bayestestR, HDInterval)
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+cat("\\newpage")
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: fig-Methods
+#| fig.cap: "Experimental design. Panel (**A**) shows the early environment manipulation procedures. (**B**) illustrates the arena where the tests were performed, while panel (**C**) indicates the measurements of the platform used for the experiments. Finally, panel (**D**) displays the types of numerical tests used and the orientation of the crickets in each test."
+
+knitr::include_graphics("./Others/NUMDIS_FIG_1.png")
+
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: data_process
+# Load the data and process it to get the main dfs
+source(here("R", "data_process.R"))
+#
+#
+#
+#| label: models
+# Fitting the model and extraction of posteriors 
+refit = FALSE
+if(refit){
+  # Get the formula and family of variables
+  lat <- bf(log(latency) ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = gaussian())
+  choice <- bf(choice ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = bernoulli(link = "logit"))
+  int <- bf(compared_interest ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch),
+    family = gaussian())
+  # Fit the model
+  model <- brm(lat + choice + int,
[…]

### 2024-11-12 ms.qmd
Updates

@@
 From social behaviour to navigating complex environments, quantitative abilities can be crucial to fitness.
-Early life conditions can impact brain development to affect quantitative competence.
-In anthropogenically modified environments, early thermal conditions or glucocorticoid concentrations (GCs) – stress-related hormones transmitted by the parents – can play a prominent role in shaping cognition through their effects on brain physiology.
-Furthermore, temperature can lead to increased levels of GCs that are passed on to offspring, emphasizing the need to study the combined effects of these two factors.
+However, early life conditions can impact brain development to affect quantitative competence.
+For example, early thermal conditions or glucocorticoid concentrations (GCs) – stress-related hormones transmitted by the parents – can play a prominent role in shaping cognition through their effects on brain physiology.
+Furthermore, temperature can lead to increased levels of GCs that can be transmitted to offspring, emphasizing the need to study the combined effects of these two factors.
 Here, we investigated the effects of elevated prenatal corticosterone (CORT) – the main GC in reptiles – and incubation temperature on numerical discrimination in the common garden skink (_Lampropholis guichenoti_).
-We then assessed numerical abilities through a spontaneous choice test with food as the relevant stimulus.
-Employing a repeated measures design, we subjected lizards to five numerical tests each differing in the ratios between the two choices (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 2 VS 4).
-Contrary to our expectations, we found no evidence for treatment effects on lizard behaviour and no use of numerical discrimination during foraging in _L.
+We assessed numerical abilities through a spontaneous choice test with food as a stimulus.
+Employing a repeated measures design, we subjected lizards to five numerical tests each differing in the ratios between the two choices (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4).
+Contrary to our predictions, we found no evidence for treatment effects on lizard behaviour and no use of numerical discrimination during foraging in _L.
 guichenoti_.
-Our findings align with previous studies indicating that reptiles feeding on live prey may have difficulty distinguishing between different prey quantities in spontaneous choice tests.
+Our findings build on previous studies suggesting that reptiles feeding on live prey may have difficulty distinguishing between different prey quantities.
+Despite our results, further investigation of numerical abilities in live-prey-feeding reptiles and the impact of early conditions on other taxa is warranted, as these areas remain understudied.
 
@@
 ## Introduction  
-The ability to discern between quantities is a fundamental cognitive skill widely observed across the animal kingdom [@stancher2015numerical; @vonk2012bears; @beran2016capuchin].
+The ability to discriminate between quantities is a fundamental cognitive skill widely observed across the animal kingdom [@stancher2015numerical; @vonk2012bears; @beran2016capuchin].
 Numerical cognition can contribute to social decisions, mate selection, and foraging behaviour, which can be adaptive in various ecological contexts [@lucon2017individual; @nieder2018evolution].
 For instance, female lions (_Panthera leo_) use numerical information to assess the risk of confronting rival groups [@mccomb1994roaring], male adult mealworms (_Tenebrio molitor_) select sites with scents of more females [@carazo2009quantity], and red-backed salamanders (_Plethodon cinereus_) prefer tubes containing larger numbers of prey [@uller2003salamanders].
-Given the adaptive advantages of numerical and quantity discrimination, research has focused extensively on exploring the ability of different species to differentiate between quantities under different contexts and paradigms.
+Given the adaptive advantages of numerical and quantity discrimination, research has focused extensively on exploring the ability of different species to differentiate between quantities employing different contexts and paradigms.
 
@@
 The latter have demonstrated that non-verbal numerical representations in mammals activates an extensive neural pathway including regions involved in eye movement, attention control, or complex learning [@nieder2005counting].
-The connection between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.
+The connection between numerical discrimination and brain physiology underscores the importance of understanding how brain development influences numerical abilities.
 
@@
 For instance, glucocorticoids (GCs), steroid hormones that mediate the physiological stress response [@sapolsky_how_2000], can alter performance in spatial, associative, or reversal learning tasks if elevated during early stages of development [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
-In ectotherms, temperature can affect GC levels in parents to affect developing offspring (i.e., maternal effects [@jessop2016multiscale; @racic_effects_2020; @mentesana2022glucocorticoids; @Crino_2023].
+In ectotherms, temperature can affect GC levels in parents, cascading into effects on the developing offspring [i.e., maternal effects; @jessop2016multiscale; @racic_effects_2020; @mentesana2022glucocorticoids; @Crino_2023].
 Prenatal exposure to GCs is typically associated with impaired learning skills, but these effects have been shown to vary depending on factors such as dose, sex, or type of cognitive test employed [@szuran_water_1994; @farrell_developmental_2015-learn; @bebus_associative_2016; @haq_prenatal_2021].
-Nevertheless, little is known about how exposure to GCs during development can affect quantitative abilities.
+Most past studies have focused on the effects of GC exposure during development on spatial, associative, or reversal learning.
+Few studies have focused on how exposure to GCs during development can affect quantitative abilities, despite the importance of numerical discrimination.
 
-The effects of prenatal GCs on learning may be modulated by other developmental effects or environmental conditions.
-For example, the early thermal environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
+Although maternally transmitted GCs are a significant factor in cognitive development, other environmental conditions can also influence cognition, either independently or through interactions with GCs.
+For instance, the early thermal environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
 In lizards, incubating eggs at elevated temperatures improves cognitive abilities consistent with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017], but impairs cognition at extreme temperatures [@dayananda_incubation_2017; @abayarathna_effects_2020].
@@
 For example, red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour with higher volumes of food [@soldati2017long], Hermann’s tortoises (_Testudo hermanni_) discriminate between large and small food quantities in a spontaneous choice test [@gazzola2018continuous], and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) rapidly learned a quantity discrimination task [@lin2021superior; @lin2024trained].
-The italian wall lizard (_Podarcis siculus_) failed to discriminate between different numbers of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity].
+The Italian wall lizard (_Podarcis siculus_) failed to discriminate between different numbers of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity].
 In contrast, Iberian rock lizards (_Iberolacerta cyreni_) preferred larger quantities of crickets when presented five versus ten items [@recio_prey_2021].
@@
 guichenoti_ eggs and incubated them under one of two temperature regimes (Cold - 23 ± 3ºC or Hot - 28 ± 3ºC) in a 2x2 factorial design.
-Then, we assessed numerical discrimination abilities of lizards unsing a spontaneous choice test approach.
+Then, we assessed numerical discrimination abilities of lizards using a spontaneous choice test approach.
 The numerical discrimination task consisted of five tests that differed in the relative number of food items: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
-We based this experiment on observations in nature that the ability to detect a change in stimulus intensity depends on the relative, not absolute, difference between the stimuli ['Weber's law'; @agrillo2014spontaneous].
+We based this experiment on observations in nature that the ability to detect a change in stimulus intensity depends on the relative, not absolute, difference between the stimuli ('Weber's law'; @agrillo2014spontaneous).
 Built upon other studies that explored the cognitive effects of prenatal GCs or thermal early environment in reptiles and other taxa, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
@@
 
-We manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23 ± 3 ºC or Hot -- 30 ± 3 ºC) in a 2x2 factorial design (@fig-Methods A).
+We manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design (@fig-Methods A).
 Eggs were allocated to one of two different developmental stress treatments: CORT-treated eggs were topically supplied with 5 µL of a CORT solution (10 pg/mL).
@@
 At least one egg per clutch to each treatment, and the remaining eggs were randomly assigned to one of the treatments.
-When we found less than four eggs in a cluthc, we assigned each egg randomly to one of the treatments.
+When we found less than four eggs in a clutch, we assigned each egg randomly to one of the treatments.
 
@@
 In contrast, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities despite showing high quantitative skills when prey size was the main cue [@miletto2017quantitative] or when tested using training procedures [@miletto2018quantity].
-Remarkably, the numbers of food items employed in @miletto2017quantitative were always fewer than four.
+Notably, the numbers of food items employed in @miletto2017quantitative were always fewer than four.
 Spontaneous choice tests assume that animals should select the most appropriate quantity, which may not necessarily be the largest one [@agrillo2014spontaneous].
@@
 For example, when presented two groups with different numbers of conspecifics or if prey handling costs are related to larger quantitative traits like size [@agrillo2014spontaneous].
-It could be that the perceived handling costs of larger numbers of prey is affecting lizard response in our study.
-Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of lizard size on their choices, as larger animals may be more adept at handling a greater number of prey items.
+It could be that the perceived handling costs of larger numbers of prey affected lizard response in our study.
+Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food.
+Additionally we would have expected body size to influence lizards choices, as larger animals may be more adept at handling a greater number of prey items.
 @miletto2017quantitative did not report any effect of body size, and age - a proxy for size - did not affect _L.
@@
 For instance, when considering hunting groups of 1 VS 4 or 2 VS 4 crickets, the probability of success might be similar.
-Furthermore, while one individual handles a prey item in the patch, other conspecifics could capture the rest.
+Furthermore, while one individual handles a prey item within the patch, other conspecifics could capture the rest.
 As a result, insectivorous lizards might not have evolved a preference for larger patches of prey, at least with small numbers [see @recio_prey_2021].
@@
 Conversely, our findings may be influenced by the experimental design.
-In our study, we rotated the position of the crickets in the choice with fewer items to create a situation where the numbers and other quantitative cues associatied with numerosity were in conflict.
+In our study, we rotated the position of the crickets in the choice with fewer items to create a situation where the numerical cues (fewer items) conflicted with other quantitative cues associated with numerosity.
 This approach was intended to control for other quantitative cues that could be used by lizards to make decisions [@cooper2024tell].
@@
 Both GCs and thermal environment can exert differential effects on distinct areas of the brain.
-For example, incubating eggs of the eastern three-lined skink (_Bassiana dupeyerri_) at elevated temperatures increased cell density in the medial cortex - linked to learning process and memory - but did not impact cell density in other nuclei [@amiel_effects_2017].
+For example, in the eastern three-lined skink (_Bassiana dupeyerri_), elevated incubation temperatures increased cell density in the medial cortex - linked to learning process and memory - but did not impact cell density in other nuclei [@amiel_effects_2017].
 Similarly, impairments in spatial learning in rats [@haq_prenatal_2021] are likely associated with the greater density of GCs receptors in the hippocampus compared to other regions of the brain [@siegel_2006].
 Researchers have proposed that non-verbal number representation is processed by two types of systems: one for small sets (≤ 4) that tracks and encodes objects as individual entities, and another for larger numerosities that operate through estimation processes [@nieder2005counting; @cantlon2009beyond; @hyde2011two].
-While it remains unknown which areas of the reptiles' brain encode numerical information, our results suggest that those brain regions involved in discrete numerical representation (≤ 4 items) might be robust to the effect of prenatal CORT and temperature in _L.
+While it remains unknown which areas of the reptiles' brain encode numerical information, our results suggest that the brain regions involved in discrete numerical representation (≤ 4 items) might be robust to the effects of prenatal CORT and temperature in _L.
 guichenoti_.
 
-Alternatively, it is plausible that treatment effects are masked by the overall poor performance across most of the tests, which could be a consequence of the species’ reliance on other types of cues for foraging other than numerical information.
+Alternatively, it is plausible that treatment effects are masked by the overall absence of discrimination between numbers of prey items across most of the tests.
 In this regard, motivation is a key factor in spontaneous choice tests [@agrillo2014spontaneous; @bisazza2014experim].
-Both developmental CORT and incubation temperature can affect metabolic rate (hereafter MR) which could affect motivation in food based learning paradigms [@crino_corticosterone_2014-learn; @noble_developmental_2018; @cossin2022effect].
+Both developmental CORT and incubation temperature can affect metabolic rate (hereafter MR) which could affect motivation in food based learning paradigms [@noble_developmental_2018; @cossin2022effect; but see @crino_corticosterone_2014-learn].
 Although still possible, this explanation is less likely in our study because we controlled for motivation by providing only one cricket per day and fasting lizards for more than 20 hours before the test.
@@
 Our findings suggest that _L.
-guichenoti_ does not rely on numerical information for foraging decisions and that their performance does not vary across different prenatal treatments.
+guichenoti_ does not rely on numerical information for foraging decisions and that their performance does not vary due to prenatal treatments.
 These results indicate that early environmental conditions, such as incubation temperature and CORT exposure, may not affect numerosity in this species.
@@
 We declare Chat GPT was used for questions related to coding and data analyses.
-All other parts of the manuscript were written by the authors and those parts where AI was used were checked and modified accordingly by the authors.
+All other parts of the manuscript were written by the authors and those parts where AI were used were checked and modified accordingly by the authors.
 
@@
 #### Prey orientation test
-Because the orientation of the crickets could affect the lizards' choice, we performed one simple choice test to investigate this.
+We performed a choice test to investigate if the orientation of crickets affected the lizards’ choice.
 We used the same platform and Petri dish as in the numerical discrimination task, but we placed one cricket on each side of the platform, one oriented vertically in respect to the shelter position (see @fig-Methods B) and the other one horizontal to the shelter.

### 2025-02-28 ms.qmd
Updates
updates

@@
 ## Abstract
-From social behaviour to navigating complex environments, quantitative abilities can be crucial to fitness.
+
+**From social behaviour to navigating complex environments, quantitative abilities can be crucial to fitness.
 However, early life conditions can impact brain development to affect quantitative competence.
@@
 Our findings build on previous studies suggesting that reptiles feeding on live prey may have difficulty distinguishing between different prey quantities.
-Despite our results, further investigation of numerical abilities in live-prey-feeding reptiles and the impact of early conditions on other taxa is warranted, as these areas remain understudied.
+Despite our results, further investigation of numerical abilities in live-prey-feeding reptiles and the impact of early conditions on other taxa is warranted, as these areas remain understudied.**
 
 ## Significance statement
-Discriminating between quantities can be crucial for individual fitness, yet the impact of early environmental factors on the development of quantitative competence remains largely unexplored.
+
+**Discriminating between quantities can be crucial for individual fitness, yet the impact of early environmental factors on the development of quantitative competence remains largely unexplored.
 In this study, we investigated, for the first time, the effects of prenatal corticosterone exposure and incubation temperature on numerical discrimination in a reptile.
@@
 These results open new avenues for exploring the evolution of numerical abilities in reptiles and identifying factors influencing the development of quantitative skills in this group.
-Further research should explore alternative methods, such as trained tasks, to better understand the cognitive mechanisms underlying quantitative competence in reptiles.
+Further research should explore alternative methods, such as trained tasks, to better understand the cognitive mechanisms underlying quantitative competence in reptiles.**
 
@@
 
-Numerical competence is often evaluated by considering the types of cues animals use [@plotnik2019elephants; @cooper2024tell], the ecological contexts in which these abilities are deployed [@mccomb1994roaring; @chittka1995can; @uller2003salamanders; @carazo2009quantity], and the neural mechanisms that underlie them [@nieder2005counting].
-The latter have demonstrated that non-verbal numerical representations in mammals activates an extensive neural pathway including regions involved in eye movement, attention control, or complex learning [@nieder2005counting].
-The connection between numerical discrimination and brain physiology underscores the importance of understanding how brain development influences numerical abilities.
+The main two approaches experimental approaches in the study of quantity discrimination are spontaneous choice tests and training procedures.
+In spontaneous choice tests, the subjects are presented two groups of biologically relevant stimuli differing in some quantitative trait (number, size, volumen, etc) and the experimenters record if the subjects select or not the most most appropriate quantity (e.g.
+the group with more conspecifics or the larger amount of food) (@agrillo2014spontaneous; @cox2016quantity).
+In training procedures, animals must learn a numerical rule to receive a reward [@agrillo2014spontaneous; @miletto2018quantity].
+Both methods are believed to be complementary to each other, and both can provide different results even in the same species [see @miletto2017quantitative and @miletto2018quantity].
+One reason is that spontaneous choice tests assume that one of the options presented is more adequate than the other when this may not always be the case.
+For example, choosing groups of conspecifics with less individuals could be advantageous when food is scarce but disadvantageous when predators are abundant [**REFs first thesis??**].
+As such, context and motivation can be important drivers of decision-making that can change the way we measure animals' abilities to discriminate between quantities [@agrillo2014spontaneous].
+The other reason is that each method may tap into different cognitive mechanisms [@agrillo2014spontaneous].
+In this sense, **Researchers have suggested that numerousness can be processed by two types of systems: one system for small sets (≤ 4) that works by tracking and encoding objects as individual entities; and an analogous system that operates with large numerosities by estimation processes (for a review, see Cantlon et al.
+(2009); Hyde (2011).**   **HOW CAN EACH BE RELATED TO MOTIVATION OR LEARNING??**
 
-Brains are particularly sensitive to environmental inputs during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004], with long-lasting effects on cognitive abilities [@zhu_prenatal_2004; @amiel_egg_2014].
-For instance, glucocorticoids (GCs), steroid hormones that mediate the physiological stress response [@sapolsky_how_2000], can alter performance in spatial, associative, or reversal learning tasks if elevated during early stages of development [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
-In ectotherms, temperature can affect GC levels in parents, cascading into effects on the developing offspring [i.e., maternal effects; @jessop2016multiscale; @racic_effects_2020; @mentesana2022glucocorticoids; @Crino_2023].
-Prenatal exposure to GCs is typically associated with impaired learning skills, but these effects have been shown to vary depending on factors such as dose, sex, or type of cognitive test employed [@szuran_water_1994; @farrell_developmental_2015-learn; @bebus_associative_2016; @haq_prenatal_2021].
-Most past studies have focused on the effects of GC exposure during development on spatial, associative, or reversal learning.
-Few studies have focused on how exposure to GCs during development can affect quantitative abilities, despite the importance of numerical discrimination.
+**Differences in test discrete (OFS) versus overall quantity/size (ANS).
+How to measure both?
+Creating NUMBER inconsistencies with the SAME SIZE (examples).** OFS, can be challenging, since it is necessary to control for other continuous variables such as area, density, or movement (@agrillo2014spontaneous; **Mehlis et al.
+2015**).
+Generally, authors increase one variable at a time without increases in the other to give conflicting information about the correlation between numerical cues and quantitative continuous traits [**REFs**].
+For example, **Look for something where they increase number controlling for size** **HOW??**
 
-Although maternally transmitted GCs are a significant factor in cognitive development, other environmental conditions can also influence cognition, either independently or through interactions with GCs.
-For instance, the early thermal environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019].
-In lizards, incubating eggs at elevated temperatures improves cognitive abilities consistent with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017], but impairs cognition at extreme temperatures [@dayananda_incubation_2017; @abayarathna_effects_2020].
-In addition, when @vila_pouca_quantity_2019 tested the effects of elevated incubation temperature on numerical discrimination in  Port-Jackson sharks (_Heterodontus portusjacksoni_) and found that sharks incubated at elevated temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019].
-However, the effects of early thermal environment on numerical abilities in other taxa remains largely unexplored.
+No other amphibian or non-avian reptile study has set out to
+manipulate these variables to a degree as thoroughly as Stancher et al.42. Future experiments
+involving amphibians and non-avian reptiles that manipulate these variables will be crucial for
+understanding the relationship between the core systems and the information available for
+discrimination
 
-Early-life GCs and thermal conditions could have interactive effects that influence cognitive development.
-Understanding how these factors influence quantitative abilities could reveal important insights into the ecological relevance of numerosity.
-Given the influence of numerical cognition on fitness, we would expect evolutionary pressures to be similar across taxa, and yet, research on numerical discrimination has been traditionally focused on mammals, birds, and fishes [@nieder2005counting; @agrillo2014spontaneous].
-However, recent studies demonstrate that reptiles are also competent in numbers.
+These studies have been traditionally focused on mammals, birds, and fishes despite the influence of numerical cognition on fitness across taxa [@nieder2005counting; @agrillo2014spontaneous].
+However, recent studies demonstrate that other vertebrates like amphibians or non-avian reptiles are also competent in numbers [@uller2003salamanders; @stancher2015numerical; @miletto2017quantitative].
 For example, red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour with higher volumes of food [@soldati2017long], Hermann’s tortoises (_Testudo hermanni_) discriminate between large and small food quantities in a spontaneous choice test [@gazzola2018continuous], and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) rapidly learned a quantity discrimination task [@lin2021superior; @lin2024trained].
 The Italian wall lizard (_Podarcis siculus_) failed to discriminate between different numbers of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity].
-In contrast, Iberian rock lizards (_Iberolacerta cyreni_) preferred larger quantities of crickets when presented five versus ten items [@recio_prey_2021].
-Indeed, the literature shows evidence for basic quantitative abilities in reptiles that may vary with the type of cue and the test employed.
+In contrast, Iberian rock lizards (_Iberolacerta cyreni_) spontaneously preferred larger quantities of crickets when presented five versus ten items [@recio_prey_2021].
+**In skinks,**
 
-Here, we investigated the effect of prenatal corticosterone (CORT) - the main GC in reptiles - and incubation temperature on numerical discrimination in a lizard.
-We manipulated CORT concentration (CORT-treated or a sham control) in _L.
+Lizards were presented with food quantities
+differing in number or size and were able to select the larger quantity of food items when they differed in number but not when items
+differed in size. We show, for the first time, superior spontaneous discrimination of items differing in number over size in a lizard spe-
+cies, which contrasts with previous findings. Our simple method, however, did not include controls for the use of continuous quantities,
+and further tests are required to determine the role of such information during quantity discrimination. [@szabo_spontaneous_2021]
+
+We subjected lizards to a Y-maze experiment presenting both discrete (comparisons involving two
+vs. four, two vs. six, two vs. eight, four vs. six, four vs. eight, and six vs. eight pieces) and equivalent continuous combinations (single
+pieces differing in size) of pumpkin pieces. Our findings revealed that sleepy lizards exhibited spontaneous quantity discrimination, fa-
+voring the larger quantity in three discrete comparisons (eight vs. two, two vs. six, four vs. six) and two continuous comparisons (eight
+vs. two, eight vs. four). However, low accuracy may be attributable to difficulty in visual acuity exacerbated by difficulty discriminating
+quantities at small ratios. [@szabo2024spontaneous]
+
+**These studies demonstrate that reptiles have cognitive abilities related to quantity discrimination that have been overlooked in the past, but that should be explored further.**
+
+Cognitive abilities, however, can be strongly influenced by developmental conditions when the brain is particularly sensitive to environmental inputs [@zhu_prenatal_2004].
+For instance, the temperature at which embryos develop can affect neuron density or learning abilities in different species of reptiles [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @dayananda_incubation_2017; @siviter_incubation_2017; @abayarathna_effects_2020].
+These thermal effects on cognition may arise directly or indirectly through impacts on parental condition [@Crino_2023].
+Temperatures outside the optimal range can trigger stress responses, leading to the release of glucocorticoids (GCs) [@sapolsky_how_2000], which may then be transmitted to the offspring [@Crino_2023].
+Elevated prenatal GCs can, in turn, influence cognition by altering brain function [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
+As such, the interaction between GCs and develpomental temperature must be relevant in shaping reptiles' cognition.
+Most studies on the effects of GC exposure or elevated temperatures during development focused on learning and memory, but few studies have investigated how early-life conditions can affect quantitative abilities despite the importance of numerical discrimination [but see @vila_pouca_quantity_2019].
+Furthermore, the interactive effects of prenatal GCs and temperature on cognitive abilities remain largely unexplored.
+
+Here, we investigated the use of numerical cues in the common garden skink (_Lampropholis guichenoti_) using a spontaneous choice test where lizards were presented with different number of food items while **controlling** for total length.
+In adition, we examined how prenatal corticosterone (CORT) - the main GC in reptiles - and incubation temperature can affect numerical discrimination and decision-making in foraging contexts.
+To do so, we manipulated CORT concentration (CORT-treated or a sham control) in _L.
 guichenoti_ eggs and incubated them under one of two temperature regimes (Cold - 23 ± 3ºC or Hot - 28 ± 3ºC) in a 2x2 factorial design.
-Then, we assessed numerical discrimination abilities of lizards using a spontaneous choice test approach.
-The numerical discrimination task consisted of five tests that differed in the relative number of food items: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
-We based this experiment on observations in nature that the ability to detect a change in stimulus intensity depends on the relative, not absolute, difference between the stimuli ('Weber's law'; @agrillo2014spontaneous).
-Built upon other studies that explored the cognitive effects of prenatal GCs or thermal early environment in reptiles and other taxa, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
+Then, we assessed numerical discrimination abilities on five tests that differed in the relative number of food items: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
+Previous studies in other skinks showed that these species can discriminater between different quantities of food, although the cues employed to do so remain unclear [@szabo_spontaneous_2021; @szabo2024spontaneous].
+We predicted _L.
+guichenoti_ to be able to discriminate between different number of crickets even when controlling by total length.
+Furthermore, we predicted lizards' decisions to be influenced by the early conditions they were subjected to.
+In this regard, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
 We predicted that increased CORT levels would impair numerical competence [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
@@
 We excluded from the analyses all the individuals that did not make any choice in three or more tests.
+All the analyses were performed by PR, who was blind to the treatment of the lizards.
 

### 2025-03-02 ms.qmd
updates

@@
 ---
-title: "Do early environmental experiences impact numerical discrimination abilities in a lizard?"
+title: "Quantity discrimination in the common garden skink and the role of early-life conditions"
 docx: 
@@
 
+```{r, data_length_trials}
+#| label: data_length_trials
+#
+source(here("R", "func.R"))
+#
+data_length <- read.csv(here("./data/measures_crickets.csv")) %>%
+  group_by(test) %>%
+  summarize(
+    mean_length_small = format_dec(mean(measure_small, na.rm = TRUE), 2),
+    mean_length_large = format_dec(mean(measure_big, na.rm = TRUE), 2),
+    mean_difference = format_dec(mean(diff_measure, na.rm = TRUE), 2),
+    low_diff = format_dec(quantile(diff_measure, 0.025, na.rm = TRUE), 2),
+    high_diff = format_dec(quantile(diff_measure, 0.975, na.rm = TRUE), 2)
+  ) %>%
+  ungroup() %>%
+  mutate(
+    `95% CI` = paste0("[", low_diff, " , ", high_diff, "]")
+  ) %>%
+  dplyr::select(test, mean_length_small, mean_length_large, mean_difference, `95% CI`)
+```
+
 Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen^3^, Amelia Y. Peardon ^1^, Daniel W.A. Noble^1^
@@
 ## Abstract
-
-**From social behaviour to navigating complex environments, quantitative abilities can be crucial to fitness.
-However, early life conditions can impact brain development to affect quantitative competence.
-For example, early thermal conditions or glucocorticoid concentrations (GCs) – stress-related hormones transmitted by the parents – can play a prominent role in shaping cognition through their effects on brain physiology.
-Furthermore, temperature can lead to increased levels of GCs that can be transmitted to offspring, emphasizing the need to study the combined effects of these two factors.
-Here, we investigated the effects of elevated prenatal corticosterone (CORT) – the main GC in reptiles – and incubation temperature on numerical discrimination in the common garden skink (_Lampropholis guichenoti_).
-We assessed numerical abilities through a spontaneous choice test with food as a stimulus.
-Employing a repeated measures design, we subjected lizards to five numerical tests each differing in the ratios between the two choices (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4).
-Contrary to our predictions, we found no evidence for treatment effects on lizard behaviour and no use of numerical discrimination during foraging in _L.
+From social behavior to navigating complex environments, quantity discrimination essential for fitness.
+Animals can use one of two systems to discriminate between different quantities, with variation across taxa.
+In reptiles, quantity discrimination remained largely unexplored until recently, and the use of one system over the other can depend on the species or the context in which they are tested.
+Additionally, early life conditions can influence brain development and, in turn, quantity discrimination.
+For example, early thermal conditions or glucocorticoid concentrations (GCs) — stress-related hormones - when transmitted by the mother, can significantly shape cognition through their effects on brain physiology.
+Here, we investigated numerical abilities in the common garden skink (_Lampropholis guichenoti_), as well as the effects of elevated prenatal GCs and incubation temperature on numerical discrimination.
+We assessed numerical abilities using a spontaneous choice test with food as a stimulus.
+Using a repeated measures design, we subjected lizards to five numerical tests (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 prey items), while controlling for continuous variables.
+Contrary to our predictions, we found no evidence of numerical discrimination in _L.
 guichenoti_.
-Our findings build on previous studies suggesting that reptiles feeding on live prey may have difficulty distinguishing between different prey quantities.
-Despite our results, further investigation of numerical abilities in live-prey-feeding reptiles and the impact of early conditions on other taxa is warranted, as these areas remain understudied.**
+Furthermore, prenatal conditions did not influence their performance or decision-making.
+Our findings contribute to the growing body of evidence suggesting that reptiles feeding on live prey may struggle to distinguish between quantities in spontaneous choice tests.
+Despite these results, further investigation into numerical abilities in live-prey-feeding reptiles and the impact of early conditions on other taxa remains an important area of research.
 
 ## Significance statement
-
-**Discriminating between quantities can be crucial for individual fitness, yet the impact of early environmental factors on the development of quantitative competence remains largely unexplored.
-In this study, we investigated, for the first time, the effects of prenatal corticosterone exposure and incubation temperature on numerical discrimination in a reptile.
-Contrary to our predictions, we did not observe any evidence of numerical discrimination and therefore no impact of early developmental environment.
-Our results align with previous studies showing that reptiles consuming live prey struggle to differentiate prey numbers in spontaneous choice tests.
-These results open new avenues for exploring the evolution of numerical abilities in reptiles and identifying factors influencing the development of quantitative skills in this group.
-Further research should explore alternative methods, such as trained tasks, to better understand the cognitive mechanisms underlying quantitative competence in reptiles.**
+Discriminating between quantities can be crucial for individual fitness, yet the ability to distinguish quantities remains poorly understood in reptiles.
+In this study, we investigated numerical discrimination in _Lampropholis guichenoti_, using a spontaneous choice test with food as the stimulus.
+Contrary to our predictions, we found no evidence of numerical discrimination in this species.
+In addition, we explored the impact of early environmental factors, specifically prenatal corticosterone exposure and incubation temperature, on numerical competence.
+However, we did not observe any effect of these early-life conditions on lizard decision-making or performance.
+These results open new avenues for exploring the evolution of numerical abilities in reptiles and identifying factors influencing the development of quantity discrimination in this group.
+Further research should explore alternative methods, such as trained tasks, to better understand the cognitive mechanisms underlying numerical competence in reptiles
 
@@
 The main two approaches experimental approaches in the study of quantity discrimination are spontaneous choice tests and training procedures.
-In spontaneous choice tests, the subjects are presented two groups of biologically relevant stimuli differing in some quantitative trait (number, size, volumen, etc) and the experimenters record if the subjects select or not the most most appropriate quantity (e.g.
+In spontaneous choice tests, the subjects are presented two groups of biologically relevant stimuli differing in some quantitative traits (number, size, volumen, etc) and the experimenters record if the subjects select or not the most most appropriate quantity (e.g.
 the group with more conspecifics or the larger amount of food) (@agrillo2014spontaneous; @cox2016quantity).
@@
 One reason is that spontaneous choice tests assume that one of the options presented is more adequate than the other when this may not always be the case.
-For example, choosing groups of conspecifics with less individuals could be advantageous when food is scarce but disadvantageous when predators are abundant [**REFs first thesis??**].
+For example, choosing groups of conspecifics with less individuals could be advantageous when food is scarce but disadvantageous when predators are abundant [@yee2013costs].
 As such, context and motivation can be important drivers of decision-making that can change the way we measure animals' abilities to discriminate between quantities [@agrillo2014spontaneous].
+
+The other reason is that each method may engage distinct cognitive mechanisms when processing numerical information [@agrillo2014spontaneous].
+In this regard, researchers propose that two core systems underlie numerical cognition: the approximate number system (ANS) and the object file system (OFS) [@hyde2011two].
+The ANS estimates the approximate magnitude of items within a set, with accuracy decreasing as the numerical difference between sets decreases — a principle known as Weber’s law [@hyde2011two].
+In contrast, the OFS allows for the precise tracking of discrete items but is thought to be limited to small quantities, typically up to three or four [@hyde2011two].
+The engagement of these systems depends not only on the numerical magnitude but also on contextual factors such as the availability of continuous variables (e.g., area, density, movement) and the nature of the task [@stancher2015numerical].
+The presence of such continuous cues typically favors ANS-based processing, while in their absence, the OFS is more likely to be employed [@agrillo2014spontaneous].
+
 The other reason is that each method may tap into different cognitive mechanisms [@agrillo2014spontaneous].
-In this sense, **Researchers have suggested that numerousness can be processed by two types of systems: one system for small sets (≤ 4) that works by tracking and encoding objects as individual entities; and an analogous system that operates with large numerosities by estimation processes (for a review, see Cantlon et al.
-(2009); Hyde (2011).**   **HOW CAN EACH BE RELATED TO MOTIVATION OR LEARNING??**
-
-**Differences in test discrete (OFS) versus overall quantity/size (ANS).
-How to measure both?
-Creating NUMBER inconsistencies with the SAME SIZE (examples).** OFS, can be challenging, since it is necessary to control for other continuous variables such as area, density, or movement (@agrillo2014spontaneous; **Mehlis et al.
-2015**).
-Generally, authors increase one variable at a time without increases in the other to give conflicting information about the correlation between numerical cues and quantitative continuous traits [**REFs**].
-For example, **Look for something where they increase number controlling for size** **HOW??**
-
-No other amphibian or non-avian reptile study has set out to
-manipulate these variables to a degree as thoroughly as Stancher et al.42. Future experiments
-involving amphibians and non-avian reptiles that manipulate these variables will be crucial for
-understanding the relationship between the core systems and the information available for
-discrimination
-
-These studies have been traditionally focused on mammals, birds, and fishes despite the influence of numerical cognition on fitness across taxa [@nieder2005counting; @agrillo2014spontaneous].
+In this sense, researchers consider two core systems within the brain that allow for the processing of quantity information: the approximate number system (ANS) and the object file system (OFS) [@hyde2011two].
+ANS encodes the approximate magnitudes of items within sets, with accuracy decreasing as the distance between numbers decreases, obeying the called Weber’s law [@hyde2011two].
+OFS, on the other hand, allows for the precise tracking of discrete items within sets, and it is thought to be limited to small numbers (typically up to three or four) [@hyde2011two].
+However, the use of either system may be dictated, not only by the magnitude of the stimulus, but also by factors such as the availability of continuous information or the context of the task [@stancher2015numerical].
+For example, the presence of continuous variables such as area, density, or movement usually entails the use of the ANS, while the OFS would be predominant in the absence of those continuos cues (@agrillo2014spontaneous).
+This often complicates the study of numerosity, as continuous traits must be carefully controlled to ensure the use of the OFS [@agrillo2014spontaneous].
+In spontaneous choice tests, for instance, animals may rely on cumulative surface area rather than numerical differences when selecting between two sets of food items [@agrillo2014spontaneous].
+To disentangle these cues, researchers often equate the total area of the sets while changing the number of items, forcing individuals to rely on numerical information alone [@stancher2015numerical].
+
+Quantity discrimination studies on both systems have been traditionally focused on mammals, birds, and fishes despite the influence of numerical cognition on fitness across taxa [@mehlis2015quantification; @nieder2018evolution].
 However, recent studies demonstrate that other vertebrates like amphibians or non-avian reptiles are also competent in numbers [@uller2003salamanders; @stancher2015numerical; @miletto2017quantitative].
@@
 In contrast, Iberian rock lizards (_Iberolacerta cyreni_) spontaneously preferred larger quantities of crickets when presented five versus ten items [@recio_prey_2021].
-**In skinks,**
-
-Lizards were presented with food quantities
-differing in number or size and were able to select the larger quantity of food items when they differed in number but not when items
-differed in size. We show, for the first time, superior spontaneous discrimination of items differing in number over size in a lizard spe-
-cies, which contrasts with previous findings. Our simple method, however, did not include controls for the use of continuous quantities,
-and further tests are required to determine the role of such information during quantity discrimination. [@szabo_spontaneous_2021]
-
-We subjected lizards to a Y-maze experiment presenting both discrete (comparisons involving two
-vs. four, two vs. six, two vs. eight, four vs. six, four vs. eight, and six vs. eight pieces) and equivalent continuous combinations (single
-pieces differing in size) of pumpkin pieces. Our findings revealed that sleepy lizards exhibited spontaneous quantity discrimination, fa-
-voring the larger quantity in three discrete comparisons (eight vs. two, two vs. six, four vs. six) and two continuous comparisons (eight
-vs. two, eight vs. four). However, low accuracy may be attributable to difficulty in visual acuity exacerbated by difficulty discriminating
-quantities at small ratios. [@szabo2024spontaneous]
-
-**These studies demonstrate that reptiles have cognitive abilities related to quantity discrimination that have been overlooked in the past, but that should be explored further.**
-
-Cognitive abilities, however, can be strongly influenced by developmental conditions when the brain is particularly sensitive to environmental inputs [@zhu_prenatal_2004].
+In skinks, Shinglebacks (_Tiliqua rugosa_) exhibited spontaneous quantity discrimination using both discrete and continuous cues [@szabo2024spontaneous], while Gidgee skinks (_Egernia stokesii_) discriminated between itemst when they differed in number, but not when size was the main quantitative cue [@szabo_spontaneous_2021].
+These studies demonstrate that reptiles have cognitive abilities related to quantity discrimination that have been overlooked in the past, and that the use of the ANS versus the OFS can differ between species and in different contexts [@szabo_spontaneous_2021].
+However, cognitive performance is not solely dictated by species differences; developmental conditions can also shape an individual's ability to process numerical information [@vila_pouca_quantity_2019].
+
+Cognition can be strongly influenced by developmental conditions because the brain is particularly sensitive to environmental inputs in this life stages [@zhu_prenatal_2004].
 For instance, the temperature at which embryos develop can affect neuron density or learning abilities in different species of reptiles [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @dayananda_incubation_2017; @siviter_incubation_2017; @abayarathna_effects_2020].
@@
 As such, the interaction between GCs and develpomental temperature must be relevant in shaping reptiles' cognition.
-Most studies on the effects of GC exposure or elevated temperatures during development focused on learning and memory, but few studies have investigated how early-life conditions can affect quantitative abilities despite the importance of numerical discrimination [but see @vila_pouca_quantity_2019].
+Most studies on the effects of GC exposure or elevated temperatures during development focused on learning and memory, but few studies have investigated how early-life conditions can affect quantity discrimination despite its importance on animals' lifes [but see @vila_pouca_quantity_2019].
 Furthermore, the interactive effects of prenatal GCs and temperature on cognitive abilities remain largely unexplored.
 
-Here, we investigated the use of numerical cues in the common garden skink (_Lampropholis guichenoti_) using a spontaneous choice test where lizards were presented with different number of food items while **controlling** for total length.
+Here, we investigated the use of numerical cues in the common garden skink (_Lampropholis guichenoti_) using a spontaneous choice test where lizards were presented with different number of food items while controlling for continuous variables.
 In adition, we examined how prenatal corticosterone (CORT) - the main GC in reptiles - and incubation temperature can affect numerical discrimination and decision-making in foraging contexts.
@@
 Then, we assessed numerical discrimination abilities on five tests that differed in the relative number of food items: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
-Previous studies in other skinks showed that these species can discriminater between different quantities of food, although the cues employed to do so remain unclear [@szabo_spontaneous_2021; @szabo2024spontaneous].
+Previous studies in other skinks showed that these species can discriminater between different quantities of food using both the ANS and the OFS [@szabo_spontaneous_2021; @szabo2024spontaneous].
 We predicted _L.
-guichenoti_ to be able to discriminate between different number of crickets even when controlling by total length.
+guichenoti_ to be able to discriminate between different number of crickets even when controlling by other continuous variables (i.e.
+by using the OFS).
 Furthermore, we predicted lizards' decisions to be influenced by the early conditions they were subjected to.
@@
 We predicted that increased CORT levels would impair numerical competence [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
-Consistent with Weber's law, we predicted that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g.
+We predicted that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g.
 3 VS 4), whereas even the lowest ratios (e.g.
@@
 
-*Hatchlings* -- Immediately after hatching, we measured Snout-Vent Length (SVL) and Tail Length (TL) with a ruler to the nearest mm (± 0.1 mm), and mass using a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
+*Hatchlings* -- Immediately after hatching, we measured snout-vent length (SVL) and tail length (TL) with a ruler to the nearest mm (± 0.1 mm), and mass using a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
 Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) provided with non-stick matting and a small water dish.
@@
 We decided to use frozen crickets to avoid movement that would affect lizard behaviour [@cooper2024tell].
-
-Lizards were fed frozen crickets for one month prior to the experiments to habituate them to eating frozen prey.
+Lizards were habituated to eating frozen crickets for one month before the experiments.
+
+The crickets used during the experiment were all the same size, but to control for some continuous traits we changed the orientation of the crickets as illustrated in @fig-Methods D.
+This method was intended to control for length while minimizing increases in convex hull as much as possible (see @stancher2015numerical).
+We tested whether the area occupied was similar in all the test by measuring the total length occupied by all the crickets in both choices for half of the animals (see _Supplementary Material - Control of size in both options_), but we did not explicitly control for convex hull.
+Our results showed that except for the 1 VS 4 and 1 VS 3 tests, the difference in length between both choices was not different from zero.
+As such, discrimination between both choices in the rest of the tests (2 VS 4, 2 VS 3, 3 VS 4 crickets) would be produced by the OFS, not the ANS [@agrillo2014spontaneous].
+In addition, after the experiments, we conducted another set of trials to test whether the orientation of the crickets could affect lizards’ choice, but we found no effect of orientation (see _Supplementary Material - Prey orientation test_).
+
 To habituate lizards to the experimental conditions, frozen crickets were placed on top of one of the platforms and Petri dishes during each feeding for two weeks prior to experimental trials.
@@
 We also counterbalanced the number of individuals per treatment tested in the same type of trial each day.
-
-The crickets used during the experiment were all the same size, but to control for other quantitative traits we changed the orientation of the crickets as illustrated in @fig-Methods D.
-This approach created a conflict of information between choices, as horizontally oriented crickets could occupy the same space as a greater number of vertically oriented crickets.
-This allowed us to determine whether the lizards relied on discrete numerical units rather than non-numerical cues, such as total length, area, or volume [@agrillo2014spontaneous].
-In addition, after the experiments, we conducted another set of trials to test whether the orientation of the crickets could affect lizards’ choice, but we found no effect of orientation (see _Supplementary Material - Prey orientation test_).
 
@@
 We excluded from the analyses all the individuals that did not make any choice in three or more tests.
-All the analyses were performed by PR, who was blind to the treatment of the lizards.
+All the analyses were performed by PR, who was blind to the treatment of the lizards, but not the type of test or the cricket's number as the videos were analysed manually.
 
@@
 #### Foraging context may influence prey number discrimination in reptiles  
-Lack of numerical discrimination in _L.
+Our aim in this study was to investigate whether _L.
+guichenoti_ could discriminate between different number of crickets in a spontaenous choice test relying exclusively on the OFS.
+This question arose from previous research showing that other species within the Scincidae family could use the OFS in spontaneous discrimination tests [@szabo_spontaneous_2021].
+To control for non numerical cues like total length and convex hull, we rotated the position of the crickets in the choice with fewer items, creating a situation where total length was similar in both choices [@agrillo2014spontaneous; @stancher2015numerical; @cooper2024tell].
+This design intended to test whether _L.
+guichenoti_ could rely on numerical information alone when the ANS was not available.
+However, this manipulation could have made the task more challenging for lizards, potentially affecting their performance.
+If _L.
+guichenoti_ primarily relies on the ANS for quantity discrimination, this could explain the absence of a numerical preference.
+Nonetheless, in two of our tests (1 VS 4 and 1 VS 3 crickets) our manipulation was not sufficient to control for total length (see _Supplementary Material - Control of size in both options_).
+Yet, we still did not find quantity discrimination in _L.
+guichenoti_.
+This suggests that the absence of numerical preferences in our study is unlikely to be an artifact of our manipulation, but rather reflects a genuine lack of quantity discrimination in this species, at least in spontaneous choice tests.
+
+In this regard, lack of numerical discrimination in _L.
 guichenoti_ may be due to insectivorous lizards relying on other cues, such as movement or size, for foraging decisions.
@@
 10)[@recio_prey_2021].
-In contrast, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities despite showing high quantitative skills when prey size was the main cue [@miletto2017quantitative] or when tested using training procedures [@miletto2018quantity].
+In contrast, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities despite showing quantity discrimination when prey size was the main cue [@miletto2017quantitative] or when tested using training procedures [@miletto2018quantity].
 Notably, the numbers of food items employed in @miletto2017quantitative were always fewer than four.
@@
 Furthermore, while one individual handles a prey item within the patch, other conspecifics could capture the rest.
-As a result, insectivorous lizards might not have evolved a preference for larger patches of prey, at least with small numbers [see @recio_prey_2021].
-
-Conversely, our findings may be influenced by the experimental design.
-In our study, we rotated the position of the crickets in the choice with fewer items to create a situation where the numerical cues (fewer items) conflicted with other quantitative cues associated with numerosity.
-This approach was intended to control for other quantitative cues that could be used by lizards to make decisions [@cooper2024tell].
-However, this manipulation could have made the task more challenging for lizards, potentially affecting their performance.
-Indeed, it is possible that _L.
-guichenoti_ uses multiple quantitative cues simultaneously during foraging decisions.
-When those cues conflict, the preference for one cue over another might vary between individuals, potentially producing the results observed here.
-Future studies should explore the use of other quantitative cues in _L.
-guichenoti_ and how they interact with numerical information.
-Such studies would enhance our understanding of how _L.
-guichenoti_ integrates different quantitative cues during foraging, providing insights into their decision-making processes and overall quantitative abilities.
+As a result, even if diet might not affect quantity discrimination abilities, insectivorous lizards might not have evolved a preference for larger patches of prey, at least with small numbers [see @recio_prey_2021].
 
@@
 Given the potent effects of early environment on cognition and brain development [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic], it is crucial to explore how early environmental conditions shape numerical abilities across taxa.
-
-In our experimental design, prenatal treatments did not influence numerical discrimination in _L.
-guichenoti_, but we cannot discount the possibility that early environmental factors may affect numerical discrimination under other approaches or contexts.
-Based on previous studies showing the effects of prenatal temperature and GCs on cognitive abilities in reptiles and other groups [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic], we predicted that lower incubation temperatures and exposure to CORT will impair lizards' performance, especially in those tests with higher ratios [@agrillo2014spontaneous; @nieder2018evolution].
-However, we did not find any effect of the experimental treatments on _L.
-guichenoti_'s behaviour when tested in a numerical task.
-Both GCs and thermal environment can exert differential effects on distinct areas of the brain.
-For example, in the eastern three-lined skink (_Bassiana dupeyerri_), elevated incubation temperatures increased cell density in the medial cortex - linked to learning process and memory - but did not impact cell density in other nuclei [@amiel_effects_2017].
-Similarly, impairments in spatial learning in rats [@haq_prenatal_2021] are likely associated with the greater density of GCs receptors in the hippocampus compared to other regions of the brain [@siegel_2006].
-Researchers have proposed that non-verbal number representation is processed by two types of systems: one for small sets (≤ 4) that tracks and encodes objects as individual entities, and another for larger numerosities that operate through estimation processes [@nieder2005counting; @cantlon2009beyond; @hyde2011two].
-While it remains unknown which areas of the reptiles' brain encode numerical information, our results suggest that the brain regions involved in discrete numerical representation (≤ 4 items) might be robust to the effects of prenatal CORT and temperature in _L.
+In our experimental design, prenatal treatments did not influence quantity discrimination in _L.
 guichenoti_.
-
-Alternatively, it is plausible that treatment effects are masked by the overall absence of discrimination between numbers of prey items across most of the tests.
-In this regard, motivation is a key factor in spontaneous choice tests [@agrillo2014spontaneous; @bisazza2014experim].
-Both developmental CORT and incubation temperature can affect metabolic rate (hereafter MR) which could affect motivation in food based learning paradigms [@noble_developmental_2018; @cossin2022effect; but see @crino_corticosterone_2014-learn].
-Although still possible, this explanation is less likely in our study because we controlled for motivation by providing only one cricket per day and fasting lizards for more than 20 hours before the test.
-In addition, we excluded tests where lizards did not interact with any of the options or did not eat the cricket provided after the test.
-Further studies should explore the effects of prenatal conditions on motivation and how it may affect the outcome of cognitive choice tests in reptiles.
+However, since none of the groups showed any kind of quantity discrimination in our tests, we cannot determine whether prenatal conditions affected the lizards' numerical abilities or not; and we cannot discount the possibility that early environmental factors may affect quantity discrimination under other approaches or contexts.
+
+Nonetheless, our results showed that decision-making was not influenced by CORT elevations or incubation temperature, as latency to make a choice did not vary across prenatal conditions.
+This suggests that _L.
+guichenoti_ may be robust to these early-life factors, at least in the context of foraging decisions.
+Previous studies have reported that pine snakes (_Pituophis melanoleucus_) incubated at low temperatures were slower and less proficient in capturing prey [@burger_effects_1991], but cold incubated bearded dragons (_Pogona vitticeps_) were faster at approaching a novel objects at least at a young age [@siviter_impact_2017].
+Maternal transmitted GCs are also known to reduce exploration and neophilia in different taxa [@koolhaas1999coping; @hope_incubation_2018].
+Additionally, both prenatal CORT and incubation temperature can influence metabolic rate (MR), which in turn could affect motivation in food-based learning paradigms [@noble_developmental_2018; @cossin2022effect; but see @trnik_persistent_2011; @crino_corticosterone_2014-learn].
+In our design, we habituated lizards to the test conditions for two weeks before the experiment to minimize neophilia.
+And to control for motivation, we standardized hunger levels by providing only one cricket per day and fasting lizards for over 20 hours before testing.
+Additionally, we excluded trials where lizards failed to interact with either option or refused to eat the cricket provided after testing.
+These measures likely minimized variation in decision latency due to differences in exploration, neophilia, or hunger levels, which could explain why we did not detect an effect of prenatal conditions.
+Interestingly, we found that males and older lizards made decisions faster, although these factors were not central to our main hypotheses.
+However, it is possible that sex and aging interact with early-life conditions in shaping behavioral responses.
+Future research should investigate potential interactions between developmental environment, sex, and age to better understand how these factors influence decision-making in reptiles.
 
 #### Conclusion
-Our study is, to the best of our knowledge, the first to explore the impact of prenatal conditions on numerical abilities in reptiles.
+Our study was focused on investigating if _L.
+guichenoti_ could use the OFS to disriminate between different number of food items.
+Aditionally, we were interested in exploring, for the first time, the impact of prenatal conditions on numerical abilities in reptiles.
 Our findings suggest that _L.
-guichenoti_ does not rely on numerical information for foraging decisions and that their performance does not vary due to prenatal treatments.
-These results indicate that early environmental conditions, such as incubation temperature and CORT exposure, may not affect numerosity in this species.
+guichenoti_ does not rely on numerical information for foraging decisions and that their performance and decision making does not vary due to prenatal treatments.
+These results align with other studies showing that insectivorous reptiles fail in quantity discrimination tasks in spontaneous choice tests.
 However, it is critical to note that the methodologies employed in this study may not have fully captured the nuances of numerical cognition in _L.
@@
 Therefore, we recommend that future studies employ a broader array of tests and paradigms, including training-based tasks, larger sets of stimuli, and other sensory modalities, to more comprehensively assess numerical competence in this and related species.
-Such work will be fundamental in unravelling early environmental factors' effects on reptiles' cognitive abilities and advancing our understanding of numerical discrimination across taxa.
+Furthermore, it would be relevant to study the effect of early environment on quantity discrimination to advance in our understanding of nuerosity across taxa.
 
@@
 
+#### Control of size in both options
+On the videos of half of the lizards, we estimated the length of both options using the ruler tool on Photoshop 2024.
+For each chioce in each test, we first calibrated the masure of one centimeter using the length of the bottom of the platform (see Fig below).
+After calibrating, we measured the length of the crickets at their maximum length, trying to keep the line parallel to the bottom of the platform (see below).
+The results are shown in Table S4.
+
+```{r, fig-S1}
+#| label: fig-S1
+#| fig.cap: "Calibration of the ruler tool in Photoshop 2024. The bottom of the platform was used to calibrate the ruler tool to measure the length of the crickets."
+#
+knitr::include_graphics("./Others/Calibration.png")
+```
+
+```{r, fig-S2}
+#| label: fig-S2
+#| fig.cap: "Measurement of the length of the crickets. The ruler tool was used to measure the length of the crickets at their maximum length."
+#
+knitr::include_graphics("./Others/Measure.png")
+```
+
+_Table S4.- Mean length (cm) of the crickets in each of the numerical tests. The difference in length between the two options is also shown, with the 95% CI._
+```{r, tbl-size}
+#| label: tbl-size
+#| tbl-cap: "Mean length (cm) of the crickets in each of the numerical tests. The difference in length between the two options is also shown, with the 95% CI."
+table_size <- flextable(data_length) %>%
+  flextable::compose(j = 1, value = as_paragraph("Test"), part = "header") %>%
+  flextable::compose(j = 2, value = as_paragraph("Mean length small (cm)"), part = "header") %>%
+  flextable::compose(j = 3, value = as_paragraph("Mean length large (cm)"), part = "header") %>%
+  flextable::compose(j = 4, value = as_paragraph("Mean differences in length"), part = "header") %>%
+  flextable::compose(j = 5, value = as_paragraph("95% CI differences in length"), part = "header") %>%
+  set_table_properties( width = 0.9, layout = "autofit") %>%
+  align(align = "center", part = "all")
+#
+table_size
+```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
 #### Prey orientation test
@@
 
-_Table S4.- Number of individuals per treatment that chose the cricket oriented horizontally (Horizontal) or vertically (Vertical) in our control tests.
+_Table S5.- Number of individuals per treatment that chose the cricket oriented horizontally (Horizontal) or vertically (Vertical) in our control tests.
 p-value indicates the result of the binomial test comparing the number of choices between both choices._

### 2025-03-02 .vdoc.r

@@ -1,747 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: packages
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats, MASS, nortest, fitdistrplus, ggh4x, PupillometryR, cowplot, png, grid, remotes, ggthemes, bayestestR, HDInterval)
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig.cap: "Experimental design. Panel (**A**) shows the early environment manipulation procedures. (**B**) illustrates the arena where the tests were performed, while panel (**C**) indicates the measurements of the platform used for the experiments. Finally, panel (**D**) displays the types of numerical tests used and the orientation of the crickets in each test."
-
-knitr::include_graphics("./Others/NUMDIS_FIG_1.png")
-
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: data_process
-# Load the data and process it to get the main dfs
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: models
-# Fitting the model and extraction of posteriors 
-refit = FALSE
-if(refit){
-  # Get the formula and family of variables
-  lat <- bf(log(latency) ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = gaussian())
-  choice <- bf(choice ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = bernoulli(link = "logit"))
-  int <- bf(compared_interest ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch),
-    family = gaussian())
-  # Fit the model
-  model <- brm(lat + choice + int,
[…]

### 2025-03-16 ms.qmd
updates

@@
 ## Abstract
-From social behavior to navigating complex environments, quantity discrimination essential for fitness.
-Animals can use one of two systems to discriminate between different quantities, with variation across taxa.
+From social behaviour to navigating complex environments, quantitative abilities is essential for fitness.
+Animals use different cognitive systems to distinguish between quantities.
+The Approximate Number System (ANS) estimates, while the Object File System (OFS) tracks discrete items.
 In reptiles, quantity discrimination remained largely unexplored until recently, and the use of one system over the other can depend on the species or the context in which they are tested.
-Additionally, early life conditions can influence brain development and, in turn, quantity discrimination.
+Additionally, early life conditions can influence brain development and, in turn, quantity discrimination and decision-making.
 For example, early thermal conditions or glucocorticoid concentrations (GCs) — stress-related hormones - when transmitted by the mother, can significantly shape cognition through their effects on brain physiology.
-Here, we investigated numerical abilities in the common garden skink (_Lampropholis guichenoti_), as well as the effects of elevated prenatal GCs and incubation temperature on numerical discrimination.
+Here, we investigated numerical abilities in the common garden skink (_Lampropholis guichenoti_), as well as the effects of elevated prenatal GCs and incubation temperature on numerical discrimination and decision-making.
 We assessed numerical abilities using a spontaneous choice test with food as a stimulus.
-Using a repeated measures design, we subjected lizards to five numerical tests (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 prey items), while controlling for continuous variables.
-Contrary to our predictions, we found no evidence of numerical discrimination in _L.
-guichenoti_.
+We subjected lizards to five numerical tests each differing in the ratios between the two choices (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4) while controlling for total length and area.
+Contrary to our predictions, we found no evidence of numerical discrimination in L.
+guichenoti.
 Furthermore, prenatal conditions did not influence their performance or decision-making.
-Our findings contribute to the growing body of evidence suggesting that reptiles feeding on live prey may struggle to distinguish between quantities in spontaneous choice tests.
-Despite these results, further investigation into numerical abilities in live-prey-feeding reptiles and the impact of early conditions on other taxa remains an important area of research.
+Despite these results, further investigation of numerical abilities in live-prey-feeding reptiles and the impact of early conditions on other taxa remains an important area of research.
 
@@
 
-The main two approaches experimental approaches in the study of quantity discrimination are spontaneous choice tests and training procedures.
-In spontaneous choice tests, the subjects are presented two groups of biologically relevant stimuli differing in some quantitative traits (number, size, volumen, etc) and the experimenters record if the subjects select or not the most most appropriate quantity (e.g.
-the group with more conspecifics or the larger amount of food) (@agrillo2014spontaneous; @cox2016quantity).
-In training procedures, animals must learn a numerical rule to receive a reward [@agrillo2014spontaneous; @miletto2018quantity].
-Both methods are believed to be complementary to each other, and both can provide different results even in the same species [see @miletto2017quantitative and @miletto2018quantity].
-One reason is that spontaneous choice tests assume that one of the options presented is more adequate than the other when this may not always be the case.
+The main two experimental approaches in the study of quantity discrimination are spontaneous choice tests and training procedures.
+In spontaneous choice tests, subjects are presented with two groups of biologically relevant stimuli differing quantitatively (number, size, volumen, etc) and ascertain if animals can distinguish them [@agrillo2014spontaneous; @cox2016quantity].
+In training procedures, animals must instead learn a numerical rule to receive a reward [@agrillo2014spontaneous; @miletto2018quantity].
+Both methods are believed to be complementary to each other, and both can provide different results even in the same species [see @miletto2017quantitative; @miletto2018quantity].
+
+Numerous reasons exist for why methods can lead to different results.
+First, spontaneous choice tests assume that one of the options presented is more adequate than the other (e.g.
+the group with more conspecifics or the larger amount of food) when this may not always be the case.
 For example, choosing groups of conspecifics with less individuals could be advantageous when food is scarce but disadvantageous when predators are abundant [@yee2013costs].
-As such, context and motivation can be important drivers of decision-making that can change the way we measure animals' abilities to discriminate between quantities [@agrillo2014spontaneous].
-
-The other reason is that each method may engage distinct cognitive mechanisms when processing numerical information [@agrillo2014spontaneous].
+As such, context and motivation can be important drivers of decision-making that can change the way animals discriminate between quantities [@agrillo2014spontaneous].
+Second, each method may engage distinct cognitive mechanisms when processing numerical information [@agrillo2014spontaneous].
+Different procedures may shape how animals discriminate between quantities, making them more responsive towards certain quantity traits over others.
+Some tasks may encourage a focus on approximate magnitudes, while others may promote attention to individual items [@agrillo2014spontaneous].
+
 In this regard, researchers propose that two core systems underlie numerical cognition: the approximate number system (ANS) and the object file system (OFS) [@hyde2011two].
@@
 The presence of such continuous cues typically favors ANS-based processing, while in their absence, the OFS is more likely to be employed [@agrillo2014spontaneous].
-
-The other reason is that each method may tap into different cognitive mechanisms [@agrillo2014spontaneous].
-In this sense, researchers consider two core systems within the brain that allow for the processing of quantity information: the approximate number system (ANS) and the object file system (OFS) [@hyde2011two].
-ANS encodes the approximate magnitudes of items within sets, with accuracy decreasing as the distance between numbers decreases, obeying the called Weber’s law [@hyde2011two].
-OFS, on the other hand, allows for the precise tracking of discrete items within sets, and it is thought to be limited to small numbers (typically up to three or four) [@hyde2011two].
-However, the use of either system may be dictated, not only by the magnitude of the stimulus, but also by factors such as the availability of continuous information or the context of the task [@stancher2015numerical].
-For example, the presence of continuous variables such as area, density, or movement usually entails the use of the ANS, while the OFS would be predominant in the absence of those continuos cues (@agrillo2014spontaneous).
 This often complicates the study of numerosity, as continuous traits must be carefully controlled to ensure the use of the OFS [@agrillo2014spontaneous].
@@
 
-Quantity discrimination studies on both systems have been traditionally focused on mammals, birds, and fishes despite the influence of numerical cognition on fitness across taxa [@mehlis2015quantification; @nieder2018evolution].
+Quantity discrimination studies have been traditionally focused on mammals, birds, and fishes despite the influence of numerical cognition on fitness across taxa [@mehlis2015quantification; @nieder2018evolution].
 However, recent studies demonstrate that other vertebrates like amphibians or non-avian reptiles are also competent in numbers [@uller2003salamanders; @stancher2015numerical; @miletto2017quantitative].
@@
 In contrast, Iberian rock lizards (_Iberolacerta cyreni_) spontaneously preferred larger quantities of crickets when presented five versus ten items [@recio_prey_2021].
-In skinks, Shinglebacks (_Tiliqua rugosa_) exhibited spontaneous quantity discrimination using both discrete and continuous cues [@szabo2024spontaneous], while Gidgee skinks (_Egernia stokesii_) discriminated between itemst when they differed in number, but not when size was the main quantitative cue [@szabo_spontaneous_2021].
+In skinks, Shinglebacks (_Tiliqua rugosa_) exhibited spontaneous quantity discrimination using both discrete and continuous cues [@szabo2024spontaneous], while Gidgee skinks (_Egernia stokesii_) discriminated between items when they differed in number, but not when size was the main quantitative cue [@szabo_spontaneous_2021].
 These studies demonstrate that reptiles have cognitive abilities related to quantity discrimination that have been overlooked in the past, and that the use of the ANS versus the OFS can differ between species and in different contexts [@szabo_spontaneous_2021].
@@
 
-Cognition can be strongly influenced by developmental conditions because the brain is particularly sensitive to environmental inputs in this life stages [@zhu_prenatal_2004].
+Cognition can be strongly influenced by developmental conditions because the brain is particularly sensitive to environmental inputs in early life stages [@zhu_prenatal_2004].
 For instance, the temperature at which embryos develop can affect neuron density or learning abilities in different species of reptiles [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @dayananda_incubation_2017; @siviter_incubation_2017; @abayarathna_effects_2020].
-These thermal effects on cognition may arise directly or indirectly through impacts on parental condition [@Crino_2023].
-Temperatures outside the optimal range can trigger stress responses, leading to the release of glucocorticoids (GCs) [@sapolsky_how_2000], which may then be transmitted to the offspring [@Crino_2023].
+Thermal effects on cognition may arise directly or indirectly through parental effects [@Crino_2023].
+For example, temperatures outside the optimal range can trigger stress responses, leading to the release of glucocorticoids (GCs) [@sapolsky_how_2000], which may then be transmitted to offspring [@Crino_2023].
 Elevated prenatal GCs can, in turn, influence cognition by altering brain function [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
-As such, the interaction between GCs and develpomental temperature must be relevant in shaping reptiles' cognition.
-Most studies on the effects of GC exposure or elevated temperatures during development focused on learning and memory, but few studies have investigated how early-life conditions can affect quantity discrimination despite its importance on animals' lifes [but see @vila_pouca_quantity_2019].
+As such, the interaction between GCs and developmental temperature may be relevant in shaping reptile cognition.
+Most studies on the effects of GC exposure or elevated temperatures during development focus on learning and memory, but few studies have investigated how early-life conditions can affect quantity discrimination despite its importance on animals' lives [but see @vila_pouca_quantity_2019].
 Furthermore, the interactive effects of prenatal GCs and temperature on cognitive abilities remain largely unexplored.
@@
 Then, we assessed numerical discrimination abilities on five tests that differed in the relative number of food items: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
-Previous studies in other skinks showed that these species can discriminater between different quantities of food using both the ANS and the OFS [@szabo_spontaneous_2021; @szabo2024spontaneous].
+Previous studies in other skinks showed that these species can discriminate between different quantities of food using both the ANS and the OFS [@szabo_spontaneous_2021; @szabo2024spontaneous].
 We predicted _L.
@@
 Furthermore, we predicted lizards' decisions to be influenced by the early conditions they were subjected to.
-In this regard, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
+In this regard, we hypothesised that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
 We predicted that increased CORT levels would impair numerical competence [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
@@
 The crickets used during the experiment were all the same size, but to control for some continuous traits we changed the orientation of the crickets as illustrated in @fig-Methods D.
-This method was intended to control for length while minimizing increases in convex hull as much as possible (see @stancher2015numerical).
-We tested whether the area occupied was similar in all the test by measuring the total length occupied by all the crickets in both choices for half of the animals (see _Supplementary Material - Control of size in both options_), but we did not explicitly control for convex hull.
+This method was intended to control for length and total area (see @stancher2015numerical).
+We tested if the area occupied by the crickets was similar in both choices (see Supplementary Material - Control of size in both options).
 Our results showed that except for the 1 VS 4 and 1 VS 3 tests, the difference in length between both choices was not different from zero.
@@
 Motivation plays a key role in animals' performance in spontaneous choice tests [@agrillo2014spontaneous; @bisazza2014experim].
-To ensure that lizards remained motivated to forage, they were provided with only one cricket during both acclimatization and the numerical task (see above).
-During acclimatization, lizards were fed at the same time each day as the experiments were conducted.
-In the numerical task, lizards received a cricket immediately after the test and then fasted for more than 20 hours, which we expected would help standardize fasting levels across individuals.
+To ensure that lizards remained motivated to forage, they were provided with only one cricket during both acclimatisation and the numerical task (see above).
+During acclimatisation, lizards were fed at the same time each day as the experiments were conducted.
+In the numerical task, lizards received a cricket immediately after the test and then fasted for more than 20 hours, which we expected would help standardise fasting levels across individuals.
 To further control for motivation, we excluded from our analyses any trials in which lizards did not interact with any options or did not eat the cricket provided after the test (n = 0 in both cases).
@@
 The different colours indicate the different treatments.
-Points and bars represent the median and 95% Higest Density Intervals (95% HPDI) of the estimates, respectively.
+Points and bars represent the median and 95% Highest Density Intervals (95% HPDI) of the estimates, respectively.
 Vertical dashed lines in Choice and Interest graphs values 0.5 and 0, respectively.
@@
 The table shows the contrasts for each predictor (Temperature = [medianHot - medianCold]; Hormone = [medianControl - medianCORT]; and their Interaction = [(medianHot-Control - medianHot-CORT) - (medianCold-Control - medianCold-CORT)]).
-95% Higest Density Intervals (95% HPDI) test the hypothesis that contrasts are different from zero._
+95% Highest Density Intervals (95% HPDI) test the hypothesis that contrasts are different from zero._
 ```{r, tbl-data}
@@
 The table shows the contrasts for each predictor (Temperature = [medianHot - medianCold]; Hormone = [medianControl - medianCORT]; and their Interaction = [(medianHot-Control - medianHot-CORT) - (medianCold-Control - medianCold-CORT)]).
-95% Higest Density Intervals (95% HPDI) test the hypothesis that contrasts are different from zero."
+95% Highest Density Intervals (95% HPDI) test the hypothesis that contrasts are different from zero."
 source(here("R", "func.R"))
@@
 Overall, these findings suggest that _L.
-guichenoti_ did not use numerical or other potential quantitative cues for foraging decisions.
-
-#### Foraging context may influence prey number discrimination in reptiles  
-Our aim in this study was to investigate whether _L.
-guichenoti_ could discriminate between different number of crickets in a spontaenous choice test relying exclusively on the OFS.
-This question arose from previous research showing that other species within the Scincidae family could use the OFS in spontaneous discrimination tests [@szabo_spontaneous_2021].
-To control for non numerical cues like total length and convex hull, we rotated the position of the crickets in the choice with fewer items, creating a situation where total length was similar in both choices [@agrillo2014spontaneous; @stancher2015numerical; @cooper2024tell].
-This design intended to test whether _L.
-guichenoti_ could rely on numerical information alone when the ANS was not available.
-However, this manipulation could have made the task more challenging for lizards, potentially affecting their performance.
-If _L.
-guichenoti_ primarily relies on the ANS for quantity discrimination, this could explain the absence of a numerical preference.
-Nonetheless, in two of our tests (1 VS 4 and 1 VS 3 crickets) our manipulation was not sufficient to control for total length (see _Supplementary Material - Control of size in both options_).
-Yet, we still did not find quantity discrimination in _L.
-guichenoti_.
-This suggests that the absence of numerical preferences in our study is unlikely to be an artifact of our manipulation, but rather reflects a genuine lack of quantity discrimination in this species, at least in spontaneous choice tests.
-
-In this regard, lack of numerical discrimination in _L.
-guichenoti_ may be due to insectivorous lizards relying on other cues, such as movement or size, for foraging decisions.
+guichenoti_ did not use numerical or other potential quantitative cues for foraging decisions and that any decision making was not affected by early environment.
+
+#### Foraging context may influence spontaneous quantity discrimination in reptiles  
+Lack of numerical discrimination in spontaneous tests may be due to _L.
+guichenoti_ being insectivorous.
 In studies where reptiles discriminated between different amounts of food, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
-10)[@recio_prey_2021].
+10) [@recio_prey_2021].
 In contrast, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities despite showing quantity discrimination when prey size was the main cue [@miletto2017quantitative] or when tested using training procedures [@miletto2018quantity].
@@
 Spontaneous choice tests assume that animals should select the most appropriate quantity, which may not necessarily be the largest one [@agrillo2014spontaneous].
-This occurrs if the costs of selecting the larger option outweigh the benefits.
-For example, when presented two groups with different numbers of conspecifics or if prey handling costs are related to larger quantitative traits like size [@agrillo2014spontaneous].
-It could be that the perceived handling costs of larger numbers of prey affected lizard response in our study.
-Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food.
-Additionally we would have expected body size to influence lizards choices, as larger animals may be more adept at handling a greater number of prey items.
-@miletto2017quantitative did not report any effect of body size, and age - a proxy for size - did not affect _L.
-guichenoti_'s _Choice_ or _Interest_ in our study.
-The absence of preference in _P.
-siculus_ and _L.
-guichenoti_ suggests that the benefits of selecting a larger number of prey are not different from selecting fewer.
-For instance, when considering hunting groups of 1 VS 4 or 2 VS 4 crickets, the probability of success might be similar.
-Furthermore, while one individual handles a prey item within the patch, other conspecifics could capture the rest.
+This occurs if the costs of selecting the larger option outweigh the benefits.
+If capturing larger prey demands more time or energy, or if success rates remain similar regardless of prey quantity, animals may not exhibit a preference for larger quantities [@agrillo2014spontaneous].
 As a result, even if diet might not affect quantity discrimination abilities, insectivorous lizards might not have evolved a preference for larger patches of prey, at least with small numbers [see @recio_prey_2021].
 
+Previous research showed that other species within the Scincidae family use exclusively the OFS in spontaneous discrimination tests [@szabo_spontaneous_2021].
+Here, we aimed to investigate whether _L.
+guichenoti_ could discriminate between different number of crickets in a spontaneous choice test relying exclusively on the OFS.
+To ensure that lizards decision relied on tracking discret items rather than using non-numerical cues, we controlled for total length and total area — two variables that could influence ANS processing.
+We did this by rotating the position of the crickets in the smaller-choice option, making the total length and total area comparable in both choices [@agrillo2014spontaneous; @stancher2015numerical; @cooper2024tell].
+Absence of discrimination can therefore be explained by _L.
+guichenoti_ relying on the ANS for quantity discrimination.
+Nonetheless, in two of our tests (1 VS 4 and 1 VS 3 crickets) cricket rotation was not sufficient to control for total length (see _Supplementary Material - Control of size in both options_).
+Yet, we still did not find quantity discrimination in _L.
+guichenoti_ which suggests that neither the OFS nor the ANS were used.
+Instead, our results likely reflect a genuine lack of quantity discrimination in this species, at least in spontaneous choice tests.
+
 #### Foraging behaviour is robust to early environmental experiences 
-Although the effects of early environment on cognition is widely studied, little attention has been paid to its impacts on numerical discrimination.
-To the best of our knowledge, there is only one study examining how developmental conditions influence numerical abilities [@vila_pouca_quantity_2019].
-Given the potent effects of early environment on cognition and brain development [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic], it is crucial to explore how early environmental conditions shape numerical abilities across taxa.
-In our experimental design, prenatal treatments did not influence quantity discrimination in _L.
-guichenoti_.
-However, since none of the groups showed any kind of quantity discrimination in our tests, we cannot determine whether prenatal conditions affected the lizards' numerical abilities or not; and we cannot discount the possibility that early environmental factors may affect quantity discrimination under other approaches or contexts.
-
-Nonetheless, our results showed that decision-making was not influenced by CORT elevations or incubation temperature, as latency to make a choice did not vary across prenatal conditions.
+Although the effects of early environment on cognition are widely studied, little attention has been paid to its impacts on numerical discrimination and decision-making.
+We found that decision-making was not influenced by CORT elevations or incubation temperature, as latency to make a choice did not vary across prenatal conditions.
 This suggests that _L.
 guichenoti_ may be robust to these early-life factors, at least in the context of foraging decisions.
-Previous studies have reported that pine snakes (_Pituophis melanoleucus_) incubated at low temperatures were slower and less proficient in capturing prey [@burger_effects_1991], but cold incubated bearded dragons (_Pogona vitticeps_) were faster at approaching a novel objects at least at a young age [@siviter_impact_2017].
+Previous studies have reported that pine snakes (_Pituophis melanoleucus_) incubated at low temperatures were slower and less proficient in capturing prey [@burger_effects_1991], but cold incubated bearded dragons (_Pogona vitticeps_) were faster at approaching a novel object at least at a young age [@siviter_impact_2017].
 Maternal transmitted GCs are also known to reduce exploration and neophilia in different taxa [@koolhaas1999coping; @hope_incubation_2018].
 Additionally, both prenatal CORT and incubation temperature can influence metabolic rate (MR), which in turn could affect motivation in food-based learning paradigms [@noble_developmental_2018; @cossin2022effect; but see @trnik_persistent_2011; @crino_corticosterone_2014-learn].
-In our design, we habituated lizards to the test conditions for two weeks before the experiment to minimize neophilia.
-And to control for motivation, we standardized hunger levels by providing only one cricket per day and fasting lizards for over 20 hours before testing.
+In our design, we habituated lizards to the test conditions for two weeks before the experiment to minimise neophilia.
+And to control for motivation, we standardised hunger levels by providing only one cricket per day and fasting lizards for over 20 hours before testing.
 Additionally, we excluded trials where lizards failed to interact with either option or refused to eat the cricket provided after testing.
-These measures likely minimized variation in decision latency due to differences in exploration, neophilia, or hunger levels, which could explain why we did not detect an effect of prenatal conditions.
+These measures likely minimised variation in decision latency due to differences in exploration, neophilia, or hunger levels, which could explain why we did not detect an effect of prenatal conditions.
 Interestingly, we found that males and older lizards made decisions faster, although these factors were not central to our main hypotheses.
-However, it is possible that sex and aging interact with early-life conditions in shaping behavioral responses.
+However, it is possible that sex and age interact with early-life conditions in shaping behavioral responses.
 Future research should investigate potential interactions between developmental environment, sex, and age to better understand how these factors influence decision-making in reptiles.
+
+Prenatal treatments did not influence quantity discrimination in _L.
+guichenoti_ in our experiment.
+However, since none of the groups showed any kind of quantity discrimination in our tests, we cannot determine whether prenatal conditions affected the lizards' numerical abilities or not; and we cannot discount the possibility that early environmental factors may affect quantity discrimination under other approaches or contexts.
+Nonetheless, to the best of our knowledge there is only one study examining how developmental conditions influence numerical abilities [@vila_pouca_quantity_2019].
+Given the potent effects of early environment on cognition and brain development [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic], it is crucial to explore how early environmental conditions shape numerical abilities across taxa.
 
@@
 Our study was focused on investigating if _L.
-guichenoti_ could use the OFS to disriminate between different number of food items.
-Aditionally, we were interested in exploring, for the first time, the impact of prenatal conditions on numerical abilities in reptiles.
+guichenoti_ could use the OFS to discriminate between different number of food items.
+Additionally, we were interested in exploring the impact of prenatal conditions on numerical abilities and decision-making in reptiles.
 Our findings suggest that _L.
-guichenoti_ does not rely on numerical information for foraging decisions and that their performance and decision making does not vary due to prenatal treatments.
-These results align with other studies showing that insectivorous reptiles fail in quantity discrimination tasks in spontaneous choice tests.
+guichenoti_ does not rely on numerical information for foraging decisions and that their performance and decision-making does not vary due to prenatal temperature or CORT exposure.
+Our findings suggest that decision-making can be robust to early environmental conditions, at least in the context of foraging.
+
+The absence of numerical discrimination in _L.
+guichenoti_ is consistent with previous studies showing that insectivorous reptiles fail in spontaneous choice tests.
 However, it is critical to note that the methodologies employed in this study may not have fully captured the nuances of numerical cognition in _L.
 guichenoti_.
-Therefore, we recommend that future studies employ a broader array of tests and paradigms, including training-based tasks, larger sets of stimuli, and other sensory modalities, to more comprehensively assess numerical competence in this and related species.
-Furthermore, it would be relevant to study the effect of early environment on quantity discrimination to advance in our understanding of nuerosity across taxa.
+Therefore, we recommend that future studies employ a broader array of tests and paradigms to more comprehensively assess numerical competence in this and related species.
+Furthermore, it would be relevant to study the effect of early environment on quantity discrimination to advance our understanding of numerosity across taxa.
 
@@
 
-# Suplementary Material
+# Supplementary Material
 
@@
 
-_Table S1.- Median and 95% Higest Density Intervals (95% HPDI) of the estimated probability of choosing first the higher amount (Choice) and the estimated interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
+_Table S1.- Median and 95% Highest Density Intervals (95% HPDI) of the estimated probability of choosing first the higher amount (Choice) and the estimated interest for the higher amount of food (Interest) per treatment group for each of the numerical tests performed.
 95% HPDI test the hypothesis that Choice = 0.5, and Interest = 0, which would indicate a preference towards one of choices._
@@
 For the variable Latency and Interest, we show the median and the 95% CI.
-For the varaible Choice, we show the proportion of individuals that chose the higher number of crickets first._
+For the variable Choice, we show the proportion of individuals that chose the higher number of crickets first._
 
@@
 On the videos of half of the lizards, we estimated the length of both options using the ruler tool on Photoshop 2024.
-For each chioce in each test, we first calibrated the masure of one centimeter using the length of the bottom of the platform (see Fig below).
+For each choice in each test, we first calibrated the masure of one centimeter using the length of the bottom of the platform (see Fig below).
 After calibrating, we measured the length of the crickets at their maximum length, trying to keep the line parallel to the bottom of the platform (see below).

### 2025-03-17 ms.qmd
updates

@@
 ^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
-^2^ Flinder's University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
+^2^ College of Science and Engineering, Flinders University, Bedford Park, SA 5042, Australia 
 ^3^ University of Wollongong, Wollongong, NSW 2500, Australia  
@@
 ## Keywords
-Numerical discrimination, Corticosterone, Temperature, Reptiles, Early life stress, Cognitive abilities
+Cognitive abilities, Corticosterone, Early life stress, Numerical discrimination, Reptiles, Temperature
 
 ## Abstract
-From social behaviour to navigating complex environments, quantitative abilities is essential for fitness.
+Quantity discrimination affects a range of behaviours that are essential for fitness including social interactions, navigation, and foraging.
 Animals use different cognitive systems to distinguish between quantities.
@@
 In reptiles, quantity discrimination remained largely unexplored until recently, and the use of one system over the other can depend on the species or the context in which they are tested.
-Additionally, early life conditions can influence brain development and, in turn, quantity discrimination and decision-making.
-For example, early thermal conditions or glucocorticoid concentrations (GCs) — stress-related hormones - when transmitted by the mother, can significantly shape cognition through their effects on brain physiology.
-Here, we investigated numerical abilities in the common garden skink (_Lampropholis guichenoti_), as well as the effects of elevated prenatal GCs and incubation temperature on numerical discrimination and decision-making.
-We assessed numerical abilities using a spontaneous choice test with food as a stimulus.
-We subjected lizards to five numerical tests each differing in the ratios between the two choices (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4) while controlling for total length and area.
-Contrary to our predictions, we found no evidence of numerical discrimination in L.
-guichenoti.
+Additionally, early life conditions can influence brain development and cognition, with potential effects on quantity discrimination and decision-making.
+Here, we tested the ability of the common garden skink (_Lampropholis guichenoti_) to discriminate between quantities and the effects of elevated prenatal GCs and incubation temperature on quantity discrimination and decision-making.
+We assessed quantity discrimination using a spontaneous choice test with food as a stimulus.
+We subjected lizards to five two-choice tests each differing in the number of items (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4).
+Contrary to our predictions, we found no evidence of quantity discrimination in _L.
+guichenoti_.
 Furthermore, prenatal conditions did not influence their performance or decision-making.
-Despite these results, further investigation of numerical abilities in live-prey-feeding reptiles and the impact of early conditions on other taxa remains an important area of research.
+Despite these results, further investigation of quantity discrimination in live-prey-feeding reptiles and the impact of early conditions on other taxa remains an important area of research.
 
 ## Significance statement
-Discriminating between quantities can be crucial for individual fitness, yet the ability to distinguish quantities remains poorly understood in reptiles.
-In this study, we investigated numerical discrimination in _Lampropholis guichenoti_, using a spontaneous choice test with food as the stimulus.
-Contrary to our predictions, we found no evidence of numerical discrimination in this species.
-In addition, we explored the impact of early environmental factors, specifically prenatal corticosterone exposure and incubation temperature, on numerical competence.
-However, we did not observe any effect of these early-life conditions on lizard decision-making or performance.
-These results open new avenues for exploring the evolution of numerical abilities in reptiles and identifying factors influencing the development of quantity discrimination in this group.
-Further research should explore alternative methods, such as trained tasks, to better understand the cognitive mechanisms underlying numerical competence in reptiles
+Discriminating between quantities can be crucial for individual fitness, yet it remains poorly understood in reptiles.
+In this study, we investigated quantity discrimination in _Lampropholis guichenoti_ using a spontaneous choice test with food as the stimulus.
+Contrary to our predictions, we found no evidence of quantity discrimination in this species.
+In addition, we explored the impact of early environmental factors, specifically prenatal corticosterone exposure and incubation temperature, on quantity discrimination and decision-making.
+However, we did not observe any effect of these early-life conditions on lizards’ behaviour.
+These results open new avenues for exploring the evolution of quantity discrimination in reptiles and identifying factors influencing cognitive development in this group.
+Further research should explore alternative methods, such as trained tasks, to better understand the mechanisms underlying quantity discrimination and decision-making in reptiles.
 
@@
 The ability to discriminate between quantities is a fundamental cognitive skill widely observed across the animal kingdom [@stancher2015numerical; @vonk2012bears; @beran2016capuchin].
-Numerical cognition can contribute to social decisions, mate selection, and foraging behaviour, which can be adaptive in various ecological contexts [@lucon2017individual; @nieder2018evolution].
+Quantity discrimination can contribute to social decisions, mate selection, and foraging behaviour, which can be adaptive in numerous ecological contexts [@lucon2017individual; @nieder2018evolution].
 For instance, female lions (_Panthera leo_) use numerical information to assess the risk of confronting rival groups [@mccomb1994roaring], male adult mealworms (_Tenebrio molitor_) select sites with scents of more females [@carazo2009quantity], and red-backed salamanders (_Plethodon cinereus_) prefer tubes containing larger numbers of prey [@uller2003salamanders].
-Given the adaptive advantages of numerical and quantity discrimination, research has focused extensively on exploring the ability of different species to differentiate between quantities employing different contexts and paradigms.
+Given the adaptive advantages of quantity discrimination, research has focused extensively on exploring the ability of different species to differentiate between quantities.
 
 The main two experimental approaches in the study of quantity discrimination are spontaneous choice tests and training procedures.
-In spontaneous choice tests, subjects are presented with two groups of biologically relevant stimuli differing quantitatively (number, size, volumen, etc) and ascertain if animals can distinguish them [@agrillo2014spontaneous; @cox2016quantity].
-In training procedures, animals must instead learn a numerical rule to receive a reward [@agrillo2014spontaneous; @miletto2018quantity].
+Spontaneous choice tests ascertain if animals can distinguish between two groups of biologically relevant stimuli that differ quantitatively (number, size, volume, etc) [@agrillo2014spontaneous; @cox2016quantity].
+In comparison, experiments that employ training procedures evaluate an animal’s ability to learn a numerical rule to receive a reward [@agrillo2014spontaneous; @miletto2018quantity].
 Both methods are believed to be complementary to each other, and both can provide different results even in the same species [see @miletto2017quantitative; @miletto2018quantity].
@@
 However, recent studies demonstrate that other vertebrates like amphibians or non-avian reptiles are also competent in numbers [@uller2003salamanders; @stancher2015numerical; @miletto2017quantitative].
-For example, red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour with higher volumes of food [@soldati2017long], Hermann’s tortoises (_Testudo hermanni_) discriminate between large and small food quantities in a spontaneous choice test [@gazzola2018continuous], and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) rapidly learned a quantity discrimination task [@lin2021superior; @lin2024trained].
-The Italian wall lizard (_Podarcis siculus_) failed to discriminate between different numbers of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity].
-In contrast, Iberian rock lizards (_Iberolacerta cyreni_) spontaneously preferred larger quantities of crickets when presented five versus ten items [@recio_prey_2021].
-In skinks, Shinglebacks (_Tiliqua rugosa_) exhibited spontaneous quantity discrimination using both discrete and continuous cues [@szabo2024spontaneous], while Gidgee skinks (_Egernia stokesii_) discriminated between items when they differed in number, but not when size was the main quantitative cue [@szabo_spontaneous_2021].
+For example, freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) and red-footed turtles (_Chelonoidis carbonaria_) can rapidly learn a quantity discrimination task [@soldati2017long; @lin2021superior; @lin2024trained], and Hermann’s tortoises (_Testudo hermanni_) can discriminate between large and small food quantities in a spontaneous choice test [@gazzola2018continuous].
+In lizards, the Italian wall lizard (_Podarcis siculus_) failed to discriminate between different numbers of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity].
+Shinglebacks (_Tiliqua rugosa_) exhibited spontaneous quantity discrimination using both discrete and continuous cues [@szabo2024spontaneous], while Gidgee skinks (_Egernia stokesii_) discriminated between items when they differed in number, but not when size was the main quantitative cue [@szabo_spontaneous_2021].
 These studies demonstrate that reptiles have cognitive abilities related to quantity discrimination that have been overlooked in the past, and that the use of the ANS versus the OFS can differ between species and in different contexts [@szabo_spontaneous_2021].
@@
 Cognition can be strongly influenced by developmental conditions because the brain is particularly sensitive to environmental inputs in early life stages [@zhu_prenatal_2004].
-For instance, the temperature at which embryos develop can affect neuron density or learning abilities in different species of reptiles [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @dayananda_incubation_2017; @siviter_incubation_2017; @abayarathna_effects_2020].
+For instance, the temperature at which embryos develop can affect neuron density and learning abilities in different species of reptiles [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @dayananda_incubation_2017; @siviter_incubation_2017; @abayarathna_effects_2020].
 Thermal effects on cognition may arise directly or indirectly through parental effects [@Crino_2023].
@@
 
-Here, we investigated the use of numerical cues in the common garden skink (_Lampropholis guichenoti_) using a spontaneous choice test where lizards were presented with different number of food items while controlling for continuous variables.
-In adition, we examined how prenatal corticosterone (CORT) - the main GC in reptiles - and incubation temperature can affect numerical discrimination and decision-making in foraging contexts.
+Here, we investigated quantity discrimination in the common garden skink (_Lampropholis guichenoti_) using a spontaneous choice test where lizards were presented with different number of food items.
+In addition, we examined how prenatal corticosterone (CORT) - the main GC in reptiles - and incubation temperature can affect quantity discrimination and decision-making in foraging contexts.
 To do so, we manipulated CORT concentration (CORT-treated or a sham control) in _L.
@@
 Then, we assessed numerical discrimination abilities on five tests that differed in the relative number of food items: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
-Previous studies in other skinks showed that these species can discriminate between different quantities of food using both the ANS and the OFS [@szabo_spontaneous_2021; @szabo2024spontaneous].
+Previous studies showed that skink species can discriminate between different quantities of food using both the ANS and the OFS [@szabo_spontaneous_2021; @szabo2024spontaneous].
 We predicted _L.
-guichenoti_ to be able to discriminate between different number of crickets even when controlling by other continuous variables (i.e.
+guichenoti_ to be able to discriminate between different number of crickets even when controlling by other continuous variables like total length and area (i.e.
 by using the OFS).
-Furthermore, we predicted lizards' decisions to be influenced by the early conditions they were subjected to.
-In this regard, we hypothesised that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
-We predicted that increased CORT levels would impair numerical competence [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
-We predicted that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g.
-3 VS 4), whereas even the lowest ratios (e.g.
-1 VS 4) could be challenging for those exposed to CORT or incubated at lower temperatures.
+However, we predicted that numerical discrimination would be affected by early-life conditions.
+Specifically, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
+Furthermore, we predicted that increased CORT levels would impair numerical discrimination [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
 
@@
 We checked for eggs in the boxes three days a week.
-Egg length and width were measured with a digital calliper to the nearest mm (± 0.1 mm), while mass was recorded with a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
+Egg length and width were measured with digital calipers to the nearest mm (± 0.1 mm), while mass was recorded with a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
 Eggs were assigned a clutch number and an individual identity.
@@
 
-Two weeks before starting the numerical discrimination task (see below), lizards were moved to the experimental arenas for acclimatisation.
+Lizards were acclimatised to the experimental arenas for two weeks before starting the numerical discrimination task (see below).
 The arenas were individual medium-size (41 L x 29.7 W x 22 H cm) plastic containers provided with a shelter (9 L x 6 W x 1.5 H cm) on one of the sides and a water dish in the center (@fig-Methods B).
@@
 The crickets used during the experiment were all the same size, but to control for some continuous traits we changed the orientation of the crickets as illustrated in @fig-Methods D.
-This method was intended to control for length and total area (see @stancher2015numerical).
-We tested if the area occupied by the crickets was similar in both choices (see Supplementary Material - Control of size in both options).
+This method was intended to control for length and total area [see @stancher2015numerical].
+We tested if the length occupied by the crickets was similar in both choices (see Supplementary Material - Control of size in both options).
 Our results showed that except for the 1 VS 4 and 1 VS 3 tests, the difference in length between both choices was not different from zero.
@@
 We excluded from the analyses all the individuals that did not make any choice in three or more tests.
-All the analyses were performed by PR, who was blind to the treatment of the lizards, but not the type of test or the cricket's number as the videos were analysed manually.
+All the analyses were by one observer (PR), who was blind to the treatment of the lizards, but not the type of test or the number of crickets because the videos were analysed manually.
 
@@
 Nonetheless, to the best of our knowledge there is only one study examining how developmental conditions influence numerical abilities [@vila_pouca_quantity_2019].
+Vila-Pouca et al (2019) found that Port Jackson sharks (_Heterodontus portusjacksoni_) incubated at warmer temperatures learned a numerical rule faster than cold-incubated individuals.
 Given the potent effects of early environment on cognition and brain development [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic], it is crucial to explore how early environmental conditions shape numerical abilities across taxa.

### 2025-03-17 ms.qmd
updates

@@
 ---
-title: "Quantity discrimination in the common garden skink and the role of early-life conditions"
+title: "Quantity discrimination, decision-making, and the role of early-life conditions in a lizard "
 docx: 
@@
 
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen^3^, Amelia Y. Peardon ^1^, Daniel W.A. Noble^1^
+Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen ^3,4^, Amelia Y. Peardon ^1^, Daniel W.A. Noble^1^
 
 ^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia
+
 ^2^ College of Science and Engineering, Flinders University, Bedford Park, SA 5042, Australia
+
 ^3^ University of Wollongong, Wollongong, NSW 2500, Australia
+
+^4^ Environmental Futures University of Wollongong, Wollongong, NSW, Australia
+
 $\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au
@@
 _Table S4.- Number of individuals per treatment that chose the right (R) or left (L) side in each of the numerical tests.
-p-value indicates the result of the binomial test comparing the number of choices between sides.
+p-value indicates the result of the binomial test comparing the number of choices between sides._
 ```{r, tbl-sides}

### 2025-03-23 ms.qmd
updates

@@
 ---
-title: "Quantity discrimination, decision-making, and the role of early-life conditions in a lizard "
+title: "Quantity discrimination, decision-making, and the role of early-life conditions in a lizard"
 docx: 
@@
 ## Abstract
-Quantity discrimination affects a range of behaviours that are essential for fitness including social interactions, navigation, and foraging.
+Quantity discrimination affects a range of behaviours essential for fitness, including social interactions, navigation, and foraging.
 Animals use different cognitive systems to distinguish between quantities.
@@
 In reptiles, quantity discrimination remained largely unexplored until recently, and the use of one system over the other can depend on the species or the context in which they are tested.
-Additionally, early life conditions can influence brain development and cognition, with potential effects on quantity discrimination and decision-making.
+Early life conditions can also influence brain development and cognition, potentially affecting quantity discrimination and decision-making.
 Here, we tested the ability of the common garden skink (_Lampropholis guichenoti_) to discriminate between quantities and the effects of elevated prenatal GCs and incubation temperature on quantity discrimination and decision-making.
 We assessed quantity discrimination using a spontaneous choice test with food as a stimulus.
-We subjected lizards to five two-choice tests each differing in the number of items (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4).
+We subjected lizards to five two-choice tests differing in the number of items (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4).
 Contrary to our predictions, we found no evidence of quantity discrimination in _L.
@@
 Discriminating between quantities can be crucial for individual fitness, yet it remains poorly understood in reptiles.
-In this study, we investigated quantity discrimination in _Lampropholis guichenoti_ using a spontaneous choice test with food as the stimulus.
+This study investigated quantity discrimination in _Lampropholis guichenoti_ using a spontaneous choice test with food as the stimulus.
 Contrary to our predictions, we found no evidence of quantity discrimination in this species.
@@
 Spontaneous choice tests ascertain if animals can distinguish between two groups of biologically relevant stimuli that differ quantitatively (number, size, volume, etc) [@agrillo2014spontaneous; @cox2016quantity].
-In comparison, experiments that employ training procedures evaluate an animal’s ability to learn a numerical rule to receive a reward [@agrillo2014spontaneous; @miletto2018quantity].
-Both methods are believed to be complementary to each other, and both can provide different results even in the same species [see @miletto2017quantitative; @miletto2018quantity].
+Experiments employing training procedures evaluate an animal’s ability to learn a numerical rule to receive a reward [@agrillo2014spontaneous; @miletto2018quantity].
+Both methods are believed to be complementary to each other and can provide different results even in the same species [see @miletto2017quantitative; @miletto2018quantity].
 
 Numerous reasons exist for why methods can lead to different results.
-First, spontaneous choice tests assume that one of the options presented is more adequate than the other (e.g.
+First, spontaneous choice tests assume that one of the options presented is more suitable than the other (e.g.
 the group with more conspecifics or the larger amount of food) when this may not always be the case.
-For example, choosing groups of conspecifics with less individuals could be advantageous when food is scarce but disadvantageous when predators are abundant [@yee2013costs].
-As such, context and motivation can be important drivers of decision-making that can change the way animals discriminate between quantities [@agrillo2014spontaneous].
+For example, choosing groups of conspecifics with fewer individuals could be advantageous when food is scarce but disadvantageous when predators are abundant [@yee2013costs].
+As such, context and motivation can be essential drivers of decision-making that can change the way animals discriminate between quantities [@agrillo2014spontaneous].
 Second, each method may engage distinct cognitive mechanisms when processing numerical information [@agrillo2014spontaneous].
@@
 The engagement of these systems depends not only on the numerical magnitude but also on contextual factors such as the availability of continuous variables (e.g., area, density, movement) and the nature of the task [@stancher2015numerical].
-The presence of such continuous cues typically favors ANS-based processing, while in their absence, the OFS is more likely to be employed [@agrillo2014spontaneous].
-This often complicates the study of numerosity, as continuous traits must be carefully controlled to ensure the use of the OFS [@agrillo2014spontaneous].
+Such continuous cues typically favor ANS-based processing, while in their absence, the OFS is more likely to be employed [@agrillo2014spontaneous], which often complicates the study of numerosity, as continuous traits must be carefully controlled to ensure the use of the OFS [@agrillo2014spontaneous].
 In spontaneous choice tests, for instance, animals may rely on cumulative surface area rather than numerical differences when selecting between two sets of food items [@agrillo2014spontaneous].
-To disentangle these cues, researchers often equate the total area of the sets while changing the number of items, forcing individuals to rely on numerical information alone [@stancher2015numerical].
+To disentangle these cues, researchers often keep the total area of the sets constant while changing the number of items, forcing individuals to rely on numerical information alone [@stancher2015numerical].
 
 Quantity discrimination studies have been traditionally focused on mammals, birds, and fishes despite the influence of numerical cognition on fitness across taxa [@mehlis2015quantification; @nieder2018evolution].
-However, recent studies demonstrate that other vertebrates like amphibians or non-avian reptiles are also competent in numbers [@uller2003salamanders; @stancher2015numerical; @miletto2017quantitative].
+However, recent studies demonstrate that other vertebrates like amphibians or non-avian reptiles are also competent in discerning numbers [@uller2003salamanders; @stancher2015numerical; @miletto2017quantitative].
 For example, freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) and red-footed turtles (_Chelonoidis carbonaria_) can rapidly learn a quantity discrimination task [@soldati2017long; @lin2021superior; @lin2024trained], and Hermann’s tortoises (_Testudo hermanni_) can discriminate between large and small food quantities in a spontaneous choice test [@gazzola2018continuous].
 In lizards, the Italian wall lizard (_Podarcis siculus_) failed to discriminate between different numbers of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity].
-Shinglebacks (_Tiliqua rugosa_) exhibited spontaneous quantity discrimination using both discrete and continuous cues [@szabo2024spontaneous], while Gidgee skinks (_Egernia stokesii_) discriminated between items when they differed in number, but not when size was the main quantitative cue [@szabo_spontaneous_2021].
+Shinglebacks (_Tiliqua rugosa_) exhibited spontaneous quantity discrimination using both discrete and continuous cues [@szabo2024spontaneous], while Gidgee skinks (_Egernia stokesii_) discriminated between items when they differed in number but not when size was the main quantitative cue [@szabo_spontaneous_2021].
 These studies demonstrate that reptiles have cognitive abilities related to quantity discrimination that have been overlooked in the past, and that the use of the ANS versus the OFS can differ between species and in different contexts [@szabo_spontaneous_2021].
@@
 
-Cognition can be strongly influenced by developmental conditions because the brain is particularly sensitive to environmental inputs in early life stages [@zhu_prenatal_2004].
+Developmental conditions can strongly influence cognition because the brain is particularly sensitive to environmental inputs in early life stages [@zhu_prenatal_2004].
 For instance, the temperature at which embryos develop can affect neuron density and learning abilities in different species of reptiles [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @dayananda_incubation_2017; @siviter_incubation_2017; @abayarathna_effects_2020].
@@
 
-Here, we investigated quantity discrimination in the common garden skink (_Lampropholis guichenoti_) using a spontaneous choice test where lizards were presented with different number of food items.
+Here, we investigated quantity discrimination in the common garden skink (_Lampropholis guichenoti_) using a spontaneous choice test where lizards were presented with different numbers of food items.
 In addition, we examined how prenatal corticosterone (CORT) - the main GC in reptiles - and incubation temperature can affect quantity discrimination and decision-making in foraging contexts.
@@
 We predicted _L.
-guichenoti_ to be able to discriminate between different number of crickets even when controlling by other continuous variables like total length and area (i.e.
+guichenoti_ to be able to discriminate between different numbers of crickets even when controlling by other continuous variables like total length and area (i.e.
 by using the OFS).
@@
 Eggs were assigned a clutch number and an individual identity.
-Then we treated eggs with CORT or a vehicle control (see _Manipulating Early Thermal and CORT Environments_ details below) and placed them in individual cups (80 mL) with moist vermiculite (12 g water to 4 g vermiculite).
+Then, we treated eggs with CORT or a vehicle control (see _Manipulating Early Thermal and CORT Environments_ details below) and placed them in individual cups (80 mL) with moist vermiculite (12 g water to 4 g vermiculite).
 We covered the cups with plastic wrap to retain moisture and left the eggs in incubators (LATWIT 2X5D-R1160) programmed to two different thermal regimes (see _Manipulating Early Thermal and CORT Environments_ details below) until hatching.
@@
 *Hatchlings* -- Immediately after hatching, we measured snout-vent length (SVL) and tail length (TL) with a ruler to the nearest mm (± 0.1 mm), and mass using a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
-Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) provided with non-stick matting and a small water dish.
+Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) with non-stick matting and a small water dish.
 During this period, they were given water daily and received 3-6 small *A.
@@
 We manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design (@fig-Methods A).
-Eggs were allocated to one of two different developmental stress treatments: CORT-treated eggs were topically supplied with 5 µL of a CORT solution (10 pg/mL).
+Eggs were allocated to one of two developmental stress treatments: CORT-treated eggs were topically supplied with 5 µL of a CORT solution (10 pg/mL).
 Control treatment eggs received an equal volume of 100% Ethanol.
@@
 C2505) in 100% ethanol.
-We selected doses based on previous studies where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs in _Lampropholis delicata_ [@crino2024eggs].
+based on previous studies, we selected doses where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs in _Lampropholis delicata_ [@crino2024eggs].
 Then, eggs were incubated in one of the two previously mentioned temperature regimes (Cold or Hot) until hatching.
-These temperatures represent the upper and lower limit of the natural incubation temperatures for this species [@qualls2000post].
+These temperatures represent the upper and lower limits of the natural incubation temperatures for this species [@qualls2000post].
 
 The number of eggs per clutch assigned to each hormone and temperature treatment was counterbalanced in a partial split-clutch design.
-At least one egg per clutch to each treatment, and the remaining eggs were randomly assigned to one of the treatments.
+At least one egg per clutch was allocated to each treatment, and the remaining eggs were randomly assigned to one of the treatments.
 When we found less than four eggs in a clutch, we assigned each egg randomly to one of the treatments.
@@
 Each rack was associated with its own CCTV system (device model DVR-HP210475), allowing us to record lizards' behaviour during the experiment.
-Despite the conditions in the experimental rooms were the same as in the colony room, we counterbalanced the number of lizards per treatment in each rack to control for any potential effects of the room or the position of the lizard in the rack.
-During acclimatisation and throughout the experiment, lizards were fed with only one cricket daily, dusted with calcium and multivitamins (see protocol below), and water was supplied *ad libitum*.
+Despite the conditions in the experimental rooms being the same as in the colony room, we counterbalanced the number of lizards per treatment in each rack to control for any potential effects of the room or the position of the lizard in the rack.
+During acclimatisation and throughout the experiment, lizards were fed only one cricket daily, dusted with calcium and multivitamins (see protocol below), and water was supplied *ad libitum*.
 The temperature of both rooms was set to 22-24ºC, but we also provided a temperature gradient (24-32°C) using a heat cord along the center of the arena and heat lamps in a 12 h light: 12 h dark cycle.
@@
 
-The crickets used during the experiment were all the same size, but to control for some continuous traits we changed the orientation of the crickets as illustrated in @fig-Methods D.
+The crickets used during the experiment were all the same size, but to control for some continuous traits, we changed the orientation of the crickets as illustrated in @fig-Methods D.
 This method was intended to control for length and total area [see @stancher2015numerical].
@@
 We considered that the lizard interacted with the food when it touched the Petri dish with its snout or when the front half of the body was on top of the platform.
-For five minutes after the first interaction with one of the options, we recorded the time spent interacting with that option for a maximum of five minutes; we did this for both amounts of food independently of what was the first choice.
-We measured the interest shown for the higher amount of food (_Interest_) by subtracting the time interacting with the option with fewer crickets to the time spent interacting with the larger amount of food.
+For five minutes after the first interaction with one of the options, we recorded the time spent interacting with that option for a maximum of five minutes; we did this for both amounts of food independently of the first choice.
+We measured the interest shown for the higher amount of food (_Interest_) by subtracting the time interacting with the option with fewer crickets from the time spent interacting with the larger amount of food.
 If there was no interaction with one of the options during the test, we considered the time spent as zero for that option.
-When there was no interaction with any of the choices, both _Latency_, _Choice_, and _Interest_ were noted as NA.
-We also considered Interest as NA in a few cases where the animals managed to lift the dish and eat the prey; Latency and Choice were recorded and employed in the analyses.
+_Latency_, _Choice_, and _Interest_ were noted as NA when there was no interaction with any of the choices.
+We also considered _Interest_ as NA in a few cases where the animals managed to lift the dish and eat the prey; _Latency_ and _Choice_ were recorded and employed in the analyses.
 We excluded from the analyses all the individuals that did not make any choice in three or more tests.
@@
 Motivation plays a key role in animals' performance in spontaneous choice tests [@agrillo2014spontaneous; @bisazza2014experim].
-To ensure that lizards remained motivated to forage, they were provided with only one cricket during both acclimatisation and the numerical task (see above).
+To ensure that lizards remained motivated to forage, they were provided only one cricket during acclimatisation and the numerical task (see above).
 During acclimatisation, lizards were fed at the same time each day as the experiments were conducted.
@@
 To further control for motivation, we excluded from our analyses any trials in which lizards did not interact with any options or did not eat the cricket provided after the test (n = 0 in both cases).
-Consequently, we do not consider motivation to be a major factor influencing our results (but see Discussion).
+Consequently, we do not consider motivation a significant factor influencing our results (but see Discussion).
 
@@
 We included test (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4), hormone (CORT versus Control), incubation temperature (Cold versus Hot), and the three-way interaction as fixed factors.
-We also included in our predictors sex, and age.
+We also included sex, and age as predictors.
 We did not find any effect of sex or age on Choice or Interest, but we found an effect on Latency (see Results).
@@
 guichenoti_ lays up to two clutches per year [@joss1985reproductive].
-Since egg collection was done during half of the breeding season, each clutch likely came from a unique mother, and so, clutch identity captures potential maternal effects.
-
-We used the posterior distributions of parameters from these models to test for between treatment differences in the ability to discriminate different numbers of crickets for each of the numerical tests.
+Since egg collection was done during half of the breeding season, each clutch likely came from a unique mother, so clutch identity captures potential maternal effects.
+
+We used the posterior distributions of parameters from these models to test for treatment differences in the ability to discriminate different numbers of crickets for each of the numerical tests.
 Because the posterior distributions were not always normally distributed, we report the median as the main measure of central tendency.
-We used the 95% Highest Posterior Density Intervals (95% HPDI) using the hdi function in bayestestR [@bayestestR] to test if the contrasts between treatments for all the variables, or the _Interest_ per each test was different form zero.
-Given Choice was Bernoulli, we modified our null hypotheses to test if differed from a probability of 0.5 as there were only two possible choices and if animals were choosing randomly we would predict each choice would have a 50% probability.
+We used the 95% Highest Posterior Density Intervals (95% HPDI) using the hdi function in bayestestR [@bayestestR] to test if the contrasts between treatments for all the variables or the _Interest_ per each test were different from zero.
+Given Choice was Bernoulli, we modified our null hypotheses to test if it differed from a probability of 0.5 as there were only two possible choices and if animals were choosing randomly we would predict each choice would have a 50% probability.
 
@@
 Experimental procedures and laboratory housing complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33).
-Upon completion of experimental trials, lizards were kept in their enclosure and will be euthanized following standard humane procedures once they reach the humane endpoint.
+Upon completion of experimental trials, lizards were kept in their enclosure and will be euthanised following standard humane procedures once they reach the humane endpoint.
 
@@
 Overall, these findings suggest that _L.
-guichenoti_ did not use numerical or other potential quantitative cues for foraging decisions and that any decision making was not affected by early environment.
+guichenoti_ did not use numerical or other potential quantitative cues for foraging decisions and that any decision-making was not affected by early environment.
 
 #### Foraging context may influence spontaneous quantity discrimination in reptiles  
-Lack of numerical discrimination in spontaneous tests may be due to _L.
+The lack of numerical discrimination in spontaneous tests may be due to _L.
 guichenoti_ being insectivorous.
@@
 10) [@recio_prey_2021].
-In contrast, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities despite showing quantity discrimination when prey size was the main cue [@miletto2017quantitative] or when tested using training procedures [@miletto2018quantity].
+In contrast, when _Podarcis siculus_ were presented with different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities despite showing quantity discrimination when prey size was the main cue [@miletto2017quantitative] or when tested using training procedures [@miletto2018quantity].
 Notably, the numbers of food items employed in @miletto2017quantitative were always fewer than four.
-Spontaneous choice tests assume that animals should select the most appropriate quantity, which may not necessarily be the largest one [@agrillo2014spontaneous].
-This occurs if the costs of selecting the larger option outweigh the benefits.
-If capturing larger prey demands more time or energy, or if success rates remain similar regardless of prey quantity, animals may not exhibit a preference for larger quantities [@agrillo2014spontaneous].
+Spontaneous choice tests assume that animals should select the most appropriate quantity, which may not necessarily be the largest one [@agrillo2014spontaneous] if the costs of selecting the larger option outweigh the benefits.
+For example, if capturing larger prey demands more time or energy or success rates remain similar regardless of prey quantity, animals may not prefer larger quantities [@agrillo2014spontaneous].
 As a result, even if diet might not affect quantity discrimination abilities, insectivorous lizards might not have evolved a preference for larger patches of prey, at least with small numbers [see @recio_prey_2021].
@@
 Here, we aimed to investigate whether _L.
-guichenoti_ could discriminate between different number of crickets in a spontaneous choice test relying exclusively on the OFS.
-To ensure that lizards decision relied on tracking discret items rather than using non-numerical cues, we controlled for total length and total area — two variables that could influence ANS processing.
+guichenoti_ could discriminate between different numbers of crickets in a spontaneous choice test relying exclusively on the OFS.
+To ensure that lizards decision relied on tracking discrete items rather than using non-numerical cues, we controlled for total length and total area — two variables that could influence ANS processing.
 We did this by rotating the position of the crickets in the smaller-choice option, making the total length and total area comparable in both choices [@agrillo2014spontaneous; @stancher2015numerical; @cooper2024tell].
-Absence of discrimination can therefore be explained by _L.
+The absence of discrimination can therefore be explained by _L.
 guichenoti_ relying on the ANS for quantity discrimination.
-Nonetheless, in two of our tests (1 VS 4 and 1 VS 3 crickets) cricket rotation was not sufficient to control for total length (see _Supplementary Material - Control of size in both options_).
+Nonetheless, in two of our tests (1 VS 4 and 1 VS 3 crickets) cricket rotation was insufficient to control for total length (see _Supplementary Material - Control of size in both options_).
 Yet, we still did not find quantity discrimination in _L.
-guichenoti_ which suggests that neither the OFS nor the ANS were used.
+guichenoti_, which suggests that neither the OFS nor the ANS were used.
 Instead, our results likely reflect a genuine lack of quantity discrimination in this species, at least in spontaneous choice tests.
@@
 Although the effects of early environment on cognition are widely studied, little attention has been paid to its impacts on numerical discrimination and decision-making.
-We found that decision-making was not influenced by CORT elevations or incubation temperature, as latency to make a choice did not vary across prenatal conditions.
-This suggests that _L.
+We found that decision-making was not influenced by CORT elevations or incubation temperature, as latency to make a choice did not vary across prenatal conditions, which suggests that _L.
 guichenoti_ may be robust to these early-life factors, at least in the context of foraging decisions.
 Previous studies have reported that pine snakes (_Pituophis melanoleucus_) incubated at low temperatures were slower and less proficient in capturing prey [@burger_effects_1991], but cold incubated bearded dragons (_Pogona vitticeps_) were faster at approaching a novel object at least at a young age [@siviter_impact_2017].
-Maternal transmitted GCs are also known to reduce exploration and neophilia in different taxa [@koolhaas1999coping; @hope_incubation_2018].
+Maternally transmitted GCs are also known to reduce exploration and neophilia in different taxa [@koolhaas1999coping; @hope_incubation_2018].
 Additionally, both prenatal CORT and incubation temperature can influence metabolic rate (MR), which in turn could affect motivation in food-based learning paradigms [@noble_developmental_2018; @cossin2022effect; but see @trnik_persistent_2011; @crino_corticosterone_2014-learn].
@@
 #### Conclusion
-Our study was focused on investigating if _L.
-guichenoti_ could use the OFS to discriminate between different number of food items.
+Our study investigated whether _L.
+guichenoti_ could use the OFS to discriminate between different numbers of food items.
 Additionally, we were interested in exploring the impact of prenatal conditions on numerical abilities and decision-making in reptiles.
 Our findings suggest that _L.
-guichenoti_ does not rely on numerical information for foraging decisions and that their performance and decision-making does not vary due to prenatal temperature or CORT exposure.
+guichenoti_ does not rely on numerical information for foraging decisions and that their performance and decision-making do not vary due to prenatal temperature or CORT exposure.
 Our findings suggest that decision-making can be robust to early environmental conditions, at least in the context of foraging.

### 2025-05-20 ms.qmd
updates

@@
 bibliography: "./bib/refs.bib"
-csl: "./bib/behavioral-ecology-and-sociobiology.csl"
+csl: "./bib/animal-cognition.csl"
 execute:
@@
 We assessed quantity discrimination using a spontaneous choice test with food as a stimulus.
-We subjected lizards to five two-choice tests differing in the number of items (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4).
+We subjected lizards to five two-choice tests differing in the number of items (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4) while controlling for total length and area.
 Contrary to our predictions, we found no evidence of quantity discrimination in _L.

### 2025-07-28 ms.qmd
updates

@@
 However, recent studies demonstrate that other vertebrates like amphibians or non-avian reptiles are also competent in discerning numbers [@uller2003salamanders; @stancher2015numerical; @miletto2017quantitative].
-For example, freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) and red-footed turtles (_Chelonoidis carbonaria_) can rapidly learn a quantity discrimination task [@soldati2017long; @lin2021superior; @lin2024trained], and Hermann’s tortoises (_Testudo hermanni_) can discriminate between large and small food quantities in a spontaneous choice test [@gazzola2018continuous].
+For example, freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) and red-footed tortoises (_Chelonoidis carbonaria_) can rapidly learn a quantity discrimination task [@soldati2017long; @lin2021superior; @lin2024trained], and Hermann’s tortoises (_Testudo hermanni_) can discriminate between large and small food quantities in a spontaneous choice test [@gazzola2018continuous].
 In lizards, the Italian wall lizard (_Podarcis siculus_) failed to discriminate between different numbers of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity].
@@
 Specifically, we compared: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
-The food consisted of small frozen crickets (*A.
-domestica*) that were thawed and dusted with calcium and multivitamins.
-We decided to use frozen crickets to avoid movement that would affect lizard behaviour [@cooper2024tell].
-Lizards were habituated to eating frozen crickets for one month before the experiments.
+Each numerical test was made only one per lizard.
+The food consisted of small thawed frozen crickets (*A.
+domestica*) that were dusted with calcium and multivitamins.
+We decided to use thawed frozen crickets to avoid movement that would affect lizard behaviour [@cooper2024tell].
+Lizards were habituated to eating thawed frozen crickets for one month before the experiments.
 
@@
 This method was intended to control for length and total area [see @stancher2015numerical].
-We tested if the length occupied by the crickets was similar in both choices (see Supplementary Material - Control of size in both options).
+We tested if the length occupied by the crickets was similar in both choices (see _Supplementary Material: Control of size in both options_).
 Our results showed that except for the 1 VS 4 and 1 VS 3 tests, the difference in length between both choices was not different from zero.
 As such, discrimination between both choices in the rest of the tests (2 VS 4, 2 VS 3, 3 VS 4 crickets) would be produced by the OFS, not the ANS [@agrillo2014spontaneous].
-In addition, after the experiments, we conducted another set of trials to test whether the orientation of the crickets could affect lizards’ choice, but we found no effect of orientation (see _Supplementary Material - Prey orientation test_).
+In addition, after the experiments, we conducted another set of trials to test whether the orientation of the crickets could affect lizards’ choice, but we found no effect of orientation (see _Supplementary Material: Prey orientation test_).
 
@@
 
-knitr::include_graphics("./Others/NUMDIS_FIG_1.png")
+knitr::include_graphics("./Others/NUMDIS_FIG_1.pdf")
 
@@
 We considered that the lizard interacted with the food when it touched the Petri dish with its snout or when the front half of the body was on top of the platform.
-For five minutes after the first interaction with one of the options, we recorded the time spent interacting with that option for a maximum of five minutes; we did this for both amounts of food independently of the first choice.
-We measured the interest shown for the higher amount of food (_Interest_) by subtracting the time interacting with the option with fewer crickets from the time spent interacting with the larger amount of food.
+After the first interaction with any of the options, we recorded the time spent interacting with that option for a maximum of five minutes.
+We recorded this time for both amounts of food independently of the first choice.
+We measured the interest shown for the higher amount of food (_Interest_) by subtracting the time spent interacting with the smaller food option from the time spent interacting with the larger one.
 If there was no interaction with one of the options during the test, we considered the time spent as zero for that option.
@@
   annotate("text", x = 0.61, y = 0.505, label = "*", size = 7, color = "black")
-ggsave("./output/figures/fig_results.png", plot=fig_results, width = 25, height = 25, units = "cm", dpi = 600)
-knitr::include_graphics("./output/figures/fig_results.png")
+ggsave("./output/figures/fig_results.pdf", plot=fig_results, width = 25, height = 25, units = "cm", dpi = 600)
+knitr::include_graphics("./output/figures/fig_results.pdf")
 ```
@@
 
-#### Foraging context may influence spontaneous quantity discrimination in reptiles  
-The lack of numerical discrimination in spontaneous tests may be due to _L.
-guichenoti_ being insectivorous.
-In studies where reptiles discriminated between different amounts of food, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs.
+#### Foraging context may influence spontaneous quantity discrimination in _L. guichenoti_  
+The lack of numerical discrimination in food-related spontaneous tests in _L.
+guichenoti_ may be due to their insectivorous diet.
+In studies where reptiles discriminated between different amounts of food, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large numbers or alive prey (e.g., 5 vs.
 10) [@recio_prey_2021].
@@
 For example, if capturing larger prey demands more time or energy or success rates remain similar regardless of prey quantity, animals may not prefer larger quantities [@agrillo2014spontaneous].
+Similarly, animals would not show any preference towards larger number of food items if the likelihood of acquiring one item is the same regardless of quantity [@agrillo2014spontaneous].
 As a result, even if diet might not affect quantity discrimination abilities, insectivorous lizards might not have evolved a preference for larger patches of prey, at least with small numbers [see @recio_prey_2021].
@@
 We did this by rotating the position of the crickets in the smaller-choice option, making the total length and total area comparable in both choices [@agrillo2014spontaneous; @stancher2015numerical; @cooper2024tell].
-The absence of discrimination can therefore be explained by _L.
+The absence of discrimination in our experiment could be explained by _L.
 guichenoti_ relying on the ANS for quantity discrimination.
-Nonetheless, in two of our tests (1 VS 4 and 1 VS 3 crickets) cricket rotation was insufficient to control for total length (see _Supplementary Material - Control of size in both options_).
+Nonetheless, in two of our tests (1 VS 4 and 1 VS 3 crickets) cricket rotation was insufficient to control for total length (see _Supplementary Material: Control of size in both options_).
 Yet, we still did not find quantity discrimination in _L.
 guichenoti_, which suggests that neither the OFS nor the ANS were used.
-Instead, our results likely reflect a genuine lack of quantity discrimination in this species, at least in spontaneous choice tests.
+Instead, our results likely reflect a genuine lack of quantity discrimination in foraging-related spontaneous choice tests.
 
@@
 guichenoti_ may be robust to these early-life factors, at least in the context of foraging decisions.
-Previous studies have reported that pine snakes (_Pituophis melanoleucus_) incubated at low temperatures were slower and less proficient in capturing prey [@burger_effects_1991], but cold incubated bearded dragons (_Pogona vitticeps_) were faster at approaching a novel object at least at a young age [@siviter_impact_2017].
+Previous studies have reported that pine snakes (_Pituophis melanoleucus_) incubated at low temperatures were slower and less proficient in capturing prey [@burger_effects_1991], and cold incubated bearded dragons (_Pogona vitticeps_) made more errors in a foraging task than warm-incubated lizards [@siviter2019egg].
 Maternally transmitted GCs are also known to reduce exploration and neophilia in different taxa [@koolhaas1999coping; @hope_incubation_2018].
@@
 The absence of numerical discrimination in _L.
-guichenoti_ is consistent with previous studies showing that insectivorous reptiles fail in spontaneous choice tests.
+guichenoti_ in food-related spontaneous tests align with previous studies on insectivorous reptiles.
 However, it is critical to note that the methodologies employed in this study may not have fully captured the nuances of numerical cognition in _L.

### 2025-09-11 ms.qmd
updates

@@
 
+```{r, sex_ratios}
+#| label: sex_ratios
+# Get the sex ratios for each treatment
+sex <- c("m", "f")
+trt <- c("A_23", "B_23", "A_28", "B_28")
+for(s in sex){
+  for(t in trt){
+    count <- data_num %>%
+      filter(day == 1, trt == t, sex == s) %>%
+      nrow()
+
+    name <- paste0(t, "_", s)   # e.g. "A_23_f"
+    assign(name, count)
+  }
+}
+```
+
 ```{r, models}
@@
 ```
+
+We recorded and analysed a total of 400 videos from 80 lizards (n = 20 per treatment, see @fig_results).
+The lizards were between `r min(data_num$age_raw)` and `r max(data_num$age_raw)` days post-hatching (see Fig.
+S5 in _Supplementary Material_).
+Final sample sizes were composed by 43 females and 37 males, with no less than 6 females per treatment and no more than 13.
 
@@
 
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+#### Lizards age by Treatment
+
+```{r, fig-age_treatment}
+#| label: fig-age
+#| fig.cap: "Distribution of the age of the lizards by treatment"
+# Plotting the distribution of the age of the lizards by treatment
+#
+plot_age_df <- data_num %>%
+  mutate(trt = factor(trt, 
+          levels = c("A_23", "B_23", "A_28", "B_28"), 
+          labels = c("Cold-Control", "Cold-CORT", "Hot-Control", "Hot-CORT")))
+#
+plot_age <- ggplot(plot_age_df, aes(x = age_raw, color = trt)) +
+  geom_density(aes(y = after_stat(count)), linewidth = 1.5) +
+  scale_color_manual(values = c("Cold-CORT" = "#00008B",
+                              "Cold-Control" = "#68bde1",
+                              "Hot-CORT" = "#b50101",
+                              "Hot-Control" = "#fa927d")) +  
+  labs(y = "Number of lizards", x = "Age (days)", color = "Treatments") +
+  theme_classic() +
+  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm"),
+    axis.text.y = element_text(size = 12, family = "Times"),   # Remove y-axis labels
+    axis.title = element_text(size = 12, family = "Times"),
+    axis.text = element_text(size = 10, family = "Times"),
+    legend.position = "right",
+    legend.title = element_text(size = 12, family = "Times"),
+    legend.text = element_text(size = 11, family = "Times")
+    ) 
+#
+ggsave("./output/figures/plot_age.png", plot=plot_age, width = 25, height = 18, units = "cm", dpi = 600)
+knitr::include_graphics("./output/figures/plot_age.png")
+```
+
+Fig. S5 - Distribution of the age of the lizards by treatment
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
 #### Checking the model plots  

### 2025-09-11 R/data_process.R

@@ -20,6 +20,7 @@ data_num <- read.csv(here("data/lizard_database_number.csv")) %>%
     labels = c("B" = "CORT", "A" = "Control"))) %>%
   mutate(test_type = factor(test_type,
     levels = c("1 VS 4", "1 VS 3", "2 VS 4", "2 VS 3", "3 VS 4"))) %>%
+  mutate(age_raw = as.numeric(age)) %>%
   mutate(age = scale(age, center = TRUE, scale =FALSE)) %>%
 data.frame()
 #

### 2025-09-24 ms.qmd
updates

@@
 
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen ^3,4^, Amelia Y. Peardon ^1^, Daniel W.A. Noble^1^
+Pablo Recio^1, 2^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen ^3,4^, Amelia Y. Peardon ^1^, Daniel W.A. Noble^1^
 
@@
 ## Abstract
-Quantity discrimination affects a range of behaviours essential for fitness, including social interactions, navigation, and foraging.
-Animals use different cognitive systems to distinguish between quantities.
-The Approximate Number System (ANS) estimates, while the Object File System (OFS) tracks discrete items.
-In reptiles, quantity discrimination remained largely unexplored until recently, and the use of one system over the other can depend on the species or the context in which they are tested.
-Early life conditions can also influence brain development and cognition, potentially affecting quantity discrimination and decision-making.
+Quantity discrimination affects a range of behaviours essential for fitness, including social interactions, navigation, and foraging and is widespread across animal taxa.
+However, in reptiles, evidence for this ability—particularly for small number contrasts—remains limited and mixed.
+Early-life conditions, such as hormonal exposure and temperature, can also shape brain development and cognitive performance, potentially affecting numerical abilities.
 Here, we tested the ability of the common garden skink (_Lampropholis guichenoti_) to discriminate between quantities and the effects of elevated prenatal GCs and incubation temperature on quantity discrimination and decision-making.
@@
 The main two experimental approaches in the study of quantity discrimination are spontaneous choice tests and training procedures.
-Spontaneous choice tests ascertain if animals can distinguish between two groups of biologically relevant stimuli that differ quantitatively (number, size, volume, etc) [@agrillo2014spontaneous; @cox2016quantity].
+Spontaneous choice tests ascertain if animals can distinguish between two groups of biologically relevant stimuli that differ quantitatively (number, size, volume) [@agrillo2014spontaneous; @cox2016quantity].
 Experiments employing training procedures evaluate an animal’s ability to learn a numerical rule to receive a reward [@agrillo2014spontaneous; @miletto2018quantity].
@@
 For example, choosing groups of conspecifics with fewer individuals could be advantageous when food is scarce but disadvantageous when predators are abundant [@yee2013costs].
-As such, context and motivation can be essential drivers of decision-making that can change the way animals discriminate between quantities [@agrillo2014spontaneous].
+As such, context and motivation can be essential drivers of decision-making that can change the way animals discriminate between quantities [@gazes2018impact].
 Second, each method may engage distinct cognitive mechanisms when processing numerical information [@agrillo2014spontaneous].
 Different procedures may shape how animals discriminate between quantities, making them more responsive towards certain quantity traits over others.
-Some tasks may encourage a focus on approximate magnitudes, while others may promote attention to individual items [@agrillo2014spontaneous].
+Some tasks may encourage a focus on approximate magnitudes, while others may promote attention to individual items [@soldati2017long].
 
@@
 In contrast, the OFS allows for the precise tracking of discrete items but is thought to be limited to small quantities, typically up to three or four [@hyde2011two].
-The engagement of these systems depends not only on the numerical magnitude but also on contextual factors such as the availability of continuous variables (e.g., area, density, movement) and the nature of the task [@stancher2015numerical].
+However, when discriminating between small quantities, the engagement of these systems also depends on contextual factors such as the availability of continuous variables (e.g., area, density, movement) and the nature of the task [@stancher2015numerical].
 Such continuous cues typically favor ANS-based processing, while in their absence, the OFS is more likely to be employed [@agrillo2014spontaneous], which often complicates the study of numerosity, as continuous traits must be carefully controlled to ensure the use of the OFS [@agrillo2014spontaneous].
-In spontaneous choice tests, for instance, animals may rely on cumulative surface area rather than numerical differences when selecting between two sets of food items [@agrillo2014spontaneous].
-To disentangle these cues, researchers often keep the total area of the sets constant while changing the number of items, forcing individuals to rely on numerical information alone [@stancher2015numerical].
+In spontaneous choice tests, for instance, animals may rely on cumulative surface area (ANS-related) rather than numerical (OFS-associated) differences when selecting between two sets of food items [@agrillo2014spontaneous].
+To disentangle these cues, researchers often keep the total area of the sets constant while changing the number of items, forcing individuals to rely on numerical information alone [@vonk2012bears; @stancher2015numerical; @beran2016capuchin].
 
@@
 Developmental conditions can strongly influence cognition because the brain is particularly sensitive to environmental inputs in early life stages [@zhu_prenatal_2004].
-For instance, the temperature at which embryos develop can affect neuron density and learning abilities in different species of reptiles [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @dayananda_incubation_2017; @siviter_incubation_2017; @abayarathna_effects_2020].
-Thermal effects on cognition may arise directly or indirectly through parental effects [@Crino_2023].
-For example, temperatures outside the optimal range can trigger stress responses, leading to the release of glucocorticoids (GCs) [@sapolsky_how_2000], which may then be transmitted to offspring [@Crino_2023].
+In ectotherms, the temperature at which embryos develop have significant effects on several traits [@noble_developmental_2018].
+For example, incubation at warm temperatures influences neuron density and learning abilities in different species of reptiles [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @dayananda_incubation_2017; @siviter_incubation_2017; @abayarathna_effects_2020].
+Importantly, thermal effects on cognition may arise directly or through parental effects [@Crino_2023].Thermal fluctuations, particularly those approaching the edges of an organism's optimal range can trigger the stress response, leading to the release of glucocorticoids (GCs) [@sapolsky_how_2000], which may then be transmitted to offspring [@Crino_2023].
 Elevated prenatal GCs can, in turn, influence cognition by altering brain function [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic].
-As such, the interaction between GCs and developmental temperature may be relevant in shaping reptile cognition.
+For instance, stressing pregnant rats (_Rattus norvegicus_) suppresses neurogenesis in the dentate gyrus of the offspring, associated with impairments in spatial learning [@lemaire_prenatal_2000].
+As such, the interaction between GCs and developmental temperature may be relevant in shaping reptile cognition, especially under the context of climate change [@Crino_2023].
 Most studies on the effects of GC exposure or elevated temperatures during development focus on learning and memory, but few studies have investigated how early-life conditions can affect quantity discrimination despite its importance on animals' lives [but see @vila_pouca_quantity_2019].
@@
 Here, we investigated quantity discrimination in the common garden skink (_Lampropholis guichenoti_) using a spontaneous choice test where lizards were presented with different numbers of food items.
+Our aim was to understand if this species could use the OFS to discriminate between small quantities of prey in a spontaneous choice test.
 In addition, we examined how prenatal corticosterone (CORT) - the main GC in reptiles - and incubation temperature can affect quantity discrimination and decision-making in foraging contexts.
@@
 We predicted _L.
-guichenoti_ to be able to discriminate between different numbers of crickets even when controlling by other continuous variables like total length and area (i.e.
+guichenoti_ to be able to discriminate between different numbers of crickets even when controlling for other continuous variables like total length and area (i.e.
 by using the OFS).
 However, we predicted that numerical discrimination would be affected by early-life conditions.
-Specifically, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
+Specifically, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task as previous studies have reported enhanced cognitive abilities ectotherms incubated under warmer conditions [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
 Furthermore, we predicted that increased CORT levels would impair numerical discrimination [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
@@
 We checked for eggs in the boxes three days a week.
-Egg length and width were measured with digital calipers to the nearest mm (± 0.1 mm), while mass was recorded with a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
 Eggs were assigned a clutch number and an individual identity.
-Then, we treated eggs with CORT or a vehicle control (see _Manipulating Early Thermal and CORT Environments_ details below) and placed them in individual cups (80 mL) with moist vermiculite (12 g water to 4 g vermiculite).
-We covered the cups with plastic wrap to retain moisture and left the eggs in incubators (LATWIT 2X5D-R1160) programmed to two different thermal regimes (see _Manipulating Early Thermal and CORT Environments_ details below) until hatching.
+Then, we treated eggs with CORT or a vehicle control (see _Manipulating Early Thermal and CORT Environments_ details below)and placed each egg in individual cups (80 mL) with moist vermiculite (12 g water to 4 g vermiculite) and covered the cups with plastic wrap to retain moisture.
+Each egg was incubated individually in incubators (LATWIT 2X5D-R1160) programmed to two different thermal regimes (see _Manipulating Early Thermal and CORT Environments_ details below) until hatching.
 Incubators were checked for the presence of hatchlings three times a week.
-
-*Hatchlings* -- Immediately after hatching, we measured snout-vent length (SVL) and tail length (TL) with a ruler to the nearest mm (± 0.1 mm), and mass using a digital scale (OHAUS, Model spx123) to the nearest g (± 0.001 g).
-Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) with non-stick matting and a small water dish.
-During this period, they were given water daily and received 3-6 small *A.
-domestica* crickets three times a week.
-All care otherwise follows similar protocols to adults (see above).
+Egg and hatchling mortality was not higher than 13% in any of the treatments (Leibold et al., In prep.).
+
+*Hatchlings* -- Hatchlings were placed in individual enclosures immediately after hatchling (18.7L x 13.2W x 6.3H cm) with non-stick matting and a small water dish and until the acclimatisation period (see below).
+During this time, they were given water daily and received 3-6 small _A.
+domestica_ crickets three times a week.
+All care otherwise followed similar protocols to adults (see above).
 
@@
 We manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design (@fig-Methods A).
-Eggs were allocated to one of two developmental stress treatments: CORT-treated eggs were topically supplied with 5 µL of a CORT solution (10 pg/mL).
-Control treatment eggs received an equal volume of 100% Ethanol.
+Eggs were allocated to one of two developmental stress treatments: CORT-treated eggs were topically supplied with 5 µL of a CORT solution (10 pg/mL) once, immediately after collection and before incubation.
 Corticosterone treatments were made by dissolving crystalline corticosterone (Sigma, Cat.
 No.
-C2505) in 100% ethanol.
-based on previous studies, we selected doses where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs in _Lampropholis delicata_ [@crino2024eggs].
-Then, eggs were incubated in one of the two previously mentioned temperature regimes (Cold or Hot) until hatching.
+C2505) in 100% ethanol [@carter2016turtle; @carter2018evidence].
+This CORT dose increased mean yolk CORT levels 2 standard deviations higher than those of control eggs in previous studies [@crino2024eggs].
+Control treatment eggs received an equal volume of 100% Ethanol.
+Then, all eggs were incubated in one of the two previously mentioned temperature regimes (Cold or Hot) until hatching.
 These temperatures represent the upper and lower limits of the natural incubation temperatures for this species [@qualls2000post].
@@
 
-Lizards were acclimatised to the experimental arenas for two weeks before starting the numerical discrimination task (see below).
+Lizards were acclimatised to the experimental arenas for two weeks before starting the experiments (see below).
 The arenas were individual medium-size (41 L x 29.7 W x 22 H cm) plastic containers provided with a shelter (9 L x 6 W x 1.5 H cm) on one of the sides and a water dish in the center (@fig-Methods B).
@@
 Each choice was placed on one of the two lateral platforms of a white 3D printed device (see @fig-Methods B, C) and then covered with a transparent Petri dish to prevent the lizard from obtaining the food.
-We performed five tests that differed in the ratio between both choices of crickets.
-Specifically, we compared: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
-Each numerical test was made only one per lizard.
-The food consisted of small thawed frozen crickets (*A.
-domestica*) that were dusted with calcium and multivitamins.
+We performed five numerical discrimination tests per lizard, each differing in the ratio between both choices of crickets: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4 crickets.
+Each numerical discrimination test was made only once per lizard.
+We counterbalanced the order of testing by ensuring that, each day, an equal number of individuals from each treatment were tested in each type of trial.
+We controlled for potential side biases by placing the higher number of crickets randomly on the right or the left platform.
+The food consisted of small thawed frozen crickets (_A.
+domestica_) that were dusted with calcium and multivitamins.
 We decided to use thawed frozen crickets to avoid movement that would affect lizard behaviour [@cooper2024tell].
-Lizards were habituated to eating thawed frozen crickets for one month before the experiments.
+After acclimatisation to the arenas and before the numerical tests, lizards were habituated to eating thawed frozen crickets for one month before the experiments.
 
@@
 We tested if the length occupied by the crickets was similar in both choices (see _Supplementary Material: Control of size in both options_).
-Our results showed that except for the 1 VS 4 and 1 VS 3 tests, the difference in length between both choices was not different from zero.
+this analysis showed that except for the 1 VS 4 and 1 VS 3 tests, the difference in length between both choices was not different from zero.
 As such, discrimination between both choices in the rest of the tests (2 VS 4, 2 VS 3, 3 VS 4 crickets) would be produced by the OFS, not the ANS [@agrillo2014spontaneous].
-In addition, after the experiments, we conducted another set of trials to test whether the orientation of the crickets could affect lizards’ choice, but we found no effect of orientation (see _Supplementary Material: Prey orientation test_).
+In addition, after the experiments, we subjected the same lizards to another set of trials to test whether the orientation of the crickets could affect lizards’ choice, but we found no effect of orientation (see _Supplementary Material: Prey orientation test_).
 
@@
 The side where the food was placed was changed every day to avoid any side bias during the trials.
-However, to further control for side biases, the higher number of crickets was randomly placed on the right or the left platform.
-We also counterbalanced the number of individuals per treatment tested in the same type of trial each day.
 
@@
 Each trial was performed daily for five days between the 17^th^ and 22^nd^ of May 2024.
+Every trial began when we left the platform with the crickets in the enclosure, and lasted for 90 minutes until we took the platform away.
 We recorded three main variables: i) _Latency_, defined as the time between the start of the test - when the platform was left in the enclosure - until the first interaction with one of the options; ii) First choice of the lizard (_Choice_), recorded as 1 if the lizard interacted first with the larger amount or 0 if not; and iii) _Interest_ shown for the higher amount of food (detailed below).
-We considered that the lizard interacted with the food when it touched the Petri dish with its snout or when the front half of the body was on top of the platform.
-After the first interaction with any of the options, we recorded the time spent interacting with that option for a maximum of five minutes.
-We recorded this time for both amounts of food independently of the first choice.
-We measured the interest shown for the higher amount of food (_Interest_) by subtracting the time spent interacting with the smaller food option from the time spent interacting with the larger one.
+Interaction was defined as touching the Petri dish with the snout or having the front half of the body on the platform.
+For each option, and regardless of _Latency_, we recorded the time the lizard spent interacting with it for five minutes after the first interaction.
+The timer was not paused if the lizard walked away, so the recorded time reflects total engagement with each option during the first five minutes following initial contact.
+We calculated _Interest_ by subtracting the time spent interacting with the smaller food option from the time spent interacting with the larger one.
 If there was no interaction with one of the options during the test, we considered the time spent as zero for that option.
-_Latency_, _Choice_, and _Interest_ were noted as NA when there was no interaction with any of the choices.
-We also considered _Interest_ as NA in a few cases where the animals managed to lift the dish and eat the prey; _Latency_ and _Choice_ were recorded and employed in the analyses.
-We excluded from the analyses all the individuals that did not make any choice in three or more tests.
+_Latency_, _Choice_, and _Interest_ were noted as NA when there was no interaction with any of the choices (n = 4 observations).
+We also considered _Interest_ as NA in a few cases where the animals managed to lift the dish and eat the prey (n = 7 observations); _Latency_ and _Choice_ were recorded and employed in the analyses.
+We excluded from the analyses all the individuals that did not make any choice in three or more tests (n = 0 individuals excluded).
 All the analyses were by one observer (PR), who was blind to the treatment of the lizards, but not the type of test or the number of crickets because the videos were analysed manually.
 
-Motivation plays a key role in animals' performance in spontaneous choice tests [@agrillo2014spontaneous; @bisazza2014experim].
-To ensure that lizards remained motivated to forage, they were provided only one cricket during acclimatisation and the numerical task (see above).
+Motivation plays a key role in animals' performance in spontaneous choice tests [@bisazza2014experim; @gazes2018impact].
+To ensure that lizards remained motivated to forage, they were provided only one cricket per day during acclimatisation and the numerical task (see above).
 During acclimatisation, lizards were fed at the same time each day as the experiments were conducted.
@@
 We included test (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4), hormone (CORT versus Control), incubation temperature (Cold versus Hot), and the three-way interaction as fixed factors.
-We also included sex, and age as predictors.
-We did not find any effect of sex or age on Choice or Interest, but we found an effect on Latency (see Results).
-Regardless, we included both factors in all models.
-Age was centered to zero in the model, and we averaged the predicted values for males and females.
-
-We included lizard identity and clutch as random factors.
+We also included sex, and age (days since hatching) as predictors.
+Age was centered to the mean to facilitate interpretation of model estimates, and posterior predictions were averaged across sexes to provide estimates independent of the reference category.
+
+We included lizard identity and clutch identity as random factors.
 _L.
@@
 
-We recorded and analysed a total of 400 videos from 80 lizards (n = 20 per treatment, see @fig_results).
+We recorded and analysed a total of 400 videos from 80 lizards (n = 20 per treatment, see @fig-results).
 The lizards were between `r min(data_num$age_raw)` and `r max(data_num$age_raw)` days post-hatching (see Fig.
@@
 
-Neither temperature, CORT, or their interaction affected performance in any of the tests, nor were there any clear patterns suggestive of an effect of the developmental environment on numerical discrimination (see Table 1 and @fig-results).
+Neither temperature, CORT, or their interaction affected _Choice_ or _Interest_ in any of the tests, nor were there any clear patterns suggestive of an effect of the developmental environment on numerical discrimination (see Table 1 and @fig-results).
 Except for Hot-CORT lizards in 1 VS 3 tests or Cold-Control lizards in 2 VS 3 tests, the predicted probability of choosing the larger number of crickets first was not different from 0.5 (@fig-results; but also see Table S1 and Table S2 in _Supplementary Material_).
@@
 #### Foraging context may influence spontaneous quantity discrimination in _L. guichenoti_  
-The lack of numerical discrimination in food-related spontaneous tests in _L.
-guichenoti_ may be due to their insectivorous diet.
-In studies where reptiles discriminated between different amounts of food, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large numbers or alive prey (e.g., 5 vs.
-10) [@recio_prey_2021].
-In contrast, when _Podarcis siculus_ were presented with different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities despite showing quantity discrimination when prey size was the main cue [@miletto2017quantitative] or when tested using training procedures [@miletto2018quantity].
-Notably, the numbers of food items employed in @miletto2017quantitative were always fewer than four.
-Spontaneous choice tests assume that animals should select the most appropriate quantity, which may not necessarily be the largest one [@agrillo2014spontaneous] if the costs of selecting the larger option outweigh the benefits.
-For example, if capturing larger prey demands more time or energy or success rates remain similar regardless of prey quantity, animals may not prefer larger quantities [@agrillo2014spontaneous].
-Similarly, animals would not show any preference towards larger number of food items if the likelihood of acquiring one item is the same regardless of quantity [@agrillo2014spontaneous].
-As a result, even if diet might not affect quantity discrimination abilities, insectivorous lizards might not have evolved a preference for larger patches of prey, at least with small numbers [see @recio_prey_2021].
+We found no evidence of numerical discrimination in any of the tests performed.
+One potential concern is that pre-test fasting could have caused indiscriminate foraging masking numerical discrimination, but it seems unlikely.
+First, if hunger alone were driving behaviour, we would expect random first choices but clear preferences during subsequent interactions, reflected in the discrimination index.
+Second, only 30 and 36 out of 400 observations showed lizards interacting with the greater or fewer option, respectively, for more than 250 seconds (out of 300 s).
+Together, these results suggest that hunger levels were not masking numerical discrimination.
 
@@
 
+The lack of spontaneous numerical discrimination in _L.
+guichenoti_ when presented with small quantities of food may be related to their insectivorous diet.
+Studies in which reptiles successfully discriminated small food quantities typically involved vegetables as stimuli [@gazzola2018continuous; @szabo_spontaneous_2021; @szabo2024spontaneous].
+In contrast, insectivorous reptiles tested in spontaneous choice tasks with fewer than four prey items generally failed to discriminate [@miletto2017quantitative].
+For example, _Podarcis siculus_ did not distinguish between small numbers (never more than four) of _Musca domestica_ larvae in a spontaneous choice test, although the same species was capable of quantity discrimination when prey size was the primary cue [@miletto2017quantitative] or when tested using a training paradigm [@miletto2018quantity].
+These results suggest that the inability to discriminate in spontaneous tests is not due to a lack of cognitive capacity, but rather to decision-making strategies shaped by diet.
+In insectivorous lizards, the actual probability of obtaining a single prey item may be similar regardless of whether one or several items are present.
+Consequently, even though insectivorous lizards can discriminate small quantities under other conditions [see @recio_prey_2021], they may not use numerical cues in foraging decisions with small prey, as selecting larger numbers may not be advantageous.
+
+_Lampropholis guichenoti_ is described as a generalist insectivore that locates prey through a combination of active foraging and sit-and-wait behaviour [@lunney1989diets; @mo2021prey].
+Although handling time may vary with prey size, field observations suggest relatively slow feeding rates.
+For example, @martin2015selective reported that when feeding ant alates, individuals consumed an average of only two alates every 15 minutes even when the prey was abundant.
+In such situations, the difference in payoff between small prey group sizes (e.g., 1 vs.
+4) may be constrained by capture probability and effort, reducing the selective advantage of relying on numerical cues in foraging decisions.
+
 #### Foraging behaviour is robust to early environmental experiences 
 Although the effects of early environment on cognition are widely studied, little attention has been paid to its impacts on numerical discrimination and decision-making.
-We found that decision-making was not influenced by CORT elevations or incubation temperature, as latency to make a choice did not vary across prenatal conditions, which suggests that _L.
+We found that teh time to make a decision (_Latency_) was not influenced by CORT elevations or incubation temperature, which suggests that _L.
 guichenoti_ may be robust to these early-life factors, at least in the context of foraging decisions.
@@
 And to control for motivation, we standardised hunger levels by providing only one cricket per day and fasting lizards for over 20 hours before testing.
-Additionally, we excluded trials where lizards failed to interact with either option or refused to eat the cricket provided after testing.
-These measures likely minimised variation in decision latency due to differences in exploration, neophilia, or hunger levels, which could explain why we did not detect an effect of prenatal conditions.
+Additionally, we observed that all lizards consumed the cricket immediately after the test, showing similar levels of motivation.
+These measures likely minimised variation in decision time due to differences in exploration, neophilia, or hunger levels, which could explain why we did not detect an effect of prenatal conditions.
 Interestingly, we found that males and older lizards made decisions faster, although these factors were not central to our main hypotheses.
-However, it is possible that sex and age interact with early-life conditions in shaping behavioral responses.
-Future research should investigate potential interactions between developmental environment, sex, and age to better understand how these factors influence decision-making in reptiles.
+Previous work suggests weak effects of sex on cognition in reptiles [@szabo2019sex], while age may have a stronger influence [@noble_age-dependent_2014].
+Nevertheless, further studies are needed to explore how age and sex interact with early-life conditions in shaping behavioral responses.
 
@@
 However, since none of the groups showed any kind of quantity discrimination in our tests, we cannot determine whether prenatal conditions affected the lizards' numerical abilities or not; and we cannot discount the possibility that early environmental factors may affect quantity discrimination under other approaches or contexts.
-Nonetheless, to the best of our knowledge there is only one study examining how developmental conditions influence numerical abilities [@vila_pouca_quantity_2019].
-Vila-Pouca et al (2019) found that Port Jackson sharks (_Heterodontus portusjacksoni_) incubated at warmer temperatures learned a numerical rule faster than cold-incubated individuals.
-Given the potent effects of early environment on cognition and brain development [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic], it is crucial to explore how early environmental conditions shape numerical abilities across taxa.
+For instance, testing numerical discrimination using learning paradigms yielded different results than spontaneous choice tests in _P.
+siculus_ [@miletto2017quantitative; @miletto2018quantity].
+In fact, the only study we found that examined how developmental conditions influence numerical abilities showed that Port Jackson sharks (_Heterodontus portusjacksoni_) incubated at warmer temperatures learned a numerical rule faster than cold-incubated individuals [@vila_pouca_quantity_2019], but discrimination in a spontaneous test was not investigated.
+Given the potent effects of early environment on cognition and brain development [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic], it is crucial to explore how early environmental conditions shape numerical abilities across taxa employing different methodologies.
 
@@
 Our findings suggest that _L.
-guichenoti_ does not rely on numerical information for foraging decisions and that their performance and decision-making do not vary due to prenatal temperature or CORT exposure.
-Our findings suggest that decision-making can be robust to early environmental conditions, at least in the context of foraging.
+guichenoti_ does not rely on numerical information for foraging decisions, which limits the potential for early-life conditions to influence this specific ability.
+However, our results show that decision time does not vary due to prenatal temperature or CORT exposure, suggesting that this aspect of decision-making can be robust to early environmental conditions, at least in the context of foraging behaviour.
 
@@
 guichenoti_.
+For example, employing learning paradigms where animals must learn to follow a specific numerical rule can yield different results.
 Therefore, we recommend that future studies employ a broader array of tests and paradigms to more comprehensively assess numerical competence in this and related species.

### 2025-09-29 ms.qmd
updates

@@
 However, we predicted that numerical discrimination would be affected by early-life conditions.
-Specifically, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task as previous studies have reported enhanced cognitive abilities ectotherms incubated under warmer conditions [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
+Specifically, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task as previous studies have reported enhanced cognitive abilities in ectotherms incubated under warmer conditions [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
 Furthermore, we predicted that increased CORT levels would impair numerical discrimination [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
@@
 Eggs were assigned a clutch number and an individual identity.
-Then, we treated eggs with CORT or a vehicle control (see _Manipulating Early Thermal and CORT Environments_ details below)and placed each egg in individual cups (80 mL) with moist vermiculite (12 g water to 4 g vermiculite) and covered the cups with plastic wrap to retain moisture.
+Then, we treated eggs with CORT or a vehicle control (see _Manipulating Early Thermal and CORT Environments_ details below) and placed each egg in individual cups (80 mL) with moist vermiculite (12 g water to 4 g vermiculite) and covered the cups with plastic wrap to retain moisture.
 Each egg was incubated individually in incubators (LATWIT 2X5D-R1160) programmed to two different thermal regimes (see _Manipulating Early Thermal and CORT Environments_ details below) until hatching.

### 2025-11-25 ms.qmd
updates

@@
 However, we predicted that numerical discrimination would be affected by early-life conditions.
-Specifically, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task as previous studies have reported enhanced cognitive abilities in ectotherms incubated under warmer conditions [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
+Specifically, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task as previous studies have reported enhanced cognitive abilities of ectotherms incubated under warmer conditions [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019].
 Furthermore, we predicted that increased CORT levels would impair numerical discrimination [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures.
@@
 
-*Hatchlings* -- Hatchlings were placed in individual enclosures immediately after hatchling (18.7L x 13.2W x 6.3H cm) with non-stick matting and a small water dish and until the acclimatisation period (see below).
+*Hatchlings* -- Hatchlings were placed in individual enclosures immediately after hatching (18.7L x 13.2W x 6.3H cm) with non-stick matting and a small water dish and until the acclimatisation period (see below).
 During this time, they were given water daily and received 3-6 small _A.
@@
 One potential concern is that pre-test fasting could have caused indiscriminate foraging masking numerical discrimination, but it seems unlikely.
-First, if hunger alone were driving behaviour, we would expect random first choices but clear preferences during subsequent interactions, reflected in the discrimination index.
+First, if hunger alone were driving behaviour, we would expect random first choices but clear preferences during subsequent interactions, reflected by differences in the variable interest.
 Second, only 30 and 36 out of 400 observations showed lizards interacting with the greater or fewer option, respectively, for more than 250 seconds (out of 300 s).
@@
 All data, data description, and R code are available in public repository [https://github.com/Pablo-Recio/CORT-Temp_Numerical](https://github.com/Pablo-Recio/CORT-Temp_Numerical).
+Data is also available at [Dryad](https://doi.org/10.5061/dryad.08kprr5gv)
 
